### 2017-07-19 RegressionPvNP.Rmd

---
title: "RegressionPvNP"
author: "Sarah Killcoyne"
date: "22 June 2017"
output: 
  html_document: 
    fig_height: 7
    fig_width: 5
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
library(biomaRt)
library(ggplot2)
library(ggrepel)
library(ggfortify)
library(GGally)
library(plyr)
library(dplyr)
library(pander)
library(reshape2)
library(gridExtra)
library(GenomicRanges)
library(glmnet)
library(CoxHD)
library(mice)

source('lib/load_patient_metadata.R')

includeDemo = FALSE # demographics

data = '~/Data/Ellie'

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

load(grep('Test_patients.Rdata', analysis.files, value=T), verbose=T)
validation.patient.data = patient.data
length(validation.patient.data)

dataset = 'Training'
load(grep('Training_patients.Rdata', analysis.files, value=T), verbose=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

patient.info = read.patient.info(patient.file, demo.file, set=dataset)$info
head(patient.info)

validation.patient.info = read.patient.info(patient.file, demo.file, set='Test')$info
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.validation.info = summarise.patient.info(validation.patient.info)

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

if (length(which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)) > 0)
  warning(paste("Patients missing data:",
                paste(names(which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)), collapse=', ')))

if (length(which(sapply(validation.patient.data, function(df) nrow(df$seg.vals)) == 0)) > 0)
  warning(paste("Patients missing data:",
                paste(names(which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)), collapse=', ')))

nor = which(sapply(patient.data, function(df) nrow(df$seg.vals)) == 0)
if (length(nor) > 0) patient.data[nor] = NULL

nor = which(sapply(validation.patient.data, function(df) nrow(df$seg.vals)) == 0)
if (length(nor) > 0) validation.patient.data[nor] = NULL

sum.patient.data = as.data.frame(subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))) ## For now

# Missing some of the samples as they aren't all sequenced yet
for (pt in names(patient.data)) {
  #print(pt)
  #print(patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)])
  patient.data[[pt]]$info = patient.data[[pt]]$info[patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)],]
}

for (pt in names(validation.patient.data)) {
  #print(pt)
  #print(patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals)[-(1:5)])
  validation.patient.data[[pt]]$info = validation.patient.data[[pt]]$info[validation.patient.data[[pt]]$info$Samplename %in% colnames(validation.patient.data[[pt]]$seg.vals)[-(1:5)],]
}

#ct = sapply(patient.data, function(pd) nrow(pd$seg.vals) )
#hist(ct)

#lens = sapply(patient.data, function(pd) mean(pd$seg.vals$end.pos - pd$seg.vals$start.pos) )
#hist(lens)

#vrs = unlist(lapply(patient.data, function(pd) apply(as.matrix(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))]), 2, var) ))
#hist( vrs )
#pd$seg.vals[,grep('^D', colnames(pd$seg.vals))] - apply(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))], 2, mean)

pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

pander(sum.validation.info[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Validation patient set')
```

The previous analysis employed AP clustering on a per-patient basis to look for evidence that it was possible to see differences between the progressors and non-progressors, and to quantify that in some sort of complexity measure.  It was able to separate them, and continued to do so after removing HGD samples, but quickly failed to model any difference as I removed timepoints from the patient.

# Note about the data

This report uses all of the genomic data as well as encoded demographic information for age at diagnosis, BMI, p53 staining status, smoking, gender, length and circumference of the Barrett's segment.  None of these appear to improve the model.  This may not be surprising as most of these were controlled for in generating the cohort.

# Encode the large data matrix

To use GLMs across the patients I merge all samples from all patients and overlap the copy number segments. Where segments from a sample get split, the value of the original sample is retained for each of the split samples.

```{r tile, echo=F, message=F, warning=F, include=F}
tile.w=5e06
tile.files = grep(as.character(tile.w), analysis.files, value=T, fixed=T)
load(tile.files[grep('Test|validation', tile.files)], verbose=T)
mergedDfValidation = mergedDf
dim(mergedDfValidation)

load(tile.files[grep('Test|validation', tile.files, invert=T)], verbose=T)
dim(mergedDf)

cache.dir = paste(data, 'Analysis',sub('\\+', '', tile.w), sep='/')
if (!dir.exists(cache.dir)) dir.create(cache.dir)
```

Using a very large segment size for tiling across all patients (`r tile.w`) I get the following binomial models.  None of them fit well, and decreasing the size of the segments results in poor (or no) fits with mostly NA coefficients, or a lot of P(0 or 1).

# y = Progressors (1) vs Non (0)

The labels are split such that all samples from progressor patients are (1) and all samples from non-progressors are (0).

```{r labelsPNP, echo=F, message=F, include=T}
## binomial: dysplasia 1, BE 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

pts = do.call(rbind, lapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  cbind(df$info[,c('Hospital.Research.ID','Samplename')])
}))
rownames(pts) = 1:nrow(pts)

# sort in label order
if (length(setdiff(colnames(mergedDf), names(labels))) > 3)
  warning("Labels vector is missing samples")

dysplasia.df = t(mergedDf[,names(labels)])
dim(dysplasia.df)

unit.var <- function(x) {
  if (sd(x) == 0) return(x)
  return((x-mean(x))/sd(x) )
}

#dysplasia.df = apply(dysplasia.df, 2, function(x) x-median(x))
dysplasia.df = apply(dysplasia.df, 2, unit.var)

```

```{r addDemoCols, echo=T, message=F, warning=F}

  ## Some sort of complexity score
  #plot(unlist(complexity.measures), col=sum.patient.data$Status)
  # t.test(
  # complexity.measures[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID],
  # complexity.measures[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID]
  # )

  score.cx <-function(pd) {
    samples = grep('^D', colnames(pd$seg.vals), value=T)
    len.var = var(pd$seg.vals$end.pos - pd$seg.vals$start.pos) # length won't vary between samples

    #extremes = apply(as.matrix(pd$seg.vals[,samples]), 2, function(col) {
    #  length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
    #})

    num.var.reg = vector(mode='integer', length=length(samples))
    if (!is.null(pd$variable.regions) && nrow(pd$variable.regions) > 0) {
      num.var.reg = apply(as.matrix(pd$variable.regions[,samples]), 2, function(col) {
        length(which(col >= 1.1 | col <= 0.9))
      })
    }

    sample.var = apply(as.matrix(pd$seg.vals[, samples]), 2, var)
    t(cbind(len.var, num.var.reg, sample.var))
  }

  complexity.measures = lapply(patient.data, score.cx)

  dysplasia.df = cbind(dysplasia.df, 'segment.length'=0, 'num.var.regions'=0, 'sample.variance'=0)
  for (pt in names(complexity.measures)) {
    for (sample in colnames(complexity.measures[[pt]])) {
      dysplasia.df[sample,c('segment.length', 'num.var.regions', 'sample.variance')] = as.numeric(complexity.measures[[pt]][,sample])
    }
    #for (sample in rownames(dysplasia.df)) {
      #dysplasia.df[sample, ] = 
        #complexity.measures[subset(patient.info, Samplename == sample)$Hospital.Research.ID]
    #}
  }
  # norm to 1 to keep within the magnitude of the genomic data
  scale.to.1<-function(x) {
    (x-min(x))/(max(x)-min(x))
  }

  dysplasia.df[,c('segment.length', 'num.var.regions', 'sample.variance')] = apply(dysplasia.df[,c('segment.length', 'num.var.regions', 'sample.variance')], 2, scale)

  #cx = dysplasia.df[,'cx.measure'] 
  #dysplasia.df[,'cx.measure'] = (log(cx)/log(max(cx)))
  #dysplasia.df[,'cx.measure'] = dysplasia.df[,'cx.measure'] - median(dysplasia.df[,'cx.measure'])
  #dysplasia.df[,'cx.measure'] = ((cx - min(cx))/(max(cx)-min(cx)))

if (includeDemo) {

  ## Add in demographics

  # sort by samples in the order of the matrix
  patient.info = patient.info[match(rownames(dysplasia.df), patient.info$Samplename),]

  patient.info[which(patient.info$BMI > 100), 'BMI'] = NA # I think this one is wrong

  demo.data = patient.info[,c('Sex','Circumference','Maximal', 'Age.at.diagnosis', 'BMI', 'p53.Status', 'Smoking')]
  demo.data$Sex = as.integer(demo.data$Sex)-1
  demo.data$Smoking = as.integer(demo.data$Smoking)-1
  demo.data$p53.Status = as.integer(demo.data$p53.Status)-1
  demo.cols = c('Sex','C','M','Age', 'BMI', 'p53.Status', 'Smoking')

  encode.age <- function(x) {
    if (x > 70) { 
      (x - 70)/6 
    } else if (x < 50) {
      (x -50)/6
    } else {
      0
    }
  }
  #demo.data$Age.at.diagnosis = scale(demo.data$Age.at.diagnosis)

  # Encode based on presumed distributions or unit normalize
  demo.data = demo.data %>% rowwise() %>% mutate( 'age.encoded'= encode.age(Age.at.diagnosis))
  demo.data$Age.at.diagnosis = demo.data$age.encoded
  demo.data$age.encoded = NULL
  demo.data$BMI = scale(demo.data$BMI)

  demo.data[,c('Circumference','Maximal')] = apply(demo.data[,c('Circumference','Maximal')], 2, function(x) (x-mean(x, na.rm=T))/sd(x, na.rm=T))

  ## Impute missing data
  imp = mice(as.matrix(demo.data), diagnostics = F)

  dysplasia.df = cbind(dysplasia.df, 'Sex'=NA,'C'=NA,'M'=NA,'Age'=NA, 'BMI'=NA, 'p53.Status'=NA, 'Smoking'=NA)
  dysplasia.df[,demo.cols] = as.matrix(complete(imp))

}
```

We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.

Example subset of the data matrix:
`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

## Build/Cross-validate by patient

Keeping in mind that the matrix is built on a sample, not patient, basis while the labels (Progressor/Non) are on a per-patient basis I generated the cross validated models for various values of alpha by running 10 or more iterations of 5-fold cross validation that pulled out all the samples for 5 patients at each fold.  1000 lambda values.

```{r cv.funcs, message=F, warning=F, echo=T, fig.height=6, fig.width=6}
pi.hat<-function(x) exp(x)/(1+exp(x))

non.zero.coef<-function(fit, s) {
  cf =  as.matrix(coef(fit, s))
  cf[which(cf != 0),][-1]
}

coef.stability<-function(opt, nz.list) {
  for (i in 1:length(nz.list)) {
    df = as.data.frame(nz.list[[i]])
    colnames(df) = i
    opt = merge(opt, df, by='row.names', all.x=T)
    rownames(opt) = opt$Row.names
    opt$Row.names = NULL
    opt[,(i+1)] = as.integer(!is.na(opt[,(i+1)]))
  }
  opt = opt[order(sapply(rownames(opt), function(x) as.numeric(unlist(strsplit(x, ':'))[1])  )),]
  return(opt)
}

create.patient.sets<-function(pts, n, splits, minR=0.2) {
  # This function just makes sure the sets don't become too unbalanced with regards to the labels.
  check.sets<-function(df, grpCol, min) {
    sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
    while ( (length(which(sets/rowSums(sets) < minR) ) >= 2 | length(which(sets/rowSums(sets) == 0)) > 0) ) {
      #print(sets/rowSums(sets))
      s = sample(rep(seq(5), length = length(unique(df$Hospital.Research.ID))))
      df2 = merge(df, cbind('Hospital.Research.ID'=unique(df$Hospital.Research.ID), 'tmpgrp'=s), by="Hospital.Research.ID")
      df[[grpCol]] = df2$tmpgrp
      sets = table(cbind.data.frame('set'=df[[grpCol]], 'labels'=labels[df$Samplename]))
    }
    return(df[[grpCol]])
  }

  s = sample(rep(seq(splits), length = length(unique(pts$Hospital.Research.ID))))
  patients = merge(pts, cbind('Hospital.Research.ID'=unique(pts$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
  colnames(patients)[3] = c('fold.1')
  patients$fold.1 = check.sets(patients, 'fold.1', minR)

  for (i in 2:n) {
    s = sample(rep(seq(5), length = length(unique(pts$Hospital.Research.ID))))
    patients = merge(patients, cbind('Hospital.Research.ID'=unique(patients$Hospital.Research.ID), 'group'=s), by="Hospital.Research.ID")  
    foldcol = grep('group',colnames(patients))
    colnames(patients)[foldcol] = paste('fold',i,sep='.')
    patients[[paste('fold',i,sep='.')]] = check.sets(patients, paste('fold',i,sep='.'))
  }

  return(patients)
}

binomial.deviance<-function(pmat, y) {
  # Binomial deviance, lifted from cv.lognet
  prob_min = 1e-05; prob_max = 1 - prob_min
  pmat = pmin(pmax(pmat, prob_min), prob_max)
  dev = apply(pmat, 2, function(p)  -2*((y==1)*log(p)+(y==0)*log(1-p)) )
  return(dev)
}

crossvalidate.by.patient<-function(x,y,lambda,pts,a=1,nfolds=10, splits=5, fit=NULL, minR=0.2, select='deviance', ...) {
  if (nfolds > 5) minR = 0.1
  x = as.matrix(x)
  message(paste("Running", splits, "splits",nfolds,"times on", paste(dim(x), collapse=':'), 'alpha=',a ))
  fit.e = list()

  if (ncol(pts) == (nfolds+2)) {
    tpts = pts
  } else {
    tpts = create.patient.sets(pts, nfolds, splits, minR)
  }

  cv.pred = (matrix(nrow=0, ncol=length(lambda)))
  cv.binomial.deviance = (matrix(nrow=0, ncol=length(lambda)))
  for (n in 1:nfolds) {  
    message(paste(n, "fold"))
    setCol = grep(paste('^fold.',n,'$',sep=''), colnames(tpts))

    cv.class = matrix(nrow=splits, ncol=length(lambda))
    deviance = matrix(nrow=splits, ncol=length(lambda))
    for (i in 1:splits) { 
      message(paste(i, "split"))
      test.rows = which(rownames(x) %in% tpts[which(tpts[,setCol] == i), 'Samplename'])
      test = x[test.rows,]
      training = x[-test.rows,]
      # pre-spec lambda seq
      fitCV = glmnet(training, y[-test.rows], lambda=lambda, alpha=a, family='binomial', ...) 
      # autoplot(fit) + theme(legend.position="none")

      # Confusion matrix: quantitiative, pos results + neg results / number of test rows
      pred <- pi.hat(predict(fitCV, test, type='link'))  # pi.hat turns these into probabilities from logit
      cv.class[i,] = apply(pred, 2, function(p) {
        (p%*%y[test.rows] + (1-p) %*% (1-y[test.rows]))/length(test.rows)
      })

      # Binomial deviance, lifted from cv.lognet
      dev = binomial.deviance(predict(fitCV, test, type='response'), as.factor(y[test.rows]))
      deviance[i,] = apply(dev, 2, weighted.mean, w=rep(1, nrow(dev)), na.rm=T)

      fit.e[[length(fit.e)+1]] = fitCV
    }
    cv.pred = rbind(cv.pred, cv.class)    
    cv.binomial.deviance = rbind(cv.binomial.deviance, deviance)
  }

  # Not really sure what to do with the binomial deviance now...
  df = cbind.data.frame('lambda.at'=1:ncol(cv.pred),
                        'mean'= colMeans(cv.pred), 
                        'sme'= apply(cv.pred, 2, sd)/sqrt(nrow(cv.pred)), 
                        'sd'= apply(cv.pred, 2, sd), 
                        'lambda'= lambda,
                        'log.lambda' = log(lambda),
                        'mean.b.dev' = colMeans(cv.binomial.deviance),
                        'sd.b.dev' = apply(cv.binomial.deviance, 2,sd),
                        'sme.b.dev' = apply(cv.binomial.deviance, 2, sd)/sqrt(nrow(cv.binomial.deviance)),
                        'lambda.min' = F, 'lambda.1se' = F)

  if (select == 'classification') {
    # Minimize the classification mean error & select min lambda
    df[which.max(subset(df, sme < median(sme) & sme > min(sme))$mean), 'lambda.min'] = T

    #df$lambda.min = df$mean == max(df$mean)
    se1 = df[df$lambda.min == T, 'mean'] - df[df$lambda.min == T, 'sme']

    lambda.1se.search = subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda) 
    if (nrow(lambda.1se.search) > 0) {
      arrange(lambda.1se.search, -mean)[1:10,]
      df[arrange(subset(df, mean < se1 & !lambda.min & log.lambda > subset(df, lambda.min)$log.lambda ), -mean)[1, 'lambda.at'], 'lambda.1se'] = T
    }
    #df[df$log.lambda < subset(df, mean == max(mean))$log.lambda-sd(df$log.lambda),][1, 'lambda-1se'] = TRUE
  } else if (select == 'deviance') {
    df[which.min(df$mean.b.dev), 'lambda.min'] = T

    lmin = subset(df, lambda.min)$lambda
    df[subset(df, log.lambda <= log(lmin) + log(sd(lambda))/2 )[1, 'lambda.at'], 'lambda.1se'] = T
  }

  lambda.min = lambda[subset(df, lambda.min == T)$lambda.at] 
  lambda.1se = lambda[subset(df, lambda.1se == T)$lambda.at]

  nzcf = lapply(fit.e, non.zero.coef, s=lambda.1se)

  plots = plot.patient.cv(df, fit)
  plots$performance = plots$performance + theme(legend.position='bottom') + scale_colour_discrete(name = "") +
    labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='mean pred.', x='log(lambda)') 
  plots$deviance = plots$deviance + theme(legend.position='bottom') + scale_colour_discrete(name="") + 
    labs(title=paste(splits,' splits, ',nfolds,' folds, alpha=',a, sep=''), y='mean Binomial Deviance', x='log(lambda)')

  return(list('max.cm'=df[df$lambda.1se ==T,'mean'], 
              'lambda.min'=lambda.min, 'lambda.1se'=lambda.1se, 'lambdas'=df, 'plot'=plots$performance, 'deviance.plot'=plots$deviance, 'non.zero.cf'=nzcf))
}

plot.patient.cv<-function(df, fit=NULL) {
  gp = ggplot(df, aes(y=mean,x=log.lambda)) + geom_point() + geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + 
    geom_point(data=subset(df, lambda.min == T), aes(y=mean, x=log.lambda, colour="lambda.min"), size=2 ) +  
    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
             label=paste(round(df[subset(df, lambda.min == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))

  gpD = ggplot(df, aes(y=mean.b.dev,x=log.lambda)) + geom_point() +
    geom_errorbar(aes(ymin=mean.b.dev-sme.b.dev, ymax=mean.b.dev+sme.b.dev)) + 
    geom_point(data=subset(df, lambda.min == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.min"), size=2 ) +          
    geom_vline(xintercept = subset(df, lambda.min == T)$log.lambda, colour='grey') +
    annotate("text", x=subset(df, lambda.min == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
             label=paste(round(df[subset(df, lambda.min == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))

  if (nrow(subset(df, lambda.1se == T)) > 0) {
    gp = gp + geom_point(data=subset(df, lambda.1se == T), aes(y=mean, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-'))

    gpD = gpD + geom_point(data=subset(df, lambda.1se == T), aes(y=mean.b.dev, x=log.lambda, colour="lambda.1se"), size=2 ) +
      geom_vline(xintercept = subset(df, lambda.1se == T)$log.lambda, colour='grey') +
      annotate("text", x=subset(df, lambda.1se == T)$log.lambda, y=min(df$mean)+sd(df$mean), 
               label=paste(round(df[subset(df, lambda.1se == T)$lambda.at,c('mean', 'sme')],3), collapse='\n+/-')) 
  }

  gp = gp + theme(legend.position='bottom') + scale_colour_discrete(name = "") 
  gpD = gpD + theme(legend.position='bottom') + scale_colour_discrete(name="")

  if (!is.null(fit)) {
    df$nzcoef = sapply(df$lambda, function(l) length(non.zero.coef(fit, l)))

    coef.min = subset(df, lambda.min == T)$nzcoef
    coef.1se = subset(df, lambda.1se == T)$nzcoef

    d = coef.1se-coef.min

    coef.text = arrange(subset(df,nzcoef %in% c(0,  coef.min, coef.1se, max(nzcoef) )), -lambda.min, -lambda.1se)
    coef.text = arrange(coef.text[!duplicated(coef.text$nzcoef),], -nzcoef)

    gp = gp + annotate("text", y=min(df$mean), x=coef.text$log.lambda, label=coef.text$nzcoef, color='darkblue')
    gpD = gpD + annotate("text", y=max(df$mean.b.dev), x=coef.text$log.lambda, label=coef.text$nzcoef, color='darkblue')
  }
  return(list('performance'=gp, 'deviance'=gpD))
}

```

```{r xvalpt, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
# dysplasia.df = cbind(dysplasia.df, 'sl.nvr'=0)
# dysplasia.df[, 'sl.nvr'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'num.var.regions'])
# 
# dysplasia.df = cbind(dysplasia.df, 'sl.nvr.sv'=0)
# dysplasia.df[, 'sl.nvr.sv'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'num.var.regions'] + dysplasia.df[, 'sample.variance'])
# 
# dysplasia.df = cbind(dysplasia.df, 'sl.sv'=0)
# dysplasia.df[, 'sl.sv'] = scale(dysplasia.df[, "segment.length"] + dysplasia.df[, 'sample.variance'])
# 
# dysplasia.df = cbind(dysplasia.df, 'nvr.sv'=0)
# dysplasia.df[, 'sl.sv'] = scale(dysplasia.df[, "num.var.regions"] + dysplasia.df[, 'sample.variance'])

nl = 1000

coefs = list(); plots = list(); performance.at.1se = list(); models = list();
folds = 10; splits = 5 
alpha.values = c(0, 0.5,0.7,0.8,0.9,1)

file = paste(cache.dir, 'patient_folds.tsv', sep='/')
if (file.exists(file)) {
  sets = read.table(file, header=T, sep='\t')
} else {
  sets = create.patient.sets(pts, folds, splits, 0.2)  
  write.table(sets, quote=F, sep='\t', row.names=F, file=file)
}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  for (a in alpha.values) {
    fit0 <- glmnet(dysplasia.df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F) # all patients
    #autoplot(fit0, xvar='lambda', main=paste('fit0, all samples, alpha=',a,sep='')) + theme(legend.position='none')

    l = fit0$lambda
    if (a == 0) {
      l = sort(c(fit0$lambda, seq(exp(-5), exp(-10), -1e-6),
            seq(exp(-10), exp(-15), -1e-8),
            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
    }
    cv.patient = crossvalidate.by.patient(x=dysplasia.df, y=labels, lambda=l, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', standardize=F)

    lambda.opt = ifelse( length(cv.patient$lambda.1se) > 0, cv.patient$lambda.1se, cv.patient$lambda.min )

    coef.opt = as.data.frame(non.zero.coef(fit0, lambda.opt))
    coefs[[as.character(a)]] = coef.stability(coef.opt, cv.patient$non.zero.cf)

    plots[[as.character(a)]] = arrangeGrob(cv.patient$plot+ggtitle('Classification'), cv.patient$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    performance.at.1se[[as.character(a)]] = subset(cv.patient$lambdas, lambda.1se == T)
    models[[as.character(a)]] = fit0
  }
  save(plots, coefs, performance.at.1se, dysplasia.df, models, file=file)
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))
```

```{r xvalpt2, message=F, warning=F, echo=T, fig.width=8, fig.height=6}

performance = do.call(rbind.data.frame, performance.at.1se)
performance$alpha = rownames(performance)

## check how often the feature(s) is selected at that lamda in each split. "stability selection"
coef.stable = lapply( coefs, function(cf) {
  sort(rowSums(cf[,-1]), decreasing=T)
})
coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))

performance = cbind(performance, do.call(rbind,  lapply(coef.stable, function(x) {
  cbind('n.Coef'=length(x), '25%'=length(which(x>=.25)), 
        '50%'=length(which(x>=.5)), '75%'=length(which(x>=.75 )), 
        '100%'=length(which(x==1)) )
})))

ggplot(performance, aes(alpha, mean)) + geom_text( aes(label=round(mean, 3)), hjust=-0.5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), color='grey') + geom_point() + ylim(0.5,0.8) +
  geom_text(aes(label=n.Coef), nudge_y=0.03) +
  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
  labs(x='Elasticnet penalty value: ridge <-> lasso', y='Model classification at lambda.1se', 
       title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))

cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=length(names(coef.stable)), 
                           dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
for (i in names(coef.stable)) 
  cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1

pander(performance[,c('n.Coef','25%','50%','75%', '100%')], justify='left', caption='Number of features stable in n% or more folds')

select.alpha = as.character(0.9)

# Genomic regions only
most.stable = grep('^\\d+', names(which(coef.stable[[select.alpha]] >= 0.75 )), value=T)

most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

most.stable$chr = as.integer(as.character(most.stable$chr))
most.stable = arrange(most.stable, chr, start)

```

Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at 0.8. They share all non-zero features as well.
`r pander(sapply(coef.stable, length), caption='Total features at each value of alpha')`

`r pander(most.stable[,c('chr','start','stop')], caption='Features identified as stable in 75% or more folds.')`

### Feature stability

`r pander(colSums(cfs), caption="Number of features found in 1, 2, or all 3 models by alpha 0.8, 0.9 and 1. While there are fewer features for higher values of alpha, the features are the same across models.")`

```{r echo=F, message=F, warning=F}
featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
```

```{r echo=F, warning=F, message=F, eval=F}
### Features in samples
## TODO Look at the coefficients, are the high pos ones "gains" and the low neg ones "losses" etc
plot(coefs[[select.alpha]][,1])
selected = coefs[[select.alpha]]

selected[which(
  selected[,1] > mean(selected[,1]) + sd(selected[,1]) | 
  selected[,1] < mean(selected[,1] - sd(selected[,1]))  
),]

## Variance in the selected features?  Not so much
coefs[[select.alpha]][,1]

v = vector(length(colnames(dysplasia.df)), mode='numeric')
names(v) = colnames(dysplasia.df)
v[rownames(selected)] = coefs[[select.alpha]][,1]

vv <- apply(dysplasia.df, 2, var)
plot(v, vv, log='y', col=(v!=0)+1)

boxplot(vv ~ (v==0), log='y', ylim=c(0.0001,1))

qqplot(vv[v==0], vv[v!=0], log='xy')

length(which( coef.stable[[select.alpha]] >= 0.5) )

regions = rownames(coefs[[select.alpha]])

pg = dysplasia.df[ labels == 1 ,regions]

nonpg = dysplasia.df[ labels == 0 ,regions]

colnames(pg)

meantests = sapply(regions, function(r)   t.test(pg[,r], nonpg[,r])$p.value )

which(meantests < 0.05)

#summary(pg[,'1:119640299-129610322'])
#summary(nonpg[,'1:119640299-129610322'])

```

### Feature Information

Using the features from the alpha=`r select.alpha` models

```{r, warning=F, message=F, eval=T}
#topF = names(which((coef.stable[['1']]) >= 0.75))
topF = names(coef.stable[[select.alpha]])

features.in.prog = dysplasia.df[names(which(labels == 1)), grep('^\\d+:', topF, value=T)]
features.nonprog = dysplasia.df[names(which(labels == 0)), grep('^\\d+:', topF, value=T)]

feat = which(colnames(dysplasia.df) %in% topF)

# Gain or loss not correlated in prog or non
genomeCols = grep('^\\d+:', colnames(dysplasia.df), value=T)
gl = cbind.data.frame(
  'P.gain' = apply(features.in.prog, 2, function(x) sum( as.integer(x > 0) ) ), 
  'P.loss' = apply(features.in.prog, 2, function(x) sum( as.integer(x < 0) ) ),
  'NP.gain' = apply(features.nonprog, 2, function(x) sum( as.integer(x > 0) ) ),
  'NP.loss' = apply(features.nonprog, 2, function(x) sum( as.integer(x < 0) ) ) )

allgl = cbind.data.frame(
  'ALL.P.gain' = apply(dysplasia.df[names(which(labels == 1)), genomeCols], 2, function(x) sum( as.integer(x > 0) ) ), 
  'ALL.P.loss' = apply(dysplasia.df[names(which(labels == 1)),genomeCols], 2, function(x) sum( as.integer(x < 0) ) ),
  'ALL.NP.gain' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x > 0) ) ), 
  'ALL.NP.loss' = apply(dysplasia.df[names(which(labels == 0)),genomeCols], 2, function(x) sum( as.integer(x < 0) ) ) )

grid.arrange(
  ggcorr(gl, label=T) + ggtitle('Selected Features'),
  ggcorr(allgl, label=T) + ggtitle('All regions'), 
top='Gains vs Losses')

# Gains/losses enriched in the top features?  Appears not
(fisher.test(rbind(colSums(gl[,c('P.gain','P.loss')]),colSums(allgl[-feat,c('ALL.P.gain','ALL.P.loss')])  )) )

(fisher.test(rbind(colSums(gl[,c('NP.gain','NP.loss')]),colSums(allgl[-feat,c('ALL.NP.gain','ALL.NP.loss')])  )) )

```

#### bp Length

```{r, echo=F, warning=F, message=F}
chromInfo = read.table('hg19_genes.txt', sep='\t', header=T)
chromInfo$chr = factor(chromInfo$chr, levels=c(1:22, 'X','Y'))

topFSegments = do.call(rbind,  strsplit( topF[grep('^\\d+:', topF)], ':|-' ))
colnames(topFSegments) = c('chr','start','end')
topFSegments = as.data.frame(apply(topFSegments, 2, as.numeric))
topFSegments$chr = factor(topFSegments$chr, levels=droplevels(chromInfo$chr[1:22]), ordered=T)
topFSegments = arrange(topFSegments, chr, start)

topFRegions = arrange(merge(chromInfo[,c('chr','length', 'protein_coding')], topFSegments %>% group_by(chr) %>% summarise('totalbp'=sum(end-start), 'n.segments'=length(chr)), all.x=T), -protein_coding)

# Number of segments is strongly correlated with the length of the chromosome
#with(topFRegions, cor.test(length, n.segments))

topFRegions$adj.info.content = topFRegions$protein_coding/(topFRegions$length^0.6)
#with(topFRegions, cor.test(protein_coding, length)) # Adjust the length so that the correlation between length and

topFRegions = transform(topFRegions, ai.cut=cut(adj.info.content, 5, labels=c(5:1), include.lowest=T))           # bin data

grid.arrange(
  #ggplot(topFRegions, aes(reorder(chr, -adj.info.content), totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
  ggplot(topFRegions, aes(chr, totalbp/length, fill=ai.cut)) + ylim(0, 0.5) +
    labs(y='Feature bp/total chr length', title='Ratio of covered base pairs', x='chr') +
    geom_bar(stat='identity') +  geom_text(aes(label=round(totalbp/length, 2)), vjust=-0.5) +
    scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom'), 
      bottom='Ratio of base pair coverage per chromosome in the top non-zero features\n selected from the model. Chromosomes are sorted\nand binned by their gene (e.g. information) content rather than by overall length.')

ct = with(topFRegions, cor.test(adj.info.content, totalbp/length))

with(topFRegions, cor.test(adj.info.content, totalbp/length))

bpCovered = with(topFRegions, sum(totalbp, na.rm=T)/sum(as.numeric(length)))
```

`r pander(ct, caption="Pearson correlation between the adjusted information content (genes) per chromosome, and the ratio of length of the features")`

`r round(bpCovered, 2)*100`% of the all bps in the genome are within the top selected features.

#### Genes?

Using Biomart get all genes that are fully within the the segment boundaries.

```{r, warning=F, message=F}
bm = NULL
tryCatch({
  mart = biomaRt::useMart("ENSEMBL_MART_ENSEMBL", host="grch37.ensembl.org")
  biomart = useDataset("hsapiens_gene_ensembl", mart)

  attr = c('ensembl_gene_id','hgnc_symbol','start_position','end_position', 'chromosome_name', 'ensembl_exon_id','percentage_gene_gc_content')

  bm = getBM(mart=biomart, attributes=attr, 
           filters=list('chromosomal_region'=sub('-', ':',topF[grep('^\\d+:', topF)]), "biotype"=c("IG_C_gene","IG_D_gene","IG_J_gene","IG_V_gene","protein_coding","TR_C_gene","TR_D_gene","TR_J_gene","TR_V_gene")))
#head(bm)

  bm = bm[-which(bm$hgnc_symbol == ""),]
  bm = bm %>% group_by(ensembl_gene_id, hgnc_symbol, chromosome_name, start_position, end_position) %>% summarise('exon_count' = length(ensembl_exon_id)) 
  exons = bm %>% group_by(chromosome_name) %>% summarise('exons'=sum(exon_count))

  write.table(bm$ensembl_gene_id, sep='\t', col.names = F, row.names = F, quote = F, file='feature_genes.txt')

  for (i in 1:nrow(topFSegments)) {
    s = topFSegments[i,]
    sbm = with(bm, which(chromosome_name == s[['chr']] & start_position >= s[['start']] & end_position <= s[['end']]))

    topFSegments[i,'genes'] = length(sbm)
    topFSegments[i,'exons'] = sum(bm[sbm,'exon_count'])

    bm[sbm, 'segment.index'] = i
  }

  # Not related to the length of the segment
  pander(with(topFSegments, cor.test(genes, (end-start))), caption="Number of genes per segment may be related to the segment length.")
}, error = function(e) {
  warning(paste("Failed to connect to Biomart service", e))
})
```

```{r echo=F, message=F, warning=F, eval=F}
#### Correlations
library(Hmisc)

flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

head(dysplasia.df)
rc = rcorr(dysplasia.df)
rc = flattenCorrMatrix(rc$r, rc$P)

rc = subset(rc, row %in% rownames(coefs[[select.alpha]]) | column %in% rownames(coefs[[select.alpha]]))
rc = subset(rc, p < 0.05)

#nrow(rc)
#head(rc)

pos = do.call(rbind.data.frame, strsplit(as.character(rc$column), ':|-'))
colnames(pos) = c('col.chr','col.start','col.end')
pos[] = lapply(pos[], function(x) as.numeric(as.character(x)))
rc = cbind(pos, rc)

pos = do.call(rbind.data.frame, strsplit(as.character(rc$row), ':|-'))
colnames(pos) = c('row.chr','row.start','row.end')
pos[] = lapply(pos[], function(x) as.numeric(as.character(x)))
rc = cbind(pos, rc)

rc$r.coef = as.integer(rc$row %in% rownames(coefs[[select.alpha]]))

#head(rc)

rc = subset(rc, cor >= 0.5)

rc = arrange(rc, row.chr, col.chr, row.start, col.start)
#head(rc)

#rc[sort(c(which(rc$r.coef == 1), which(rc$r.coef == 1)+1)),]

rc1  = subset(rc, row.chr == 16 & col.chr == 16)

rc1 = transform(rc1, row=reorder(row, -row.start) ) 
rc1 = transform(rc1, column=reorder(column, col.start) )

subset(rc1, column == '16:85599240-90354752' | row == '16:85599240-90354752')

#head(rc1)

ggplot(rc1, aes(column, row, fill=cor)) + geom_tile()  +  #scale_fill_gradient(low = "white", high = "red") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

## closer regions are more liklye to be correlated
cor.test(rc1$col.start - rc1$row.end , rc1$cor)

```

### Hazards

Using the model for alpha=`r select.alpha`.

```{r echo=F, warning=F, message=F,eval=T, fig.height=8, fig.width=8}

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})
ch = as.data.frame(hazards)
ch$coef = coefs[[select.alpha]][,1]
ch$label = rownames(ch)
ch$stability = featureStability

ggplot(ch, aes(coef, hazards, col=hazards > quantile(hazards)[['75%']])) + 
  #xlim(min(coefGR$coef)-5, max(coefGR$coef)+10) +
  geom_point(alpha=0.5) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards > quantile(hazards)[['75%']], label, '')),show.legend=F) +
  labs(title='Top features coef vs hazard') + theme(legend.position = 'bottom')
```

```{r echo=F, warning=F, message=F,eval=F, fig.height=12, fig.width=8}
ch = arrange(ch, -stability, -hazards, -coef)
ch$hazards = round(ch$hazards, 3)
ch$coef = round(ch$coef, 3)
write.table(ch[c('label', 'stability','hazards','coef')], row.names = F, quote=F, sep="\t", file='features.txt')

ch = arrange(ch, -hazards)

melt.by.reg<-function(x) {
  r = x[['label']]

  reg = cbind.data.frame( 'value'=dysplasia.df[ ,r], labels)
  reg$labels = factor(reg$labels)
  reg$labels = revalue(reg$labels, c("1"="P", "0"="NP"))

  tt = wilcox.test(dysplasia.df[ which(labels == 0) , r ], dysplasia.df[ which(labels == 1) , r ])

  reg$pval = tt$p.value
  reg$region = r
  reg$hazard = as.numeric(x[['hazards']])
  reg$coef = as.numeric(x[['coef']])
  return(reg)
}

plot.reg<-function(x) {
  reg = melt.by.reg(x)

  pv = ifelse(reg$pval[1] < .001, '< .001', format.pval(unique(reg$pval[1])))

#  ggplot(reg, aes(value, fill=labels) ) + geom_histogram(bins=50, color='grey') + 
#    labs(x='seg. value', title=r, subtitle=paste('haz=',x[['hazards']], ' coef=', x[['coef']], ' MW pval: ',pv,sep=''))
  ggplot(reg, aes(x=labels, y=value, group=labels, fill=labels) ) + geom_jitter(color='grey', show.legend = F) + geom_violin(show.legend = F, alpha=0.5) +
    geom_hline(yintercept = median(reg$value) + sd(reg$value)) + geom_hline(yintercept = median(reg$value) - sd(reg$value)) +
    labs(x='', y='seg. value', title=x[['label']], subtitle=paste('haz= ',x[['hazards']], ' coef= ',x[['coef']], '\nMW pval ',pv,sep=''))
}

ch = arrange(ch, -hazards)

regions = apply(ch, 1, melt.by.reg)
names(regions) = ch$label

do.call(grid.arrange, c(lapply(regions[1:15], function(reg) {
  print(reg$region[1])
  cols = c('indianred3','slateblue2')
  if ( as.numeric(reg$coef[1]) < 0) cols = c('green3', 'darkviolet')

  ggplot(reg, aes(labels, value*coef, fill=labels, color=labels)) + 
    geom_jitter(aes(labels, value), width=0.05, color='yellow4', alpha=0.3) +
    geom_jitter(aes(labels, value*coef, color=labels), width=0.05, alpha=0.5) +
    geom_text_repel(data=subset(reg, value >= quantile(reg$value)[['100%']]), aes(labels, value, label=round(value, 3)), color='yellow4') +
    geom_text_repel(data=subset(reg, value >= quantile(reg$value)[['100%']]), aes(labels, value*coef, label=round(value, 3),color=labels)) + geom_violin(alpha=0.2) + 
    scale_fill_manual(values=cols) + scale_color_manual(values=cols) +
    theme(legend.position = 'none') + labs(title=unique(reg$region)[1], x='')
})))

    #geom_violin(aes(labels, value, fill=labels), alpha=0.5) + theme(legend.position = 'none')

#  x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value > 0))),
#                               'loss' = length(which(with(subset(reg, labels == 'P'), value < 0))) ),

#                        'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value > 0))),
#                                'loss' = length(which(with(subset(reg, labels == 'NP'), value < 0))) ) )

    # x = rbind.data.frame('P'= cbind( 'gain' = length(which(with(subset(reg, labels == 'P'), value >= median(value) + sd(value)*se))),
  #                              'loss' = length(which(with(subset(reg, labels == 'P'), value <= median(value) - sd(value)*se))) ),
  # 
  #                 'NP' = cbind( 'gain' = length(which(with(subset(reg, labels == 'NP'), value >= median(value) + sd(value)*se))),
  #                               'loss' = length(which(with(subset(reg, labels == 'NP'), value <= median(value) - sd(value)*se))) ) )
  # x = rbind.data.frame('P'= cbind( 'high' = length(which(with(subset(reg, labels == 'P'), value >= 1.1))),
  #                                  'low' = length(which(with(subset(reg, labels == 'P'), value <= 0.9))) ),
  # 
  #                 'NP'= cbind( 'high' = length(which(with(subset(reg, labels == 'NP'), value >= 1.1))),
  #                                  'low' = length(which(with(subset(reg, labels == 'NP'), value <= 0.9))) ))

  #cols = c('indianred3','slateblue2')
  #if ( as.numeric(reg$coef[1]) < 0) cols = c('plum', 'skyblue3')

  #ggplot(melt(as.matrix(x)), aes(Var2, value, fill=Var1)) + facet_grid(~Var1) +
  #  geom_col( show.legend = F) + labs(x='', y='', title=unique(reg$region)[1]) + scale_fill_manual(values=cols)

#})))

## Individual hazards vs CN values?  exp(coef) -- odds ratio
# plot( exp( coefs[[select.alpha]][ch$label[1],1]  * dysplasia.df[,ch$label[1]]),  dysplasia.df[,ch$label[1]] )
# 
# plot(  exp( t(dysplasia.df[, ch$label[1]]) *  coefs[[select.alpha]][ch$label[1],1]  ), dysplasia.df[,ch$label[1]]  ) ## pos coef
# plot(exp(t(dysplasia.df[, ch$label[3]]) *  coefs[[select.alpha]][ch$label[3],1]), dysplasia.df[,ch$label[3]]) ## pos coef
# 
# plot(exp(t(dysplasia.df[, ch$label[2]]) *  coefs[[select.alpha]][ch$label[2],1]), dysplasia.df[,ch$label[2]]) ## neg coef
# plot(exp(t(dysplasia.df[, ch$label[4]]) *  coefs[[select.alpha]][ch$label[4],1]), dysplasia.df[,ch$label[4]]) ## neg coef
# 
# 
# labels[which( dysplasia.df[, ch$label[2]] <= 0.9)]
# 
# labels[which( dysplasia.df[, ch$label[1]] >= 1.1)]

#do.call(grid.arrange, c(apply(ch[1:10,],1, plot.reg), ncol=2))

#do.call(grid.arrange, c(apply(ch[11:20,],1, plot.reg), ncol=2))

#do.call(grid.arrange, c(apply(ch[21:30,],1, plot.reg), ncol=2))

```

```{r}

cols = intersect(rownames(ch), grep('\\d+:', colnames(dysplasia.df), invert=T, value=T))
if (length(cols) > 0)
  pander(ch[cols,1:2], caption='Non-zero demographic hazard ratios')
```

`r pander(ch[order(-ch$hazards),][1:10,1:2], caption='Top 10 features by hazard ratio.')`

```{r echo=F, warning=F, message=F}
if (!is.null(bm)) {
  topFRegions = merge(topFRegions, (bm %>% group_by(chromosome_name) %>% summarise('gene.cnt'=length(unique(hgnc_symbol)), 'exon.cnt'=sum(exon_count))), by.x='chr', by.y='chromosome_name', all.x=T)
  colnames(topFRegions) = sub('\\.(x|y)', '', colnames(topFRegions)  )

  bmGR = makeGRangesFromDataFrame(bm, start.field = 'start_position', end.field = 'end_position', keep.extra.columns = T)

  # Merge COSMIC genes
  ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
  ccgenes = ccgenes[-grep('X,Y', ccgenes$Genome.Location),]

  ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),] # drop the translocations
  #nrow(ccgenes)

  bmGR$COSMIC = bmGR$hgnc_symbol %in% ccgenes$Gene.Symbol

  grcfs = grep('^\\d+:', rownames(coefs[[select.alpha]]), value=T)
  topF = do.call(rbind.data.frame, strsplit(grcfs , ':|-'))

  colnames(topF) = c('chr','start','end')
  coefGR = makeGRangesFromDataFrame(topF)
  coefGR$coef = coefs[[select.alpha]][grcfs,1]

  coefGR$hazard = hazards[grcfs]
  coefGR$gene = ''
  coefGR$ensembl_ids = ''
  coefGR$COSMIC = ''
  coefGR$COSMIC_ensembl_ids = ''

  ov = findOverlaps(bmGR, coefGR)
  for (hit in unique(subjectHits(ov))) {
    tmp = bmGR[ queryHits(ov[which(subjectHits(ov) == hit)]) ]

    coefGR[hit]$COSMIC = paste(tmp[which(tmp$COSMIC)]$hgnc_symbol, collapse=', ')
    coefGR[hit]$COSMIC_ensembl_ids = paste(tmp[which(tmp$COSMIC)]$ensembl_gene_id, collapse=', ')

    coefGR[hit]$gene = paste(tmp$hgnc_symbol, collapse=', ')
    coefGR[hit]$ensembl_ids = paste(tmp$ensembl_gene_id, collapse=', ')
  }

  coefGR = coefGR[order(coefGR$hazard)]
  highest = coefGR[coefGR$hazard > quantile(coefGR$hazard)[['75%']]]

  write.table(unlist(strsplit(unique(highest$ensembl_ids), ', ')), quote=F, row.names = F, col.names = F, file='highest_coef_genes.txt')

  topFRegions$topCoefHaz = 0
  topFRegions$topCoefHaz[1:22] = table(seqnames(highest))
  grid.arrange(ggplot(topFRegions, aes(reorder(chr, -adj.info.content), gene.cnt/protein_coding, fill=ai.cut)) + geom_bar(stat='identity') +
    labs(y='Gene Ratio', x='chr', title='Feature genes vs all protein coding genes') +
    geom_text(aes(label=ifelse(topCoefHaz > 0, '*', ''), color=ai.cut), nudge_y=0.01, show.legend=F, size=10) +
    geom_text(aes(label=round(gene.cnt/protein_coding, 2)), vjust=-0.5) +
    scale_fill_discrete(name='Length adjusted information content') + theme(legend.position = 'bottom'), 
    bottom='Ratio of genes found within the boundaries of the top features vs all genes on the chromosome.\nChromosomes are ordered and binned by adjusted gene (e.g. information) content rather than length.')

  write.table(unlist(strsplit(unique(coefGR$COSMIC_ensembl_ids), ',')), quote=F, row.names = F, col.names = F, file='feature_cosmic.txt')
}
#However, the `r round(bpCovered, 2)*100`% of bases appear to include `r round(length(unique(bm$hgnc_symbol))/sum(chromInfo$protein_coding), 2)*100`% (`r length(unique(bm$hgnc_symbol))`) of known protein-coding genes.  Similarly, `r round(length(unlist(strsplit(unique(coefGR$COSMIC), ',')))/length(ccgenes$Gene.Symbol),2)*100`% of the `r length(ccgenes$Gene.Symbol)` COSMIC genes are contained within these regions.

```

### Gains vs Losses

try

gains = pmax(0, x-median(x))
losses = pmin(0, x-median(x))

```{r warning=F, message=F, fig.height=8, fig.width=8}
run.cv<-function(x, y, alphas=c(0,0.2,0.5,1), ...) {
  plots = list(); coefs = list(); performance.at.1se = list()
  for (a in alphas) {
    fitg <- glmnet(x, y, alpha=a, nlambda=nl, family='binomial', standardize = F) # all patients

    l = fitg$lambda
    if (a == 0) {
      l = sort(c(fitg$lambda, seq(exp(-5), exp(-10), -1e-6),
            seq(exp(-10), exp(-15), -1e-8),
            seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
    }

    cvp = crossvalidate.by.patient(x, y, fit=fitg, lambda=l, a=a, pts=sets, standardize=F)
    plots[[as.character(a)]] = arrangeGrob(cvp$plot+ggtitle('Classification'), cvp$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    coef.1se = as.data.frame(non.zero.coef(fitg, cvp$lambda.1se))
    coefs[[as.character(a)]] = coef.stability(coef.1se, cvp$non.zero.cf)

    performance.at.1se[[as.character(a)]] = subset(cvp$lambdas, lambda.1se == T)

  }
  return(list('plots'=plots, 'coefs'=coefs, 'performance'=performance.at.1se))
}

file = paste(cache.dir, 'gain.loss.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  gain.df = as.data.frame(matrix(nrow=nrow(dysplasia.df), ncol=ncol(dysplasia.df),data=0, dimnames=list(rownames(dysplasia.df), colnames(dysplasia.df))))
  loss.df = gain.df

  for (i in 1:nrow(dysplasia.df)) {
     #gain.df[i,] = as.integer(dysplasia.df[i,] >= 1.1)
     #loss.df[i,] = as.integer(dysplasia.df[i,] <= 0.9)
     #all.binary[i,] = as.integer(dysplasia.df[i,] >= 1.1 | dysplasia.df[i,] <= 0.9)
    gain.df[i,] = pmax(0, dysplasia.df[i,] - median(dysplasia.df[i,]))
    loss.df[i,] = pmin(0, dysplasia.df[i,] - median(dysplasia.df[i,]))
  }

  gp = run.cv(x=as.matrix(gain.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0.1,0.5,0.8,1))
  lp = run.cv(x=as.matrix(loss.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0.1,0.5,0.8,1))

  colnames(gain.df) = paste(colnames(gain.df), '_GAIN', sep='')
  colnames(loss.df) = paste(colnames(loss.df), '_LOSS', sep='')

  gl.df = merge(gain.df, loss.df, by='row.names')
  rownames(gl.df) = gl.df$Row.names
  gl.df$Row.names = NULL

  both = run.cv(x=as.matrix(gl.df), y=labels, pts=pts, nfolds=folds, splits=splits, alphas=c(0.1,0.5,0.8,1))

  save(gp, lp, both, file=file)
}

do.call(grid.arrange, c(gp$plots, top='All samples, GAINS 10fold, 5 splits'))
do.call(grid.arrange, c(lp$plots, top='All samples, LOSSES 10fold, 5 splits'))
do.call(grid.arrange, c(both$plots, top='All samples, gains & losses 10fold, 5 splits'))

```

### Remove HGD/IMC Samples

Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features.  What happens if we retrain it without HGD?

```{r noHGD, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
`%nin%` <- Negate(`%in%`)

info = do.call(rbind, lapply(patient.data, function(df) df$info))

no.hgd.plots = list(); coefs = list(); performance.at.1se = list()

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  # No HGD/IMC
  samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC'))$Samplename)
  for (a in alpha.values) {
    # all patients
    fitNoHGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl, standardize = F) 
    cv.nohgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoHGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoHGD, standardize = F)

    no.hgd.plots[[as.character(a)]] = arrangeGrob(cv.nohgd$plot+ggtitle('Classification'), cv.nohgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    coef.1se = as.data.frame(non.zero.coef(fitNoHGD, cv.nohgd$lambda.1se))
    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nohgd$non.zero.cf)

    performance.at.1se[[as.character(a)]] = subset(cv.nohgd$lambdas, lambda.1se == T)
  }
  save(no.hgd.plots, coefs, performance.at.1se, file=file)
}

do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
```

```{r noHGD2, echo=T, message=F, warning=F}
performance.at.1se = do.call(rbind, performance.at.1se)
performance.at.1se$alpha = rownames(performance.at.1se)

performance.at.1se = cbind(performance.at.1se, do.call(rbind, lapply(coefs, function(x) cbind('n.Coef'=nrow(x), '75%'=length(which(rowSums(x)/50 >= 0.75))))))

ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() + geom_text( aes(label=round(mean, 2)), hjust=-0.5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) + 
  geom_text(aes(label=n.Coef), nudge_y=0.03) +
  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
       title=paste('No HGD/IMC samples,', folds, 'folds,', splits, 'patient splits'))

## check how often the feature(s) is selected at that lamda in each split. "stability selection"
coef.stable = lapply( coefs[c('0.8','0.9','1')], function(cf) {
  sort(rowSums(cf[,-1]), decreasing=T)
})

cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=3, 
                           dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))

for (i in names(coef.stable)) {
  if (is.null(names(coef.stable$`0.8`))) next
  cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
}

coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))

stability = do.call(rbind, lapply(coef.stable, function(cf) {
  cbind('>=75%'=length(which(cf >= 0.75)),  '>=50%'=length(which(cf >= 0.5)),  '>=50%'=length(which(cf >= 0.25)), 'Total'=length(cf))
}))
rownames(stability) = names(coef.stable)
```

#### Feature Stability

`r pander(stability, caption='Total features found at each value of alpha in 50% or more of the folds run')`

They also share all of the non-zero coefficients (`r which.max(sapply(coef.stable, length))` identifies the most non-zero coefs `r max(sapply(coef.stable, length))`).

### Remove HGD/IMC & LGD Samples

Now remove all LGD samples from the progressor's and retrain.

```{r noLGD, echo=T, message=F, warning=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')

nolgd.plots = list(); coefs = list(); performance.at.1se = list()
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  # No LGD
  samples = intersect(rownames(dysplasia.df), c(subset(info, Status == 'NP')$Samplename, subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD') & Status == 'P')$Samplename))

  # No HGD/IMC/LGD in all patients
  samples = intersect(rownames(dysplasia.df), subset(info, Pathology %nin% c('HGD', 'IMC', 'LGD'))$Samplename)

  for (a in alpha.values) {
    fitNoLGD <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, family='binomial', nlambda=nl, standardize = F) # all patients

    cv.nolgd = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitNoLGD$lambda, pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitNoLGD, standardize = F)

    nolgd.plots[[as.character(a)]] = arrangeGrob(cv.nolgd$plot+ggtitle('Classification'), cv.nolgd$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)

    coef.1se = as.data.frame(non.zero.coef(fitNoLGD, cv.nolgd$lambda.1se))
    coefs[[as.character(a)]] = coef.stability(coef.1se, cv.nolgd$non.zero.cf)

    performance.at.1se[[as.character(a)]] = subset(cv.nolgd$lambdas, lambda.1se == T)
  }
  save(nolgd.plots, coefs, performance.at.1se, file=file)
}

do.call(grid.arrange, c(nolgd.plots, top="No HGD/IMC/LGD samples", ncol=2))
```

```{r noLGD2, echo=T, message=F, warning=F}
if (class(performance.at.1se) == 'list')
performance.at.1se = do.call(rbind, performance.at.1se)

performance.at.1se$alpha = rownames(performance.at.1se)

performance.at.1se = cbind(performance.at.1se, do.call(rbind, lapply(coefs, function(x) cbind('n.Coef'=nrow(x), '75%'=length(which(rowSums(x)/50 >= 0.75))))))

ggplot(performance.at.1se, aes(alpha, mean)) + geom_point() + geom_text( aes(label=round(mean, 2)), hjust=-0.5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme)) + ylim(0.5,0.8) +
  geom_text(aes(label=n.Coef), nudge_y=0.03) +
  geom_text(aes(label=paste('(',`75%`,')',sep='')), nudge_y=0.02) +
  labs(x='Elasticnet penalty value', y='Model performance at lambda.1se', 
       title=paste('No HGD/IMC/LGD samples,', folds, 'folds,', splits, 'patient splits'))

## check how often the feature(s) is selected at that lamda in each split. "stability selection"
coef.stable = lapply( coefs[c('0.8','0.9','1')], function(cf) {
  sort(rowSums(cf[,-1]), decreasing=T)
})

cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=3, 
                           dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))

for (i in names(coef.stable)) 
  cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1

coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))

stability = do.call(rbind, lapply(coef.stable, function(cf) {
  cbind('>=75%'=length(which(cf >= 0.75)),  '>=50%'=length(which(cf >= 0.5)),  '>=50%'=length(which(cf >= 0.25)), 'Total'=length(cf))
}))
rownames(stability) = names(coef.stable)

```

#### Feature Stability

`r pander(stability, caption='Total features found at each value of alpha in 50% or more of the folds run')`

They also share all of the non-zero coefficients (`r which.max(sapply(coef.stable, length))` identifies the most non-zero coefs `r max(sapply(coef.stable, length))`).

```{r manyfew, echo=T, message=F, warning=F, fig.height=16, fig.width=16, eval=F}
## Compare progressors with many samples vs few
m = median(subset(sum.patient.data, Status == 'P')$total.samples)
nonps = as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$info$Samplename)))

file = paste(cache.dir, 'few.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  # Few
  samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples < m)$Hospital.Research.ID], function(pt) pt$info$Samplename))))

  few.plots = list()
  for (a in alpha.values) {
    fitF <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial', standardize = F) # all patients
    cv.back.few = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitF$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=folds, a=a, fit=fitF, minR=0.1, standardize = F)
    #few.plots[[as.character(a)]] = cv.back.few$plot + ggtitle(paste('alpha=',a,sep=''))

    few.plots[[as.character(a)]] = arrangeGrob(cv.back.few$plot+ggtitle('Classification'), cv.back.few$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
  }
  save(few.plots, file=file)
} 
do.call(grid.arrange, c(few.plots, ncol=2, top=paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples < m))," Progressors with < median # samples (",m,")")))

# Many
file = paste(cache.dir, 'many.Rdata', sep='/')

if (file.exists(file)) {
  load(file, verbose=T)
} else {
  samples = c(nonps, as.vector(unlist(lapply(patient.data[subset(sum.patient.data, Status == 'P' & total.samples >= m)$Hospital.Research.ID], function(pt) pt$info$Samplename))))

  many.plots = list()
  for (a in alpha.values) {
    fitM <- glmnet(dysplasia.df[samples,], labels[samples], alpha=a, nlambda=nl, family='binomial', standardize = F) # all patients

    cv.back.many = crossvalidate.by.patient(x=dysplasia.df[samples,], y=labels[samples], lambda=fitM$lambda, pts=subset(pts, Samplename %in% samples), splits=5, nfolds=20, a=a, fit=fitM, standardize = F)

    many.plots[[as.character(a)]] = arrangeGrob(cv.back.many$plot+ggtitle('Classification'), cv.back.many$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
  }
  save(many.plots, file=file)
}

do.call(grid.arrange, c(many.plots, ncol=2, top=paste(nrow(subset(sum.patient.data, Status == 'P' & total.samples >= m))," Progressors with >= median # samples (",m,")")))

```

## Randomize the labels

By patient.  This should break the models...and as would be expected, the models are really poor with perfomance at 50% or less (coin toss).

```{r, echo=T, warning=F, message=F, fig.height=12, fig.width=12}

file = paste(cache.dir, 'rand.Rdata', sep='/')

if (file.exists(file)) {
  load(file, verbose=T)
} else {
  rand.patients = sum.patient.data
  rand.patients$Status = sample( c('NP','P'), nrow(sum.patient.data), replace=T, prob=c(0.5,0.5))

  r.labels = unlist(apply(rand.patients[,c('Hospital.Research.ID','Status')],1, function(st) {
    info = patient.data[[ st['Hospital.Research.ID'] ]]$info
    lbl = rep(as.integer( st['Status'] == 'P' ), nrow(info))
    names(lbl) = info$Samplename
    return(lbl)
  }))
  names(r.labels) = sub('.*\\.', '',  names(r.labels))

  # sort in label order
  r.dysplasia.df = t(mergedDf[,names(r.labels)])

  r.coefs = list(); r.plots = list()
  folds = 10; splits=5
  for (a in c(0,0.5,1)) {
    fitR <- glmnet(r.dysplasia.df, r.labels, alpha=a, nlambda=nl, family='binomial', standardize = F) # all patients

    cv.R = crossvalidate.by.patient(x=r.dysplasia.df, y=r.labels, lambda=fitR$lambda, pts=pts, a=a, nfolds=folds, splits=splits, fit=fitR, standardize = F)

    lambda.opt = ifelse( length(cv.R$lambda.1se) > 0, cv.R$lambda.1se, cv.R$lambda.min )

    coef.opt = as.data.frame(non.zero.coef(fitR, lambda.opt))
    r.coefs[[as.character(a)]] = coef.stability(coef.opt, cv.R$non.zero.cf)

    r.plots[[as.character(a)]] = arrangeGrob(cv.R$plot+ggtitle('Classification'), cv.R$deviance.plot+ggtitle('Binomial Deviance'), top=paste('alpha=',a,sep=''), ncol=2)
  }
  save(r.plots, r.coefs, file=file) 
}

do.call(grid.arrange, c(r.plots, top='All samples, Randomized labels (by patient)'))

```

# Leave one out and evaluate predictions

Leave out each patient (both P and NP) and run a new CV fit each time then predict the samples for the patient that was left out.  Fitted model includes HGD samples.

```{r leaveoneout, echo=T, message=F, warning=F}

info = do.call(rbind, lapply(patient.data, function(df) df$info))
pg.samp = lapply(patient.data, function(pt) {
  info = pt$info
  info$SampleSD = NA
  info$SampleMEAN = NA
  if (length(info$Samplename) > 1) {
    info$SampleSD = apply(pt$seg.vals[,info$Samplename], 2, sd)
    info$SampleMEAN =  apply(pt$seg.vals[,info$Samplename], 2, mean)
  }
  info$Prediction = NA
  info$Prediction.Dev.Resid = NA
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})

file = paste(cache.dir, 'loo.Rdata', sep='/')

if (file.exists(file)) {
  load(file, verbose=T)
} else {
  performance.at.1se = c(); coefs = list(); plots = list(); fits = list()
  # Remove each patient (LOO)
  for (pt in names(pg.samp)) {
    print(pt)
    tmp.patient.data = patient.data[subset(sum.patient.data, Hospital.Research.ID != pt)$Hospital.Research.ID]
    samples = as.vector(unlist(sapply(tmp.patient.data, function(df) df$info$Samplename )))

    train.rows = which(rownames(dysplasia.df) %in% samples)
    training = dysplasia.df[train.rows,]
    test = as.matrix(dysplasia.df[-train.rows,])
    if ( nrow(test) == ncol(dysplasia.df) ) test = t(test)

    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
    sparsed_test_data <- Matrix(data=0, nrow=ifelse(length(pg.samp[[pt]]$Samplename) > 1, nrow(test), 1),  ncol=ncol(training),
                                dimnames=list(pg.samp[[pt]]$Samplename,colnames(training)), sparse=T)
    for(i in colnames(dysplasia.df)) sparsed_test_data[,i] = test[,i]

    # Fit generated on all samples, including HGD
    a = select.alpha
    fitLOO <- glmnet(training, labels[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
    #l = sort(c(fitLOO$lambda, seq(exp(-5), exp(-10), -0.001),
    #      seq(exp(-10), exp(-15), -1e-7), 
    #      seq(exp(-15), exp(-20), -1e-9)), decreasing=T)
    l = fitLOO$lambda
    cv = crossvalidate.by.patient(x=training, y=labels[train.rows], lambda=l, 
                                  pts=subset(pts, Samplename %in% samples), a=a, nfolds=folds, splits=splits, fit=fitLOO, standardize=F)
    #plots[[pt]] = cv$plot + ggtitle(pt)

    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)

    fits[[pt]] = cv

    if ( length(cv$lambda.1se) > 0 ) {
      performance.at.1se = c(performance.at.1se, subset(cv$lambdas, lambda == cv$lambda.1se)$mean)

      coef.1se = as.data.frame(non.zero.coef(fitLOO, cv$lambda.1se))
      coefs[[pt]] = coef.stability(coef.1se, cv$non.zero.cf)

      logit <- function(p){log(p/(1-p))}
      inverse.logit <- function(or){1/(1 + exp(-or))}

      pm = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='response')
      sy = as.matrix(sqrt(binomial.deviance(pm, labels[pg.samp[[pt]]$Samplename])))

      #df = cbind.data.frame(pm, sy, 1:length(sy))
      #colnames(df) = c('pred','dev','x')
      #ggplot(df, aes(x, pred)) + geom_point() + geom_errorbar(aes(ymin=pred-dev, ymax=pred+dev))

      pg.samp[[pt]]$Prediction = pm[,1]
      pg.samp[[pt]]$Prediction.Dev.Resid = sy[,1]

    } else {
      warning(paste("Hospital.Research.ID", pt, "did not have a 1se"))
    }
  }
  save(plots, performance.at.1se, coefs, fits, pg.samp, file=file)
}

ggplot(as.data.frame(performance.at.1se), aes(y=performance.at.1se, x='LOO')) + ylim(0.5,0.9) +
  geom_boxplot( fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA) + geom_jitter() +
  labs(y='performance at lambda.1se', x='', title='Performance for Leave One (patient) Out, at alpha=1')

```

### Prediction ROC curves

Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=10, fig.width=6}
preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(df) df$Prediction))
np.preds = unlist(lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], function(df) df$Prediction))

roc50 = pROC::roc(as.integer(c(preds, np.preds) > 0.5), 
                c(rep(1, length(preds)), rep(0, length(np.preds))),
                main='Progressors & Non-progressors, cutoff 0.5')

df = cbind.data.frame('Specificity'=rev(roc50$specificities), 'Sensitivity'=rev(roc50$sensitivities))
g1 = ggplot(df, aes(Specificity, Sensitivity)) + geom_line() +
  scale_x_reverse() + 
  geom_label(data=t(as.data.frame(round(pROC::coords(roc50, "best"),2))), 
             aes(x=specificity, y=sensitivity, label=paste('prob=',threshold, ' (FPR ', specificity*100, '%, TPR ', sensitivity*100,'%)\nAUC:', sprintf("%.2f",roc50$auc*100), '%', sep='')), nudge_x=.25) +
  labs(title='Probability threshold = 0.5', x='Specificity (FPR)', y='Sensitivity (TPR)')

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, ci=T, of='thresholds')

df = cbind.data.frame('Specificity'=rev(roc$specificities), 'Sensitivity'=rev(roc$sensitivities))
g2 = ggplot(df, aes(Specificity, Sensitivity)) + geom_line() +
  scale_x_reverse() + 
  geom_label(data=t(as.data.frame(round(pROC::coords(roc, "best"),2))), aes(x=specificity, y=sensitivity, label=paste('prob=',threshold, ' (FPR ', specificity*100, '%, TPR ', sensitivity*100,'%)\nAUC:', sprintf("%.2f",roc$auc*100), '%', sep='')), nudge_x=.25) +
  labs(title='Probability threshold calculated by ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')

grid.arrange(g1, g2, nrow=2, top="ROC - Progressors vs Non")

cutoff = pROC::coords(roc, 'best')[['threshold']]
cutoff = 0.5
```

This argues that we should set the prediction probability cutoff to `r pROC::coords(roc, 'best')[['threshold']]`.  However, for now the predictions from here on are based on the default `r cutoff` value.

***

__All predictions from here onwards are based on the `r cutoff` cutoff, not the ROC curve__

***

## How do the predictions look at each time point?

Providing a single sample at a time to predict is unnecessary as each sample is independently predicted from the test matrix. Predictions are exactly the same even when a single sample is provided (checked).

```{r timepoints, echo=F, message=F, warning=F}
# For HGD/IMC samples, progressors only
hgd = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
  hgd = subset(pt, Pathology %in% c('HGD', 'IMC'))
  sum(as.integer(hgd$Prediction > cutoff))/nrow(hgd)
})

hgd = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
  hgd = subset(pt, Pathology %in% c('HGD', 'IMC'))
  cbind.data.frame('correctly.predicted'=sum(as.integer(hgd$Prediction > cutoff)), 'total.hgd.samples'=nrow(hgd))
}))
rownames(hgd) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(hgd))$Patient

hgd = subset(hgd, total.hgd.samples > 0) # Some patients had no hgd samples

hgd$pred.ratio = hgd$correctly.predicted/hgd$total.hgd.samples

hgd$Patients = rownames(hgd)
hgd = transform(hgd, Patients=reorder(Patients, -pred.ratio) )

ggplot(hgd[,c('total.hgd.samples','pred.ratio', 'Patients')], aes(x=Patients, y=pred.ratio)) + 
  geom_bar(stat='identity', fill='darkblue', alpha=0.6) + geom_label( aes(label=total.hgd.samples) ) +
  coord_flip() + labs(title='Correctly predicted HGD samples', y='Correct prediction ratio')

global.by.path = do.call(rbind.data.frame,
                         lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
  final.endo = subset(pt, Endoscopy.Year == max(Endoscopy.Year))
  final.endo = final.endo[nrow(final.endo),]

  #pred <- function(rows) length(which(rows$Prediction > 0.5))/nrow(rows)
  pred <- function(rows) length(which(rows$Prediction > cutoff))

  h = subset(pt, Pathology %in% c('IMC','HGD'))
  l = subset(pt, Pathology == 'LGD' & PID != final.endo$PID)
  id = subset(pt, Pathology == 'ID' & PID != final.endo$PID)
  be = subset(pt, Pathology %in% c('BE') & PID != final.endo$PID)

  cbind('HGD'=pred(h), 'total.h'=nrow(h), 
        'LGD'=pred(l), 'total.l'=nrow(l), 
        'ID'=pred(id), 'total.id'=nrow(id),
        'NDBE'=pred(be), 'total.be'=nrow(be))
})
)

global.by.path = colSums(global.by.path)

global.path = as.data.frame(matrix(global.by.path, byrow=T, nrow=4, dimnames=list(c('HGD','LGD','ID','NDBE'),c('pred', 'total'))))
global.path$ratio = (global.path$pred/global.path$total)*100
global.path$path = rownames(global.path)

global.path$path = factor(global.path$path, levels=c('HGD','LGD','ID','NDBE'))

ggplot(melt(global.path[,c('ratio','total','path')], measure.vars='ratio', id.vars=c('path', 'total')), aes(path, value, group=path)) + 
  geom_bar(stat='identity', fill='dodgerblue3') + 
  geom_text(aes(y=value+2,label=paste(round(value, 1), '%', sep=''))) +
  geom_text(aes(y=3,label=paste('n=', total, sep=''))) +
  ylim(0,100) + labs(title='Samples predicted by pathology', x='Pathology', y='Percent identified')
```

```{r echo=F, message=F, warning=F}
lgd = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
  e = subset(pt, Pathology  == 'LGD')
  cbind.data.frame('predicted'=sum(as.integer(e$Prediction > cutoff)), 'total.lgd'=nrow(e))
}))
rownames(lgd) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(lgd))$Patient

lgd = subset(lgd, total.lgd > 0)

lgd$pred.ratio = round(lgd$predicted/lgd$total.lgd, 2)

lgd$Patients = rownames(lgd)
lgd = transform(lgd, Patients=reorder(Patients, -pred.ratio) )

ggplot(lgd[,c('total.lgd','pred.ratio', 'Patients')], aes(x=Patients, y=pred.ratio)) + 
  geom_bar(stat='identity', fill='darkgreen', alpha=0.6) + geom_label( aes(label=total.lgd) ) +
  coord_flip() + labs(title='Correctly predicted LGD samples', y='Correct prediction ratio')
```

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=10}
# All samples
all.predictions = sapply(pg.samp, function(pt) {
  if (unique(pt$Status) == 'NP') {
    as.integer(pt$Prediction < cutoff)
  } else {
    as.integer(pt$Prediction > cutoff)
  }
})

all.predictions = do.call(rbind, lapply(all.predictions, function(x) {
  cbind.data.frame('correctly.predicted'=sum(x), 'total.samples'=length(x))
}))

all.predictions$Status = ifelse (rownames(all.predictions) %in% subset(sum.patient.data, Status == 'P')$Hospital.Research.ID, 'P', 'NP')

rownames(all.predictions) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(all.predictions))$Patient

#ggplot(melt(all.predictions, id.vars='Status'), aes(y=value, x=Status, fill=variable)) +
#  geom_bar(stat='identity', position='dodge') + labs(x='', y='Sample Counts', main='Total Sample Predictions')

all.predictions$pred.ratio = with(all.predictions, correctly.predicted/total.samples)
all.predictions$Patient = rownames(all.predictions)
all.predictions = transform(all.predictions, Patient=reorder(Patient, pred.ratio) )

p = subset(all.predictions, Status == 'P')
np = subset(all.predictions, Status == 'NP')

m = melt(all.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))

grid.arrange(
  ggplot() + geom_bar(data=subset(m, Status == 'P'), aes(Patient, value), fill='darkblue', color='grey', stat='identity') +  
    geom_label( aes(x=nrow(p)/2+nrow(p)/3, y=0.5, 
                    label=paste(round(nrow(p[p$pred.ratio == 1,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
    geom_label( aes(x=nrow(p)/3, y=0.5, 
                    label=paste(round(nrow(p[p$pred.ratio < 1 & p$pred.ratio > 0,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
        geom_text( aes(x=2, y=0.5, 
                    label=paste(round(nrow(p[p$pred.ratio == 0,])/nrow(p),2)*100,'%',sep='')), size=8,  fontface='bold' ) +
    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
    labs(title='Progressors', x='', y='Ratio predicted samples'),

  ggplot() + geom_bar(data=subset(m, Status == 'NP'), aes(Patient, value), fill='darkgreen', color='grey', stat='identity') + 
        geom_label( aes(x=nrow(np)/2+nrow(np)/4, y=0.5, 
                    label=paste(round(nrow(np[np$pred.ratio == 1,])/nrow(np),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
    geom_label( aes(x=nrow(np)/4, y=0.5, 
                    label=paste(round(nrow(np[np$pred.ratio < 1 & np$pred.ratio > 0,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
        geom_text( aes(x=1.2, y=0.5, 
                    label=paste(round(nrow(np[np$pred.ratio == 0,])/nrow(np),2)*100,'%',sep='')), size=8,  fontface='bold' ) +
    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + coord_flip() +
    labs(title='Non Progressors', x='', y='Ratio predicted samples'),
  ncol=2)

```

### Progressors

In progressors, `r round(sum(hgd$correctly.predicted)/sum(hgd$total.hgd.samples), 2)*100`% of the HGD samples were correctly predicted.

When predicting each sample individually `r round(nrow(subset(p, pred.ratio == 1))/nrow(p)*100, 1)`% patients were correctly predicted in all samples. `r round(nrow(subset(p, pred.ratio < 1 & pred.ratio > 0))/nrow(p)*100, 1)`% of patients were correctly predicted in at least one sample.

`r nrow(subset(p, pred.ratio == 0))` patients had no correctly predicted samples.

### Non-Progressors

In non-progressors `r round(nrow(subset(np, pred.ratio == 1))/nrow(np), 2)*100`% of patients were predicted to be non-progressors in _all_ of their samples. `r nrow(subset(np, pred.ratio >= 0.7))` patients were predicted correctly in 70% or more of their samples. Only `r nrow(subset(np, pred.ratio < 0.7 & pred.ratio > 0))` patients had fewer than 70% of their samples predicted correctly, and both of these have consistently been predicting _progression_ across multiple endoscopies.

```{r timepoints2, echo=T, message=F, warning=F, fig.height=10, fig.width=8}
# Only BE, Progressors
ndbe.only = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
  final.endo = subset(pt, Pathology %in% c('IMC','HGD'))
  e = subset(pt, Pathology %in% c('BE', '', '?'))
  if (nrow(final.endo) > 0) # exclude BE at the HGD endoscopy
    e = subset(e, PID != final.endo[nrow(final.endo), 'PID'])

  cbind.data.frame('predicted'=sum(as.integer(e$Prediction > cutoff)), 'total.be'=nrow(e))
}))
ndbe.only = subset(ndbe.only, total.be > 0) # Exclude patients without any BE
ndbe.only$pred.ratio = round(ndbe.only$predicted/ndbe.only$total.be, 2)

rownames(ndbe.only) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(ndbe.only))$Patient

ndbe.only$Patients = rownames(ndbe.only)
ndbe.only = transform(ndbe.only, Patients=reorder(Patients, -pred.ratio) )

ggplot(ndbe.only[,c('total.be','pred.ratio', 'Patients')], aes(x=Patients, y=pred.ratio)) + 
  geom_bar(stat='identity', fill='darkred', alpha=0.6) + geom_label( aes(label=total.be) ) +
  coord_flip() + labs(title='Correctly predicted NDBE samples', subtitle='Excluding samples at HGD endoscopy', y='Correct prediction ratio')

total = do.call(rbind, lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
    cbind.data.frame('predicted'=sum(as.integer(pt$Prediction > cutoff)), 'total'=nrow(pt))
}))

total$pred.ratio = round(total$predicted/total$total, 2)

rownames(total) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(total))$Patient

total$Patients = rownames(total)
total = transform(total, Patients=reorder(Patients, -pred.ratio) ) 
ggplot(total[,c('total','pred.ratio', 'Patients')], aes(x=Patients, y=pred.ratio)) + 
  geom_bar(stat='identity', fill='purple', alpha=0.6) + geom_label( aes(label=total) ) +
  coord_flip() + labs(title='Correctly predicted Progressor samples', y='Correct prediction ratio')

# For all pre-HGD/IMC/LGD in progressors
pre.hgd = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], function(pt) {
  e = subset(pt, Pathology %nin% c('HGD', 'IMC', 'LGD'))
  e$Prediction
  sum(as.integer(e$Prediction > cutoff))/nrow(e)
})

pred.by.endo<-function(pt, nyears) {
  e = subset(pt, PID %in% unique(pt$PID)[nyears])
  if (  length(which(e$Pathology %in% c('HGD','IMC'))) > 0 ) return(NA)
  p = as.vector(as.integer(e$Prediction > cutoff))
  names(p) = e$Endoscopy.Year
  return(p)
}

predicted.progression<-function(x, pos=1) { 
  if(length(which(is.na(x))) == length(x)) return(NA)
  as.integer(length(which(x == pos)) > 0)  
}

cols = c('%.Predicted','n.Endoscopies','n.Patients', 'FNR','FPR', 'median.pred.samp','median.total.samp', 'max.samples', 'sd.samples')
p.predictions = as.data.frame(matrix(data=0,nrow=6, ncol=length(cols), dimnames=list(c(), cols)))
for (i in 1:nrow(p.predictions)) {
  p = sapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], pred.by.endo, i)
  pp = sapply(p, predicted.progression)
  pp = pp[!is.na(pp)]

  fpr = sapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID], pred.by.endo, i)
  fpr = sapply(fpr, predicted.progression)
  fpr = fpr[!is.na(fpr)]

  p.predictions[i,] =  rbind(round(sum(pp[!is.na(pp)]/length(pp[!is.na(pp)])), 2), i, length(pp), 
                             round(length(which(pp == 0))/length(pp), 2),
                             round(sum(fpr)/length(fpr), 2), 
                             median(sapply(p[names(which(pp > 0) )], function(x) length(which(x == 1)) )),
                             median(sapply(p[names(which(pp > 0))], length)), 
                             max(sapply(p[names(which(pp > 0))], length)),
                             sd(sapply(p[names(which(pp > 0))], length))
                             )
}

ggplot(p.predictions, aes(y=`%.Predicted`,x=factor(n.Endoscopies))) + 
  geom_bar(stat="identity", fill='dodgerblue3') + 
  geom_text(aes(label=paste(`%.Predicted`*100, '%', sep='')), nudge_y=-0.02) + 
  geom_text(aes(label=paste(n.Patients, 'patients')), nudge_y=0.01) + 
  geom_bar(stat='identity', fill='darkred', alpha=0.3, aes(y=FPR)) +
  geom_text(aes(label=FPR, y=FPR)) +
  labs(title="Progression predicted per endoscopy", x='Number of Endoscopies', y='% Patients') +
  theme(text = element_text(size = 14))

```

## Backwards from HGD

Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.

```{r, warning=F, message=F, echo=F}
transform.pt.by.time<-function(pt) {
  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
                  'pred'=sum(as.integer(Prediction > cutoff)), 'samples'=length(PID), 
                  #'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
                  'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12, 
                  'max.path'=max(Pathology))
  arrange(pt, -Endoscopy.Year, PID)
}

back = lapply(pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID], transform.pt.by.time)

endos = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
  df = do.call(rbind, lapply(back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
  }))
  rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient

  df$Patient = rownames(df)
  return(df)
}))

endos = endos[complete.cases(endos),]
endos$i = endos$i-1

endos$pred.ratio = round(endos$pred/endos$samples, 2)
endos = transform(endos, Patient=reorder(Patient, -months.before.final) )

endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
    breaks=c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 12*12, 15*12))

endos$labels = gsub('\\[|\\(|\\]', '',endos$brks)
endos$labels = sub(',', '-',endos$labels)

m = endos %>% group_by(brks, labels) %>% summarise( 'pred.r'= length(which(pred > 0))/length(brks), 'n'=length((Patient)) )
ggplot(m, aes(brks, pred.r*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='dodgerblue3') + 
  geom_text(aes(y=25, label=paste(round(pred.r*100), '%', sep='')), nudge_y=-0.02) + 
  geom_text(aes(y=pred.r*100+1.5, label=paste('n=',n, sep='')), nudge_y=0.02) + 
  scale_x_discrete(labels=m$labels) +
  labs(x='Months', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = number of endoscopies')

plot.endo<-function(df, minM=NULL, maxM=NULL, bin=F) {
  if (is.null(minM)) minM = min(df$months.before.final)
  if (is.null(maxM)) maxM = max(df$months.before.final)

  df = arrange(df %>% group_by(Patient) %>% mutate('overall.pred'=sum(as.integer(pred > 0))/length(Patient)), Patient)

  df = transform(df, Patient=reorder(Patient, -overall.pred) )

  if (bin) {
    df$endo.pred = ifelse(df$pred > 0, 'Pos','Neg')
    ggplot(df, aes(months.before.final, endo.pred, group=1,color=Patient)) + facet_grid(Patient ~.) + 
      scale_x_continuous(breaks=seq(minM, maxM, by = 12)) +
      geom_line() + geom_point() +
      geom_label_repel(aes(label=max.path, fill=Patient, color=Patient), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines')) +
      labs(x='Months before diagnosis', y='Predictive Endoscopy') + 
      theme(text=element_text(size=14, face="bold"), axis.title=element_text(size=16), legend.position='none')
  } else {
  ggplot(df, aes(months.before.final, pred.ratio, group=1, color=Patient, fill=Patient)) + facet_grid(Patient ~.) +
      ylim(0,1) + geom_abline(slope=0, intercept=seq(.25,.75,.25), color='grey', linetype='dashed') +
      geom_col(position='dodge', aes(y=pred.ratio)) + geom_line() +
      scale_x_continuous(breaks=seq(minM, maxM, by = 12)) + 
      geom_label(aes(y=ifelse(pred.ratio==1, 0.75, pred.ratio), label=max.path, fill=Patient), color='white', fontface='bold', size=4, label.size=0, label.r=unit(0.25, 'lines'), label.padding=unit(0.15, 'lines'), nudge_y=0.1) +
      labs(x='Months before diagnosis', y='Sample prediction ratio') + 
      theme(text=element_text(size=14, face="bold"), axis.title=element_text(size=16), legend.position='none')
  }
}

endos$months.before.final = endos$months.before.final*-1
me = endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))
```

```{r, echo=F, warning=F, message=F, fig.height=22, fig.width=10}
print.png<-function(p, name, ...) {
  png(filename=name,...)
  print(p)
  dev.off()
}

p = plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=T) + 
  labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
print.png(p, 'progressors_1.png', width=900, height=1200)
grid.arrange(p)

p1a = plot.endo( subset(endos, Patient %in% subset(me, months <= -96)$Patient), bin=T) + 
  labs(title="Progressors", subtitle=paste(96/12, "years or more follow-up"))
print.png(p1a, 'progressors_1a.png', width=600, height=800)

p1b = plot.endo( subset(endos, Patient %in% subset(me, months <= -72 & months > -96)$Patient), bin=T) + 
  labs(title="Progressors", subtitle=paste(72/12, "-",96/12,"years or more follow-up"))
print.png(p1b, 'progressors_1b.png', width=600, height=800)

p1 = plot.endo( subset(endos, Patient %in% subset(me, months <= -72)$Patient), bin=F) +
  labs(title="Progressors", subtitle=paste(72/12, "years or more follow-up"))
print.png(p1, 'progressors_1_bar.png', width=900, height=1200)

```

```{r, echo=F, warning=F, message=F, fig.height=22, fig.width=10}
p1 = plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=T) +
  labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
print.png(p1, 'progressors_2.png', width=400, height=600)

p1 = plot.endo( subset(endos, Patient %in% subset(me, months <= -48 & months > -72)$Patient), bin=F) +
  labs(title="Progressors", subtitle=paste(48/12, "-",72/12,  "years or more follow-up"))
print.png(p1, 'progressors_2_bar.png', width=400, height=600)

p2 = plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=T) +
  labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
print.png(p2, 'progressors_3.png', width=400, height=600)

p2 = plot.endo( subset(endos, Patient %in% subset(me, months <= -24 & months > -48)$Patient), bin=F) +
  labs(title="Progressors", subtitle=paste(24/12, "-",48/12,  "years or more follow-up"))
print.png(p2, 'progressors_3_bar.png', width=400, height=600)

p3 = plot.endo( subset(endos, Patient %in% subset(me, months >= -12)$Patient), bin=T) + 
  labs(title="Progressors", subtitle=paste("1 year follow-up"))
print.png(p3, 'progressors_4.png', width=400, height=600)

p3 = plot.endo( subset(endos, Patient %in% subset(me, months >= -12)$Patient), bin=F) +
  labs(title="Progressors", subtitle=paste("1 year follow-up"))
print.png(p3, 'progressors_4_bar.png', width=400, height=600)

grid.arrange(p1,p2,p3,ncol=1, top='Progressor predictions prior to diagnosis')

```

Just for completeness, we can plot non-progressors similarly:

```{r echo=F, warning=F, message=F, fig.height=42, fig.width=10}
nonp.back = lapply(pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID],transform.pt.by.time)

nonp.endos = do.call(rbind, lapply(1:max(sapply(nonp.back,nrow)), function(i) {
  df = do.call(rbind, lapply(nonp.back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
  }))
  rownames(df) = subset(sum.patient.data, Hospital.Research.ID %in% rownames(df))$Patient

  df$Patient = rownames(df)

  return(df)
}))
nonp.endos = nonp.endos[complete.cases(nonp.endos),]
#nonp.endos$i = nonp.endos$i-1

nonp.endos$pred.ratio = round(nonp.endos$pred/nonp.endos$samples, 2)
nonp.endos = transform(nonp.endos, Patient=reorder(Patient, -months.before.final) ) 
nonp.endos$months.before.final = nonp.endos$months.before.final*-1

me = nonp.endos %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))

plot.endo(subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=F)

p = plot.endo( subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=T) + 
  labs(title="Non-Progressors", subtitle=paste("Patients with predictive endoscopies, max", abs(min(me$months)/12), 'years follow-up'))
print.png(p, 'non-progressors.png', width=600, height=800)

p1 = plot.endo( subset(nonp.endos, Patient %in% subset(nonp.endos, pred.ratio > 0)$Patient), bin=F) + 
  labs(title="Non-Progressors", subtitle=paste("Patients with predictive endoscopies, max", abs(min(me$months)/12), 'years follow-up'))
print.png(p1, 'non-progressors_bar.png', width=600, height=800)

grid.arrange(p)
```

```{r, warning=F, message=F, fig.height=4, fig.width=4, eval=F}
### How many samples?  TBD

#Clinicians asked how many samples they needed to take.

# How do I evaluate the predictions?

#Basically I need to to a FNR and FPR calculation that is extremely conservative and use that somehow

all = do.call(rbind,lapply(pg.samp, function(df) {
  df[,c('Status','Prediction','Prediction.Dev.Resid')]
}))

ggplot(all, aes(Prediction.Dev.Resid, Prediction, color=Status)) + geom_point()  + 
  labs(y='Predicted (response)', x='Binomial dev residual?')
```

## Data Quality

### By Batch first

```{r dq_batch, echo=T, warning=F, message=F}
pg.samp = lapply(pg.samp, function(df){ df$Batch.Name = sub('_', '', df$Batch.Name ); df })

all = do.call(rbind, pg.samp)

prog = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'P')$Hospital.Research.ID])
nonp = do.call(rbind, pg.samp[subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID])

## Similar numbers of each in these batches so how did we do?
tfPR = sapply(as.character(unique(patient.info$Batch.Name)), function(b) {
  pred = subset(prog, Batch.Name == b)$Prediction
  predB = subset(nonp, Batch.Name == b)$Prediction

  # Only for progressors because we know for certain they progress, the non-ps may just not have progressed *yet*
  tpr = length(which(pred > cutoff))/length(pred)
  fpr = length(which(predB > cutoff))/length(predB)
  cbind.data.frame('TPR'=tpr, 'FPR'=fpr)
})
```

`r pander(t(tfPR), justify='left', caption=paste('True positive rate is between', paste(round(range(tfPR['TPR',], na.rm=T), 2), collapse='-')))`

```{r dq_batch2, echo=F, warning=F, message=F}

pilot = pg.samp[subset(patient.info, grepl('nppilotstudy', Batch.Name, ignore.case=T))$Hospital.Research.ID]
pilot = do.call(rbind, pilot[ which(sapply(pilot, nrow) > 0) ])
pilot = subset(pilot, Status == 'NP')
length(which(pilot$Prediction > cutoff))/nrow(pilot)

p2 = subset(pilot, Patient %nin% c('AD0591', 'AH0329'))
length(which(p2$Prediction > cutoff))/nrow(p2)

exome =  pg.samp[subset(patient.info, grepl('exomesubcohort', Batch.Name, ignore.case=T))$Hospital.Research.ID]
exome = do.call(rbind, exome[ which(sapply(exome, nrow) > 0) ])
exome = subset(exome, Status == 'NP')
length(which(exome$Prediction > cutoff))/nrow(exome)

e2 = subset(exome, Patient %nin% c('AD0591', 'AH0329'))
length(which(e2$Prediction > cutoff))/nrow(e2)

```

The false positive rate in both the exome and NP_pilot batches is driven by two patients that we predict to be progressors in nearly every sample: AD0591 and AH0329.  Both patients had samples in the exome batch, while AH0329 also had samples in the NP_Pilot batch.  Without those patients the FPR in the exome batch was `r round(length(which(e2$Prediction > cutoff))/nrow(e2), 2)*100`% for `r length(unique(e2$Patient))` patient(s), and in the NP_Pilot batch the FPR was `r round(length(which(p2$Prediction > cutoff))/nrow(p2), 2)*100`% for `r length(unique(p2$Patient))` patient(s).

The somewhat lower Batch_1 prediction rate is driven by the samples from AHM0277 as only `r length(which(pg.samp$AHM0277$Prediction > cutoff))/nrow(pg.samp$AHM0277)*100`% of this patient's samples were predicted to progress.  Batch_3 is driven by three other patients that have low prediction rates: PR1_ADH_076, PR1_HIN_046, and PR1_WSH_026.

## Cellularity or general data quality measures?

```{r, echo=F, warning=F, message=F}
all$SampleCoV = with(all, SampleSD/SampleMEAN)

prog = subset(all, Status == 'P', select=c('Prediction', 'Prediction.Dev.Resid', 'SampleMEAN', 'SampleSD', 'SampleCoV', 'Total.Reads'))
nonp = subset(all, Status == 'NP', select=c('Prediction', 'Prediction.Dev.Resid', 'SampleMEAN', 'SampleSD', 'SampleCoV', 'Total.Reads'))

#GGally::ggcorr(data=prog, method=c('pairwise','pearson'), label=T) + labs(title='Pearson correlation for several QC measures')

library(Hmisc)
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}

rc = rcorr( as.matrix(all[, c('Prediction', 'Prediction.Dev.Resid', 'SampleMEAN', 'SampleSD','SampleCoV', 'Total.Reads')]) )

frc = flattenCorrMatrix(rc$r, rc$P)
pander(frc[which(p.adjust(frc$p, 'bonferroni') < 0.05),], caption='Prediction and SampleSD have a small but significant correlation after Bonferroni correction. This could indicate cellularity or sample quality is influencing the predictions.')
```

```{r}   
# Compare poorly predicted to well predicted based on the sample SD measures, no difference
wilcox.test(subset(prog, Prediction > cutoff)$SampleSD, subset(prog, Prediction < cutoff)$SampleSD)
wilcox.test(subset(nonp, Prediction > cutoff)$SampleSD, subset(nonp, Prediction < cutoff)$SampleSD)

wilcox.test(prog$SampleSD, nonp$SampleSD) # Appears to be a small difference in this measure, where is it from?

## Progressors
tprows = which(prog$Prediction > cutoff)

cor.test(prog[tprows, 'SampleSD'], prog[tprows, 'Prediction'] ) # Correlated for True Pos predictions
cor.test(prog[-tprows, 'SampleSD'], prog[-tprows, 'Prediction'] ) # But not for the poorly predicted

badrows = which(prog$Prediction < 0.3)
cor.test(prog[badrows, 'SampleSD'], prog[badrows, 'Prediction'] ) # But not for the poorly predicted

# There's no real difference between the sample 'cellularity' or quality between the good or bad predictions in the progressors
wilcox.test(prog[tprows, 'SampleSD'], prog[-tprows, 'SampleSD']) 
wilcox.test(prog[tprows, 'SampleCoV'], prog[-tprows, 'SampleCoV'])

## Non-Progressors, this is a little more complicated as we cannot be sure of True Positives but we'll look anyhow
tprowsNP = which(nonp$Prediction < cutoff)
cor.test(nonp[tprowsNP, 'SampleSD'], nonp[tprowsNP, 'Prediction'] ) # Correlated for True Pos predictions
cor.test(nonp[-tprowsNP, 'SampleSD'], nonp[-tprowsNP, 'Prediction'] ) # But not for the poorly predicted

badrows = which(nonp$Prediction > 0.7)
cor.test(nonp[badrows, 'SampleSD'], nonp[badrows, 'Prediction'] ) # But not for the poorly predicted

# There's no real difference between the sample 'cellularity' or quality between the good or bad predictions in the progressors
wilcox.test(nonp[tprowsNP, 'SampleSD'], nonp[-tprowsNP, 'SampleSD']) 
wilcox.test(nonp[tprowsNP, 'SampleCoV'], nonp[-tprowsNP, 'SampleCoV'])

## How about the TPs for Progressors and the predicted Non-P?
wilcox.test(nonp[-tprowsNP, 'SampleSD'], prog[tprows, 'SampleSD']) # Basically the same
wilcox.test(nonp[tprowsNP, 'SampleSD'], prog[-tprows, 'SampleSD']) # Basically the same

ggplot(prog, aes(x=SampleSD)) + 
  geom_histogram(data=subset(prog, Prediction > cutoff), aes(SampleSD, fill='TP'), alpha=0.5) +
  geom_histogram(data=subset(prog, Prediction < cutoff), aes(SampleSD, fill='FN'), alpha=0.5) + 
  labs(title='SD of Samples vs True positive ~ False negative Progressors')

```

# Validation Dataset

20 Patients (10P, 10NP) set aside for validation purposes

```{r validation, message=F, warning=F, echo=F, fig.height=8, fig.width=6}

score.cx <-function(pd) { ## OLD method
#  print(pd$info)
  len.var = (var(pd$seg.vals$end.pos - pd$seg.vals$start.pos))/1e12
  extreme.values = max( apply(as.matrix(pd$seg.vals[,grep('^D', colnames(pd$seg.vals))]), 2, function(col) {
    length(which(col >= (median(col)+2*sd(col)))) + length(which(col <= (median(col)-2*sd(col)))) 
  }) )

  (len.var + extreme.values)
}

validation.df = t(mergedDfValidation[, grep('^D', colnames(mergedDfValidation))])
validation.df = apply(validation.df, 2, unit.var)

complexity.measures = sapply(validation.patient.data, score.cx)
validation.df = cbind(validation.df, 'cx.measure'=NA)
for (sample in rownames(validation.df)) {
  validation.df[sample, 'cx.measure'] = complexity.measures[subset(validation.patient.info, Samplename == sample)$Hospital.Research.ID]
}
# 
# # norm to 1 to keep within the magnitude of the genomic data
cx = validation.df[,'cx.measure']
validation.df[,'cx.measure'] = (log(cx)/log(max(cx)))

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
message(paste("loading", file))
load(file, verbose=T)

#fitV <- glmnet(dysplasia.df, labels, alpha=select.alpha, nlambda=nl, family='binomial') # all patients
#autoplot(fitV)
fitV = models[[select.alpha]]
lambda.opt = performance[select.alpha, 'lambda']

pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')

vpd = lapply(validation.patient.data, function(pt) {
  info = pt$info
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})

samples = intersect(rownames(pred), validation.patient.info$Samplename)
validation.patient.info$Prediciton = NA
validation.patient.info[which(validation.patient.info$Samplename %in% samples), 'Prediction'] = pred[samples,]

for (pt in names(vpd)) {
  vpd[[pt]]$Prediction[which(vpd[[pt]]$Samplename %in% rownames(pred))] = pred[ intersect(rownames(pred), vpd[[pt]]$Samplename), ]
}

v.predictions = sapply(vpd, function(pt) {
  if (unique(pt$Status) == 'NP') {
    as.integer(pt$Prediction < cutoff)
  } else {
    as.integer(pt$Prediction > cutoff)
  }
})

names(v.predictions) = subset(sum.validation.info, Hospital.Research.ID %in% names(v.predictions))$Patient

v.predictions = do.call(rbind, lapply(v.predictions, function(x) {
  cbind.data.frame('correctly.predicted'=sum(x,na.rm=T), 'total.samples'=length(x))
}))

v.predictions$Status = ifelse (rownames(v.predictions) %in% subset(sum.validation.info, Status == 'P')$Patient, 'P', 'NP')

v.predictions$pred.ratio = with(v.predictions, correctly.predicted/total.samples)
v.predictions$Patient = rownames(v.predictions)
v.predictions = transform(v.predictions, Patient=reorder(Patient, pred.ratio) )

p = arrange(subset(v.predictions, Status == 'P'), -pred.ratio)
np = arrange(subset(v.predictions, Status == 'NP'), -pred.ratio)

m = melt(v.predictions, id.vars=c('Patient','Status'), measure.vars=c('pred.ratio'))

p1 = ggplot(data=subset(v.predictions, Status == 'P'), aes(Patient, pred.ratio)) + geom_col(fill='darkblue', color='grey') + coord_flip() +
    geom_label( aes(x=p[nrow(subset(p, pred.ratio == 1))/2,'Patient'], y=0.5, label=paste(round(nrow(p[p$pred.ratio == 1,])/nrow(p),2)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
    geom_label( aes(x='77', y=0.5, label=paste(round(nrow(p[p$pred.ratio < 1 & p$pred.ratio > 0,])/nrow(p),3)*100,'%',sep='')), size=8, fill='darkblue', color='white', fontface='bold' ) +
    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + 
    labs(title='Progressors', x='', y='Ratio predicted samples')
if (nrow(subset(p,pred.ratio == 0)) > 0) {
  p1 = p1 +  geom_text( aes(x=subset(p,pred.ratio == 0)$Patient[1], y=0.5, label=paste(round(nrow(p[p$pred.ratio <= 0,])/nrow(p),3)*100,'%',sep='')), size=8, fontface='bold' ) 
}

p2 = ggplot(data=subset(v.predictions, Status == 'NP'), aes(Patient, pred.ratio)) + geom_col(fill='darkgreen', color='grey') + coord_flip() +
    geom_label( aes(x=np[nrow(subset(np, pred.ratio == 1))/2,'Patient'], y=0.5, 
                    label=paste(round(nrow(np[np$pred.ratio == 1,])/nrow(np),2)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
    geom_label( aes(x='41', y=0.5, label=paste(round(nrow(np[np$pred.ratio < 1 & np$pred.ratio > 0,])/nrow(np),3)*100,'%',sep='')), size=8, fill='darkgreen', color='white', fontface='bold' ) +
    theme(legend.position='none', axis.text.x = element_text(angle = 45, hjust = 1)) + 
    labs(title='Non Progressors', x='', y='Ratio predicted samples')
if (nrow(subset(np,pred.ratio == 0)) > 0) {
  p2 = p2 +  geom_text( aes(x=subset(np,pred.ratio == 0)$Patient[1], y=0.5, label=paste(round(nrow(pn[pn$pred.ratio <= 0,])/nrow(np),3)*100,'%',sep='')), size=8, fontface='bold' ) 
}

grid.arrange(p1, p2, ncol=2, top='Validation data set')

```

```{r, echo=F, message=F, warning=F, fig.width=8, fig.height=12}

back = lapply(vpd, transform.pt.by.time)
vpi = do.call(rbind, lapply(1:max(sapply(back,nrow)), function(i) {
  df = do.call(rbind, lapply(back, function(x) {
    if (i > nrow(x)) return(NA)
    cbind(x[i, c('Endoscopy.Year','pred','samples', 'months.before.final', 'max.path')], i)
  }))
  rownames(df) = subset(sum.validation.info, Hospital.Research.ID %in% rownames(df))$Patient
  df$Patient = rownames(df)
  return(df)
}))

vpi = vpi[complete.cases(vpi),]
vpi$pred.ratio = round(vpi$pred/vpi$samples, 2)
vpi = transform(vpi, Patient=reorder(Patient, -months.before.final) ) 
vpi$months.before.final = vpi$months.before.final*-1

me = vpi %>% group_by(Patient) %>% summarise('max.endo' = max(i), 'months' = min(months.before.final))

p1 = plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=T) + labs(title="Validation - Progressors")
print.png(p1, 'validation_progressors.png', width=600, height=800)

print.png(plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'P')$Patient), bin=F) + labs(title="Validation - Progressors"), 'validation_progressors_bar.png', width=600, height=800)

p2 = plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=T) + labs(title="Validation - Non Progressors")

print.png(p2,'validation_Nonprogressors.png', width=600, height=800)
print.png(plot.endo(subset(vpi, Patient %in% subset(sum.validation.info, Status == 'NP')$Patient), bin=F) + labs(title="Validation - Non Progressors"),'validation_Nonprogressors_bar.png', width=600, height=800)

grid.arrange(p1, p2,ncol=1)
```

```{r echo=F, warning=F, message=F}

# TPR
tpr = round(length(which(subset(validation.patient.info, Status == 'P')$Prediction > cutoff))/nrow(subset(validation.patient.info, Status == 'P')), 2)
# FNR
fnr = round(length(which(subset(validation.patient.info, Status == 'P')$Prediction < cutoff))/nrow(subset(validation.patient.info, Status == 'P')), 2)

# FPR
fpr = round(length(which(subset(validation.patient.info, Status == 'NP')$Prediction > cutoff))/nrow(subset(validation.patient.info, Status == 'NP')), 2)
# TNR
tnr = round(length(which(subset(validation.patient.info, Status == 'NP')$Prediction < cutoff))/nrow(subset(validation.patient.info, Status == 'NP')), 2)

m = melt(validation.patient.info, id.vars = c('Status'), measure.vars = c('Prediction'))
p3 = ggplot(m, aes(Status, value, fill=Status)) + 
  geom_jitter(aes(color=Status), width=0.2, show.legend = F) +  geom_violin(alpha=0.8) +
  geom_label(aes(x='P', y=0.75, label=paste('TPR=',tpr*100, '%', sep='')), fill='white') + 
  geom_label(aes(x='P', y=0.25, label=paste('FNR=',fnr*100, '%', sep='')), fill='white') + 
  geom_label(aes(x='NP', y=0.55, label=paste('FPR=',fpr*100, '%', sep='')), fill='white') + 
  geom_label(aes(x='NP', y=0.25, label=paste('TNR=',tnr*100, '%', sep='')), fill='white') + 
  geom_hline(yintercept = cutoff, linetype = 'dashed', color = 'grey27') + 
  labs(title="Validation set predictions per sample", subtitle=paste("Prediction cutoff =",cutoff))

grid.arrange(p3)

print.png(p3, 'validation_predictions.png', width=600, height=800)

```