### 2017-11-03 features-OAC.Rmd
updating glm pipeline

---
title: "Progression Features & OAC"
author: "Sarah Killcoyne"
date: "29 June 2017"
output: 
  html_document: 
    fig_height: 7
    fig_width: 5
    toc: yes
    toc_depth: 4
---

```{r, include=FALSE, warning=F, message=F, echo=F}

library(pander)
library(gridExtra)
library(plyr)
library(ggplot2)
library(reshape2)
library(GenomicRanges)
library(ggrepel)
library(ggfortify)
library(readxl)

source('lib/load_patient_metadata.R')

plot.theme = theme(text=element_text(size=14, face='bold'), panel.background=element_blank(), axis.line=element_line(color='black'), panel.grid.major=element_line(color='grey90'))

## Lists the sequenced patients
sequenced = read_excel('~/Documents/OneDrive/OneDrive - MRC Cancer Unit at University of Cambridge/skillcoyne/OC Data/WGS_Phase 1 complete - cleaned.xlsx', trim_ws=T)
sequenced = subset(sequenced, grepl('^OC', OCCAMS_ID)  )[,c(1:8)]
sequenced$Sample_type = as.factor(tolower(sequenced$Sample_type))
sequenced = subset(sequenced, grepl('barret|normal|tumour', Sample_type))

be.seq = subset(sequenced, grepl('barret', Sample_type))

# n patients sequenced
length(unique(sequenced$OCCAMS_ID))
# ------------------------------ #

# OCCAMS ID to Illumina ID
paired = read.table('~/Documents/OneDrive/MRC Cancer Unit at University of Cambridge/icgc group - icgc_shared/icgc-group/batch_progress/reports/g37/PairedOverviewSummary_g37_bwa.txt', header=T, sep='\t', stringsAsFactors = F)
paired = subset(paired, aligned == '*')
pairedg37 = subset(paired, OCCAMS %in% unique(sequenced$OCCAMS_ID), select=c('OCCAMS','SampleId'))
length(unique(pairedg37$OCCAMS))

paired = read.table('~/Documents/OneDrive/MRC Cancer Unit at University of Cambridge/icgc group - icgc_shared/icgc-group/batch_progress/reports/g1k//PairedOverviewSummary_g1k_bwa.txt', header=T, sep='\t', stringsAsFactors = F)
paired = subset(paired, aligned == '*')
pairedg1k = subset(paired, OCCAMS %in% unique(sequenced$OCCAMS_ID), select=c('OCCAMS','SampleId'))

paired = rbind(pairedg37, pairedg1k)
paired = unique(paired)

length(unique(paired$OCCAMS))

paired[] = lapply(paired[], strip.whitespace)
length(unique(paired$OCCAMS))

sp = do.call(rbind, lapply(paired$SampleId, function(x) unlist(strsplit(x, '_vs_'))))
colnames(sp) = c('tb.sample', 'normal')

paired = cbind(paired, sp)

barretts.pairs = subset(paired, tb.sample %in% be.seq$Illumina_Barcode)

paired = paired[-which(paired$tb.sample %in% be.seq$Illumina_Barcode),]

# ------------------------------ #
# Clinical data
load('~/Data/ICGC/OAC/clinical/labkey_20172706.Rdata', verbose=T)
demo = occams$patients
cols = c(grep('Reason',colnames(demo), value=T),
         grep('Planned',colnames(demo),value=T),
         'FE.EndPoint','RP.Tumor.Size','ST.MainSurgery','RP.HistoryOfNeoAdjuvantTherapy',
         'TR.PalliativeTreatmentModality','TR.CurativeTreatmentModality')
demo[cols] = lapply(demo[cols], as.null)
#demo = oac.demo[unique(sequenced$OCCAMS_ID),]
# ------------------------------ #

# Barrett's adjacency updates from Caitrona
be.updates = data.table::fread("~/Documents/OneDrive/OneDrive - MRC Cancer Unit at University of Cambridge/skillcoyne/OC Data/Barret's Confirmed Data-Caitrona_30092016.txt", stringsAsFactors=F, colClasses=rep('character', 7) )
ba = merge(demo[intersect(rownames(demo), be.updates$`OCCAMS/ID`), c('StudySite','RP.BarrettsAdjacent')], be.updates, by.x='row.names',by.y='OCCAMS/ID') 
rownames(ba) = ba$Row.names

ba = ba[c('RP.BarrettsAdjacent',"Barret's Confirmed")]
colnames(ba) = c('labkey','caitrona')

ba[which(ba$caitrona == ""), 'caitrona'] = NA
ba[which(ba$caitrona == "Y"), 'caitrona'] = "yes"
ba[which(ba$caitrona == "N"), 'caitrona'] = "no"
changed = which(ba[,'labkey'] != ba[,'caitrona'] | is.na(ba[,'labkey']) & !is.na(ba[,'caitrona']))

demo[rownames(ba[changed,]), 'RP.BarrettsAdjacent'] = ba[changed,'caitrona']
# ------------------------------ #

# Copy number data - tumour
read.caveman<-function(s) {
  files = sample.list[[s]]
  if (length(files) < 2) {
    warning(s)
    stop()
  }
  stats = read.table(grep('samplestatistics.csv', files, value=T), row.names=1)

  caveman = read.csv(grep('copynumber.caveman.csv$', files, value=T))
  caveman$ploidy = stats['Ploidy',]
  caveman$rho = stats['rho',]
  caveman$contamination = stats['NormalContamination',]

  caveman$Illumina.Barcode.full = s
  caveman
}

# Tumours
dir = '~/Data/ICGC/BE_OAC_Cohort/tumour/data/ascat'
files = list.files(dir, full.names=T, recursive=T)
samples = list.files(dir, full.names=F)

sample.list = lapply(samples, function(x) {
  grep(x, files, value=T)
} )
names(sample.list) = samples

cnv = do.call(rbind, lapply(samples, read.caveman))
totalCNV = merge(cnv, paired, by.x='Illumina.Barcode.full', by.y='SampleId')

demo.cols = c('StudySubjectID','RP.TStage.PrimaryTumour','RP.BarrettsAdjacent','Weeks.Survival','RP.TumourDifferentiation','DI.ageAtDiagnosis','DI.PatientGender','FE.Patient.Died')
totalCNV = merge(totalCNV, demo[,demo.cols], by.y='StudySubjectID', by.x='OCCAMS')

oacCNV = makeGRangesFromDataFrame(totalCNV, keep.extra.columns = T)

# Barretts ADJACENT SAMPLES
dir = '~/Data/ICGC/BE_OAC_Cohort/BA/data/ascat'
files = list.files(dir, full.names=T, recursive=T)
samples = list.files(dir, full.names=F)

sample.list = lapply(samples, function(x) {
  grep(x, files, value=T)
} )
names(sample.list) = samples

cnv = do.call(rbind, lapply(samples, read.caveman))
beCNV = merge(cnv, barretts.pairs, by.x='Illumina.Barcode.full', by.y='SampleId')
beCNV = merge(beCNV, demo[,demo.cols], by.y='StudySubjectID', by.x='OCCAMS')
beCNV = makeGRangesFromDataFrame(beCNV, keep.extra.columns = T)
# ------------------------------ #

oac.demo = subset(demo, StudySubjectID %in% unique(oacCNV$OCCAMS))
be.demo = subset(demo, StudySubjectID %in% unique(beCNV$OCCAMS))

## Shallow WGS data 
load('~/Data/Ellie/Analysis/5e06/loo.Rdata', verbose=T)
swgsPredictions = pg.samp
rm(pg.samp)

load('~/Data/Ellie/Analysis/5e06/all.pt.alpha.Rdata', verbose=T)
rawSWGSdf = dysplasia.df
a = '0.9'
coefs = coefs[[a]]; model = models[[a]]

hazards = apply( exp(t(rawSWGSdf[, rownames(coefs)])*coefs[,1]), 1, function(x) { sd(x)/mean(x) })
rm(dysplasia.df,fits,plots,performance.at.1se)
# ------------------------------ #

ch = as.data.frame(hazards)
ch$coef = coefs[,1]
ch$stability = rowSums(coefs[,-1])/50
ch$label = rownames(ch)

analysis.files = list.files('~/Data/Ellie/Analysis', full.names=T)
load(grep('Training_patients.Rdata', analysis.files, value=T), verbose=T)

labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

labels = labels[which(names(labels) %in% rownames(rawSWGSdf))]

rm(patient.data)
```

# Questions in the Shallow WGS data

Load the coefficients from the selected model and merge with the copynumber regions identified in (some) of the OAC patients
```{r, echo=F}
regions = do.call(rbind.data.frame, c(strsplit(grep('\\d+', ch$label, value=T), ':|-'), stringsAsFactors=F))
colnames(regions) = c('chr','start','stop')
regions[colnames(regions)] = lapply(regions[colnames(regions)], function(x) as.numeric(x))

# Turn SWGs features into grange object
swgRegions = makeGRangesFromDataFrame(cbind.data.frame(regions, ch[grep('\\d+', ch$label),]), keep.extra.columns = T)
swgRegions = swgRegions[order(-swgRegions$hazards)]

ggplot(ch, aes(coef, hazards, col=hazards > quantile(hazards)[['75%']])) + 
  geom_point(alpha=0.5) + 
  geom_text_repel(aes(label=ifelse(hazards > quantile(hazards)[['75%']], label, '')),show.legend=F) +
  labs(title='Top features coef vs hazard') + theme(legend.position = 'none')
```

## Are gains/losses enriched in the coefficients?
We would expect them to be since these are the regions that the model selects as being the best combination to predict P vs NP.

Certainly they are found at a different rate in gains for P vs NP. Which we would expect.

```{r}
# positive coefficients were related to gains in progressors
gains = subset(ch, coef > 0)

gain = do.call(rbind, lapply(gains$label, function(reg) {
  values = rawSWGSdf[,reg] 
  table(labels[which(values > 1)])
}))

t.test(gain[,1], gain[,2])

# negative coefficients were related to losses in non-progressors
losses = subset(ch, coef < 0)

loss = do.call(rbind, lapply(losses$label, function(reg) {
  values = rawSWGSdf[,reg] 
  table(labels[which(values < -1)])
}))

t.test(loss[,1], loss[,2])

```

This is true of randomly selected regions as well, which we also expect based on the general trend towards disorder we see in P vs NP.

```{r}
sample.reg = sample(setdiff(colnames(rawSWGSdf), ch$label), nrow(ch))

randomGains = do.call(rbind, lapply(sample.reg, function(reg) {
  values = rawSWGSdf[,reg] 
  table(labels[which(values > 1)])
}))

t.test(randomGains[,1], randomGains[,2])

randomLosses = do.call(rbind, lapply(sample.reg, function(reg) {
  values = rawSWGSdf[,reg] 
  table(labels[which(values < -1)])
}))

t.test(randomLosses[,1], randomLosses[,2])

```

However, while there appears to be no difference in the gains found in NP for randomly selected regions vs the coefficients:

pander(wilcox.test(gain[,'0'], randomGains[,'0']), caption="Non progressors, coeffiecents vs randomly selected")

pander(wilcox.test(loss[,'0'], randomLosses[,'0']), caption="Non progressors, coeffiecents vs randomly selected")

There is a significant difference in the selected regions for progressors. Again this is mostly a sanity check:

pander(wilcox.test(gain[,'1'], randomGains[,'1']), caption="Progressors, coeffiecents vs randomly selected gains")

pander(wilcox.test(loss[,'1'], randomLosses[,'1']), caption="Progressors, coeffiecents vs randomly selected losses")

## Two groups of patients?

There appears to be two groups of patients.

Group 1 is patients that show progression consistently ("born bad")
Group 2 is patients that show sudden progression in the final or nearly final sample

So is there a difference between the two in regards to the alterations in the features drawn from the model?
Or in randomly drawn regions?

```{r}
cutoff = 0.5
status = sapply(swgsPredictions, function(x) unique(x$Status))
transform.pt.by.time<-function(pt) {
  pt = pt %>% group_by(PID, Endoscopy.Year) %>% summarise(
    'pred'=sum(as.integer(Prediction > cutoff)), 'samples'=length(PID), 
    #'final.endo'= length(which(grepl('HGD|IMC', Pathology))) > 0,
    'months.before.final'= (max(pt$Endoscopy.Year) - unique(Endoscopy.Year))*12, 
    'max.path'=max(Pathology))
  arrange(pt, -Endoscopy.Year, PID)
}

back = lapply(swgsPredictions[which(status == 'P')], transform.pt.by.time)

# By sample
fvpS = do.call(rbind.data.frame, lapply(back, function(df) {
  last.sample = (df[1,'pred']/df[1,'samples'])[[1]]
  if (nrow(df) <= 1 | max(df$months.before.final) < 12 ) return(NA)
  cbind('final'=last.sample,  'previous'=sum(df[-1, 'pred']/df[-1, 'samples'])/(nrow(df)-1) )
}))

# Do it by endoscopy? Not sample? And max.path should matter
final.vs.prev = do.call(rbind.data.frame, lapply(back, function(df) {
  last.endo = as.integer(df[1,'pred'] > 0)
  if (nrow(df) <= 1 | max(df$months.before.final) < 12 ) return(NA)

  previous = subset(df[-1,], max.path %in% c('BE','ID'))
  cbind('final'=last.endo,  'previous'=sum(as.integer(previous$pred > 0)/(nrow(previous))))
}))

final.vs.prev = final.vs.prev[complete.cases(final.vs.prev),]

m = melt(as.matrix(final.vs.prev))
m = transform(m, Var1=reorder(Var1, -value))

```

First we want to separate out the patient groups we have. In this table we only include patients that have more than a single sample, and with follow up samples 1 or more years prior to their final (HGD/IMC) sample.

```{r, echo=F}
ggplot(m, aes(x=Var1, y=value,fill=Var2)) + 
  geom_bar(stat="identity", width=.5, position = "dodge") + coord_flip() + 
  labs(title='Ratio of final sample predictions vs all previous samples')
```

For ~~samples~~ endoscopies predicted prior to the final our mean ratio is `r mean(final.vs.prev$previous)` so most are predicted. 
```{r, echo=F}
hist(final.vs.prev$previous)
```

When the final ~~sample~~ endoscopy was not predicted correctly the prediction for previous samples also declines to a mean of `r mean(summary(subset(final.vs.prev, final <= 0)$previous))`.

So the two groups are:

Sudden - predicted at endpoint, but not consistently prior 
```{r}
sudden = subset(final.vs.prev, final > 0 & previous < mean(previous))

pander(sudden, caption='Patients whose final endoscopies were predicted, but prior endoscopies were less clear.')
```

Born bad - predicted consistently prior to endpoint

```{r}
born.bad = subset(final.vs.prev, final > 0 & previous > mean(previous))

pander(born.bad, caption='Patients whose endoscopies were predicted throughout')
```

So how can we compare these?
```{r}
count.losses <- function(x) { length(which(x < -1)) }
count.gains <- function(x) { length(which(x > 1)) }

annotate.by.pt<-function(df) {
  df = df[,c('Prediction','Pathology', 'PID', 'Endoscopy.Year', 'Samplename')]
  df$Prediction = as.integer(df$Prediction > cutoff)

  df$Final = df$Pathology %in% c('HGD','IMC') & df$Endoscopy.Year == df$Endoscopy.Year
  df[,c('Prediction','Pathology','Samplename','Final')]
}

sd.pts = do.call(rbind, lapply(swgsPredictions[rownames(sudden)],annotate.by.pt))
bb.pts = do.call(rbind, lapply(swgsPredictions[rownames(born.bad)],annotate.by.pt))

# Can't randomly sample patients here because it's very likely that I'll end up with many of the same patients in both random and sudden/born bad groups
# Could randomly select samples that are not HGD/IMC though...
#rd.pts = do.call(rbind, lapply(swgsPredictions[names(status[which(status == 'P')])],annotate.by.pt))

rd.samp = subset(do.call(rbind, swgsPredictions[names(status[which(status == 'P')])]), Pathology %in% c('BE','ID','LGD'), select=c('Prediction','Pathology','Samplename'))
rd.samp = rd.samp[sample(nrow(rd.samp), 120),]

gains.by.group = cbind.data.frame(
  'sudden'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,gains$label], 2, count.gains),
  'born.bad'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,gains$label], 2, count.gains),
  'random'=apply(rawSWGSdf[rd.samp$Samplename,gains$label], 2, count.gains)
)

losses.by.group = cbind.data.frame(
  'sudden'=apply(rawSWGSdf[subset(sd.pts, !Final)$Samplename,losses$label], 2, count.losses),
  'born.bad'=apply(rawSWGSdf[subset(bb.pts, !Final)$Samplename,losses$label], 2, count.losses),
  'random'=apply(rawSWGSdf[rd.samp$Samplename,losses$label], 2, count.losses)
)
```

This may be driven by the number of samples. I'm not sure if this is correct or how to check at the moment.
In all of these I exclude the "final" sample (typically HGD/IMC).

```{r}
# "born bad" and sudden are different
wilcox.test(gains.by.group$sudden, gains.by.group$born.bad)
wilcox.test(losses.by.group$sudden, losses.by.group$born.bad)

# Random and sudden are different
wilcox.test(gains.by.group$sudden, gains.by.group$random)
wilcox.test(losses.by.group$sudden, losses.by.group$random)

# random and born bad aren't?
wilcox.test(gains.by.group$born.bad, gains.by.group$random) ## not so different
wilcox.test(losses.by.group$born.bad, losses.by.group$random) ## not so different

```

```{r echo=F}
# one-way anova by ranks -- basically the same as wilcox I think
one.way.aov<-function(a, b) {
  df = rbind(cbind.data.frame('counts'=a, 'class'='a'),
             cbind.data.frame('counts'=b, 'class'='b')) 
  kruskal.test(counts~class, data=df)
  #chisq.test(df$counts, df$class)
  #glm(counts~class, data=df, family='poisson')
}
```

So why would random and born bad groups be the same?  Sudden appears to be the only group that stands out in any way, being different from both born bad and random groups.  Sudden has half the samples the other two do, but then random and sudden are different so samplesize should be at issue here.

```{r, echo=F, eval=F}
gains = gains[grep('\\d+',gains$label),]
losses = losses[grep('\\d+',losses$label),]

sd.samples = unlist(lapply(swgsPredictions[rownames(sudden)], function(df) df$Samplename))
bb.samples = unlist(lapply(swgsPredictions[rownames(born.bad)], function(df) df$Samplename))
random.samples = unlist(lapply(swgsPredictions[sample(which(status == 'P'), 15)], function(df) df$Samplename ))

# Born bad losses/gains in features
bbL = apply(rawSWGSdf[bb.samples, losses$label], 2, count.losses)
bbG = apply(rawSWGSdf[bb.samples, gains$label], 2, count.gains)

# sudden losses/gains in features
snL = apply(rawSWGSdf[sd.samples, losses$label], 2, count.losses)
snG = apply(rawSWGSdf[sd.samples, gains$label], 2, count.gains)

# random losses/gains in features
rrL = apply(rawSWGSdf[random.samples, losses$label], 2, count.losses)
rrG = apply(rawSWGSdf[random.samples, gains$label], 2, count.gains)

one.way.aov(bbG, snG)
one.way.aov(bbG, rrG)

one.way.aov(bbL, snL)
one.way.aov(bbL, rrL)

one.way.aov(snL, rrL)
one.way.aov(snG, rrG)
```

```{r, warning=F, message=F, echo=F, fig.width=7, fig.height=10}

totals = melt(cbind.data.frame(
  'sudden'=nrow(subset(sd.pts, !Final)),
  'born.bad'=nrow(subset(bb.pts, !Final)),
  'random'=nrow(rd.samp)))

grid.arrange(
  arrangeGrob(
    ggplot(melt(gains.by.group), aes(variable, value, color=variable, fill=variable)) + geom_jitter(width=0.2) + 
    geom_boxplot(alpha=0.5, outlier.color = NA) + labs(title='Gains', x='', y='counts') + theme(legend.position = 'none'),
  ggplot(melt(losses.by.group), aes(variable, value, color=variable, fill=variable)) + geom_jitter(width=0.2) + 
    geom_boxplot(alpha=0.5, outlier.color = NA) + labs(title='Losses', x='', y='counts') + theme(legend.position = 'none'), ncol=2),

  tableGrob(  rbind(cbind('Group'='Born Bad','# Samples'=nrow(subset(bb.pts, !Final))),
                    cbind('Group'='Sudden','# Samples'=nrow(subset(sd.pts, !Final))),
                    cbind('Group'='Random','# Samples'=nrow(rd.samp)))   ) 
  )

```

### The two groups don't actually exist

TODO  I don't recall what these tests were looking at, need to redo this with the new data

```{r, eval=F}

gains = gains[grep('\\d+',gains$label),]
losses = losses[grep('\\d+',losses$label),]

km = kmeans(rawSWGSdf[c(bb.samples,sd.samples),ch$label], 2)
table(km$cluster)

clusPred = sapply(swgsPredictions[which(status == 'P')], function(df) {
  sum(as.integer(subset(df, Samplename %in% names(which(km$cluster == 1)))$Prediction > cutoff))
})

sum(clusPred)

binom.test(17,17,.72)
names(which(clusPred > 0))
which(rownames(sudden) %in% names(which(clusPred > 0)))
which(rownames(born.bad) %in% names(which(clusPred > 0)))

```

# Questions in OAC data

## Gains

These may be more nuanced, as we need to be aware of gains that reflect WGD vs gains that are new amplifications. At the moment there's no distinction

```{r, fig.height=16, fig.width=16}
plot.gains<-function(df) {
  df = lapply(df, function(x) {
    x = arrange(x, -width)
    x$order = factor(1:nrow(x))
    return(x)
  })

  m = do.call(rbind.data.frame, df)
  m = m %>% group_by(seqnames) %>% ungroup() %>% arrange(segment, seqnames, width, start) %>% mutate(segorder = row_number())
m$isWGD = as.factor(m$isWGD)

m$CN = cut(m$Total_CN, include.lowest=T, ordered_result=T, breaks=c(3, 5, 7, 9, 10, max(m$Total_CN)))
ggplot(data=m, aes(x=order, y=start)) +  
  facet_wrap(~segment, scales='free') +
  geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
  geom_segment(aes(x=0,xend=0, y=seg.start, yend=seg.end), color='red', size=2) +
  scale_colour_brewer(palette='YlGnBu', direction=-1) +
  coord_flip() + plot.theme +
  theme(legend.position='bottom', axis.text=element_blank(),panel.grid.major = element_blank()) + labs(y='chr position', x='Patient', title='Gains')
}

# WGD
oacCNV$isWGD = oacCNV$ploidy > 3.2 
beCNV$isWGD = beCNV$ploidy > 3.2

#average genome ploidy <= 2.7 AND total copy number >= 5
#OR average genome ploidy > 2.7 AND total copy number >= 9

gain.for<-function(gg, cnv) {
  ov = findOverlaps(cnv, gg)  
  tmp = cnv[queryHits(ov)]

  if (length(tmp) > 0) {
    #tmp = subset(tmp, ploidy > 2 & Total_CN > round(ploidy))
    tmp = subset(tmp, 
       (!isWGD & Total_CN > round(ploidy)) | 
         (isWGD & Total_CN > round(ploidy+3.2)) )
    if (length(tmp) > 0) {
      tmp$segment = as.character(gg)
      tmp$seg.start = start(gg)
      tmp$seg.end = end(gg)
    }
  }
  as.data.frame(tmp)
}

gr = which(swgRegions$coef > 0)

## gains in patients -- issues with WGD samples?
gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV)
names(gains) = as.character(swgRegions[gr])
plot.gains(gains) + labs(title='Gains in OAC samples')

be.gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')])
names(be.gains) = as.character(swgRegions[gr])
plot.gains(be.gains[sapply(be.gains, nrow) > 0]) + labs(title='Gains in BE Adjacent samples')

non.be.gains = lapply(swgRegions[gr], gain.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')])
names(non.be.gains) = as.character(swgRegions[gr])
plot.gains(non.be.gains[sapply(non.be.gains, nrow) > 0]) + labs(title='Gains in NO BE Adjacent samples')

```

```{r, fig.height=8, fig.width=12}

# Per patient
gfreq = sort(sapply(gains, function(x) length(unique(x$OCCAMS)))/length(unique(sequenced$OCCAMS_ID)))
ggplot(melt(as.matrix(gfreq)), aes(Var1, value)) + geom_col(fill='darkblue') + ylim(0,0.5) + coord_flip() + 
  labs(x='', y='Frequency', title='Gain frequency in OAC') + theme_bw() + theme(text=element_text(size=14))

gfreq1 = sort(sapply(be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')]$OCCAMS)))
gfreq1 = gfreq1[names(gfreq)]

gfreq2 = sort(sapply(non.be.gains, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')]$OCCAMS)))
gfreq2 = gfreq2[names(gfreq)]

m = cbind.data.frame('yes'=gfreq1, 'no'=gfreq2)
m$region = rownames(m)
m = transform(m, region=reorder(region, yes))

ggplot(melt(m), aes(region, value, group=variable, fill=variable)) + facet_grid(~variable) + geom_bar(stat='identity', position=position_dodge(), alpha=0.8) + 
  scale_fill_manual(values=c('dodgerblue3','firebrick3'), name='Barretts Adjacent') +
  coord_flip() + labs(x='', y='Frequency', title="Gain frequency BE vs non") + plot.theme + theme(legend.position = 'bottom')

```

## Losses

```{r, fig.height=12, fig.width=12}
plot.losses<-function(df) {
  if (!is.data.frame(df)) {
    df = lapply(df, function(x) {
      if(nrow(x) <= 0) return(NULL)
      x = arrange(x, -width)
      x$order = factor(1:nrow(x))
      return(x)
    })
    m = do.call(rbind.data.frame, df)
  } else {
    m = df
    m = arrange(m, -width)
    m$order = factor(1:nrow(m))
  }
  #m$segnames = ordered(m$seqnames, levels=c(1:22))

  m = m %>% group_by(seqnames) %>% ungroup() %>% arrange(segment, seqnames, width, start) %>% mutate(segorder = row_number())
  m$isWGD = as.factor(m$isWGD)
  m$CN = as.factor(m$Total_CN)
  #m = transform(m, segment=reorder(segment, segnames))
  p = ggplot(data=m, aes(x=order, y=start))

  if (length(unique(m$segment)) > 1)
    p = p + facet_wrap(~segment, scales='free')

  p + geom_segment(aes(x=order, xend=order, y=start, yend=end, color=CN)) + 
    geom_segment(aes(x=0,xend=0, y=seg.start, yend=seg.end), color='red', size=2) +
    scale_colour_brewer(palette='YlGnBu', direction=1) +
    coord_flip() + plot.theme +
    theme(legend.position='bottom', axis.text=element_blank(), panel.grid.major=element_blank()) + labs(y='chr position', x='Patient', title='Losses in OAC Patients')
}

# Losses in patients -- complicated by WGD
##average genome ploidy <= 2.7 AND total copy number = 0
#OR average genome ploidy > 2.7 AND total copy number < ( average genome ploidy - 2.7 )
loss.for<-function(gg, cnv){
  ov = findOverlaps(cnv, gg)  
  tmp = cnv[queryHits(ov)]
  if (length(tmp) > 0) {
    tmp = subset(tmp, (!isWGD & Total_CN == 0) |
                    (isWGD & Total_CN < round(ploidy)) )
    #tmp = tmp[tmp$Total_CN < 2]
    if (length(tmp) > 0) {
      tmp$segment = as.character(gg)
      tmp$seg.start = start(gg)
      tmp$seg.end = end(gg)
    }
  }
  as.data.frame(tmp)
}

lr = which(swgRegions$coef < 0)

losses = lapply(swgRegions[lr], loss.for, oacCNV)
names(losses) = as.character(swgRegions[lr])
plot.losses(losses)

be.losses = lapply(swgRegions[lr], loss.for, oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')])
names(be.losses) = as.character(swgRegions[lr])
plot.losses(be.losses) + labs(title="Losses in (YES) BE Adjacent samples")

non.be.losses = lapply(swgRegions[lr], loss.for, cnv=oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')])
names(non.be.losses) = as.character(swgRegions[lr])
plot.losses(non.be.losses) + labs(title='Losses in (NO) BE Adjacent samples')

# grid.arrange(plot.losses(be.losses$`16:76088214-80843726`) + labs(title='BE loss'),
#              plot.losses(non.be.losses$`16:76088214-80843726`) + labs(title='(NO) BE loss') )

```

```{r, fig.height=8, fig.width=12}

lfreq = sort(sapply(losses, function(x) length(unique(x$OCCAMS)))/length(unique(sequenced$OCCAMS_ID)))
ggplot(melt(as.matrix(lfreq)), aes(Var1, value)) + geom_col(fill='darkgreen') + ylim(0,1) + coord_flip() + 
  labs(x='', y='Frequency', title='Loss frequency in OAC') + theme_bw() + theme(text=element_text(size=14))

lfreq1 = sapply(be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'yes')]$OCCAMS))
lfreq1 = lfreq1[names(lfreq)]

lfreq2 = sort(sapply(non.be.losses, function(x) length(unique(x$OCCAMS)))/length(unique(oacCNV[which(oacCNV$RP.BarrettsAdjacent == 'no')]$OCCAMS)))
lfreq2 = lfreq2[names(lfreq)]

m = cbind.data.frame('yes'=lfreq1, 'no'=lfreq2)
m$region = rownames(m)
m = transform(m, region=reorder(region, yes))

ggplot(melt(m), aes(region, value, group=variable, fill=variable)) + facet_grid(~variable) + geom_bar(stat='identity', position=position_dodge(), alpha=0.8) + 
  scale_fill_manual(values=c('dodgerblue3','firebrick3'), name='Barretts Adjacent') +
  coord_flip() + labs(x='', y='Frequency', title="Loss frequency BE vs non") + plot.theme + theme(legend.position = 'bottom')

```

```{r fig.height=6,fig.width=8}

## Note it's a subset of the 321 patients (66) that have this
chr17 = get.chr.lengths()[17,]

#chr17losses = lapply(makeGRangesFromDataFrame( cbind.data.frame('start'=0, 'end'=chr17$chr.length, 'chr'='17') ), loss.for)[[1]]
#m = arrange(chr17losses, width, start)

plot.losses(losses$`17:1-4776188`) + geom_hline(data=chr17, aes(yintercept=(chr.cent-cent.gap)), color='red', linetype='dashed') +
  labs(y='chr17 position', x='Patient', title='Chr17 Losses in OAC Patients', subtitle=paste(round(length(unique(losses$`17:1-4776188`$OCCAMS))/length(unique(oacCNV$OCCAMS)), 2)*100, '% of OAC patients', sep=''))

```

So is there a difference between the gains/losses in the selected regions?  
```{r}
wilcox.test(sapply(gains, nrow), sapply(losses, nrow))
one.way.aov(sapply(gains, nrow), sapply(losses, nrow))
```

Not between be/non be either
```{r}
wilcox.test(sapply(non.be.gains, nrow), sapply(be.gains, nrow))
wilcox.test(sapply(non.be.losses, nrow), sapply(be.losses, nrow))
```

## Random

Compare those to gains/losses in randomly selected regions of the same size across the genome

```{r, echo=F, warning=F, message=F}
chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths$start = 1

genome = makeGRangesFromDataFrame(chr.lengths[1:22,], seqnames.field = 'chrom', end.field='chr.length')
tiles = slidingWindows(genome, width=5e6, step=2.5e6)
chr.samples = sample(length(tiles), length(tiles)/2) # half of the chromosomes
sample.regions = do.call(c, lapply(tiles[chr.samples], function(x) sample(x, length(x)/4)  )) # grab some portion of each
```

### Gains

These may be more nuanced, as we need to be aware of gains that reflect WGD vs gains that are new amplifications. At the moment there's no distinction

```{r}
randomGains = lapply(sample.regions, gain.for, cnv=oacCNV)
names(randomGains) = as.character(sample.regions)

plot.gains(randomGains[sample(length(randomGains), 20)])

# Maybe a difference in # gains in the selected features vs random
wilcox.test(sapply(randomGains, nrow), sapply(gains, nrow))

wilcox.test(sapply(randomGains, nrow), sapply(be.gains,nrow))
wilcox.test(sapply(randomGains, nrow), sapply(non.be.gains,nrow))

# Not a difference here though
wilcox.test(sapply(gains, function(df) length(unique(df$OCCAMS))), 
            sapply(randomGains, function(df) length(unique(df$OCCAMS))))

df = rbind(cbind.data.frame('counts'=sapply(gains, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomGains, function(df) length(unique(df$OCCAMS))), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Patient counts by region') + plot.theme

df = rbind(cbind.data.frame('counts'=sapply(gains, nrow), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomGains, nrow), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Gain counts per region') + plot.theme

```

```{r, echo=T}
df = rbind(cbind.data.frame('counts'=sapply(non.be.gains, nrow), 'type'='non.be'),
           cbind.data.frame('counts'=sapply(be.gains, nrow), 'type'='be'),
           cbind.data.frame('counts'=sapply(gains, nrow), 'type'='all.selected'),
           cbind.data.frame('counts'=sapply(randomGains, nrow), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type, color=type)) + geom_boxplot(alpha=0.5, outlier.colour = NA) + 
  geom_jitter(width=0.3) + labs(title='Gain counts per region') + plot.theme

wilcox.test(subset(df, type == 'random')$counts, subset(df, type == 'be')$counts)
wilcox.test(subset(df, type == 'random')$counts, subset(df, type == 'non.be')$counts)
wilcox.test(subset(df, type == 'random')$counts, subset(df, type == 'all.selected')$counts)

```

### Losses

```{r}
randomLosses = lapply(sample.regions, loss.for, cnv=oacCNV)
names(randomLosses) = as.character(sample.regions)

plot.losses(randomLosses[sample(length(randomLosses), 15)])

# Definite difference between the numbers of gains and losses in randomly selected regions
wilcox.test(sapply(randomGains, nrow), sapply(randomLosses, nrow))

# No difference in # losses in the selected features
wilcox.test(sapply(randomLosses, nrow), sapply(losses, nrow))

wilcox.test(sapply(randomLosses, nrow), sapply(be.losses,nrow))
wilcox.test(sapply(randomLosses, nrow), sapply(non.be.losses,nrow))

# No difference between the two
df = rbind(cbind.data.frame('counts'=sapply(losses, function(df) length(unique(df$OCCAMS))), 'type'='selected') , 
      cbind.data.frame('counts'=sapply(randomLosses, function(df) length(unique(df$OCCAMS))), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type)) + geom_boxplot() + geom_jitter(width=0.3) + labs(title='Loss counts by patient') + plot.theme

df = rbind(cbind.data.frame('counts'=sapply(non.be.losses, nrow), 'type'='non.be'),
           cbind.data.frame('counts'=sapply(be.losses, nrow), 'type'='be'),
           cbind.data.frame('counts'=sapply(losses, nrow), 'type'='all.selected'),
           cbind.data.frame('counts'=sapply(randomLosses, nrow), 'type'='random'))
ggplot(df, aes(type, counts, group=type, fill=type, color=type)) + geom_boxplot(alpha=0.5) + 
  geom_jitter(width=0.3) + labs(title='Loss counts per region') + plot.theme

```

## Cox

```{r}
library(survival)
library(GGally)

pt.gains = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
  rbind(sapply(gains, function(g)  nrow(subset(g, OCCAMS == id))))
}))
rownames(pt.gains) = oac.demo$StudySubjectID
colnames(pt.gains) = paste('gain_', gsub(':|-', '\\.', colnames(pt.gains)), sep='')

pt.losses = do.call(rbind, lapply(oac.demo$StudySubjectID, function(id) {
  rbind(sapply(losses, function(g)  nrow(subset(g, OCCAMS == id))))
}))
rownames(pt.losses) = oac.demo$StudySubjectID
colnames(pt.losses) = paste('loss_',gsub(':|-', '\\.', colnames(pt.losses)), sep='')

x = merge(pt.gains, pt.losses, by='row.names', all=T)

x = merge(x,oac.demo[,demo.cols], by.x='Row.names', by.y='StudySubjectID')
x$Row.names = NULL

time = 'Weeks.Survival';  event='FE.Patient.Died'; type = '"right"'
x[c(time,event)] = lapply(x[c(time,event)], as.numeric)

x$RP.TStage.PrimaryTumour = as.character(x$RP.TStage.PrimaryTumour)

x[grepl('T1', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T1'
x[grepl('T4', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T4'
x[grepl('Tx', x$RP.TStage.PrimaryTumour), 'RP.TStage.PrimaryTumour'] = 'T0'

terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~ RP.BarrettsAdjacent", sep=""))
sfit = survfit(terms, data=x,conf.int=0.95, conf.type='log', na.action=na.omit)
ggsurv(sfit)

f1 = paste(demo.cols[c(3,5:7)], collapse='+')
f2 = paste(paste('`', grep('^gain|loss', colnames(x), value=T), '`', sep=''), collapse='+')

f = paste(f1,f2, sep='+')

terms = as.formula(paste("Surv(",time,",",event,", type=",type,") ~", f, sep=""))
cox = coxph(terms, data=x,na.action=na.omit,ties='breslow')
cfit = survfit(cox)
ggsurv(cfit) + labs(title=paste(length(grep('gain|loss', colnames(x))), 'copy number + demo'))

pander(cox)

anova(cox)
```

```{r}
library(CoxHD)

x = x[,c(grep('gain|loss',colnames(x), invert=T), grep('gain|loss',colnames(x)))]
xs = x

surv = with(xs, Surv(Weeks.Survival, FE.Patient.Died))
ggsurv(survfit(surv~1,data=xs))

xs = xs[,-grep('Weeks|Died',colnames(xs))]

xs$RP.BarrettsAdjacent = as.integer(xs$RP.BarrettsAdjacent)-1
xs$RP.TumourDifferentiation = as.integer(as.factor(xs$RP.TumourDifferentiation))-1
xs$DI.PatientGender = as.integer(as.factor(xs$DI.PatientGender))-1

groups = c(rep('clin', 4), rep('cn',(ncol(pt.gains)+ncol(pt.losses)) ))

rfx = CoxRFX(xs, surv, groups, nu=1, verbose=T)
summary(rfx)

plot(rfx)

VarianceComponents(rfx)

WaldTest(rfx)

PartialC(rfx)

PlotVarianceComponents(rfx)

```

## Focal vs Large CN

It's worth looking to see if focal/large changes are enriched in the gains particularly as the plots seemed to suggest they may be.

#### Gains

```{r, fig.width=8}
size = unlist(sapply(gains, function(df) df$width))
hist(size)
abline(v = 3e6, col='red')

#chr.lengths$chr.length*.01

# Focal changes 1% or the chromosome?
countBySize <-function(df) {
  s = round(subset(chr.lengths, chrom == unique(df$seqnames))$chr.length *.01)
  cbind('focal'=nrow(subset(df, width < s)),
        'large'=nrow(subset(df, width >= s)))
}

sizebyPatient<-function(df) {
  s = round(subset(chr.lengths, chrom == unique(df$seqnames))$chr.length *.01)
  cbind('focal'=length(unique(subset(df, width < s)$OCCAMS)),
        'large'=length(unique(subset(df, width >= s)$OCCAMS)))
}

# sizebyPatient<-function(x) {
#   cbind('focal'=length(unique(subset(x, width <= quantile(width)[['25%']])$OCCAMS)),
#         'large'=length(unique(subset(x, width > quantile(width)[['25%']])$OCCAMS)) )
# }

# countBySize<-function(x) {
#   cbind('focal'=nrow(subset(x, width < quantile(width)[['25%']])),
#         'large'=nrow(subset(x, width >= quantile(width)[['25%']])) )
# }

sizeGainsByPatient = do.call(rbind.data.frame,lapply(gains, sizebyPatient))
sizeGains = do.call(rbind.data.frame,lapply(gains, countBySize))

#  enrichment for some focal changes
x = which(p.adjust(apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
apply(sizeGains, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )[,x]

## Not sure what to do with this really
sizeGainsBE = do.call(rbind.data.frame, lapply(be.gains, countBySize))
which(p.adjust(apply(sizeGainsBE, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
sizeGainsNonBE = do.call(rbind.data.frame, lapply(non.be.gains, countBySize))
which(p.adjust(apply(sizeGainsNonBE, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)

# Same is true for gains of type by patient
xx = which(p.adjust(apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
apply(sizeGainsByPatient, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )[,xx]

## Not sure what to do with this really
sizeGainsBEByPt = do.call(rbind.data.frame, lapply(be.gains, sizebyPatient))
which(p.adjust(apply(sizeGainsBEByPt, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)
sizeGainsNonBEByPt = do.call(rbind.data.frame, lapply(non.be.gains, sizebyPatient))
which(p.adjust(apply(sizeGainsNonBEByPt, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value )) < 0.05)

# Clear difference between focal/large
wilcox.test(sizeGainsByPatient$focal, sizeGainsByPatient$large)
one.way.aov(sizeGainsByPatient$focal, sizeGainsByPatient$large)

sizeGainsRatio = sizeGains %>% rowwise() %>% summarise( 'focal.r'=focal/sum(focal,large), 'large.r'=large/sum(focal,large) )
sizeGainsRatio$feature = rownames(sizeGains)

# No difference in the overall number of gains per region
ggplot(melt(sizeGains), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) 
  labs(title="Types of gain per region", subtitle=paste('p=',format.pval(one.way.aov(sizeGains$focal, sizeGains$large)$p.value,3), sep=''), y='patients per region', x='')

grid.arrange(
  ggplot(melt(sizeGainsRatio), aes(feature, value, fill=variable) ) + geom_col() + coord_flip() + labs(title='Ratio of overall gains per region'),
  ggplot(melt(sizeGainsByPatient), aes(variable, value, group=variable, fill=variable)) + 
    geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
    labs(title="Patients per region", subtitle=paste('p=',format.pval(one.way.aov(sizeGainsByPatient$focal, sizeGainsByPatient$large)$p.value,digits=3), sep=''), y='patients per region', x=''),
top='Gains in selected regions', ncol=2)
```

#### Losses

```{r, fig.width=8}
sizeLossesByPatients = do.call(rbind.data.frame,lapply(losses, sizebyPatient))
sizeLosses = do.call(rbind.data.frame,lapply(losses, countBySize))

sizeLossBE = do.call(rbind.data.frame, lapply(be.losses, countBySize))
sizeLossNoBE = do.call(rbind.data.frame, lapply(non.be.losses, countBySize))

wilcox.test(sizeLossBE$large, sizeLossNoBE$large)

# True for losses as well
p.adjust(apply(sizeLosses, 1, function(x) binom.test(x[1], sum(x), 0.5)$p.value ))
apply(sizeLosses, 1, function(x) binom.test(x[1], sum(x), 0.5)$conf.int )

wilcox.test(sizeLosses$focal, sizeLosses$large)
one.way.aov(sizeLosses$focal, sizeLosses$large)

sizeLossRatio = sizeLosses %>% rowwise() %>% summarise( 'focal.r'=focal/sum(focal,large), 'large.r'=large/sum(focal,large) )
sizeLossRatio$feature = rownames(sizeLosses)

ggplot(melt(sizeLossRatio), aes(feature, value, fill=variable) ) + geom_col() + coord_flip() + labs(title='Overall losses')

grid.arrange(
ggplot(melt(sizeLosses), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Counts per region", subtitle=paste('p=',format.pval(one.way.aov(sizeLosses$focal, sizeLosses$large)$p.value,3), sep=''), y='counts per region', x=''),

ggplot(melt(sizeLossesByPatients), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Patients per region", subtitle=paste('p=',format.pval(one.way.aov(sizeLossesByPatients$focal, sizeLossesByPatients$large)$p.value,3), sep=''), y='patients per region', x=''),
top='Losses in selected regions', ncol=2)

wilcox.test(sizeLossesByPatients$focal, sizeLossesByPatients$large)
```

### Compared to random focal/large changes?

#### Losses
```{r, fig.width=8}
randomLossesByPt = do.call(rbind.data.frame,lapply(randomLosses, sizebyPatient))
randomLossesBySize = do.call(rbind.data.frame,lapply(randomLosses, countBySize))

grid.arrange(
ggplot(melt(randomLossesBySize), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Counts per region", subtitle=paste('p=',format.pval(one.way.aov(randomLossesBySize$focal, randomLossesBySize$large)$p.value,digits=3), sep=''), y='counts per region', x='') + theme(legend.position = 'none'),

ggplot(melt(randomLossesByPt), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Patients per region", subtitle=paste('p=',format.pval(one.way.aov(randomLossesByPt$focal, randomLossesByPt$large)$p.value,digits=3), sep=''), y='patients per region', x='') + theme(legend.position = 'none'),
top='Losses in randomly selected regions', ncol=2)

# No real difference in randomly selected regions for focal losses
wilcox.test(randomLossesBySize$focal, sizeLosses$focal) 
one.way.aov(randomLossesBySize$focal, sizeLosses$focal) 
wilcox.test(randomLossesByPt$focal, sizeLossesByPatients$focal)

## Difference in large losses
wilcox.test(randomLossesBySize$large, sizeLosses$large)
one.way.aov(randomLossesBySize$large, sizeLosses$large)
wilcox.test(randomLossesByPt$large, sizeLossesByPatients$large) 
```

#### Gains

```{r, fig.width=8}
randomGainsByPt = do.call(rbind.data.frame,lapply(randomGains, sizebyPatient))
randomGainsBySize = do.call(rbind.data.frame,lapply(randomGains, countBySize))

grid.arrange(
ggplot(melt(randomGainsBySize), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Counts per region", subtitle=paste('p=',format.pval(one.way.aov(randomGainsBySize$focal, randomGainsBySize$large)$p.value,3), sep=''), y='counts per region', x='') + theme(legend.position = 'none'),

ggplot(melt(randomGainsByPt), aes(variable, value, group=variable, fill=variable)) + 
  geom_boxplot(alpha=0.5, outlier.colour = NA) + geom_jitter(width=0.2, aes(color=variable)) + 
  labs(title="Patients per region", subtitle=paste('p=',format.pval(one.way.aov(randomGainsByPt$focal, randomGainsByPt$large)$p.value,3), sep=''), y='patients per region', x='') + theme(legend.position = 'none'),
top='Gains in randomly selected regions', ncol=2)

# But maybe a difference in random regions vs selected ones?
wilcox.test(randomGainsBySize$focal, sizeGains$focal) 
wilcox.test(randomGainsBySize$large, sizeGains$large)

wilcox.test(randomGainsByPt$focal, sizeGainsByPatient$focal) # maybe
wilcox.test(randomGainsByPt$large, sizeGainsByPatient$large) # maybe
```

*** 
# Summary to this point

Between the regions selected by the SWGs work and randomly selected regions of the same size across the genome only losses appear to show a difference in the rate of CN suggesting some enrichment in the selected regions.  It's possible that gains need to be broken down differently, for instance separating out gains after WGD vs gains pre-WGD.  For this I need to apply a clonal analysis timing (Mortiz's work).

There's a significant difference in the rate of large-scale alterations vs focal for both gains and losses, but again no difference in gains between selected and random regions, only in losses.  This should be reanalysed based on WGD timing as well.

***

# Barrett's Adjacency & CNA

Is there a relationship between BA patients and CNA in selected regions?

## Selected Regions

```{r, eval=F}
# A lot more gains than losses as expected
sapply(gains, nrow)
sapply(losses, nrow)

population = table(oac.demo$RP.BarrettsAdjacent)

# Check for enrichment of gains/losses in BA
gainsBAPts = as.data.frame(do.call(rbind, lapply(gains, function(g) {
 table(g$RP.BarrettsAdjacent)
})))

wilcox.test(gainsBAPts$no, gainsBAPts$yes)

# Separated by BA status
gainsBA = lapply(gains, function(g) {
  subset(g, RP.BarrettsAdjacent == 'yes')
})

gainsNoBA = lapply(gains, function(g) {
  subset(g, RP.BarrettsAdjacent == 'no')
})

wilcox.test( sapply(be.gains, nrow), sapply(non.be.gains, nrow))
# No difference
one.way.aov(sapply(gainsBA, nrow) , sapply(gainsNoBA, nrow))
# TODO Is this due to numberof patients? Might be
one.way.aov(sapply(gainsBA, nrow)/gainsBAPts[,'yes'] , sapply(gainsNoBA, nrow)/gainsBAPts[,'no'] )

# No enrichment of gains in selected regions for BA 
p.adjust( apply(gainsBAPts, 1, function(x) {
  bt = binom.test(x[['no']], sum(x), population[['no']]/sum(population))  
  bt$p.value
}) )

cnBA = do.call(rbind, lapply(subset(oac.demo, !is.na(RP.BarrettsAdjacent))$StudySubjectID, function(s) {
  cbind.data.frame(
    'ID'=s,
    'BarrettsAdjacent'=subset(oac.demo, StudySubjectID == s)$RP.BarrettsAdjacent,
    'TotalCN'=length(subset(oacCNV, OCCAMS == s)),
    'Gains'=length(subset(oacCNV, OCCAMS == s  & ploidy > 2 & Total_CN > round(ploidy))),
    'Losses'=length(subset(oacCNV, OCCAMS == s  & Total_CN < 2)) 
  )
}))

arrange(cnBA, -TotalCN)

populationGains = apply(oacCNV, 1, function(x) {
  yes = subset(oac.demo, StudySubjectID %in% x$OCCAMS & RP.BarrettsAdjacent == 'yes')$StudySubjectID
  subset(g, OCCAMS %in% yes)

})

apply(cbind(sapply(gainsBA, nrow), sapply(gainsNoBA, nrow)), 1, function(x) {
  print(x)
  bt = binom.test(x[1], sum(x), .5)
  bt$p.value
})

# But there is a significant difference between adjacent BA -- but is this true across random selected regions as well?
gainCtBA = as.data.frame(do.call(rbind, lapply(losses, function(g) {
 table(subset(oac.demo, StudySubjectID %in% g$OCCAMS)$RP.BarrettsAdjacent)
})))

one.way.aov(gainCtBA$no, gainCtBA$yes)
wilcox.test(gainCtBA$no, gainCtBA$yes) 
summary(aov(gainCtBA$no~gainCtBA$yes)) # significant difference between adjacent BA -- but is this true across random selected regions as well?
# But not when adjusted for the population size, though not sure I should do that.  Binom instead?
one.way.aov(gainCtBA$no/population[['no']], gainCtBA$yes/population[['yes']])

lossCtBA = as.data.frame(do.call(rbind, lapply(losses, function(g) {
 table(subset(oac.demo, StudySubjectID %in% g$OCCAMS)$RP.BarrettsAdjacent)
})))

wilcox.test(lossCtBA$no, lossCtBA$yes) # significant difference between adjacent BA -- but is this true across random selected regions as well?
summary(aov(lossCtBA$no~lossCtBA$yes)) # significant difference between adjacent BA -- but is this true across random selected regions as well?

overlaps = rbind(gainCtBA, lossCtBA) %>% rowwise() %>% mutate(
   'yes.CI.1'=binom.test(yes, sum(yes,no), population[['yes']]/sum(population) )$conf.int[1],
   'yes.CI.2'=binom.test(yes, sum(yes,no), population[['yes']]/sum(population) )$conf.int[2],
   'yes.pval'=binom.test(yes, sum(yes,no), population[['yes']]/sum(population) )$p.value,
   'no.pval'=binom.test(no, sum(yes,no), population[['no']]/sum(population) )$p.value
) 
overlaps$regions = c(rownames(gainsBA), rownames(lossBA))

# not significant though
p.adjust(overlaps$yes.pval, method='bonferroni', n=nrow(overlaps))
p.adjust(overlaps$no.pval, method='bonferroni', n=nrow(overlaps))

df = transform(overlaps, regions=reorder(regions, (yes.CI.2-yes.CI.1)) )
ggplot(df , aes(x=regions)) + 
    geom_segment(data=df, aes(y=yes.CI.1, yend=yes.CI.2, x=regions,xend=regions), color='blue') +
    coord_flip() + labs(x='', y='CI (yes)')

```

General gains vs losses
```{r,eval=F}

head(perms)
# No difference in the numbers of gains/losses
wilcox.test(sapply(gains, nrow), perms$amp)
wilcox.test(sapply(losses, nrow), perms$del)

# There is a difference (these are the sampled regions, not selected) between the BA categories, yes -> more disordered??
wilcox.test(perms$amp.no, perms$amp.yes)
wilcox.test(perms$del.no, perms$del.yes)

ggplot(melt(perms[,c('amp.yes','amp.no')]), aes(variable, value, group=variable, fill=variable)) + geom_boxplot()
ggplot(melt(perms[,c('del.yes','del.no')]), aes(variable, value, group=variable, fill=variable)) + geom_boxplot()

random = as.data.frame(t(apply(perms[,grep('yes|no', colnames(perms))], 1, function(row) {

  yes = sum(row[c('amp.yes','del.yes')])
  no = sum(row[c('amp.no','del.no')])

  cbind('yes' = binom.test(yes, sum(yes,no), population[['yes']]/sum(population))$conf.int,
        'no' = binom.test(no, sum(yes,no), population[['yes']]/sum(population))$conf.int)
  # binom.test(yes, sum(yes,no), population[['yes']]/sum(population))$conf.int
})))
colnames(random) = c('yes.CI.1','yes.CI.2','no.CI.1','no.CI.2')

summary( with(random, yes.CI.2-yes.CI.1)  )
summary( with(overlaps, yes.CI.2-yes.CI.1)  )

summary( with(random, no.CI.2-no.CI.1)  )
summary( with(overlaps, no.CI.2-no.CI.1)  )

# No difference between the features selected and random regions though
wilcox.test(gainsBA$yes, perms$amp.yes)
wilcox.test(gainsBA$no, perms$amp.no)

# Nope
colSums(lossBA)/nrow(lossBA)
colSums(perms[,c('del.no','del.yes')])/nrow(perms)

t.test(lossBA$yes, perms$del.yes)
t.test(lossBA$no, perms$del.no)

wilcox.test(gfreq, perms$freq.gain)
wilcox.test(lfreq, perms$freq.loss)

m = rbind(
cbind.data.frame('freq'=gfreq, 'type'='gain', 'feature'='swgs'),
cbind.data.frame('freq'=lfreq, 'type'='loss', 'feature'='swgs'),

cbind.data.frame('freq'=perms$freq.gain, 'type'='gain', 'feature'='random'),
cbind.data.frame('freq'=perms$freq.loss, 'type'='loss', 'feature'='random') 
)

# No difference
ggplot(m, aes(feature, freq, group=feature, fill=feature)) + facet_grid(~type) + geom_jitter(width=0.2) +  geom_violin(alpha=0.8) 
```

```{r, eval=F}
all = swgRegions

swgRegions = swgRegions[swgRegions$hazard > quantile(swgRegions$hazard)['75%']]

cols = c('RP.BarrettsAdjacent', 'RP.TStage.PrimaryTumour', 'Weeks.Survival', 'FE.Patient.Died')
tmp = oac.demo[,cols]

tmp = tmp[-which(is.na(tmp$Weeks.Survival)),]

swgRegions = swgRegions[order(-swgRegions$hazard)]

```

```{r, eval=F}

#summary(as.data.frame(swgRegions)[,c('num.amp.CN','num.del.CN')])

library(gridExtra)
source('../ICGC-Clinical/libs/surv.R')

tmp$FE.Patient.Died = as.numeric(tmp$FE.Patient.Died)-1

tmp[which(is.na(tmp$FE.Patient.Died)), 'FE.Patient.Died'] = 0

sp = survival.plots(tmp, 'RP.BarrettsAdjacent', verbose=F)
grid.arrange(sp$kapmei)

time='Weeks.Survival'; event='FE.Patient.Died'
f = paste('`',paste(grep('AMP|DEL',colnames(tmp), value=T), collapse='`+`'), '`', sep='')

#f = paste('`AMP_', as.character(seqnames(swgRegions[1])), ':', start(swgRegions[1]), '-', end(swgRegions[1]), '`', sep='')

# surv =  Surv(tmp[[time]],tmp[[event]],type='right')
# terms = as.formula(paste("Surv(",time,",",event,", type='right') ~", 'RP.BarrettsAdjacent', sep=""))
# 
# cox = coxph(terms, data=tmp,na.action=na.omit,ties='breslow')
# cfit = survfit(cox)
#  
# ggsurv(cfit,cens.shape=3,lty.est=1,back.white=F, xlab="Time (weeks)") + 
#       annotate("text", x=200, y=.8,label=paste('Score (logrank) pvalue =', signif(cox.pvalue(cox),2),"\n", nrow(tmp), 'samples'))
# summary(cox)

```