### 2018-01-05 coxRFX-forSWGS.Rmd

---
title: "coxRFX for SWGS"
output: html_notebook
---

```{r, include=FALSE}

library(ggplot2)
library(ggrepel)
library(CoxHD)
library(plyr)

plot.waldtest<-function(wt) {
  p = ggplot(wt, aes(coef.x, coef.y, color=p.value<0.05))
  if (length(grep('hazards', colnames(wt))) > 0) {
      p = p + geom_point(alpha=0.5, aes(size=hazards)) + 
        geom_point(aes(coef.x, coef.y, shape=gl, color=p.value<0.05, size=hazards), show.legend = F) +
        #scale_shape_manual(values=c(17,18), na.value=21) +
        scale_size(range=c(2, 10), breaks=c(0, 5, 10, 15)) +
        #geom_text_repel( aes( label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, p.value < 0.05 & !is.na(gl)), aes(coef.x,coef.y, label=label), show.legend=F) + 
        geom_text_repel(data=subset(wt, is.na(stability)), aes(coef.x, coef.y, label=label, color=p.value<0.05), show.legend=F, color='grey39') 
  } else {
    p = p + geom_point(alpha=0.5) + 
      geom_text_repel(data=subset(wt, p.value < 0.05), aes(coef.x,coef.y, label=label), show.legend=F)  
  }
  return(p + labs(title='Wald test'))
}

```

```{r}
load('swgs.Rdata', verbose=T)

# limited to GLM coefs
df = df[,c('tstart','tstop','Diagnosis','Pathology', ch$label)]
df[1:4,1:8]
```

```{r}
gcols = colnames(df)

cols = grep('^\\d+', colnames(df))
colnames(df)[cols] = paste('chr', gsub(':|-','.', grep('^\\d+', colnames(df), value=T)), sep='')

pathology = do.call(cbind.data.frame, lapply(unique(df$Pathology), function(p) {
  cbind( as.integer(df$Pathology == p))
}))
colnames(pathology) = unique(df$Pathology)

# as an ordinal
df$Pathology = as.integer(factor(df$Pathology))

surv = with(df, Surv(tstart, tstop, Diagnosis))

groups = factor(c('clinical',rep('genomic', length(grep('\\d+', colnames(df))))))
rfx = CoxRFX(df[,-c(1:3)], surv, groups, verbose=F, nu=0.1)
summary(rfx)
```

```{r}
VarianceComponents(rfx)
```

```{r}
plot(rfx)
```

```{r}
wt = WaldTest(rfx)
survConcordance(surv ~ predict(rfx))
```

```{r}
## Compare to GLM selected features

wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))

wt = merge(wt, ch[,c('label','hazards','coef', 'stability','gl')], by='label', all=T)

wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0

plot.waldtest(wt) + labs(x='RFX coef', y='GLM coef')
```

```{r}
df2 = cbind(df, pathology[,c('BE','ID')])
df2$Pathology = NULL

surv = with(df, Surv(tstart, tstop, Diagnosis))

groups = factor(c(rep('genomic', length(grep('\\d+', colnames(df)))),rep('clinical', (ncol(pathology)-1) ) ) )

rfx2 = CoxRFX(df2[,-c(1:3)], surv, groups, verbose=F)
summary(rfx2)
```

```{r}
plot(rfx2)
```

```{r}
VarianceComponents(rfx2)
```

```{r}
survConcordance(surv ~ predict(rfx2))
```

```{r}
wt = WaldTest(rfx2)
wt$label = sub('\\.', '-', sub('\\.', ':', sub('chr', '', rownames(wt))))

wt = merge(wt, ch[,c('label','hazards','coef', 'stability','gl')], by='label', all=T)

head(wt)

wt[is.na(wt$hazards), c('hazards', 'coef.y')] = 0

plot.waldtest(wt) + labs(x='RFX coef', y='GLM coef')

```