### 2018-03-06 Misclassified.Rmd

---
title: "Untitled"
author: "Sarah Killcoyne"
date: "2/21/2018"
output: html_document
---

```{r setup, include=FALSE}
library(GenomicRanges)
library(ggplot2)
library(ggrepel)
library(GGally)
library(plyr) 
library(pander)
library(reshape2)
library(gridExtra)
library(plyr)
library(dplyr)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')
source('lib/data_func.R')

assign.risk <- function(pred) {
  cuts = seq(0,1,0.1)

  pred$quants = with(pred, cut(Prediction, breaks=cuts))
  qt = as.data.frame.matrix(table(pred$quants, pred$Status))
  qt$quant = rownames(qt)

  ft.fun<-function(NP,P) {
   f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
   cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
  }

  qt = cbind(qt, pred %>% group_by(quants) %>% summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
  pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% mutate( 
    'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

  pred.confidence[c('r1','r2')] = NA
  for (i in 1:(length(cuts)-1)) pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

  pred.confidence$Risk = 'Moderate'
  pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
  pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

  pred.confidence = bind_cols(pred.confidence, 
                              data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
                                        ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))

  #grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
  pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

  riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]
ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) +
  scale_fill_manual(values=riskCols) +
  geom_point() + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none', text=element_text(size=10))

  pred = merge(pred, pred.confidence, by='quants')
  pred = arrange(pred, Patient, Endoscopy.Year, PID, Pathology)
  return(pred)
}

data = '~/Data/Ellie'
info.dir = '~/Data/Ellie/QDNAseq'

patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)

if (length(patient.file) != 1 | length(demo.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", info.dir))

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
head(all.patient.info)

patient.info = all.patient.info
#patient.info = subset(all.patient.info, Set == 'Training')

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

```

# Why are the patient sin the validaiton cohort driving down prediction?

1. Look at the patients (and samples) that are misclassified in the LOO predictions within the 'D' only set.
  a. Do they stand out for anything else like very quiet genomes?
  b. Batches?

2. Look at coef within the 'D' model, and in the 'all' model
  a. Which coef change?
  b. Which samples change?

```{r disc_data, echo=F}
disc.dir = paste(data, 'Analysis/5e6_arms_disc', sep='/')
file = paste(disc.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=T)
disc.dysplasia.df = dysplasia.df
disc.labels = labels

length(labels) == nrow(patient.info)

file = paste(disc.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
disc.coefs = coefs$`0.9`

file = paste(disc.dir, 'loo.Rdata', sep='/')
load(file, verbose=T)

disc.pred = do.call(rbind, pg.samp)
disc.performance = performance.at.1se
names(disc.performance) = names(pg.samp)

m = melt(as.matrix(disc.performance))
m$Var2 = 'LOO'
ggplot(m, aes(Var2, value, group=Var2)) + geom_boxplot(outlier.colour = NA) + 
  geom_jitter(width=0.1) + labs(x='', y='Model classification performance', title="LOO, 'D'")
disc.loo.coefs = nzcoefs

rm(dysplasia.df, fits, plots, pg.samp, allDf, nzcoefs, coefs)
```

## Misclassified in 'Discovery'

1. Look at the patients (and samples) that are misclassified in the LOO predictions within the 'D' only set.
  a. Do they stand out for anything else like very quiet genomes?
  b. Batches?

```{r, echo=T}

preds = disc.pred[c('Status','Prediction')]
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)

ggplot(disc.pred, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + 
  scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='LOO Prediction', y='n Samples', x='Probability') + plot.theme

disc.pred = assign.risk(disc.pred)

prg = subset(disc.pred, Status == 'P' & Risk == 'High')
non = subset(disc.pred, Status == 'NP' & Risk == 'Low')

## Low
falseNeg = subset(disc.pred, Risk == 'Low' & Status == 'P')
table(falseNeg$Hospital.Research.ID)/table(subset(patient.info, Hospital.Research.ID %in% unique(falseNeg$Hospital.Research.ID))$Hospital.Research.ID )

# Mod
mods = subset(disc.pred, Risk == 'Moderate')
table(falseNeg$Hospital.Research.ID)/table(subset(patient.info, Hospital.Research.ID %in% unique(falseNeg$Hospital.Research.ID))$Hospital.Research.ID )

# High
falsePos = subset(disc.pred, Risk == 'High'  & Status == 'NP')
table(falsePos$Hospital.Research.ID)/table(subset(patient.info, Hospital.Research.ID %in% unique(falsePos$Hospital.Research.ID))$Hospital.Research.ID )

intersect(mods$Hospital.Research.ID, falseNeg$Hospital.Research.ID)
intersect(mods$Hospital.Research.ID, falsePos$Hospital.Research.ID)

```

Patients with the highest false pos/neg rate: AHM0121, AHM0277, AD0252, AD0591

```{r all_data, echo=F}
all.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')
file = paste(all.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=T)
all.dysplasia.df = dysplasia.df
all.labels = labels

file = paste(all.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
all.coefs = coefs$`0.9`
all.performance = performance.at.1se$`0.9`

file = paste(all.dir, 'loo.Rdata', sep='/')
load(file, verbose=T)
all.pred = do.call(rbind, pg.samp)
all.performance.loo = performance.at.1se
names(all.performance.loo) = names(pg.samp)

all.loo.coefs = coefs

m = melt(as.matrix(all.performance.loo))
m$Var2 = 'LOO'
ggplot(m, aes(Var2, value, group=Var2)) + geom_boxplot(outlier.colour = NA) + 
  geom_jitter(width=0.1) + labs(x='', y='Model classification performance', title="LOO, 'All'")

rm(coefs, nzcoefs, dysplasia.df, fits, plots, pg.samp, allDf)
```
# What are their coefficients like?

Look at coef within the 'D' model, and in the 'all' model
  a. Which coef change?
  b. Which samples change?

```{r}
# Coefs not in the 'D' model
setdiff(rownames(disc.coefs), rownames(all.coefs))
ucf = union(rownames(disc.coefs), rownames(all.coefs))

coefs = data.frame(matrix(ncol=2, nrow=length(ucf), dimnames=list(ucf, c('D','A'))))
coefs[rownames(disc.coefs),'D'] = disc.coefs[,1]
coefs[rownames(all.coefs),'A'] = all.coefs[,1]
#coefs[is.na(coefs)] = 0

abs(coefs$D) - abs(coefs$A)

```

# Re run cv w/o those patients?

```{r}
file = list.files(data, pattern='patient_folds.tsv', recursive=T, full.names=T)[1]
if (length(file) == 1) {
  message(paste("Reading folds file", file))
  sets = read.table(file, header=T, sep='\t')
} else {
  sets = create.patient.sets(patient.info[c('Hospital.Research.ID','Samplename','Status')], folds, splits, 0.2)  
}

rows = which(rownames(all.dysplasia.df) %in% subset(patient.info, Hospital.Research.ID %in% c('AHM0121', 'AHM0277', 'AD0252','AD0591'))$Samplename)

df = all.dysplasia.df[-rows,]
labels = all.labels[rownames(df)]

a = 0.9; nl=1000; folds=10; splits=5
fit0 <- glmnet(df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F)    
autoplot(fit0) + theme(legend.position = 'none')

cv = crossvalidate.by.patient(x=df, y=labels, lambda=fit0$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit0, select='deviance', opt=-1, standardize=F)

lambda.opt = cv$lambda.1se
subset(cv$lambdas, lambda == lambda.opt)$mean

pm = predict(fit0, newx=all.dysplasia.df[rows,], s=cv$lambda.1se, type='response')
colnames(pm) = c("Prediction")
pred = merge(pm, patient.info, by.x='row.names', by.y='Samplename')

# Prediction for those three don't improve
table(subset(pred, Status == 'P')$Prediction > 0.5)

table(subset(pred, Status == 'NP')$Prediction < 0.5)

```

# Do it with the 'D' model only

```{r}
`%nin%` = Negate('%in%')

exsamples = subset(patient.info, Hospital.Research.ID %in% c('AHM0121', 'AHM0277', 'AD0252','AD0591'))$Samplename
samples = subset(patient.info, Hospital.Research.ID %nin% c('AHM0121', 'AHM0277', 'AD0252','AD0591'))$Samplename

df = disc.dysplasia.df[intersect(samples, rownames(disc.dysplasia.df)),]
labels = disc.labels[ rownames(df) ]
length(labels) == nrow(df)

a = 0.9; nl=1000; folds=10; splits=5
fit1 <- glmnet(df, labels, alpha=a, nlambda=nl, family='binomial', standardize=F)    
autoplot(fit1) + theme(legend.position = 'none')

cv = crossvalidate.by.patient(x=df, y=labels, lambda=fit1$lambda, pts=sets, a=a, nfolds=folds, splits=splits, fit=fit1, select='deviance', opt=-1, standardize=F)

pm = predict(fit1, newx=disc.dysplasia.df[exsamples,], s=cv$lambda.1se, type='response')
colnames(pm) = c("Prediction")
pred1 = merge(pm, patient.info, by.x='row.names', by.y='Samplename')

# Prediction for those three don't improve
table(subset(pred1, Status == 'P')$Prediction > 0.5)
table(subset(pred1, Status == 'NP')$Prediction < 0.5)

```

## What about the validation patients?

```{r}
vpi = subset(patient.info, Set == 'Test')

df = all.dysplasia.df[vpi$Samplename,]
labels = all.labels[rownames(df)]

pm = predict(fit1, newx=df, s=cv$lambda.1se, type='response')
colnames(pm) = c("Prediction")
pred2 = merge(pm, all.patient.info, by.x='row.names', by.y='Samplename')

table(subset(pred2, Status == 'P')$Prediction > min(subset(disc.pred, Risk == "High")$r1) )
table(subset(pred2, Status == 'NP')$Prediction < max(subset(disc.pred, Risk == "Low")$r2))

roc = pROC::roc(Status ~ Prediction, data= pred2[c('Status','Prediction')], auc=T, ci=T, of='thresholds')
roc.plot(roc)

```

## LOO better trained w/o those three?
```{r}
sets = create.patient.sets(patient.info[c('Hospital.Research.ID','Samplename','Status')], folds, splits, 0.2)

pi = subset(patient.info, Hospital.Research.ID %nin% c('AHM0121', 'AHM0277', 'AD0252','AD0591') )

pg.samp = lapply(unique(pi$Hospital.Research.ID), function(pt) {
  info = subset(pi, Hospital.Research.ID == pt)
  info$Prediction = NA
  info$Prediction.Dev.Resid = NA
  info$OR = NA
  info$PID = unlist(lapply(info$Path.ID, function(x) unlist(strsplit(x, 'B'))[1]))
  return(info)
})
names(pg.samp) = unique(pi$Hospital.Research.ID)

secf = as.data.frame(all.coefs[,1])
rownames(secf) = rownames(all.coefs)
performance.at.1se = c(); coefs = list(); plots = list(); fits = list(); nzcoefs = list()
  # Remove each patient (LOO)
for (pt in names(pg.samp)) {
    print(pt)
    samples = subset(patient.info, !Hospital.Research.ID %in% c('AHM0121', 'AHM0277', 'AD0252', pt))$Samplename

    train.rows = which(rownames(all.dysplasia.df) %in% samples)
    training = all.dysplasia.df[train.rows,]

    exsamples = subset(patient.info, Hospital.Research.ID == pt)$Samplename
    test = as.matrix(all.dysplasia.df[which(rownames(all.dysplasia.df) %in% exsamples),])
    #if (ncol(test) <= 1) next
    if ( nrow(test) == ncol(all.dysplasia.df) ) test = t(test)

    # Predict function giving me difficulty when I have only a single sample, this ensures the dimensions are the same
    sparsed_test_data <- Matrix(data=0, nrow=ifelse(length(pg.samp[[pt]]$Samplename) > 1, nrow(test), 1),  ncol=ncol(training),
                                dimnames=list(pg.samp[[pt]]$Samplename,colnames(training)), sparse=T)
    for(i in colnames(all.dysplasia.df)) sparsed_test_data[,i] = test[,i]

    # Fit generated on all samples, including HGD
    fitLOO <- glmnet(training, all.labels[train.rows], alpha=a, family='binomial', nlambda=nl, standardize=F) # all patients
    l = fitLOO$lambda

    cv = crossvalidate.by.patient(x=training, y=all.labels[train.rows], lambda=l, a=a, nfolds=folds, splits=splits,
                                  pts=subset(sets, Samplename %in% samples), fit=fitLOO, standardize=F)

    plots[[pt]] = arrangeGrob(cv$plot+ggtitle('Classification'), cv$deviance.plot+ggtitle('Binomial Deviance'), top=pt, ncol=2)

    fits[[pt]] = cv

    if ( length(cv$lambda.1se) > 0 ) {
      performance.at.1se = c(performance.at.1se, subset(cv$lambdas, lambda == cv$lambda.1se)$mean)

      #coef.1se = coef(fitLOO, cv$lambda.1se)[rownames(secf),]

      nzcoefs[[pt]] = as.data.frame(non.zero.coef(fitLOO, cv$lambda.1se))

      coefs[[pt]] = coef(fitLOO, cv$lambda.1se)[rownames(secf),]
      #coef.stability(coef.1se, cv$non.zero.cf)

      logit <- function(p){log(p/(1-p))}
      inverse.logit <- function(or){1/(1 + exp(-or))}

      pm = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='response')
      or = predict(fitLOO, newx=sparsed_test_data, s=cv$lambda.1se, type='link')
      sy = as.matrix(sqrt(binomial.deviance(pm, labels[pg.samp[[pt]]$Samplename])))

      pg.samp[[pt]]$Prediction = pm[,1]
      pg.samp[[pt]]$Prediction.Dev.Resid = sy[,1] 
      pg.samp[[pt]]$OR = or[,1]

    } else {
      warning(paste("Hospital.Research.ID", pt, "did not have a 1se"))
    }
  }

back = do.call(rbind, pg.samp)

hist(back$Prediction)

```