### 2018-03-08 AllSamplesModel.Rmd
...

---
title: "All Samples Model"
author: "Sarah Killcoyne"
date: "16 Feb 2018"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

---

```{r setup, include=FALSE}
library(biomaRt)
library(GenomicRanges)
library(stringr)
library(ggplot2)
library(ggrepel)
library(GGally)
library(pander)
library(reshape2)
library(gridExtra)
library(CoxHD)
library(plyr)
library(dplyr)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')

data = '~/Data/Ellie'

cache.dir = paste(data, 'Analysis/5e6_arms_all', sep='/')

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
#analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info
head(all.patient.info)

patient.info = arrange(all.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

pander(sum.patient.data[,c('Hospital.Research.ID', 'Status', 'start.year', 'end.year','total.samples','total.endos','highest.path')], justify='left', caption='Training patient set')

```

# Segment size: 5e06 with arms

## Quick look at our demographic information

```{r echo=F}
pander(
  as.data.frame(sum.patient.data %>% dplyr::summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range'=paste(range(end.year-start.year), collapse='-'),
  'Median Years'=paste(median(end.year-start.year),'+/-',round(sd(end.year-start.year),1)),
  'Median Endoscopies'=paste(median(total.endos),'+/-',round(sd(total.endos),1)),
  'Median age'=paste(median(age.diagnosis),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':')
)), caption='full cohort', justify='left')
```

```{r echo=F}
cohortSummary = as.data.frame(sum.patient.data %>% group_by(Status) %>% dplyr::summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  'Range endoscopies'=paste(range(total.endos), collapse='-'),
  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':'),
  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
  'naC'=length(which(is.na(C)))
))

cohortSummary = cbind(cohortSummary,
  rbind(table(subset(all.patient.info, Status == 'NP')$Pathology),
        table(subset(all.patient.info, Status == 'P')$Pathology)))
  rbind(table(subset(patient.info, Status == 'NP')$Pathology),
        table(subset(patient.info, Status == 'P')$Pathology))

```

`r pander(t(cohortSummary), caption='Full cohort, split by status', justify='left')`

We tested models that included age, BMI, smoking status, sex, length & circumference of the BE segment, and p53 IHC status.  While none of these contributed we can report on some here.

```{r echo=F, warning=F, message=F, fig.height=4, fig.width=4}
ggplot(patient.info, aes(Age.at.diagnosis, fill=Status)) + geom_histogram(binwidth=3) + labs(title='Age at diagnosis') + plot.theme

ggplot(patient.info, aes(BMI, fill=Status)) + geom_histogram(binwidth=1) + labs(title='BMI') + plot.theme

ggplot(patient.info, aes(Maximal,Circumference, color=Status)) + geom_point() + labs(title='C vs M') + plot.theme

p53Path = table(subset(patient.info, Status == 'P' & p53.Status == 1)$Pathology)
p53.pred.byPath = round(p53Path/table(subset(patient.info, Status == 'P' & !is.na(p53.Status))$Pathology), 2)

```

```{r, echo=F}
grid.arrange(
  ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining') + plot.theme,
  tableGrob( 
    rbind('Non-progressor'=summary(subset(patient.info, Status == 'NP')$p53.Status), 
           'Progressor'=summary(subset(patient.info, Status == 'P')$p53.Status))), 
  nrow=2, as.table=T, heights=c(4,1) )
```

Because we're looking at both a binary classifier and binary data the AUC looks very good, there are no false positives after all. However of the progressor samples that were stained only `r round((table(subset(patient.info, Status == 'P')$p53.Status)/sum(table(subset(patient.info, Status == 'P')$p53.Status)))[2], 2)*100`% positively stained for p53. If we look at pathology the breakdown is even more obvious.  `r p53.pred.byPath['HGD']*100`% of HGD samples were identified, but only `r p53.pred.byPath['BE']*100`% NDBE samples were identified in progressors.

`r pander(p53.pred.byPath, caption='Ratio of progressors predicted by p53 staining per pathology where p53 staining was done (excluding N/As)')`

# Data set up

To use GLMs across the patients I merge all samples from all patients and overlap the copy number segments. Where segments from a sample get split, the value of the original sample is retained for each of the split samples.

## y( Progressors (1) vs Non (0) )

The labels are split such that all samples from progressor patients are (1) and all samples from non-progressors are (0).

```{r labelsPNP, echo=F, message=F, include=T}
file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=T)
dim(dysplasia.df)
length(labels)

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=F)
rm(plots, coefs, performance.at.1se, models)

riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]
```

We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.

Example subset of the data matrix:
`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

## Global copy number plots

```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
chr.lengths = get.chr.lengths()
chr.lengths$chrom = sub('chr','',chr.lengths$chrom)
chr.lengths = subset(chr.lengths, chrom %in% c(1:22))
chr.lengths$chrom = factor(chr.lengths$chrom, levels=c(1:22), ordered = T)

arm.loc = arrange(rbind(
  (chr.lengths %>% group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% group_by(chrom) %>% dplyr::mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)

globalCNdf = t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))])

locs = do.call(rbind.data.frame, lapply(rownames(globalCNdf), function(x) unlist(strsplit(x,':|-'))))
colnames(locs) = c('chr','start','end')
locs[] = lapply(locs[], function(x) as.numeric(as.character(x)))

globalCNdf = cbind.data.frame(locs, globalCNdf)
globalCNdf = melt(globalCNdf, id.vars=colnames(locs))

globalCNdf$seg.type = '5MB'
globalCNdf[with(globalCNdf, which( start %in% arm.loc$start & end %in% arm.loc$end)), 'seg.type'] = 'arms'
globalCNdf = merge(globalCNdf, chr.lengths[,c('chrom','chr.length')], by.x='chr', by.y='chrom')

globalCNdf = merge(globalCNdf, patient.info[,c('Hospital.Research.ID','Patient','Samplename','Pathology','Status')], by.x='variable', by.y='Samplename')
globalCNdf$Status = ifelse(globalCNdf$Status == 'P', 'Progressor', 'Non-Progressor')

globalCNdf$cn = factor(sapply(globalCNdf$value, cn), levels=c('loss','neutral','gain'), ordered = T)

n.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')', sep=''))

# Average cn value
globalCNdf = globalCNdf %>% group_by(Status, start, end) %>% dplyr::mutate( 'mean.value' = value/length(unique(variable)) )
globalCNdf$mean.value = globalCNdf$value # but not going to use it

globalCNdf$Status = factor(globalCNdf$Status, levels=sort(unique(globalCNdf$Status), decreasing=T), ordered=T)
globalCN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms'), labeller(Status=n.status, chr=label_value))
cnLegend = get.legend(globalCN.plot)
#globalCN.plot = globalCN.plot + theme(legend.position = 'none')

## NDBE
nb.status <- c('Progressor'=paste('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')', sep=''),
              'Non-Progressor'=paste('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')', sep=''))
globalBECN.plot = cn.mtn.plot(subset(globalCNdf, seg.type != 'arms' & Pathology == 'BE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')
ggsave(filename='plots/globalCN.png', plot=gp, width=16, height=4.6, units='in', dpi=300)
gp

gbe = globalBECN.plot + theme(legend.position = 'none') + labs(y='mean adj.seg.', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=gbe, width=16, height=4.6, units='in', dpi=300)
gbe

rm(locs, gp, gbe)
```

```{r loaddata, message=F, warning=F, echo=F}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

cx = dysplasia.df[,'cx']

x = cbind.data.frame('Status'=labels, 'cx'=cx[names(labels)])
x$Status = factor(x$Status, labels = c('Non-Progressor','Progressor'))
x = merge(x,patient.info[,c('Samplename','Pathology','Patient')],by.x='row.names',by.y='Samplename')

tt = t.test(subset(x, Status == 'Progressor')$cx, subset(x, Status == 'Non-Progressor')$cx)
pv = ggplot(x, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv

rocX = pROC::roc(Status ~ cx, data=x, auc=T, ci=T, of='thresholds')
roc.plot(rocX, title='Per sample cx') + plot.theme

ggsave('plots/per_sample_cx.png', plot=pv, width=4, height=4, units='in', dpi=300)
ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme , width=4, height=4, units='in', dpi=300)

```

# Building the model

We built the model using a 50-fold cross-validation process on the patients (rather than samples) for various lambda and alpha values to find the best fitting model.

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
# file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
# if (file.exists(file)) {
#   message(paste("loading", file))
#   load(file, verbose=T)
# } else {
#   stop(paste("No model file", file, "found"))
# }

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))

ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)

rm(plots, cvs)
```

## Model performance

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot

ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

select.alpha = as.character(0.9)
# Genomic regions only
most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)

most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

most.stable$chr = as.integer(as.character(most.stable$chr))
most.stable = arrange(most.stable, chr, start)

classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

```

`r pander(mp$performance[,c('n.Coef','25%','50%','75%', '100%')], justify='left', caption='Number of features stable in n% or more folds')`

Values that are closer to full lasso (alpha=1) are pretty similar, however at full lasso we see fewer features than if we regularize it at 0.8. They share all non-zero features as well.
`r pander(sapply(mp$stable.coeffients, length), caption='Total features at each value of alpha')`

`r pander(most.stable[,c('chr','start','stop')], caption='Features identified as stable in 75% or more folds.')`

## Feature Stability

They share all of the non-zero coefficients (`r which.max(sapply(mp$stable.coeffients, length))` identifies the most non-zero coefs `r max(sapply(mp$stable.coeffients, length))`).

## Remove HGD/IMC & LGD Samples and check the model

Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors.

### Remove HGD/IMC

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=T}
rm(performance.at.1se, coefs, mp)

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}
nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot

ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp, no.hgd.plots)
```

### Remove LGD

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot

ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=30)

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp, nolgd.plots)
```

#### LOO minus HGD

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}
```

```{r, echo=F, message=F, fig.width=5, fig.height=5}
pg.sampNOHGD = do.call(rbind, pg.sampNOHGD)

perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))

p = ggplot(melt(perf), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) )), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
  #geom_text_repel(aes(label=pt)) +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
p
ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds')
roc.plot(rocNOHGD, 'No HGD')  
```

```{r, echo=F, fig.width=6, fig.height=6}
ggplot(pg.sampNOHGD, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') + labs(title='Prediction using model w/o HGD', y='n Samples', x='Probability') + plot.theme

cuts = seq(0,1,0.1) 
pg.sampNOHGD$quants = with(pg.sampNOHGD, cut(Prediction, breaks=cuts))

noHGD.pred.confidence = as.data.frame.matrix(table(pg.sampNOHGD$quants, pg.sampNOHGD$Status))
noHGD.pred.confidence = cbind(noHGD.pred.confidence, pg.sampNOHGD %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
noHGD.pred.confidence$quant = rownames(noHGD.pred.confidence)

ft.fun<-function(NP,P) {
  print(paste(NP, P))
  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=round(f$p.value, 5), 'odds.ratio'=round(f$estimate, 4))
}

noHGD.pred.confidence = noHGD.pred.confidence %>% as.data.frame.matrix %>% group_by(quant) %>% 
  dplyr::mutate( 'perc'=P/sum(NP,P),'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

noHGD.pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
  noHGD.pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

noHGD.pred.confidence$Risk = 'Moderate'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
noHGD.pred.confidence$Risk[ which(with(noHGD.pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

noHGD.pred.confidence = bind_cols(noHGD.pred.confidence, with(noHGD.pred.confidence, 
    data.frame(ci.low=qbeta(0.025, shape1=P+.5, shape2=NP+.5),
               ci.high=qbeta(0.975, shape1=P+.5, shape2=NP+.5))))

noHGD.pred.confidence$Risk = factor(noHGD.pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

ggplot(noHGD.pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration', subtitle='HGD excluded model') + plot.theme + theme(legend.position = 'none')

# Predict HGD samples
file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)

  fitNOHGD = models[[select.alpha]]

  lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

  hgd.samples = subset(patient.info, Pathology %in% c('HGD','IMC'))
  hgd.df = dysplasia.df[hgd.samples$Samplename,]

  pred = predict(fitNOHGD, newx=hgd.df, s=lambda.opt, type='response')
  hgd.samples$Prediction = pred

  hgd.pred = hgd.samples %>% group_by(Patient, Hospital.Research.ID, Status, Endoscopy.Year) %>% 
    dplyr::summarise( max=max(Pathology), pred=Prediction[which.max(Pathology)] )

  rm(no.hgd.plots, coefs, performance.at.1se, models)
}

#hist(hgd.samples$Prediction)

rm(plots, nzcoefs, fits)
```

## Overall

Even without dysplastic samples our model is well able to classify samples as either progressive or non.

```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
#classification.rate$x = 1:nrow(classification.rate)

p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p

#ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

# Features

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})
ch = as.data.frame(hazards)
ch$coef = coefs[[select.alpha]][,1]
ch$stability = featureStability
ch$label = rownames(ch)
ch = arrange(ch, -hazards)

qq = quantile(ch$hazards, seq(0,1,.15))
p = ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') + plot.theme 
p

#ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)

## Arms...
if (exists('arm.locs')) {
  ch[which(ch$label %in% with(arm.locs, paste(chr, ':', start, '-', end, sep=''))),]
}

```

`r pander(ch, caption='Selected coefficients')`

```{r nohgd_coefs, message=F, warning=F, echo=F}
setdiff(ch$label, rownames(nohgd.coefs))

```

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
genomeFeat = grep('^\\d+:', ch$label)
featureAnn = data.frame(apply(do.call(rbind, strsplit(ch$label[genomeFeat], ':|-')), 2, as.numeric))
colnames(featureAnn) = c('chr','start','end')
featureAnn[] = lapply(featureAnn[], function(x) as.numeric(as.character(x)))

chh = cbind(ch[genomeFeat,], featureAnn)
chh = subset(chh, !is.na(chr))

chh = merge(chh, chr.lengths[,c('chrom','chr.length')], by.y='chrom', by.x='chr')
chh$chr = factor(chh$chr, levels=c(1:22))

chh$ymax = ifelse(chh$coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value))
chh$gl = factor(ifelse(chh$coef > 0, 'gain', 'loss'))

custom.max<-function(cf, hazards, ymax) {
  if (cf > 0) {
    max = hazards*ymax
    max = ifelse(max > ymax, ymax, max)
  } else {
    max = hazards*ymax
    max = ifelse(max < ymax, ymax, max)
  }
  return(max)
}
chh = chh %>% rowwise() %>% dplyr::mutate('max'=custom.max(coef, hazards, ymax))

p = globalCN.plot + geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
p

#ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

p = globalBECN.plot +  geom_rect(data=subset(chh, gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=subset(chh, gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')

#ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

```

# Nested leave-one-out predictions

Leave out each patient (both P and NP) and run a new CV fit each time then predict the samples for the patient that was left out.  Fitted model includes HGD samples.

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

names(performance.at.1se) = names(pg.samp)
df = as.data.frame(performance.at.1se)

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p
ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

oddPt = names(pg.samp)[which.max(performance.at.1se)]
subset(sum.patient.data, Hospital.Research.ID == oddPt)
pg.samp[[oddPt]]$Prediction

# Not the pt with the most samples or endos
#sum.patient.data[which.max(sum.patient.data$total.samples),]
```

### Prediction ROC curves

Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.

```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}

load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)

```

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'
roc$is = 'all'

ggsave(filename='plots/discovery_roc.png', plot=roc.plot(roc), width=6, height=4, units='in', dpi=300)

insilico.rocs[['all']] = roc

df = do.call(rbind, lapply(insilico.rocs, function(r) 
    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'auc'=rep(r$auc, length(r$sensitivities)),'model'=r$model,'is'=r$is) 
))

p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, is == 'all'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
p
ggsave(filename='plots/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

best = do.call(rbind, lapply(insilico.rocs, function(r) {
  cbind.data.frame(t(as.data.frame(pROC::coords(r, 'best'))), 'auc'=r$auc)
}))
best$model = rownames(best)

rr = ggplot(subset(df, is == 'all' | auc %in% c(range(auc))), aes(Specificity, Sensitivity)) + 
  geom_line(aes(color=factor(model)), size=1) + 
  geom_line(data=subset(df, auc == min(auc)), color='darkblue', linetype='dotted') + 
  geom_line(data=subset(df, auc == max(auc)), color='darkblue', linetype='dotted') +
  scale_x_reverse() + scale_color_brewer(palette='Paired') + 
  geom_label(data=subset(best, model == 'all' | auc %in% c(min(auc), max(auc)) ), aes(specificity, sensitivity, label=paste('AUC ',round(auc, 2), sep='\n'))) +
  labs(title='ROC', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
  theme(legend.position = 'none')
rr  
ggsave(filename='plots/roc_CI.png', plot=rr, width=6, height=6, units='in', dpi=300)

cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)
```

### p53 ROC curve

```{r p53roc, echo=F, fig.height=2, fig.width=2}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')

ggsave(filename='plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC') + plot.theme, width=6, height=6, units='in', dpi=300)
```

### Pathology ROC curve

Based on HGD/IMC/LGD

```{r pathroc, echo=F, fig.height=4, fig.width=10 }
pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'HGD & LGD'
ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'
ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)

grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
```

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID))

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
rocEndoM$model = 'Per Endoscopy Max'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
rocEndoA$model = 'Per Endoscopy Avg'
ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

lastEndo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[nrow(pt), 'PID'])$Prediction))
}))

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
pROC::coords(rocLastEndo, 'best')
rocLastEndo$model = 'Last endoscopy'
ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

firstEndo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[1, 'PID'])$Prediction))
}))

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
pROC::coords(rocFirstEndo, 'best')
rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

```

### Per Patient ROC

```{r echo=F, fig.height=4, fig.width=4}
#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'mean'=mean(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction), 'max'=max(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction))
}))

ptendo = ptendo[-which(is.na(ptendo$mean)),]

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtM, 'best')
rocPtM$model = 'Per Patient Max'
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtA, 'best')
rocPtA$model = 'Per Patient Avg'
ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

```

### Compare ROCs

```{r echo=F, fig.height=8, fig.width=6}
get.roc.stat<-function(roc, x='best') {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )

  sens = c(pROC::coords(roc, x)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x)[['specificity']], conf.level=0.95))
  spec = c(pROC::coords(roc, x)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x)[['sensitivity']], conf.level=0.95))

  stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
  colnames(stat) = c ('value','CI.min','CI.max')
  stat = as.data.frame(stat)
  stat$model = roc$model
  stat$Measure = rownames(stat)
  stat
}

get.auc.at<-function(roc, foc='spec', p=c(1,0.99)) {
  # To find the auc at a given spec or sens
  pROC::auc(roc, partial.auc=p, partial.auc.focus=foc, partial.auc.correct=T)
}

roc$model = "CN Model"
m = rbind.data.frame(
  get.roc.stat(p53roc),
  get.roc.stat(hgdpathroc),
  get.roc.stat(pathroc),
  get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC', 'HGD Only','HGD & LGD','CN Model'), ordered = T)

rocList = list(roc,pathroc,hgdpathroc, p53roc)
multi.roc.plot(rocList)

cols = brewer.pal(4, "PRGn")
dodge = position_dodge(width=0.9)

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none') + labs(x='', y='AUC', title='ROC Comparison') 
comp

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=6, height=8, units="in", limitsize=F, dpi = 300)

```

```{r echo=F, fig.height=8, fig.width=6}
roc$model = "Per Sample"

m = rbind.data.frame(
  get.roc.stat(roc),
  get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
  get.roc.stat(rocPtM), get.roc.stat(rocPtA),
  get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
m$model = factor(m$model)

cols = brewer.pal(11, 'Paired')
dodge = position_dodge(width=0.9)
mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
p

ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)

multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', colors=rev(cols[1:7]))
```

## Prediction probability distribution

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction', 'OR')] %>% group_by(PID, Samplename) %>% dplyr::mutate(
    'months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)

  pt = arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

  tb = with(pt, table(Endoscopy.Year, PID))
  tb[tb>0] = 1
  for (yr in names(which(rowSums(tb) > 1))) {
    tmp = pt[which(pt$Endoscopy.Year == yr),]
    tmp = tmp %>% group_by(PID) %>% summarise(max(Pathology))

    match = which(pt$Endoscopy.Year == yr & pt$PID == as.character(tmp[which.min(tmp$`max(Pathology)`), 'PID']))
    pt[match,'months.before.final'] = pt[match,'months.before.final'] + 6
  }

  pt$nl = (pt$Block-1)*2

  tb = table(pt[,c('months.before.final', 'PID')])
  for (i in 1:nrow(tb)) {
    pids = tb[i,]
    yrs = which(pids > 0)
    if (length(yrs) > 1) {
      for (j in 2:length(yrs)) 
        pt[which(pt$PID == names(yrs)[j]), 'nl'] = pt[which(pt$PID == names(yrs)[j]), 'nl'] + 1
    }
  }
  arrange(pt, -Endoscopy.Year, PID)
}

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
back = merge(back, patient.info[,c('Samplename','p53.Status')])

sq = c(0, seq(1,132,12),max(back$months.before.final))
back$brks = cut(back$months.before.final, 
                breaks=sq, 
                labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
                include.lowest = T, ordered_result = T)
back$brks = factor(back$brks, rev(levels(back$brks)))

back$Pathology = as.character(back$Pathology)
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)

```

The labels in each decile are the ratio of progessor samples that have probabilities within that region. The deciles that are greyed out show insigificant enrichment for the NP/P ratio.

```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
cuts = seq(0,1,0.1)

back$quants = with(back, cut(Prediction, breaks=cuts))
qt = as.data.frame.matrix(table(back$quants, back$Status))
qt$quant = rownames(qt)

ft.fun<-function(NP,P) {
  f = fisher.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'odds'=ft.fun(NP,P)$odds.ratio, 'conf'=ifelse(p.value < 0.05, '*', '') )

pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
  pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

pred.confidence = bind_cols(pred.confidence, 
                            data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
                                       ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))

annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% dplyr::summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = riskCols, col=NA, alpha=0.2),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="black", fontface=4L)),
  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
  padding=unit(c(5,5),'mm'))

#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)
```

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + scale_fill_manual(values=riskCols) +
  geom_point() + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), width=0.01) + labs(x='Mean prediction', y='Ratio of `P` patients', title='Prediction calibration') +
  scale_color_manual(values=riskCols) + plot.theme + theme(legend.position = 'none')
pcal

ggsave(filename='plots/pred_calib.png', plot=pcal, width=3, height=3, units='in', dpi=300)

back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% group_by(Risk) %>% dplyr::mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )

mod = subset(back, Risk == 'Moderate')

```

```{r relrisk, echo=F, fig.width=6, fig.height=6}
myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))

unadjRR = ggplot(back, aes(OR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  #scale_fill_distiller(palette = 'RdYlBu', type='div',name='P(P)') +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR

ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)

predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
predhist

ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)

ggplot(back, aes(OR, Prediction, color=Prediction)) + geom_point(show.legend = F) + scale_color_gradientn(colors = myPal,  name='')  + plot.theme
```

## Jacknife coefficients

The idea here is to get some sort of CI for the coefficients, I'm not entirely sure this is correct though.

Could just as easily get a jacknife CI per patient

```{r jacknife, echo=T, warning=F, message=F, fig.height=10, fig.width=12}

coefs.per.pt = data.frame(matrix(ncol=length(pg.samp), nrow=nrow(ch), dimnames=list( ch$label, names(pg.samp))))
for (pt in names(pg.samp)) {
lcf = loo.coefs[[pt]][,1]
names(lcf) = rownames( loo.coefs[[pt]] )
sharedcfs = intersect(names(lcf), ch$label)
coefs.per.pt[sharedcfs, pt] = lcf[sharedcfs]
}
coefs.per.pt[is.na(coefs.per.pt)] = 0

library(bootstrap)

# per patient
# jack.se = apply(coefs.per.pt, 2, function(x) {
#   jk = jackknife(x, var)
#   jk$jack.se
# })
# js = cbind.data.frame(performance.at.1se, jack.se)
# ggplot(js, aes(rownames(js), performance.at.1se, color=(performance.at.1se > mean(performance.at.1se)+sd(performance.at.1se) | performance.at.1se < mean(performance.at.1se)-sd(performance.at.1se)  ))) + ylim(0.6,1) + geom_point() + 
#   geom_errorbar(aes(ymin=performance.at.1se-jack.se, ymax=performance.at.1se+jack.se)) + theme(legend.position = 'none', axis.text.x = element_text(angle=90, hjust=1))

# in this case jack.se is the SE of the variance 
# per coefficient
ch$jack.se = jack =  apply(coefs.per.pt, 1, function(x) {
jk = jackknife(x, var)
jk$jack.se
})
p = ggplot(ch, aes(reorder(label, coef), coef)) + geom_point() + 
  geom_errorbar(aes(ymin=coef-jack.se, ymax=coef+jack.se)) + coord_flip() +
  plot.theme + labs(x='', title='Jackknifed coefficient CI')
p

ggsave(filename='plots/feature_jacknife.png', plot=p, width=5, height=6, unit='in', dpi=300)

```

## How do the predictions look at each time point?

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

pathTotal = back %>% group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
riskPath$Risk = factor(riskPath$Risk, levels=c('Low','Moderate','High'), ordered = T)

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology)), p53Path)
#p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

#riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
#riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)

pathTotal = riskPath %>% group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))

rp = ggplot(riskPath, aes(Pathology, ratio*100)) + geom_bar(aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols) + scale_color_manual(values=c('white','black','white')) +
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101.5), position=position_stack(), size=4) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), size=5, show.legend = F) +
  facet_grid(~Status, scales='free_x', space = 'free_x') + plot.theme + labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp

ggsave(filename='plots/pred_by_path.png', plot=rp, width=8, height=8, units='in', dpi=300)

```

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = back %>% group_by(Status, Patient) %>% dplyr::summarise( 'Low'=sum(length(which(Risk == "Low"))),
'Moderate'=sum(length(which(Risk == "Moderate"))),
'High'=sum(length(which(Risk == "High"))),
'n.samples'=length(Patient) )

all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))
head(all.preds)

# ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
#   geom_bar(stat='identity', position = position_stack()) +
#   scale_fill_manual(values=riskCols) + 
#   geom_text(aes(y=1.01, label=n.samples)) +
#   facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
#   plot.theme

```

## Backwards from HGD

Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.

Take per endoscopy average prediction.

There are a lot more moderate predictions in this model...how does this work with this plot??

```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}

endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
  'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
  'n.endos'=length(unique(PID)), 
  'n.samples'=length(PID),
  'mean.pred'=mean(Prediction),
  'pred.samples'=length(which(Risk == 'High')),
  'max.path'=max(Pathology))

endos$quant =  cut(endos$mean.pred, breaks = cuts)
endos = merge(endos, pred.confidence[c('quant', 'Risk')])

sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,
breaks=sq,
labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

endos.m = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

tmp = subset(back, Status == 'P')
tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,
labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))
tmp = tmp %>% group_by(brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

endos.m = arrange(merge(endos.m,tmp), brks)

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,105) +
  geom_bar(stat='identity', fill='darkred', alpha=0.8, show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=9, height=7, units='in', dpi=300)

#endos$months.before.final = endos$months.before.final*-1
#me = endos %>% group_by(Patient) %>% dplyr::summarise('max.endo' = max(i), 'months' = min(months.before.final))
```

`r pander(as.data.frame(pred.confidence))`

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
back$nl = factor(back$nl, ordered=T)
back$Risk = factor(back$Risk, levels=c('Low','Moderate','High'), ordered = T)
back$Patient = factor(as.numeric(back$Patient), ordered = T)

p = pred.tiles(subset(back, Patient %in% c(6)), ncol=1, cols=myPal, probs=T) + labs(title='')
tilesLegend = get.legend(p)
p + theme(legend.position = 'none')

#pred.tiles(subset(back, Patient %in% c(48)), ncol=1, cols=myPal, probs=T) + labs(title='') + theme(legend.position = 'none')
#pred.tiles(subset(back, Patient %in% c(33)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
#pred.tiles(subset(back, Patient %in% c(89)), ncol=1, cols=myPal, probs=T) + labs(title='')+ theme(legend.position = 'none')
```

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15 }
ga = grid.arrange(
  pred.tiles(subset(back, Patient %in% c(13,20,34,56,60,48) & Status == 'P'), ncol=1, cols=myPal, probs=T) + labs(title='Progressors') + theme(legend.position = 'none'),
  pred.tiles(subset(back, Patient %in% c(15,16,21,3,45,19) & Status == 'NP'), ncol=1, cols=myPal, probs=T) + labs(title='Non-Progressors') + theme(legend.position = 'none'),
  ncol=2)
ga

#ggsave('plots/ex_pt_heatmap.png', plot=ga, width=6, height=15, units='in', dpi=300)

table.path.risk<-function(df) {
  summary.t = as.data.frame.matrix(with(df, table(Pathology, Risk)))
  summary.t$Pathology = rownames(summary.t)

  summary.t = summary.t %>% group_by(Pathology) %>% dplyr::mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
                                                            'Low.r'=round(Low/sum(High,Low,Moderate),3),
                                                            'Moderate.r'=round(Moderate/sum(High,Low,Moderate),3))
  for (g in unique(df$Risk)) {
    for (i in 1:nrow(summary.t)) {
      summary.t[i,g] = paste(summary.t[i,g], ' (', summary.t[i,paste(g,'.r',sep='')]*100, '%)',sep='')
    }
  }
  summary.t[,c('Pathology',levels(df$Risk))]
}

summary.t = table.path.risk(back)
```

`r pander(as.data.frame(summary.t))`

```{r, echo=F, message=F, warning=F, width=26, height=8}
back = back %>% group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))

p1 = pred.tiles(subset(back, Status == 'NP'), cols=myPal, probs=T, nrow=2) + labs(title='Non-Progressors') 
riskLegend = get.legend(p1)
p1 = p1 + theme(legend.position = 'bottom')
riskLegendLong = get.legend(p1)

p1 = p1 + theme(legend.position = 'none')
p2 = pred.tiles(subset(back, Status == 'P'), probs=T, cols=myPal,nrow=2) + labs(title='Progressors') + theme(legend.position = 'none')

ggsave('plots/pred_heatmap_ALL.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
ggsave('plots/npg_heatmap.png', plot=p1, width=52, height=8, units="in", limitsize=F, dpi = 300)
ggsave('plots/pg_heatmap.png', plot=p2, width=52, height=8, units="in", limitsize=F, dpi = 300)
ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)
```