### 2018-03-16 Insilico Splits.Rmd

---
title: "Insilico Splits"
author: "Sarah Killcoyne"
date: "1/10/2018"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(ggrepel)
library(pander)
library(reshape2)
library(gridExtra)
library(plyr) 
library(dplyr)
library(pROC)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')

data = '~/Data/Ellie'
tile.w=c(5e06, 'arms')

includeDemo = F

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info

#sum.pt = summarise.patient.info(all.patient.info)

sum.pt = all.patient.info %>% group_by(Set, Patient) %>% summarize(
  med.reads=median(Total.Reads),
  sd.reads=sd(Total.Reads),
  n.endoscopy=length(unique(PID)),
  n.years=unique(Final.Endoscopy-Initial.Endoscopy),
  n.samples=length(Samplename)
)

sp = sum.pt %>% group_by(Set) %>% summarize(
  med.endo=median(n.endoscopy),
  sd.endo=sd(n.endoscopy),
  med.years=median(n.years),
  med.samples=median(n.samples),
  sd.samples=sd(n.samples)
)

```

`r pander(as.data.frame(sp))`

```{r echo=F, warning=F, message=F}

#cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')
model.dir = paste(data, 'Analysis/5e6_arms_all_exAHM0320', sep='/')
split.dir = '/Volumes/research1/gerstung/skillcoyne/Ellie/Analysis/5e6_arms_splits_exAHM0320'

performance = data.frame(); rocs = list()
select.alpha = '0.9'

file = paste(model.dir, 'insilico_splits.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
  rocs = insilico.rocs
  performance = insilico.perf
} else {

  split.dirs = list.files(split.dir, full.names = T)
  for (split in 1:length(split.dirs)) {
    print(split)  
    load(list.files(split.dirs[split],full.names = T, pattern='all.*Rdata'), verbose=T)
    performance = rbind(performance, performance.at.1se[[select.alpha]])
    rm(plots,coefs,performance.at.1se,models,cvs)

    load(list.files(split.dirs[split],full.names = T, pattern='loo.*Rdata'), verbose=T)
    rm(plots, performance.at.1se,coefs,nzcoefs,fits)

    predictions = do.call(rbind, pg.samp)

    preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
    roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
    #rocs = rbind(rocs,c(ci.auc(roc), coords(roc, 'best') ))
    roc$model = 'in-silico'
    roc$is = split
    rocs[[split]] = roc

    rm(dysplasia.df, labels)
  }
  #colnames(rocs) = c('auc.ci.min','auc','auc.ci.max','threshold','specificity','sensitivity')

  insilico.rocs=rocs
  insilico.perf=performance
  save(insilico.rocs, insilico.perf, file=paste(model.dir, 'insilico_splits.Rdata', sep='/'))
}

n.splits = nrow(performance)

load(list.files(model.dir, full.names = T, pattern = 'all.*Rdata'))

performance = rbind(performance, performance.at.1se[[select.alpha]])
performance$split = c(paste('split',1:n.splits,sep='.'), 'discovery')
rm(plots,coefs,performance.at.1se,models,cvs,dysplasia.df,labels)

load(list.files(model.dir,full.names = T, pattern='loo.Rdata'))
rm(plots, performance.at.1se,coefs,nzcoefs,fits)

predictions = do.call(rbind, pg.samp)

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'discovery'
roc$is = 'discovery'

#rocs = rbind(rocs,c(ci.auc(roc), coords(roc, 'best')) )

#rocs$split = c(basename(ins_splits),'discovery')

```

Original "discovery" cohort vs insilico generated discovery cohorts.  Original is noted in red, we appear to have gotten lucky in our patient selection.

## Model Performance

```{r echo=F, fig.align='left'}
performance$name = 'insilico'
performance$name[which(performance$split == 'discovery')] = 'discovery'
m = melt(performance, measure.vars = 'mean')
ggplot(m, aes(variable,value)) + ylim(0,1) +
  geom_boxplot(fill='mintcream', color='grey55') + geom_jitter(width=0.2, aes(color=name), size=1.5) + 
  geom_text(data=subset(m, value %in% range(value)), aes(label=round(value,2)), vjust=0.02, hjust=0.2) +
  plot.theme + labs(x='CV Models',y='Mean model performance',title='Mean performance for CV model', subtitle='20 insilico 20:69 splits')
```

Median performance `r median(performance$mean)`, SD `r sd(performance$mean)`

## AUC Plot

```{r echo=FALSE, fig.align='left'}
rocs[['discovery']] = roc

mrplot <- function(rocList) {
  aucs = do.call(rbind, lapply(rocList,function(r) 
    cbind.data.frame('AUC'=r$auc*100,'model'=r$model,'Specificity'=pROC::coords(r,'best')[['specificity']], 'Sensitivity'=pROC::coords(r,'best')[['sensitivity']])))
  aucs$y = aucs$Sensitivity
  aucs$x = aucs$Specificity

  df = do.call(rbind, lapply(rocList, function(r) 
    cbind.data.frame('Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=r$is) 
  ))

  ggplot(df, aes(Specificity, Sensitivity,color=is)) + geom_line(size=1) + 
    geom_line(data=subset(df, model == 'discovery'), color='red', size=1) +
    scale_x_reverse() +
    scale_color_brewer(palette='Blues') +
    labs(title='', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
}

mrplot(rocs)

```
Median AUC `r median(rocs$auc)`, SD `r sd(rocs$auc)`