### 2018-05-02 ElliesThesisFigs.Rmd

---
title: "ThesisStuff"
author: "Sarah Killcoyne"
date: "4/12/2018"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4

---
```{r setup, include=FALSE}
#library(biomaRt)
#library(GenomicRanges)
#library(stringr)
library(ggplot2)
library(ggrepel)
library(GGally)
library(pander)
library(reshape2)
library(gridExtra)
#library(CoxHD)
library(plyr)
library(dplyr)

source('lib/load_patient_metadata.R')
source('lib/cv-pt-glm.R')
source('lib/common_plots.R')
source('lib/data_func.R')

dir.create('thesis', showWarnings = F)

model.performance<-function(p1se, coeffs, lims=NULL) {
  performance = do.call(rbind.data.frame, p1se)
  performance$alpha = rownames(performance)

  if (grepl('chr', performance$alpha[1])) {
    performance$alpha = factor(performance$alpha, levels=c(paste('chr',c(1:22), sep=''), setdiff(rownames(performance), paste('chr',c(1:22),sep=''))) )
  }

  ## check how often the feature(s) is selected at that lamda in each split. "stability selection"
  coef.stable = lapply( coeffs, function(cf) {
    sort(rowSums(cf[,-1]), decreasing=T)
  })
  coef.stable = coef.stable[names(which(sapply(coef.stable, length) > 0))]
  coef.stable = lapply(coef.stable, function(cf) cf/(folds*splits))

  performance = cbind(performance, do.call(rbind,  lapply(coef.stable, function(x) {
    cbind('n.Coef'=length(x), '25%'=length(which(x>=.25)), 
          '50%'=length(which(x>=.5)), '75%'=length(which(x>=.75 )), 
          '100%'=length(which(x==1)) )
  })))

  p = ggplot(performance, aes(alpha, mean, fill=(alpha != 0.9))) + ylim(0,1) + geom_col() +
    geom_text( aes(label=round(mean, 3)), vjust=3) +
    geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme), width=0.3, size=1, color='grey39') + 
    geom_text(aes(y=0.1, label=n.Coef), ) +
    geom_text(aes(label=paste('(',`75%`,')',sep=''), y=0.05)) + 
    scale_fill_brewer(palette='Paired') +
    #scale_fill_manual(values=c('red', 'grey39')) + 
    theme(legend.position = 'none') +
    labs(x='Elasticnet penalty value: ridge <-> lasso', y='Model classification at lambda-1se')

  if (!is.null(lims))
    p = p+ylim(lims)

  cfs = as.data.frame(matrix(data=0,nrow=length(unique(names(table(unlist(lapply(coef.stable, names)))))), ncol=length(names(coef.stable)), 
                             dimnames=list( unique(names(table(unlist(lapply(coef.stable, names))))), names(coef.stable) )))
  for (i in names(coef.stable)) 
    cfs[intersect(rownames(cfs), names(coef.stable[[i]])), i] = 1
  return(list('performance'=performance, 'stable.coeffients'=coef.stable, 'cf.per.feature'=cfs,'plot'=p))
}

data = '~/Data/Ellie'

cache.dir = paste(data, 'Analysis/5e06-genomic', sep='/')

data.files = list.files(paste(data, 'QDNAseq',sep='/'), full.names=T)
analysis.files = list.files(paste(data, 'Analysis', sep='/'), full.names=T)

## Hospital.Research.ID info file
patient.file = grep('All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('Demographics_full.xlsx', data.files, value=T)

all.patient.info = read.patient.info(patient.file, demo.file, set='all')$info

patient.info = subset(all.patient.info, Set == 'Training')

patient.info = arrange(all.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = summarise.patient.info(patient.info)

load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)

val.data = patient.data[subset(sum.patient.data, Set == 'Test')$Hospital.Research.ID]
patient.data = patient.data[subset(sum.patient.data, Set == 'Training')$Hospital.Research.ID]

```

# Segment size: 5e06 with arms

# Data set up

To use GLMs across the patients I merge all samples from all patients and overlap the copy number segments. Where segments from a sample get split, the value of the original sample is retained for each of the split samples.

## y( Progressors (1) vs Non (0) )

The labels are split such that all samples from progressor patients are (1) and all samples from non-progressors are (0).

```{r labelsPNP, echo=F, message=F, include=T}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
#rm(plots, coefs, performance.at.1se, models)

## binomial: prog 1, np 0
labels = unlist(sapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  label = as.integer(df$info$Status == 'P') #as.integer(df$info$Pathology %in% c('HGD', 'IMC'))
  names(label) = df$info$Samplename
  return(label)
}))
names(labels) = sub('.*\\.', '',  names(labels))

pts = do.call(rbind, lapply(patient.data, function(df) {
  df$info = arrange(df$info, Endoscopy.Year, Pathology)
  cbind(df$info[,c('Hospital.Research.ID','Samplename')])
}))
rownames(pts) = 1:nrow(pts)

# sort in label order
if (length(setdiff(rownames(dysplasia.df), names(labels))) > 3)
  warning("Labels vector is missing samples")

riskCols = RColorBrewer::brewer.pal(11, "RdYlBu")[c(10, 1)]
```

We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.

Example subset of the data matrix:
`r pander(dysplasia.df[1:5, 1:5], caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', ')), justify='left')`

## Build/Cross-validate by patient

Keeping in mind that the matrix is built on a sample, not patient, basis while the labels (Progressor/Non) are on a per-patient basis I generated the cross validated models for various values of alpha by running 10 or more iterations of 5-fold cross validation that pulled out all the samples for 5 patients at each fold.  1000 lambda values.

```{r, echo=F, message=F, warning=F}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))

```

## Model performance

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
folds = 10; splits = 5 
classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot
ggsave('thesis/all_perf.png', plot=mp$plot, height=3.5, width=3.5, units='in', dpi=300, scale=1.2)

select.alpha = as.character(0.9)
# Genomic regions only
most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)

most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

most.stable$chr = as.integer(as.character(most.stable$chr))
most.stable = arrange(most.stable, chr, start)

classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

```

`r pander(mp$performance[,c('n.Coef','25%','50%','75%', '100%')], justify='left', caption='Number of features stable in n% or more folds')`

## Feature Stability

They share all of the non-zero coefficients (`r which.max(sapply(mp$stable.coeffients, length))` identifies the most non-zero coefs `r max(sapply(mp$stable.coeffients, length))`).

## Remove HGD/IMC & LGD Samples and check the model

Just as a point, the initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors.

### Remove HGD/IMC

```{r noHGD, echo=F, message=F, warning=F, fig.width=6, fig.height=6}
rm(performance.at.1se, coefs, mp)

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}
nohgd.coefs = coefs[[select.alpha]]

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot
ggsave('thesis/no_hgd_perf.png', plot=mp$plot, height=3.5, width=3.5, units='in', dpi=300, scale=1.2)

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp, no.hgd.plots)
```

### Remove LGD

```{r noLGD, echo=F, warning=F, message=F, fig.width=6, fig.height=6, eval=T}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

mp = model.performance(performance.at.1se, coefs)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot
ggsave('thesis/no_lgd_perf.png', plot=mp$plot, height=3.5, width=3.5, units='in', dpi=300, scale=1.2)

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp, nolgd.plots)
```

## Overall

Even without dysplastic samples our model is well able to classify samples as either progressive or non.

```{r, echo=F, warning=F, message=F, fig.width=4, fig.height=6}
p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p
ggsave('thesis/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

# Features

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)

hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})
ch = as.data.frame(hazards)
ch$coef = coefs[[select.alpha]][,1]
ch$stability = featureStability
ch$label = rownames(ch)
ch = arrange(ch, -hazards)

qq = quantile(ch$hazards, seq(0,1,.15))
p = ggplot(ch, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by hazard ratio', x='Coefficient value', y='Hazard') 
p
ggsave('thesis/hazards.png', plot=p, width=5, height=5, units='in', dpi=300)

chtable = ch[,c('hazards','coef','stability','label')]
chtable = chtable[,c('label','hazards', 'coef')]
chtable$coef[which(chtable$coef > 0)] = 'Gain'
chtable$coef[which(chtable$coef != 'Gain')] = 'Loss'
colnames(chtable) = c('Location','Hazard', 'Gain/Loss')
chtable$Hazard = round(chtable$Hazard, 3)
#write.table(chtable, sep='\t', quote=F, row.names = F, file='hazards.txt')

```

`r pander(ch, caption='Selected coefficients')`

# Nested leave-one-out predictions

Leave out each patient (both P and NP) and run a new CV fit each time then predict the samples for the patient that was left out.  Fitted model includes HGD samples.

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop(paste("Missing model file", file))
}
#loo.coefs = nzcoefs

names(performance.at.1se) = names(pg.samp)
df = as.data.frame(performance.at.1se)

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for leave-one-out', subtitle='All samples model')
p
ggsave('thesis/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

```

### Prediction ROC curves

Based on these curves we get our best AUC at ~0.36.  Currently we are using 0.5 by default.

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'CN Model'

pROC::ci.auc(roc)

roc$is = 'all'

ggsave(filename='thesis/all_roc.png', plot=roc.plot(roc), width=6, height=4, units='in', dpi=300)

cutoff = round(pROC::coords(roc, 'best')[['threshold']], 2)
```

### p53 ROC curve

```{r p53roc, echo=F, fig.height=2, fig.width=2}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

pROC::ci.auc(p53roc)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')

ggsave(filename='thesis/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC') + plot.theme, width=6, height=6, units='in', dpi=300)
```

### Pathology ROC curve

Based on HGD/IMC/LGD

```{r pathroc, echo=F, fig.height=4, fig.width=10 }
pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'HGD & LGD'

pROC::ci.auc(pathroc)

ggsave('thesis/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'

pROC::ci.auc(hgdpathroc)

ggsave('thesis/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)

grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
```

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = do.call(rbind, pg.samp) %>% group_by(Patient, Status, Hospital.Research.ID, PID ) %>% dplyr::summarise(
  'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID))

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoM, 'best')
rocEndoM$model = 'Per Endoscopy Max'
ggsave('thesis/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
pROC::coords(rocEndoA, 'best')
rocEndoA$model = 'Per Endoscopy Avg'
ggsave('thesis/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

lastEndo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[nrow(pt), 'PID'])$Prediction))
}))

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
pROC::coords(rocLastEndo, 'best')
rocLastEndo$model = 'Last endoscopy'
ggsave('thesis/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

firstEndo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'avg'=mean(subset(pt, PID == pt[1, 'PID'])$Prediction))
}))

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')
pROC::coords(rocFirstEndo, 'best')
rocFirstEndo$model = 'First endoscopy'
ggsave('thesis/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy') , width=6, height=6, units="in", limitsize=F, dpi = 300)

```

### Per Patient ROC

```{r echo=F, fig.height=4, fig.width=4}
#ptendo = do.call(rbind, pg.samp) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise('max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(Patient))
ptendo = do.call(rbind, lapply(pg.samp, function(pt) {
  cbind(pt[1,c('Patient','Hospital.Research.ID', 'Status')],  'mean'=mean(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction), 'max'=max(subset(pt, PID != pt[nrow(pt), 'PID'])$Prediction))
}))

ptendo = ptendo[-which(is.na(ptendo$mean)),]

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtM, 'best')
rocPtM$model = 'Per Patient Max'
ggsave('thesis/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
pROC::coords(rocPtA, 'best')
rocPtA$model = 'Per Patient Avg'
ggsave('thesis/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

```

### Compare ROCs

```{r echo=F, fig.height=8, fig.width=6}
get.roc.stat<-function(roc, x='best') {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )

  sens = c(pROC::coords(roc, x)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x)[['specificity']], conf.level=0.95))
  spec = c(pROC::coords(roc, x)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x)[['sensitivity']], conf.level=0.95))

  stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
  colnames(stat) = c ('value','CI.min','CI.max')
  stat = as.data.frame(stat)
  stat$model = roc$model
  stat$Measure = rownames(stat)
  stat
}
roc$model = "CN Model"
m = rbind.data.frame(
  get.roc.stat(p53roc),
  get.roc.stat(hgdpathroc),
  get.roc.stat(pathroc),
  get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC', 'HGD Only','HGD & LGD','CN Model'), ordered = T)

rocList = list(roc,pathroc,hgdpathroc, p53roc)
multi.roc.plot(rocList)

cols = RColorBrewer::brewer.pal(4, "PRGn")
dodge = position_dodge(width=0.9)

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  ylim(0,100) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', axis.text.x = element_text(hjust=1,angle=45)) + labs(x='', y='AUC', title='') 
comp

ggsave('thesis/path_p53_cn_auc.png', plot=comp, width=3.1, height=3.9, units="in", limitsize=F, dpi = 300, scale=1)

```

```{r echo=F, fig.height=8, fig.width=6}
roc$model = "Per Sample"

m = rbind.data.frame(
  get.roc.stat(roc),
  get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
  get.roc.stat(rocPtM), get.roc.stat(rocPtA),
  get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
m$model = factor(m$model)

subset(m, model %in% c(rocEndoA$model, rocEndoM$model) & Measure == 'AUC')

cols = RColorBrewer::brewer.pal(11, 'Paired')
dodge = position_dodge(width=0.9)
mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
p

ggsave('thesis/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)

multi.roc.plot(list(roc, rocEndoM, rocEndoA, rocPtM, rocPtA,rocFirstEndo,rocLastEndo), 'ROC Comparisons', colors=rev(cols[1:7]))
```

## Prediction probability distribution

```{r back, echo=F, message=F, warning=F}
transform.by.time<-function(pt, info) {
  pt = pt[,c('Status','Patient','Hospital.Research.ID','Path.ID','PID','Pathology','Samplename','Final.Endoscopy','Endoscopy.Year','Prediction')] %>% group_by(PID, Samplename) %>% dplyr::mutate('months.before.final' = (Final.Endoscopy - Endoscopy.Year)*12)
  pt = dplyr::arrange(merge(pt, subset(info, Hospital.Research.ID == pt$Hospital.Research.ID[1], select=c('Samplename','Block')), by='Samplename'), Pathology, PID, Endoscopy.Year)

  tb = with(pt, table(months.before.final, PID))
  tb[tb>0] = 1
  for (mnths in names(which(rowSums(tb) > 1))) {
    tmp = pt[which(pt$months.before.final == mnths),]
    tmp = arrange(tmp, rev(Pathology), PID)

    mt = tmp$months.before.final[1] 
    interval = length(tmp$Endoscopy.Year)/(length(unique(tmp$PID))-1)
    for (pid in unique(tmp$PID)) {
      rows = which(tmp$PID == pid)
      if (length(grep('HGD|IMC',tmp[rows,'Pathology'])) > 0 & mt != 0) {
        mt = 0
        tmp$months.before.final[rows] = mt
      } else {
        mt = mt + interval
        tmp$months.before.final[rows] = mt
      }
    }
    pt[which(pt$months.before.final == mnths),] = tmp
  }

  pt$nl = (pt$Block-1)*2
  arrange(pt, -Endoscopy.Year, PID)
}

patient.info = subset(patient.info, Patient %in% sum.patient.data$Patient)

back = do.call(rbind.data.frame,lapply(pg.samp, transform.by.time, info=patient.info))
back = merge(back, patient.info[,c('Samplename','p53.Status')])

sq = c(0, seq(1,132,12),max(back$months.before.final))
back$brks = cut(back$months.before.final, 
                breaks=sq, 
                labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),
                include.lowest = T, ordered_result = T)
back$brks = factor(back$brks, rev(levels(back$brks)))

back$Pathology = as.character(back$Pathology)
back$Pathology = factor(back$Pathology, levels=c('BE','ID','LGD','HGD', 'IMC'), ordered = T)

```

The labels in each decile are the ratio of progessor samples that have probabilities within that region. The deciles that are greyed out show insigificant enrichment for the NP/P ratio.

```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
cuts = seq(0,1,0.1)

back$quants = with(back, cut(Prediction, breaks=cuts))
qt = as.data.frame.matrix(table(back$quants, back$Status))
qt$quant = rownames(qt)

ft.fun<-function(NP,P) {
  f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value)
  #cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value,  'conf'=ifelse(p.value < 0.05, '*', '') )

pred.confidence[c('r1','r2')] = NA
for (i in 1:(length(cuts)-1)) 
  pred.confidence[i, c('r1','r2')] = round(range(cuts[i:(i+1)]), 3)

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

pred.confidence = bind_cols(pred.confidence, 
                            data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
                                       ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))

annotation.table = arrange(pred.confidence %>% group_by(Risk) %>% dplyr::summarise('n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2)), min)

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = riskCols, col=NA, alpha=0.2),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="black", fontface=4L)),
  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
  padding=unit(c(5,5),'mm'))

#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

lows = chisq.test(rbind(colSums(pred.confidence[1:5,c('NP','P')]),table(sum.patient.data$Status)))
highs = chisq.test(rbind(colSums(pred.confidence[6:10,c('NP','P')]),table(sum.patient.data$Status)))

mods = chisq.test(rbind(colSums(pred.confidence[4:7,c('NP','P')]),table(sum.patient.data$Status)))
```

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
#pred.confidence$Risk[1:5] = 'Low'
#pred.confidence$Risk[6:10] = 'High'
#pred.confidence$Risk[4:7] = 'Moderate'

pcal = ggplot(pred.confidence, aes(mn, perc)) + 
  geom_rect(data=pred.confidence, aes(xmin=r1, xmax=r2, ymin=0,ymax=1, fill=Risk), alpha=0.6) + 
#  scale_fill_manual(values=riskCols, limits=levels(pred.confidence$Risk) ) +
  geom_point() + geom_abline(slope=1, intercept=0, color='grey39', linetype='dotted', size=0.5) + #geom_text(aes(label=paste(r1,r2,sep='-')), nudge_x=0.05) +
  geom_errorbar(aes(ymin=ci.low, ymax=ci.high), size=0.5, width=0.01) + labs(x='P(P)', y='NP:P', title='') +
#  scale_color_manual(values=riskCols,limits=levels(pred.confidence$Risk) ) + 
  scale_x_continuous(limits=c(0,1), breaks = cuts, labels = c(0.0, '','',0.3,'','','',0.7,'','',1.0)) +
  plot.theme + theme(legend.position = 'none')
#pcal

#ggsave(filename='thesis/pred_calib.png', plot=pcal, width=2.2, height=2.2, units='in', dpi=300, scale=1.5)

back = merge(back, pred.confidence[c('quant','Risk')], by.x='quants',by.y='quant')
back = back %>% group_by(Risk) %>% dplyr::mutate('se(Risk)'=(sd(Prediction)/sqrt(length(Risk)))*1.96 )

```

```{r relrisk, echo=F, fig.width=6, fig.height=6}
myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))

predhist = ggplot(back, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
predhist

ggsave(filename='thesis/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
```

At each quintile are the percentage of progressor samples that were predicted within that range, and the hypergeometric p-value (Fishers) for the significance within that quintile. Then our predictions are at best a coin flip in the middle, but highly significant at the outer edges. This would also push the prediction for "high" risk up to `r min(subset(pred.confidence, labels == 'High' )$r1)` while "low" risk is below `r max(subset(pred.confidence, labels == 'Low' )$r2)`.

`r pander(as.data.frame(pred.confidence))`

## How do the predictions look at each time point?

Here we'll just use the cutoff 0.5 (I think that's what we used in the thesis?).

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
back[back$Prediction > 0.5, 'Risk'] = 'High'
back[back$Prediction < 0.5, 'Risk'] = 'Low'
back$Risk = droplevels(back$Risk)

riskPath = rbind( cbind.data.frame( table(subset(back, Status == 'P', select=c('Pathology','Risk'))), 'Status'='Progressor' ), 
cbind.data.frame( table(subset(back, Status == 'NP', select=c('Pathology','Risk'))), 'Status'='Non-Progressor' ))

pathTotal = back %>% group_by(Status, Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology))
pathTotal$Status = ifelse(pathTotal$Status == 'P', 'Progressor','Non-Progressor')

riskPath = merge(riskPath, pathTotal)
riskPath$ratio = with(riskPath, round(Freq/nPath, 3))
riskPath$Risk = factor(riskPath$Risk, levels=levels(back$Risk), ordered = T)

p53Path = cbind.data.frame(table(subset(back, Status == 'P', select=c('Pathology','p53.Status'))), 'Status'='p53 IHC')
p53Path = merge( subset(back, Status == 'P' & !is.na(p53.Status)) %>% group_by(Pathology) %>% dplyr::summarise( 'nPath'=length(Pathology)), p53Path)
#p53Path$Risk = with(p53Path, ifelse(p53.Status == 1, 'High','Low'))
p53Path$ratio = with(p53Path, round(Freq/nPath, 2))

#riskPath = rbind(riskPath, p53Path[colnames(riskPath)])
#riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor','p53 IHC'), ordered = T)
riskPath$Status = factor(riskPath$Status, levels=c('Non-Progressor','Progressor'), ordered = T)

pathTotal = riskPath %>% group_by(Status, Pathology) %>% dplyr::summarise('nPath'=unique(nPath))

#riskPath$Pathology = as.character(riskPath$Pathology)
#riskPath$Pathology[riskPath$Pathology == 'BE'] = 'NDBE'
#riskPath$Pathology = factor(riskPath$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered=T)

rp = ggplot(riskPath, aes(Pathology, ratio*100)) + 
  geom_bar(data=riskPath, aes(group=Risk, fill=Risk),stat='identity', position=position_stack()) + 
  scale_fill_manual(values=riskCols, limits=levels(riskPath$Risk)) + 
  scale_color_manual(values=c('black','black','black'), limits=levels(pred.confidence$Risk) ) +
  scale_x_discrete(labels=c('NDBE','ID','LGD','HGD','IMC')) +
  geom_text(data=pathTotal, aes(label=paste('n=',nPath,sep=''), x=Pathology, y=102), position=position_stack()) +
  geom_text(data=subset(riskPath, ratio>0), aes(group=Risk,label=paste(ratio*100,'%',sep=''), color=Risk), position=position_stack(vjust = 0.5), show.legend = F) +
  facet_grid(~Status, scales='free_x', space = 'free_x') +  
  labs(title='Samples predicted by pathology', y='% Predicted', x='Pathology') + theme(legend.position = 'none')
rp

ggsave(filename='thesis/pred_by_path.png', plot=rp, width=5.8, height=4.5, units='in', dpi=300, scale=1)

```

```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path = arrange(p53Path, Pathology, p53.Status)
col = RColorBrewer::brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  scale_x_discrete(labels=c('NDBE','ID','LGD','HGD','IMC')) +
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=103)) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  labs(title='', y='% Predicted', x='Pathology') + theme(legend.position = 'bottom')
rp
ggsave(filename='thesis/p53_pred.png', plot=rp, width=4.5, height=3.6, units='in', dpi=300, scale=1.2)

```

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = back %>% group_by(Status, Patient) %>% dplyr::summarise( 'Low'=sum(length(which(Risk == "Low"))),
#'Moderate'=sum(length(which(Risk == "Moderate"))),
'High'=sum(length(which(Risk == "High"))),
'n.samples'=length(Patient) )

all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))
head(all.preds)

ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) +
  geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=riskCols) +
  geom_text(aes(y=1.01, label=n.samples)) +
  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') +
  plot.theme

```

## Backwards from HGD

Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.

Take per endoscopy average prediction.

There are a lot more moderate predictions in this model...how does this work with this plot??

```{r enods, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
endos = back %>% group_by(Status, months.before.final, brks, PID) %>% dplyr::summarise(
  'pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
  'n.endos'=length(unique(PID)),
  'n.samples'=length(Samplename),
  'mean.pred'=mean(Prediction),
  'pred.samples'=length(which(Risk == 'High')),
  'max.path'=max(Pathology)
  )

sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
mlab = c('Endpoint','1-24','24-48','48','48-72','72-96','96-120','>120')

# months = 12
# sq = c(0, 1, 1*months, 2*months, 4*months, 6*months, 8*months, 10*months, 12*months)
# mlab = c()
# for (i in 1:length(sq)) {
#   mlab = c(mlab, paste(sq[i], sq[i+1], sep='-'))
# }
# mlab[1] = 'Endpoint'
# mlab = mlab[1:(length(sq)-1)]
# mlab[length(mlab)] = paste('>',sq[(length(sq)-1)],sep='')
#

endos$brks = cut(endos$months.before.final, include.lowest = T, ordered_result = T,breaks=sq ,labels=mlab)
endos$brks = with(endos, factor(brks, rev(levels(brks)))) # invert order

endos.r = endos %>% group_by(Status, brks) %>% dplyr::summarise( 
  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

#endos.r = subset(endos, Status == 'P') %>% group_by(brks) %>% dplyr::summarise( 
#  'pred.r'=sum(pred.endo)/sum(n.endos),'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos))

tmp = back # subset(back, Status == 'P')
tmp$brks = cut(tmp$months.before.final, include.lowest = T, ordered_result = T,breaks=sq,labels=mlab)
tmp = tmp %>% group_by(Status, brks) %>% dplyr::summarise( 'mean(P)'=mean(Prediction),'sem(P)'=sd(Prediction)/sqrt(length(brks)) )

endos.r = arrange(merge(endos.r,tmp), brks)
endos.r$Status = plyr::revalue(endos.r$Status, c('NP'='Non-Progressor', 'P'='Progressor'))

rp = ggplot(endos.r, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) + ylim(0,110) +
  facet_grid(~Status) +
  geom_bar(stat='identity', show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2, alpha=0.5) +
  geom_text(aes(y=pred.r*100+2, label=paste(round(pred.r*100), '%', sep='')), color='black') +
  geom_text(aes(y=10, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='High risk predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') +
  theme( axis.text.x=element_text(angle=45,hjust=1) ) 
rp
ggsave(filename='thesis/pred_endo.png', plot=rp, width=11, height=5, units='in', dpi=300, scale=1)

```

```{r heatmaps, echo=F, warning=F, message=F}
back$nl = factor(back$nl, ordered=T)
back$Risk = factor(back$Risk, levels=c('Low','High'), ordered = T)
back$Patient = factor(as.numeric(back$Patient), ordered = T)

table.path.risk<-function(df) {
  summary.t = as.data.frame.matrix(with(df, table(Pathology, Risk)))
  summary.t$Pathology = rownames(summary.t)

  summary.t = summary.t %>% group_by(Pathology) %>% dplyr::mutate( 'High.r'=round(High/sum(High,Low),3), 
                                                            'Low.r'=round(Low/sum(High,Low),3))
  for (g in unique(df$Risk)) {
    for (i in 1:nrow(summary.t)) {
      summary.t[i,g] = paste(summary.t[i,g], ' (', summary.t[i,paste(g,'.r',sep='')]*100, '%)',sep='')
    }
  }
  summary.t[,c('Pathology',levels(df$Risk))]
}

summary.t = table.path.risk(back)
```

`r pander(as.data.frame(summary.t))`

```{r, echo=F, message=F, warning=F, width=26, height=8}
back = back %>% group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2), 'n.endos'=length(unique(months.before.final)), 'max.block'=max(Block))
back = transform(back, Patient=reorder(Patient, max.block))
back$Block = factor(back$Block, ordered = T)

#pred.tiles(subset(back, Patient %in% c(1,20,34,21,3, 19)), cols=riskCols, probs=F, space='free', scales='free_x', drop=T) + theme(legend.position = 'none')

pred.tiles(subset(back, Status == 'P'), cols=riskCols, probs=F, space='free', scales='free_x', drop=T) + labs(title='Progressor')
pred.tiles(subset(back, Status == 'NP'), cols=riskCols, probs=F, space='free', scales='free_x', drop=T) + labs(title="Non-Progressor")

#all = pred.tiles(back, cols=riskCols, path=4, probs=F, scales='free', space='free') + labs(title='', x='', y='') + #theme(legend.position = 'bottom')
#riskLegend = get.legend(all)
```

# AP Clustering

```{r apclust-func, echo=F, message=F, warning=F}
apclust.data<-function(segdata, samples) {
  x1 = as.data.frame(segdata[,intersect(colnames(segdata), samples)])
  #x1 = x1[, samples]
  colnames(x1) = intersect(colnames(segdata), samples)
  rownames(x1) = (segdata[,c(1:4)] %>%
    rowwise() %>%
    mutate(location=paste(paste(chrom, arm, sep=''), '.', start.pos, '-', end.pos, sep='')))$location
  return(x1)  
}

apclust<-function(segdata) {
  x1 = apclust.data(segdata)
  # q=0 minimize off-diagonal similarity
  ac = apcluster(negDistMat(r=2), x1, details=T, convits=25, q=0)
  return(list('apres'=ac, 'data'=x1))
}

net.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@netsim
}

sum.similarity<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@dpsim
}

clusters<-function(ac) {
  if (is.list(ac))
    ac = ac$apres
  ac@exemplars
}

plotAC<-function(aclist) {
  plot(aclist$apres, aclist$data)
}

heatmapAC<-function(aclist) {
  heatmap(aclist$apres)
}

# First vs last samples
#x1 = apclust.data(patient.data[[pt]]$seg.vals[,c(1:4, 6, ncol(patient.data[[pt]]$seg.vals))], patient.data[[pt]]$info$Samplename)

```

### Raw segmentation values

```{r apclust, echo=F, message=F, warning=F, fig.height=10, fig.width=10, fig.align='left'}
library(apcluster)
sample.vals = list()
q=0.2

sum.patient.data = subset(arrange(sum.patient.data, Status), Set == 'Training')

## Using the segmentation values
sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))
sim = as.data.frame(sum.patient.data[,c('Status','years')])
rownames(sim) = sum.patient.data$Hospital.Research.ID
for (pt in sum.patient.data$Hospital.Research.ID) {

  patient.data[[pt]]$info = arrange(patient.data[[pt]]$info,  Pathology, Endoscopy.Year)
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)

  gg = ggplot(melt(x1), aes(x=variable, y=value, group=variable, fill=variable)) + 
    geom_boxplot(notch=T, show.legend=F) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), legend.position='none') +
    labs(title=paste(pt, patient.data[[pt]]$info$Status[1], sep=' - ')) + 
    theme(axis.title.x=element_blank(), axis.text.x=element_blank(),axis.ticks.x=element_blank())

  sample.vals[[pt]] = gg

  info = patient.data[[pt]]$info[which(colnames(x1) %in% patient.data[[pt]]$info$Samplename),]
  colnames(x1) = paste(info$Endoscopy.Year, '(', info$Pathology, ')', sep='' )

  ac = apcluster(negDistMat(x1, r=2), q=q, details=T, includeSim=T)

  last = ncol(x1)
  if (patient.data[[pt]]$info$Status[1] == 'P' & length(which(grepl('HGD|IMC', info$Pathology)) > 0)  )
    last = sort(grep('HGD|IMC', info$Pathology), decreasing=T)[1]
  if (ncol(x1) <= 15) {
    #plot(ac, x1[,c(c(1:ncol(x1))[-last], last)], main=pt)
  } else {
    #plot(ac, x1[,c(1, sort(sample(c(1:(ncol(x1)-1))[c(-1,-last)], 8)) ,last) ], main=pt)
  }

  sim[pt, 'ns'] = net.similarity(ac)
  sim[pt, 'ss'] = sum.similarity(ac)
  sim[pt, 'clusters'] = length(clusters(ac))
  sim[pt, 'max.sd'] = max(sapply(x1, sd))

  tmp = patient.data[[pt]]$variable.regions[,-(1:5)]
  if (!is.null(tmp)) {
    sim[pt, 'n.var'] = nrow(patient.data[[pt]]$variable.regions[,-(1:5)])
    if (nrow(tmp) <= 0) {
      sim[pt,c('gains', 'loss')] = c(0,0)
    } else {
      sim[pt,'gains'] = length(tmp[tmp>=1.1])
      sim[pt,'loss'] =  length(tmp[tmp<=0.9])
    }
  } else {
    sim[pt, 'n.var'] = 0
    sim[pt,c('gains', 'loss')] = c(0,0)
  }

}
sim$id = rownames(sim)
rc = cor.test(sim$max.sd, sim$clusters)

#do.call(grid.arrange, c(sample.vals))

# May be a difference
ttps = wilcox.test( subset(sim, Status == 'NP')$ss, subset(sim, Status == 'P')$ss )$p.value
ttpc = wilcox.test( subset(sim, Status == 'NP')$clusters, subset(sim, Status == 'P')$clusters )$p.value

## Number of variable regions isn't so good, but maybe...
#sim %>%
#  group_by(Status) %>%
#  summarise( mean = mean(n.var), sd = sd(n.var), med = median(n.var))
#t.test( subset(sim, Status == 'NP')$n.var, subset(sim, Status == 'P')$n.var )$p.value

g1 = ggplot(sim[,c('Status','ss')], aes(y=ss, x=Status, fill=Status)) + geom_boxplot() + 
  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
  labs(title='All samples', subtitle=paste("p-value", signif(ttps, 3)), y='sum similarity')

g2 = ggplot(sim[,c('Status','clusters')], aes(y=clusters, x=Status, fill=Status)) + geom_boxplot() + 
  geom_jitter(width=0.2) + stat_boxplot(geom ='errorbar') +
  labs(title='All samples', subtitle=paste("p-value", signif(ttpc, 3)), y='num. clusters')

```

This plot compares the two possible measures of overall complexity from AP. The summed similarity value (top) and the number of clusters (bottom). In this case, the number of clusters appears to be show a difference between the two groups.

```{r echo=F, message=F, warning=F, fig.width=6, fig.height=12}
grid.arrange(g1, g2, top='Raw seg values')
```

## No HGD samples

There is still a difference between the progressors and non, though the p-value does increase. Some of the progressors do start looking more like NPs in their similarity measure but not all of them.

```{r nofinal, echo=F, message=F, warning=F, fig.height=5, fig.width=5}
simX = sum.patient.data[,c('Status','years')]
rownames(simX) = sum.patient.data[,'Hospital.Research.ID']
# Leave out HPC/IMC samples
for (pt in sum.patient.data$Hospital.Research.ID) {
  x1 = apclust.data(patient.data[[pt]]$seg.vals, patient.data[[pt]]$info$Samplename)
  hgdsamples = subset(patient.data[[pt]]$info, Pathology %in% c('HGD','IMC'))$Samplename

  rows = rownames(x1)
  col.nm = colnames(x1)
  if (length(hgdsamples) > 0) {
    cols = grep(paste(hgdsamples, collapse='|'), colnames(x1), invert=T)
    x1 = as.data.frame(x1[,cols])
    colnames(x1) = col.nm[cols]
    rownames(x1) = rows
  }

  if (ncol(x1) > 0) {
    ac = apcluster(negDistMat(x1, r=2), q=q)

    simX[pt, 'ss'] = sum.similarity(ac)
    simX[pt, 'clusters'] = length(clusters(ac))

    tmp = patient.data[[pt]]$variable.regions[,-(1:5)]
    simX[pt, 'n.var'] = ifelse (!is.null(tmp), nrow(patient.data[[pt]]$variable.regions[,-(1:5)]), 0)
  }
}

# Correlation between the number of variable regions and clusters, suggesting the complexity measure is reasonable
ct = cor.test(simX$clusters, simX$n.var)

ttpc = wilcox.test( subset(simX, Status == 'NP')$clusters, subset(simX, Status == 'P')$clusters )$p.value

ggplot(melt(data.frame(simX[c('Status','clusters')])), aes(x=Status, y=value, fill=Status)) +  
    geom_boxplot() + geom_jitter(shape=16, position=position_jitter(0.2)) +
    stat_boxplot(geom ='errorbar') + labs(y='num. clusters', x='', title='Excluding HGD/IMC', subtitle=paste('p-value', signif(ttpc, 3)))

```

# Validation Set

```{r echo=F, warning=F, message=F}

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("No model file", file, "found"))
}

validation.patient.info =  subset(all.patient.info, Set == 'Test')
validation.patient.info = arrange(validation.patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)

cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
cleaned = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), cleaned, value=T)
cleaned = grep('arms', cleaned, invert=T, value=T)

#raw = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='raw', full.names=T, recursive=T)
#raw = grep(paste(unique(validation.patient.info$Hospital.Research.ID), collapse='|'), raw, value=T)

fitV = models[[0.9]]
lambda.opt = performance.at.1se[[0.9]][, 'lambda']

validation.df = matrix(nrow=0, ncol=ncol(dysplasia.df), dimnames=list(c(), colnames(dysplasia.df)))
for (fi in cleaned) {
  x = read.table(fi, sep='\t', header=T)
  segs = paste(x$chr, ':', x$start, '-', x$end, sep='')

  x = t(x[,-c(1:3)])
  colnames(x) = segs

  validation.df = rbind(validation.df, x)  
}

validation.df = validation.df[validation.patient.info$Samplename, ]
validation.df = apply(validation.df, 2, unit.var)

pred = predict(fitV, newx=validation.df, s=lambda.opt, type='response')
or = predict(fitV, newx=validation.df, s=lambda.opt, type='link')

vpd = validation.patient.info

if (rownames(pred) != vpd$Samplename)
  stop('Fix this')

vpd$Prediction = pred
vpd$OR = or

table(subset(vpd, Prediction > cutoff)$Status)
table(subset(vpd, Prediction < cutoff)$Status)

```

## Technical Replicates

Predicting the replicates that were left out of the model building process.  Though maybe they shouldn't have been!

Anyhow...

```{r echo=F, warning=F, message=F}

#normals = subset(read_xlsx(patient.file, sheet='All combined'), Pathology == 'normal')
repeats = read_xlsx(patient.file, sheet='Technical Repeats')

samples = 
  c(gsub('-', '_', paste(repeats$`Plate Index`, sub('SLX-','', repeats$`SLX ID`), sep='_')))
  #gsub('-', '_', paste(normals$`Plate Index`, sub('SLX\\s?-\\s?', '',  normals$`SLX ID 1`), sep='_')))
samples = sub('10729', '10725_10729', samples)

repeats = repeats[,c(1:6,8,9)]
repeats$Samplename = samples

cleaned = list.files(path=paste(data, 'Cleaned', sep='/'), pattern='tiled', full.names=T, recursive=T)
cleaned = grep(paste(unique(repeats$`Hospital Research ID`), collapse='|'), cleaned, value=T)
cleaned = grep('arms', cleaned, invert=T, value=T)

repeat.df = matrix(nrow=0, ncol=ncol(dysplasia.df), dimnames=list(c(), colnames(dysplasia.df)))
for (fi in cleaned) {
  x = read.table(fi, sep='\t', header=T)
  segs = paste(x$chr, ':', x$start, '-', x$end, sep='')

  x = t(x[,-c(1:3)])
  colnames(x) = segs

  repeat.df = rbind(repeat.df, x)  
}

rows = which(rownames(repeat.df) %in% samples)
repeat.df = repeat.df[rows,]

repeat.df = apply(repeat.df, 2, unit.var)

pred = predict(fitV, newx=repeat.df, s=lambda.opt, type='response')
or = predict(fitV, newx=repeat.df, s=lambda.opt, type='link')

repeats$Probability = pred[repeats$Samplename,]
repeats$OR = or[repeats$Samplename,]

```

`r pander(repeats)`