### 2018-05-17 SNP_sWGS_comparison.Rmd

---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r, include=FALSE}
library(copynumber)
library(tibble)
library(dplyr)

```

```{r}

relCNV = read.table('~/Data/Ellie/Cleaned/AD0027_segmentedCoverage_fitted_gamma250/raw_probefiltered_segvals.txt', sep='\t', header=T)
head(relCNV)
#counts = read.table('~/Data/Ellie/QDNAseq/SLX-12452/SLX-12452.binSize15.readCounts.txt')
#corrected = read.table('~/Data/Ellie/QDNAseq/SLX-12452/SLX-12452.binSize15.correctedReadCounts.txt')

summary(relCNV)

segCNV = read.table('~/Data/Ellie/Cleaned/AD0027_segmentedCoverage_fitted_gamma250/5e+06_cleaned_tiled_segvals.txt', sep='\t', header=T)
head(segCNV)

```

```{r, warnings=F}
sa = data.table::fread('~/Downloads/GSE99429_normalized_data.txt')
#colnames(sa)
# baseline, hgd  final, EA
infocol = grep('SNP|Chr|Pos', colnames(sa), value=T)

ptcol = grep('391', colnames(sa), value=T)

pt = sa[, c(infocol, ptcol), with=F]

samplenames = unique( sub('\\.[Log|B\\.Allele].*', '', ptcol))
perpt = lapply (samplenames , function(sname) {
  x = pt[, c(infocol,grep(paste(sname, '.*Log', sep=''), ptcol, value=T),grep(paste(sname, '.*Allele', sep=''), ptcol, value=T)), with=F  ]
  colnames(x) = c('rs.id', 'chr','position', 'Log.R','B.Allele.Freq')

  x = subset(x, !chr %in% c('0','XY','MT') & !is.na(Log.R))

  x[which(x$chr == 'X'), 'chr'] = 23
  if (length(which(x$chr == 'YY')) > 0)
    x[which(x$chr == 'Y'), 'chr'] = 24

  x$chr = as.numeric(x$chr)
  x = dplyr::arrange(x, chr, position)

  as_tibble(cbind('sample'=sname, x))
})
names(perpt) = samplenames

data = perpt$`391_67K_SS6_2_2_22654`[,c('chr','position','Log.R')]
colnames(data)[3] = '391_67K_SS6_2_2_22654'
data$chr = as.numeric(data$chr)
segdata = copynumber::pcf(as.data.frame(subset(data, chr %in% c(1:22))), gamma=28, fast=T, verbose=T)
hist(segdata$mean)

head(perpt$`391_67K_SS6_2_2_22654`)
head(perpt$`391_73P__normal_5350N_5ng`)

hist(perpt$`391_67K_SS6_2_2_22654`$Log.R)
ks.test(perpt$`391_67K_SS6_2_2_22654`$Log.R, 'pnorm')

ks.test(rnorm(500, mean(perpt$`391_67K_SS6_2_2_22654`$Log.R), sd(perpt$`391_67K_SS6_2_2_22654`$Log.R)), 'pnorm')

ks.test(perpt$`391_67K_SS6_2_2_22654`$Log.R, rnorm(500, mean(perpt$`391_67K_SS6_2_2_22654`$Log.R), sd(perpt$`391_67K_SS6_2_2_22654`$Log.R)))
?ks.test

```