### 2018-05-22 SNP_sWGS_comparison.Rmd
testing SNP data

---
title: "R Notebook"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
library(copynumber)
library(tibble)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(glmnet)

suppressPackageStartupMessages( source('lib/data_func.R') )
```

Single sample from shallow WGS dataset.
```{r, warning=F, message=F}
relCNV = read.table('~/Data/Ellie/Cleaned/AH0254_segmentedCoverage_fitted_gamma250/raw_probefiltered_segvals.txt', sep='\t', header=T)

segCNV = load.segment.matrix('~/Data/Ellie/Cleaned/AH0254_segmentedCoverage_fitted_gamma250/5e+06_cleaned_tiled_segvals.txt')
unlogged = segCNV
segCNV = t(apply(segCNV, 1, logTransform))

grid.arrange(
  ggplot(relCNV, aes(x=1:nrow(relCNV), y=D702_D506_10725_10729)) + geom_point(color='darkblue') + labs(title='Raw', x=''),
  ggplot(melt(segCNV['D702_D506_10725_10729',]), aes(x=1:ncol(segCNV), y=value)) + geom_point(color='darkblue') + labs(title='Tiled & Logged', y='D702_D506_10725_10729 (IMC)', x='') )

grid.arrange(
  ggplot(relCNV, aes(x=1:nrow(relCNV), y=D706_D507_9242)) + geom_point(color='darkblue') + labs(title='Raw', x=''),
  ggplot(melt(segCNV['D706_D507_9242',]), aes(x=1:ncol(segCNV), y=value)) + geom_point(color='darkblue') + labs(title='Tiled & Logged', y='D706_D507_9242 (NDBE)', x='') )

grid.arrange(  
  ggplot(relCNV, aes(D702_D506_10725_10729)) + geom_histogram(color='lightblue', bins=25) + labs(title='Raw'),
  ggplot(melt(unlogged['D702_D506_10725_10729',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Segmented'),
  ggplot(melt(segCNV['D702_D506_10725_10729',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Logged & Segmented'), top='IMC'
)

grid.arrange(  
  ggplot(relCNV, aes(D706_D507_9242)) + geom_histogram(color='lightblue', bins=25) + labs(title='Raw'),
  ggplot(melt(unlogged['D706_D507_9242',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Segmented'),
  ggplot(melt(segCNV['D706_D507_9242',]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Logged & Segmented'), top='NDBE'
)

# All samples
load('~/Data/Ellie/Analysis/5e6_arms_all_logR/model_data.Rdata', verbose=T)
load('~/Data/Ellie/Analysis/5e6_arms_all_logR/loo.Rdata', verbose=T)
subset(pg.samp$AH0254, Samplename %in% c('D706_D507_9242','D702_D506_10725_10729'))

```

GSE99429 LogR data
```{r, warning=F, message=F}
filelist = list.files('~/Data/Ellie/arrays/SNP', pattern='txt', full.names=T)
samplelist = lapply(filelist, function(f) colnames(as_tibble(data.table::fread(f, nrows=2))))

whole_epi = lapply(samplelist, function(s) grep('*whole.*Log', s, value=T)  )
names(whole_epi) = basename(filelist)

normals = lapply(samplelist, function(s) grep('*normal.*Log', s, value=T)  )
names(normals) = basename(filelist)

fix.chr<- function(df) {
  df = subset(df, Chr %in% c(1:22))
  df$Chr = as.numeric(df$Chr)
  dplyr::arrange(df, Chr, Position)
}

dt = as_tibble(data.table::fread(filelist[6]))
epi = dt[,c(grep('SNP|Chr|Pos', colnames(dt), value=T),whole_epi[[6]])]
nm = dt[,c(grep('SNP|Chr|Pos', colnames(dt), value=T),normals[[6]])]

epi = fix.chr(epi)
nm = fix.chr(nm)
rm(dt)

```

```{r, warning=F, message=F}

process.patients<-function(df, outfile=NULL) {
  predictions = data.frame(matrix(nrow=0,ncol=2,dimnames=list(c(), c('Prob','RR'))))
  allptsdata = list()

  infocols = grep('Chr|Pos', colnames(df), value=T)
  pts = grep('SNP|Chr|Pos', colnames(df), invert=T, value=T)

  for (pt in pts)  {
    print(pt)    
    data = df[,c(infocols,pt)]
    segdata = copynumber::pcf(as.data.frame(data), gamma=28, fast=T, verbose=F)  
    head(segdata)

    #ggplot(segdata, aes(x=1:nrow(segdata), y=mean)) + geom_point() + labs(title='LogR whole epi', x='', y='')

    # this is the LogR, doesn't need to be logged again
    tiled = segment.matrix(tile.segmented.data(segdata[,-1], chr.info=get.chr.lengths()))
    tiled.noUV = tiled

    for (i in 1:ncol(tiled)) 
      tiled[,i] = unit.var(tiled[,i], z.mean[i], z.sd[i]) 
    tiled[is.na(tiled)] = mean(tiled, na.rm=T) #0

    cx = score.cx(tiled, 1)

    tiled.arms = segment.matrix(tile.segmented.data(segdata[,-1], size='arms', chr.info=get.chr.lengths()))
    tiled.armsNoUV = tiled.arms

    for (i in 1:ncol(tiled.arms)) 
      tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i]) 
    tiled.arms[is.na(tiled.arms)] = mean(tiled, na.rm=T) #0

    arrayDf = subtract.arms(tiled, tiled.arms)
    arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx ,sd.cx))

    # Whole epi
    predictions[pt,] = c(predict(fitV, newx=arrayDf, s=l, type='response')[,1], 'RR'=predict(fitV, newx=arrayDf, s=l, type='link')[,1] )

    allptsdata[[pt]] = list('seg'=segdata, 'tile'=tiled, 'tileNoUV'=tiled.noUV, 'tile.arms'=tiled.arms, 'tile.arms.noUV'=tiled.armsNoUV, 'arrayDf'=arrayDf, 'cx'=cx)
  }  
  if(!is.null(outfile))   save(wholeepi, predictions, file=outfile)

return(predictions)
}

load('~/Data/Ellie/Analysis/5e6_arms_all_logR/all.pt.alpha.Rdata', verbose = T)
fitV = models$`0.9`
l = performance.at.1se$`0.9`$lambda
rm(plots,performance.at.1se,dysplasia.df,models,cvs,labels)

tmp = paste('/tmp/', sub('\\.txt','',basename(filelist[6])), '_epi.Rdata', sep='')
if (file.exists(tmp)) {
  load(tmp, verbose=T)
} else {
  predictions = data.frame(matrix(nrow=0,ncol=2,dimnames=list(c(), c('Prob','RR'))))
  wholeepi = list()

  infocols = grep('Chr|Pos', colnames(epi), value=T)
  pts = grep('SNP|Chr|Pos', colnames(epi), invert=T, value=T)
  pt = pts[1]

  for (pt in pts)  {
    print(pt)    
    data = epi[,c(infocols,pt)]
    segdata = copynumber::pcf(as.data.frame(data), gamma=28, fast=T, verbose=F)  
    head(segdata)

    #ggplot(segdata, aes(x=1:nrow(segdata), y=mean)) + geom_point() + labs(title='LogR whole epi', x='', y='')

    # this is the LogR, doesn't need to be logged again
    tiled = segment.matrix(tile.segmented.data(segdata[,-1], chr.info=get.chr.lengths()))
    tiled.noUV = tiled

    for (i in 1:ncol(tiled)) 
      tiled[,i] = unit.var(tiled[,i], z.mean[i], z.sd[i]) 
    tiled[is.na(tiled)] = mean(tiled, na.rm=T) #0

    cx = score.cx(tiled, 1)

    tiled.arms = segment.matrix(tile.segmented.data(segdata[,-1], size='arms', chr.info=get.chr.lengths()))
    tiled.armsNoUV = tiled.arms

    for (i in 1:ncol(tiled.arms)) 
      tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i]) 
    tiled.arms[is.na(tiled.arms)] = mean(tiled, na.rm=T) #0

    arrayDf = subtract.arms(tiled, tiled.arms)
    arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx ,sd.cx))

    # Whole epi
    predictions[pt,] = c(predict(fitV, newx=arrayDf, s=l, type='response')[,1], 'RR'=predict(fitV, newx=arrayDf, s=l, type='link')[,1] )

    wholeepi[[pt]] = list('seg'=segdata, 'tile'=tiled, 'tileNoUV'=tiled.noUV, 'tile.arms'=tiled.arms, 'tile.arms.noUV'=tiled.armsNoUV, 'arrayDf'=arrayDf, 'cx'=cx)
    }

  save(wholeepi, predictions, file=tmp)
}

```

Normal sample?
```{r, warning=F, message=F}

npg = '911_5U__normal_10397N_200ng'
pg = '848_85O_SS1_normal_18780N_200ng'

tmp = '/tmp/normal.Rdata'
if (file.exists(tmp)) {
  load(tmp, verbose = T)
} else {

  # this is the LogR, doesn't need to be logged again
  normals = list()
  for (pt in c(npg,pg)) {
    data = normpt[[pg]][,c('chr','position','Log.R')]
    colnames(data)[3] = pt
    data$chr = as.numeric(data$chr)
    segdata = copynumber::pcf(as.data.frame(subset(data, chr %in% c(1:22))), gamma=28, fast=T, verbose=F)

    tiledN = tile.segmented.data(segdata[,-1], chr.info=get.chr.lengths())
    tiledN = segment.matrix(tiledN)

    tiledN.noUV = tiledN

    for (i in 1:ncol(tiledN)) 
      tiledN[,i] = unit.var(tiledN[,i], z.mean[i], z.sd[i]) 
    tiledN[is.na(tiledN)] = mean(tiledN, na.rm=T) #0

    cx = score.cx(tiledN, 1)

    tiled.armsN = tile.segmented.data(segdata[,-1], size='arms', chr.info=get.chr.lengths())
    tiled.armsN = segment.matrix(tiled.armsN)

    tiled.armsN.noUV = tiled.armsN

    for (i in 1:ncol(tiled.armsN)) 
      tiled.armsN[,i] = unit.var(tiled.armsN[,i], z.arms.mean[i], z.arms.sd[i]) 
    tiled.armsN[is.na(tiled.armsN)] = mean(tiled.armsN, na.rm=T) #0

    arrayDfN = subtract.arms(tiledN, tiled.armsN)
    arrayDfN = cbind(arrayDfN, 'cx'=unit.var(cx, mn.cx ,sd.cx))

    normals[[pt]] = list('seg'=segdata, 'tile'=tiledN, 'tileNoUV'=tiledN.noUV, 'tile.arms'=tiled.armsN, 'tile.arms.noUV'=tiled.armsNoUV, 'arrayDf'=arrayDfN, 'cx'=cx)
  }
  save(normals, file=tmp)
}

#normseg = segdata

```

Norm vs whole ep
The extreme tail on the normal sample is due to the 'cx' coefficient and needs to be adjusted.  For now I will set it to the mean of the cx value in the sWGS dataset.
```{r}

head(wholeepseg)

grid.arrange(
  ggplot(wholeepseg, aes(x=1:nrow(wholeepseg), y=mean)) + geom_point() + labs(title='Whole epi, segmented LogR'),
  ggplot(normseg, aes(x=1:nrow(normseg), y=mean)) + geom_point() + labs(title='Normal, segmented LogR'))

grid.arrange(
  ggplot(wholeepseg, aes(mean)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Whole epi'),
  ggplot(normseg, aes(mean)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Normal sample')
)

summary(wholeepseg$mean)
summary(normseg$mean)

grid.arrange(
  ggplot(melt(arrayDf), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, tiled', subtitle='Whole epi, 5e6'),
  ggplot(melt(arrayDfN), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, tiled', subtitle='Normal sample, 5e6')
)

arrayDf[,'cx'] = mean(allDf[,'cx'])
arrayDfN[,'cx'] = mean(allDf[,'cx'])

```

Compare to all samples.
```{r warning=F, message=F}

ggplot(melt(allDf), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='all sWGS')

ggplot(melt(allDf[sample(1:nrow(allDf),1),]), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Random sWGS sample')

grid.arrange(
  ggplot(melt(arrayDf), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Whole epi'),
  ggplot(melt(arrayDfN), aes(value)) + geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Normal sample'))

s = which(apply(allDf, 1, function(i) ks.test(i, arrayDf)$p.value) >= 0.05)

# Normal
predict(fitV, newx=arrayDfN, s=l, type='response')
predict(fitV, newx=arrayDfN, s=l, type='link')

```

Quantile normalize?
```{r, warning=F, message=F}
by = 3
mx = mean(arrayDf, na.rm=T)+sd(arrayDf, na.rm=T)*by
mn = mean(arrayDf, na.rm=T)-sd(arrayDf, na.rm=T)*by

ggplot(melt(arrayDf), aes(value)) + 
  geom_histogram(color='lightblue', bins=25) + labs(title='Array LogR, Segmented', subtitle='Whole epi?') +
  geom_vline(xintercept = c(mx,mn), color='red')

## Get the distribution as below across ALL methylation samples then sub the mean and sd for ALL instead of for one

newsd = sd(arrayDf[arrayDf <= mx & arrayDf >= mn], na.rm=T)
newmean = mean(arrayDf[arrayDf <= mx & arrayDf >= mn], na.rm=T)

#xx = qnorm(pnorm(segMethy, mean(segMethy), sd(segMethy)), mean(allDf[,1:ncol(segMethy)]), sd(allDf[,1:ncol(segMethy)]))
#hist(xx)

xx = qnorm(pnorm(arrayDf, newmean, newsd, lower.tail=F), mean(allDf[,1:ncol(arrayDf)]), sd(allDf[,1:ncol(arrayDf)]), lower.tail=F)

# less extreme, but still incorrect
xx = qnorm(pnorm(arrayDf, newmean, newsd, lower.tail=F), median(apply(allDf, 1, mean, na.rm=T)), sd(apply(allDf, 1, sd, na.rm=T)), lower.tail=F)

hist(xx)

predict(fitV, newx=xx, s=l, type='response')
predict(fitV, newx=xx, s=l, type='link')
```

## Try processing the IDAT files directly.  Perhaps if they aren't normalized together the LogR becomes more comparable.
```{r}

library(illuminaio)

file = '~/Data/Ellie/arrays/SNP/GSM2643883_8832987072_R03C01_Grn.idat'
idat <- readIDAT(file)
names(idat)

```