### 2018-09-14 snp/SNP_Prediction_Analysis.Rmd
adding exclusions for low SCA

---
title: "SNP Analysis"
author: "Sarah Killcoyne"
date: "6/13/2018"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
library(ggplot2)
library(gridExtra)
library(glmnet)
library(pander)
library(readxl)
library(pROC)
library(ggfortify)
library(preprocessCore)

source('~/workspace/shallowwgs_pipeline/lib/data_func.R')
source('~/workspace/shallowwgs_pipeline/lib/common_plots.R')

chr.info = get.chr.lengths(file='/tmp/hg19_info.txt')

adjust.cols<-function(mtx, means=NULL, sds=NULL, na.replace=0) {
  if (!is.null(means)) {
    if (ncol(mtx) != length(means))
      stop("Vector of means needs to match the columns in the matrix")
    for (i in 1:ncol(mtx)) 
      mtx[,i] = unit.var(mtx[,i], means[i], sds[i])
  } else {
    for (i in 1:ncol(mtx)) 
      mtx[,i] = unit.var(mtx[,i])
  }
  mtx[is.na(mtx)] = na.replace

  return(mtx)
}

load('~/Data/Ellie/Analysis/5e6_arms_all-NOQC/model_data.Rdata', verbose = T)
load('~/Data/Ellie/Analysis/5e6_arms_all-NOQC/all.pt.alpha.Rdata', verbose = T)
#alpha = '0'
alpha = '0.9'
fitV = models[[alpha]]
l = performance.at.1se[[alpha]]$lambda
features = coefs[[alpha]]

swgs_labels = labels

rm(plots,performance.at.1se,models,cvs,labels)

pv = var(dysplasia.df[swgs_labels == 1])  # P
npv = var(dysplasia.df[swgs_labels == 0]) # NP

range(dysplasia.df[swgs_labels == 1])
range(dysplasia.df[swgs_labels == 0])

load('~/Data/Reid_SNP/PerPatient/allpts_ascat.Rdata', verbose=T)

qcdata = qcdata %>% rowwise %>% dplyr::mutate(
  PatientID = unlist(strsplit(Samplename, '_'))[1],
  SampleID = unlist(strsplit(Samplename, '_'))[2],
  EndoID = unlist(strsplit(Samplename, '_'))[3],
  Level = unlist(strsplit(Samplename, '_'))[4]
)
segments.list = lapply(segments.list, function(pt) {
  pt$sample = sub('\\.LogR','',pt$sample)
  pt
})

qcdata$ASCAT.SCA.ratio = apply(qcdata,1,function(s) {
  smp = subset(segments.list[[ s[['PatientID']] ]], sample == s[['Samplename']] & chr %in% c(1:22))
  smp = smp %>% rowwise %>% dplyr::mutate(
      #'Total' = nAraw + nBraw,
      'Total' = nMajor + nMinor,
      'CNV' = round(Total) - round(as.numeric(s[['Ploidy']])) )

  x = subset(smp, CNV != 0 & chr %in% c(1:22))
  sum(x$endpos - x$startpos) / chr.info[22,'genome.length']
})

qcdata$SampleType = 'BE'
qcdata$SampleType[grep('BLD',qcdata$Level, ignore.case=T)] = 'Blood Normal'
qcdata$SampleType[grep('gastric',qcdata$Level, ignore.case=T)] = 'Gastric Normal'

patient.info = as.data.frame(read_xlsx('~/Data/Ellie/Analysis/SNP/metadata_T1T2.xlsx'))
patient.info$UniqueSampleID = paste(patient.info$PatientID, patient.info$`Timepoint Code`, sep='_')
patient.info$Path.Status = patient.info$Status
patient.info[patient.info$PatientID %in% subset(patient.info, Pathology %in% c('IMC','HGD'), select='PatientID')[,1], 'Path.Status'] = 'P'

sample.list = read_xlsx('~/Data/Ellie/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2)
nrow(sample.list)
colnames(sample.list)[3] = 'Total.SCA'

sample.list$SCA.Ratio = sample.list$Total.SCA/max(chr.info$genome.length)

sample.info = qcdata[,c('PatientID','SampleID','EndoID','Level','Samplename','ASCAT.SCA.ratio')]

message(paste(length(unique(sample.info$PatientID)), 'patients listed in metadata file'))
message(paste(nrow(qcdata), 'samples available'))

sample.info = base::merge(sample.info, patient.info[,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')], by.x=c('PatientID','EndoID'), by.y=c('PatientID','Timepoint Code'))
head(sample.info)

get.ratio<-function(PID, SID) {
  subset(sample.list, PatientID == PID & SampleNum == SID)$SCA.Ratio
}

sample.info = sample.info %>% rowwise() %>% dplyr::mutate(
  SCA.Ratio = ifelse(Level %in% c('BLD','GASTRIC'), 0, get.ratio(PatientID, SampleID))
)

patient.info = patient.info %>% rowwise() %>% dplyr::mutate( Endoscopy = ifelse(Timepoint == 'T1', 1, 2) )

sample.info
ggplot(sample.info, aes(ASCAT.SCA.ratio, SCA.Ratio)) + geom_point() + facet_grid(~Status)

# ASCAT and Xiahong's estimates correlate in 86% of samples, that suggests that the 14% missing is due to the LOH I'm not trying to call
cor.test(sample.info$ASCAT.SCA.ratio, sample.info$SCA.Ratio)

```

## Patient info

`r nrow(patient.info)` total endoscopies, `r pander(table(patient.info$Status), caption="Split between P and NP")`

Histology breakdown:

```{r, echo=F}
sm = patient.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) )
sm$Pathology = factor(sm$Pathology, levels=c('NDBE','ID','LGD','HGD','IMC'), ordered = T)

ggplot(sm, aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + facet_grid(~Status) + 
  geom_col(position = position_dodge()) + ylim(0,1) + labs(title='Pathology frequency per timepoint')
```

There's a significant difference in the numbers of NDBE or HGD samples between the groups
```{r, echo=F}
pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['HGD',]), caption='HGD')
pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['NDBE',]), caption='NDBE')

# Most progressors are HGD at timepoint 1
pander( round(with(subset(patient.info, Endoscopy == 1), table(Pathology, Status))['HGD',]/table(subset(patient.info, Endoscopy == 1)$Status), 3)*100, caption='% at T1')
```

# Quick QC 
In the sWGS data progressor samples show a variance of `r round(pv, 3)` and non-prog variance is `r round(npv,3)`

Purity, ploidy and goodnessoffit from single sample (not paired) ASCAT.

```{r, echo=F, warning=F, message=F}
qcdata = base::merge(qcdata, sample.info[,c('Samplename','Status','Timepoint','SCA.Ratio')], by='Samplename')

ggplot(qcdata, aes(Status, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_boxplot() + facet_grid(~SampleType)

low = subset(qcdata, SampleType == 'BE' & ASCAT.SCA.ratio <= median(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio))

anno = cbind.data.frame(table(subset(qcdata, SampleType == 'BE' & ASCAT.SCA.ratio <= median(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio))$Status)/table(subset(qcdata, SampleType == 'BE')$Status))
colnames(anno) = c('Status','Ratio')
anno$label='median(normal SCA)'
anno$x = 0.4
anno$y = median(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio)

ggplot(subset(qcdata,SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_point() + facet_grid(~Status, scales = 'free') + ylim(0,0.005) + 
  geom_hline(yintercept = x) + geom_text(data=anno, aes(x, y, label=label), nudge_y = 0.0001) +
  geom_text(data=anno, aes(x, y, label=paste(round(Ratio,2)*100,'% samples',sep='')), nudge_y = -0.0001)

# ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, SCA.Ratio, color=PatientID, group=PatientID)) + geom_point() + geom_line() + 
#   facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')

ggplot(qcdata, aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + geom_point() + facet_grid(~SampleType)

ggplot(qcdata, aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + geom_point() + facet_grid(~Status)

ggplot(subset(qcdata,SampleType == 'BE'), aes(Timepoint, ASCAT.SCA.ratio, color=PatientID, group=Timepoint)) + geom_boxplot() + 
  facet_wrap(~Status, scales='free_x', ncol=2) + labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')

ggplot(qcdata, aes(round(Ploidy,2))) + geom_histogram() + 
  facet_wrap(~SampleType, ncol=1) + labs(x='Ploidy', title='Ploidy, normal vs BE')

ggplot(qcdata, aes(round(Purity,2))) + geom_histogram() + 
  facet_wrap(~SampleType, scales='free', ncol=1) + labs(x='Purity', title='Purity, normal vs BE')

#x=median(qcdata$SCA.Ratio,na.rm=T)+sd(qcdata$SCA.Ratio,na.rm=T)
x=0.05
y=median(qcdata$Purity,na.rm=T)-sd(qcdata$Purity,na.rm=T)/2

ggplot(qcdata, aes(Purity, ASCAT.SCA.ratio, color=ASCAT.SCA.ratio < 0.2)) + geom_point() + 
         facet_wrap(~Status, ncol=1) + labs(title='Purity~SCA.Ratio') + geom_vline(xintercept=y, color='grey')

ggplot(qcdata, aes(ASCAT.SCA.ratio, Purity, color=Status)) + geom_point() + 
  facet_wrap(~Status) + labs(x='SCA', title='SCA vs Purity') + 
  geom_vline(xintercept = x, color='grey') + geom_hline(yintercept = y, color='grey')

table(unique(subset(qcdata, Purity > y & ASCAT.SCA.ratio < x & SampleType == 'BE', select=c('Status','PatientID')))$Status)

ggplot(qcdata, aes(Purity)) + geom_histogram() + geom_vline(xintercept = y, col='red') + 
  facet_wrap(~SampleType, scales = 'free') + labs(title='Purity by sampletype')

ggplot(qcdata, aes(Purity, Goodnessoffit, color=Status)) + geom_point() + 
         facet_wrap(~SampleType, ncol=1) + labs(title='Purity~fit, normal vs BE') + geom_vline(xintercept=y, color='grey')

ggplot(subset(qcdata, SampleType == 'BE'), aes(round(Ploidy,2), Purity, color=Status)) + geom_point() + facet_grid(~Status) + 
  labs(title='BE only, purity vs ploidy') + geom_hline(yintercept=y, color='grey')

ggplot(qcdata, aes(ASCAT.SCA.ratio, color=Status, fill=Status)) + facet_wrap(~Status, scales='free_y',nrow=1) +  geom_density(alpha=0.5) + labs(x='SCA Ratio', title='SCA Ratio, NP vs P')

ggplot(qcdata, aes(ASCAT.SCA.ratio, color=Status, fill=Status)) + facet_wrap(Status~SampleType, scales='free_y',nrow=2) +  geom_density(alpha=0.5) + labs(title='SCA Ratio, NP vs P')

summary(subset(qcdata, SampleType == 'BE')$ASCAT.SCA.ratio)
summary(subset(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio)

ggplot(subset(qcdata, SampleType == 'BE'), aes(Purity, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) +
  geom_point() + geom_vline(xintercept = y, color='darkgreen') + geom_hline(yintercept = x, color='red') + 
  labs(title='Barretts only')

ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
  geom_hline(yintercept = 0.02, color='red') + labs(title='Barretts only')

## PER ENDOSCOPY
qcdata.endo = qcdata %>% group_by(PatientID, EndoID, Status, Timepoint, SampleType) %>% dplyr::summarise_if(is.numeric, c('mean','max','min','sd')) %>% mutate_if(is.numeric, round, digits=3)
qcdata %>% dplyr::group_by( Status, SampleType) %>% dplyr::summarise( median(ASCAT.SCA.ratio), sd(ASCAT.SCA.ratio) )

#ggplot(subset(qcdata.endo, SampleType == 'BE'), aes(ASCAT.SCA.ratio_max, Purity_max)) + geom_point() + facet_grid(~Status) + geom_hline(yintercept = 0.95)  + geom_vline(xintercept = 0.001)

lowsca = subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.01 & Purity_max > 0.98 )

table(unique(lowsca[,c('PatientID','Status')])$Status)/table(unique(qcdata.endo[,c('PatientID','Status')])$Status)

table(unique(subset(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_max <= 0.001, select=c('PatientID','Status')))$Status)

ggplot(subset(qcdata.endo,SampleType == 'BE'), aes(Status, ASCAT.SCA.ratio_max, group=Status)) + geom_boxplot() + geom_hline(yintercept = 0.002, color='red')

qcdata %>% dplyr::group_by( Status, SampleType) %>% dplyr::summarise( median(ASCAT.SCA.ratio))

```

```{r, echo=F, warning=F, fig.height=10, fig.width=8,eval=F}

summary.file = '~/Data/Annalise/NonP/summary_logR.txt'
smy = read.table(summary.file, sep = '\t', header = T)
smy = subset(smy, Copy.number <= 12)

a = as.data.frame(smy[,c('Copy.number','var.LRR')])
colnames(a) = c('x','y')
b = as.data.frame(smy[,c('Copy.number','median.LRR')])
colnames(b) = c('x','y')
b[b$x == 2,'y'] = 0

fitSD = lm(x~y, data=a)
autoplot(fitSD)
fitM = lm(x~y, data=b)
autoplot(fitM)

fitCN = lm(y~x, data=b)

# Adjust based on the summary values per CN that we get in WGS (mostly) NP Barrett's cases
adjust.segraw<-function(segraw) {
  segraw = segraw %>% rowwise() %>% dplyr::mutate( total = nAraw+nBraw )
  segraw = subset(segraw, chr %in% c(1:22))

  segraw$adjLRR = NA
  #for (cn in unique(segraw$total)) {
  #  rows = which(segraw$total == cn)
  #  values = seg[rows, 'medLRR', drop=T]

    #newM = mean(values)
    #newSD = sd(values)

    #if (cn != 2) {
      #newM = predict(fitM, newdata=cbind.data.frame('x'=cn))
      #newSD = predict(fitSD, newdata=cbind.data.frame('x'=cn))

    #newV = mean(values)

    #}
    #segraw[rows,][['adjLRR']] = newM + (values - mean(values)) * (newSD/sd(values))
   # smy
    #median(predict(fitM, newdata=cbind.data.frame('y'=values)))-cn

    new = predict(fitCN, newdata=cbind.data.frame('x'=segraw$total))
    #plot(scale(new, center = T))

    segraw[,'adjLRR'] = scale(new,center=T)

   # segraw[rows,'adjLRR'] = (values * (predict(fitM, newdata=cbind.data.frame('y'=values)) - cn))
  #}  
segraw  
}

pt = '403'
pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
load(grep('ascat.Rdata', pfiles, value=T), verbose=F)

seg = ascat.output$segments_raw
seg = adjust.segraw(seg)

m = (melt(seg, measure.vars=c('medLRR','adjLRR')))
m$total = round(round(m$total, 1))
ggplot(m, aes(total, value, group=total)) + facet_grid(~variable) + geom_boxplot() + labs(title=pt)

seg$chr = factor(seg$chr, levels=c(1:22))
grid.arrange(
  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Low SCA, medLRR', subtitle=paste('patient',pt)),
  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
    geom_segment(aes(x=startpos, xend=endpos, y=adjLRR, yend=adjLRR), color='darkred', size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Low SCA, adjLRR', subtitle=paste('patient',pt))
)

pt = '512'
pfiles = list.files(paste('~/Data/Reid_SNP/PerPatient', pt, sep='/'), full.names = T)
load(grep('ascat.Rdata', pfiles, value=T), verbose=F)

seg = ascat.output$segments_raw
seg = adjust.segraw(seg)

m = (melt(seg, measure.vars=c('medLRR','adjLRR')))
m$total = round(round(m$total, 1))
ggplot(m, aes(total, value, group=total)) + facet_grid(~variable) + geom_boxplot() + labs(title=pt)

seg$chr = factor(seg$chr, levels=c(1:22))
grid.arrange(
  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), color='darkgreen', size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA, medLRR', subtitle=paste('patient',pt)),
  ggplot(seg) + facet_grid(sample~chr, space='free_x', scales='free_x') + 
    geom_segment(aes(x=startpos, xend=endpos, y=adjLRR, yend=adjLRR), color='darkred', size=3) + 
    theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA, adjLRR', subtitle=paste('patient',pt))
)

```

# Load tiled data

CN Adjustment uses Annalise's whole-genome non-progressor samples to generate a set of mean and SD values per copy-number state. The median LogR values are adjusted using the mean and SD based on the CN state.  The adjusted LogR values are then winsorize per sample, and tiled in 5Mb bins.

```{r, warning=F, echo=F, message=F, fig.height=10, fig.width=8}

if (file.exists('~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata')) {
  load('~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata', verbose=T) 
} else {

  mergedSegs = NULL
  mergedArms = NULL
  length(ptdirs)

  for (pt in ptdirs) {
    print(pt)
    if (length(list.files(pt, '*wins_tiled.txt', full.names=T)) <= 0) {
      message(paste("No tiled files for",pt))
      next
    }

    segvals = as.data.frame(data.table::fread(list.files(pt, '*wins_tiled.txt', full.names=T)))
    armvals = as.data.frame(data.table::fread(list.files(pt, '*arms_tiled.txt', full.names=T)))

    segvals = segment.matrix(segvals)
    segvals[is.na(segvals)] = mean(segvals,na.rm=T)

    armvals = segment.matrix(armvals)
    armvals[is.na(armvals)] = mean(armvals,na.rm=T)

    if (is.null(segvals) | is.null(armvals))
      stop(paste("Missing values in ", pt))

    if (is.null(mergedSegs)) {
      mergedSegs = segvals
      mergedArms = armvals
    } else {
      mergedSegs = rbind(mergedSegs, segvals)    
      mergedArms = rbind(mergedArms, armvals)    
    }
  }
  nrow(mergedSegs) == nrow(mergedArms)
  #setdiff(rownames(mergedSegs), rownames(mergedArms))
  dim(mergedSegs)

  save(mergedSegs, mergedArms, file='tmp_seg_pt.Rdata')
}
rownames(mergedSegs) = sub('\\.LogR','', rownames(mergedSegs))
rownames(mergedArms) = sub('\\.LogR','', rownames(mergedArms))

dim(mergedSegs)
dim(mergedArms)

# Adjust for ploidy
for (i in 1:nrow(mergedSegs)) {
  sample = rownames(mergedSegs)[i]
  x = unlist(strsplit(sample, '_'))
  #ploidy = mean(subset(qcdata, SampleType == 'BE' & PatientID == x[1] & EndoID == x[2])$Ploidy)
  ploidy = subset(qcdata.endo, PatientID == x[1] & EndoID == x[2] & SampleType == 'BE')$Ploidy_mean
  if (grepl('BLD|gastric', sample, ignore.case = T)) ploidy = 2
  mergedSegs[i,] = mergedSegs[i,]-(ploidy-1)
  mergedArms[i,] = mergedArms[i,]-(ploidy-1)
}

segmentVariance = apply(mergedSegs, 1, var)

nm.rows = grep('BLD|GASTRIC', rownames(mergedSegs))
#normalSegs = mergedSegs[nm.rows,]
#normalArms = mergedArms[nm.rows,]

# nrow(normalSegs) == nrow(normalArms)
# 
# mergedSegs = mergedSegs[-nm.rows,]
# mergedArms = mergedArms[-nm.rows,]
# 
# nrow(mergedSegs) == nrow(mergedArms)

lowSeg = mergedSegs[grep('^403', rownames(mergedSegs), value=T),,drop=F]
p1 = ggplot(melt(lowSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 403')

highSeg = mergedSegs[grep('^512', rownames(mergedSegs), value=T),,drop=F]
p2 = ggplot(melt(highSeg), aes(Var2, value)) + facet_grid(~Var1, space='free_x', scales='free') + geom_point() +
  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512')

grid.arrange(p1,p2,ncol=1)
```

# Values in the binned segments

Blood/gastric normals have very little variance compared to all of the BE samples (no idea which is P vs NP)

```{r normals, message=F, fig.height=10, fig.width=8}
#nm = melt(normalSegs)
nm = melt(mergedSegs[nm.rows,])
be = melt(mergedSegs[-nm.rows,])

lims = range(nm$value, be$value, na.rm=T)

grid.arrange(
  ggplot(nm, aes(Var2, value, group=Var2)) + geom_point() + labs(title='Blood/gastric',subtitle=paste('variance=',var(nm$value),sep=''),x='') + 
    theme(axis.text.x = element_blank()) + ylim(lims),
  ggplot(be, aes(Var2, value, group=Var2)) + geom_point() + labs(title='BE samples',subtitle=paste('variance=',var(be$value,na.rm=T),sep=''),x='') + 
    theme(axis.text.x = element_blank()) + ylim(lims)
)

```

## Compare the binned values to the median LRR from ASCAT

Using a random normal sample and a known BE progresser sample

```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8, eval=T}
randNormal = sample( grep('BLD',rownames(mergedSegs)[nm.rows], value=T), 1)
pt = unlist(strsplit(randNormal, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*', randNormal),sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

lims = range(segs$medLRR, mergedSegs[randNormal,])
# grid.arrange(
#   ggplot(segs) + facet_grid(~chr, space='free_x', scales='free') + ylim(lims) +
#     geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
#     theme(axis.text.x=element_blank() ) + labs(title='ASCAT median LRR', subtitle=pt),
#   ggplot(melt(as.matrix(normalSegs[randNormal,])), aes(Var1, value)) + geom_point() + ylim(lims) +
#     labs(title='CN Adjusted & winsorized') + theme(axis.text.x=element_blank()),
# top='Random BLD or gastric normal', bottom=randNormal)

#randBE = '512_133R'
randBE = '1006_190V'
pt = unlist(strsplit(randBE, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*',randBE), sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

#lims = range(segs$medLRR, mergedSegs[sub('\\.LogR','',randBE),])
# grid.arrange(
#   ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
#     geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) + 
#     theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt), 
#   ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() +
#     theme(axis.text.x=element_blank()) + labs(title='Winsorized LRR, binned'),  
#  ggplot(melt(adjust.cols(mergedSegs[sub('\\.LogR','',randBE),,drop=F], z.mean, z.sd)), aes(Var2, value)) + geom_point() + theme(axis.text.x=element_blank()) + labs(title='Col adjusted LRR, binned'),  
#top='BE', bottom=randBE)

```

# Set up values to predict

## Adjust per column

Adjust using the RAW values...

```{r raw.adjust, fig.height=16, fig.width=4, echo=T}
q.norm.by<-function(df, target) {
  dfq = do.call(cbind.data.frame,lapply(1:ncol(df), function(i) {
    as.data.frame(normalize.quantiles.use.target(df[,i,drop=F], as.vector(target[,i])))
  }))

  colnames(dfq) = colnames(df)
  rownames(dfq) = rownames(df)

  return(as.matrix(dfq))
}

copySegs = q.norm.by(mergedSegs, raw.segs)
copyArms = q.norm.by(mergedArms, raw.arms)

grid.arrange(
  ggplot( melt(raw.segs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, sWGS bins (no arms)'),
  ggplot( melt(mergedSegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins (no arms)'),
  ggplot( melt(copySegs), aes(sample=value)) + stat_qq() + labs(title='Normal Q-Q plot, SNP bins q-normed')
)

```

Adjust the variance of each genomic region by the mean and sd extracted when I trained on the sWGS data. This seems to be a more fair comparison than scaling all the new data to itself, but the predictions don't really change.

Using the mean bin values (merged during tiling).
```{r, warning=F, message=F, fig.width=10, fig.height=8}

#ns = adjust.cols(normalSegs, z.mean, z.sd)
#ns = adjust.cols(normalSegs)
#range(ns)

#na = adjust.cols(normalArms, z.arms.mean, z.arms.sd)
#na = adjust.cols(normalArms)
#range(na)

#ms = adjust.cols(mergedSegs, z.mean, z.sd)
#ms = adjust.cols(mergedSegs)
ms = adjust.cols(copySegs)
range(ms)
dim(ms)

# mergedSegs['997_122V',which(mergedSegs['997_122V',] <= 0)]
# ms['997_122V',which(mergedSegs['997_122V',] <= 0)]

#ma = adjust.cols(mergedArms, z.arms.mean, z.arms.sd)
#ma = adjust.cols(mergedArms)
ma = adjust.cols(copyArms)
range(ma)
dim(ma)

nm = melt(ms[nm.rows,])
be = melt(ms[-nm.rows,])

lims = range(nm$value, be$value, na.rm = T)
grid.arrange(
  ggplot(nm, aes(Var2, value, group=Var2, color=Var1)) + geom_point() + labs(title='Blood/gastric') + 
    theme(axis.text.x=element_blank(),legend.position='none') +ylim(lims),
  ggplot(be, aes(Var2, value, group=Var2, color=Var1)) + geom_point() + labs(title='BE samples') + 
    theme(axis.text.x=element_blank(),legend.position='none') + ylim(lims), top='Unit scaled')

```

```{r echo=F,message=F, fig.width=10, fig.height=8}
#cxN = score.cx(ns, 1)
#arrayDfNorm = subtract.arms(ns, na)
#arrayDfNorm = cbind(ns, na)
#arrayDfNorm = cbind(arrayDfNorm, 'cx'=unit.var(cxN, mn.cx, sd.cx))

#probN = predict(fitV, newx=arrayDfNorm, s=l, type='response')
#length(which(probN > 0.5))/nrow(probN) # FPR
#length(which(probN <= 0.5))/nrow(probN)

#preds.normal = cbind.data.frame(rownames(probN),probN, predict(fitV, newx=arrayDfNorm, s=l, type='link'))
#colnames(preds.normal) = c('Samplename','Prob', 'RR')

cx = score.cx(ms, 1)
#arrayDf = cbind(ms,ma)
arrayDf = subtract.arms(ms, ma)
#arrayDf = cbind(arrayDf, 'cx'=unit.var(cx, mn.cx, sd.cx))
arrayDf = cbind(arrayDf, 'cx'=unit.var(cx))

lims=range(arrayDf, dysplasia.df,na.rm=T)
grid.arrange(
  ggplot(melt(arrayDf), aes(Var2, value)) + geom_point() + ylim(lims) +
    geom_point(data=melt(arrayDf[,589:634]), col='red') + labs(title='SNP'),
  ggplot(melt(dysplasia.df), aes(Var2, value)) + geom_point() + ylim(lims) +
      geom_point(data=melt(dysplasia.df[,589:634]), col='red') + labs(title='sWGS'),
nrow=1)

```

# Predictions

```{r, warning=F, fig.height=8}

prob = predict(fitV, newx=arrayDf, s=l, type='response')
rr = predict(fitV, newx=arrayDf, s=l, type='link')

preds = cbind.data.frame(rownames(prob),prob, rr, stringsAsFactors=F)
colnames(preds) = c('Samplename','Prob', 'RR')

# offset = log(0.0225)
# preds = preds %>% dplyr::mutate(
#   'Adj. RR'=RR+offset,
#   'Adj. Prob'=1/(1+exp(-RR+abs(offset)) )
# )

pnr = grep('BLD|gastric', preds$Samplename, ignore.case = T)
preds.normal = preds[pnr,]

length(which(preds.normal$Prob > 0.5))/nrow(preds.normal) # FPR
length(which(preds.normal$Prob < 0.5))/nrow(preds.normal)

myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
grid.arrange(
  ggplot(preds, aes(RR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
    labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk'),
  # ggplot(preds, aes(`Adj. RR`)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  #   scale_fill_gradientn(colors = myPal,  name='') +
  #   labs(y='n Samples', x='Relative Risk', title='Adjusted relative risk'),
top='Relative risk')

grid.arrange(
  ggplot(preds, aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
    labs(title='Predictions, all samples model', y='n Samples', x='Unadjusted Probability'),
  # ggplot(preds, aes(`Adj. Prob`)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  #   scale_fill_gradientn(colors = myPal,  name='') +
  #   labs(title='Predictions, all samples model', y='n Samples', x='Adjusted Probability'),
top='Probability')

#randNormal = paste(unlist(strsplit(randNormal, '_'))[c(1,3)], collapse='_')
#randBE = paste(unlist(strsplit(randBE, '_'))[c(1,3)], collapse='_')

```

A quick look at the two random samples from earlier.  
`r pander(subset(preds.normal, Samplename == randNormal), caption='Random normal')`

The random BE should show a very high probability, but for some reason is missed entirely:
`r pander(subset(preds, Samplename == randBE), caption='Random BE')`

In the gastric/blood normals `r signif(length(which(preds.normal$Prob < 0.5))/nrow(preds.normal),2)` are predicted to have <50% probability (TPR).

For blood normals only `r signif(nrow(subset(preds.normal, grepl('BLD',Samplename) & Prob < 0.5))/nrow(subset(preds.normal, grepl('BLD',Samplename) )),2)` have a <50% probabilitye (TPR).

## Evaluating predictions

```{r, warning=F, fig.height=8}
#qcdata.be = subset(qcdata, SampleType == 'BE')
# get.qcinfo<-function(vec) {
#   endo = vec[2]
#   pt = vec[1]
#   y = subset(qcdata.be, PatientID == pt & EndoID == endo)
#   y = y %>% group_by(Status, Timepoint) %>%  dplyr::summarise( mean(Ploidy), mean(Purity), mean(Goodnessoffit), mean(SCA.Ratio) )
#   cbind.data.frame('PatientID'=pt, 'EndoID'=endo, y, stringsAsFactors = F)
# }
preds = preds %>% dplyr::group_by(Samplename) %>% dplyr::mutate(
  PatientID = unlist(strsplit(Samplename,'_'))[1],
  EndoID = unlist(strsplit(Samplename,'_'))[2]
)

pred.qc = merge(preds,qcdata.endo, by=c('PatientID','EndoID'))

#dim(pred.qc)
#table(pred.qc$Status)

grid.arrange(
  ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Prob, color=Status)) + geom_point(),
  ggplot(pred.qc, aes(Ploidy_max, Prob, color=Status)) + geom_point(),
ncol=2)

pander(cor.test(pred.qc$ASCAT.SCA.ratio_mean, pred.qc$Prob))
```

```{r, warning=F, fig.height=8}
prog = subset(pred.qc, Status == 'P')
p1 = ggplot(prog, aes(Prob)) + facet_wrap(~Timepoint, scales='free_y',nrow=1) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability')

nonprog = subset(pred.qc, Status == 'NP')
p2 = ggplot(nonprog, aes(Prob)) + facet_wrap(~Timepoint, scales='free_y', nrow=1) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on non-progressors', y='n Samples', x='Unadjusted Probability')

preds.normal = preds.normal %>% dplyr::mutate(Type = ifelse(grepl('BLD',Samplename), 'Blood','Gastric'))

p3 = ggplot(preds.normal, aes(Prob)) + facet_wrap(~Type, ncol=2, scales = 'free') +
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') +
  labs(title='Predictions on BLD/Gastric', y='n Samples', x='Unadjusted Probability')

grid.arrange(p1,p2,p3,ncol=1)

```

# Compare predictions to known status

Using my thresholds

```{r roc, echo=T}
pred.qc = base::merge( pred.qc, patient.info[,c('PatientID', 'Timepoint Code','Pathology')], by.x=c('PatientID', 'EndoID'), by.y=c('PatientID','Timepoint Code') )

# Three groups, prog, non prog (max LGD), non prog (max HGD)
# pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate( 
#   exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0 | `mean(Ploidy)` > 2.7 )) )
pred.qc$Status = factor(pred.qc$Status)

pred.qc = pred.qc %>% group_by(Status, PatientID) %>% dplyr::mutate(
  exclude=(Status == 'NP' & (length(which(Pathology == 'HGD')) > 0)),
  wgd = (Status == 'NP' & Ploidy_max > 2.7 )
)

# pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
#   lowsca = ( Status == 'P' & (`mean(SCA.Ratio)` < 0.03 & `mean(Purity)` >= 0.95))
# )

pred.qc = pred.qc %>% group_by(Status, Samplename) %>% dplyr::mutate(
  lowsca = ( ASCAT.SCA.ratio_mean < 0.02 & Purity_mean > 0.95)
)

ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Purity_mean, color=lowsca)) + geom_point() + facet_grid(~Status) + 
  geom_vline(xintercept=0.02) + geom_hline(yintercept = 0.95)

pander(
  rbind(rbind('n.Samples'=table(subset(pred.qc, !lowsca )$Status)),
  rbind('n.Patients'=table(unique(subset(pred.qc, !lowsca, select=c('PatientID','Status')))$Status))),
caption='Number of samples and patients that pass the low SCA:high purity thresholds')

pander(
  rbind(rbind('n.Samples'=table(subset(pred.qc, !lowsca & !exclude & !wgd)$Status)),
  rbind('n.Patients'=table(unique(subset(pred.qc, !lowsca  & !exclude & !wgd, select=c('PatientID','Status')))$Status))),
caption='Number of samples and patients that pass the low SCA:high purity & NP thresholds')

```

```{r echo=F}
pred.qc.ex = subset(pred.qc, SampleType == 'BE' & !exclude & !wgd & !lowsca)
cuts = seq(0,1,0.1)

pred.qc.ex$quants = with(pred.qc.ex, cut(Prob, breaks=cuts))
qt = as.data.frame.matrix(table(pred.qc.ex$quants, pred.qc.ex$Status))
qt$quant = rownames(qt)
qt$P = round(qt$P/sum(qt$P),2)
qt$NP = round(qt$NP/sum(qt$NP), 2)
pander(qt, caption='Excluding HGD NP, WGD NP, low SCA P')

# These are the cutoffs from the sWGS data
get.risk<-function(prob) {
  if (prob < 0.3) return('Low')
  if (prob >= 0.3 & prob < 0.7) return('Moderate')
  if (prob >= 0.7) return('High')
}

# All data
pred.qc = pred.qc %>% dplyr::rowwise() %>% dplyr::mutate(
  Risk = get.risk(Prob)
)
pred.qc$Risk = factor(pred.qc$Risk, levels=c('Low','Moderate','High'), ordered = T)

norm.pred = subset(pred.qc, SampleType != 'BE')

pred.qc = subset(pred.qc, SampleType == 'BE')

roc = pROC::roc(Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='All, Prob')
pander(pROC::coords(roc, 'best'), caption='All samples, ROC best')

# Exclude HGD NP patients
roc = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude), auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='Excluded HGD NP patients')

# Exclude the WGD NP
rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !wgd), auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Both endos - excluded WGD NP')
pROC::coords(rocpp, 'best')

roc = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude & !wgd), auc=T, ci=T, of='thresholds')
roc.plot(roc) + labs(title='Excluded HGD & WGD NP patients')

# Exclude the low SCA patients
rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !lowsca), auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Both endos - excluded low SCA')
pROC::coords(rocpp, 'best')

# Exclude all of the above
rocpp = pROC::roc(Status ~ Prob, data=subset(pred.qc, !exclude & !lowsca & !wgd), auc=T, ci=T, of='thresholds')
roc.plot(rocpp) + labs(title='Both endos - excluded low SCA P & HGD NP & WGD NP')
pROC::coords(rocpp, 'best')

# Timepoint 2 only
roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !exclude & !lowsca & !wgd), auc=T, ci=T, of='thresholds')
roc.plot(roct2) + labs(title='T2, Prob')

ggplot(subset(pred.qc, !exclude & !wgd & !lowsca), aes(ASCAT.SCA.ratio_mean,Prob, color=Risk, shape=Status)) + 
  facet_grid(~Status) + geom_point()

ex.pred = subset(pred.qc, !exclude & !wgd & !lowsca)

write.table(subset(pred.qc, lowsca, select=c('PatientID','EndoID','Status','ASCAT.SCA.ratio_mean','SCA.Ratio_mean','Purity_mean')),
            sep='\t', quote=F, row.names=F, file='lowsca_pts.txt')

head(qcdata)

x = subset(qcdata, ASCAT.SCA.ratio < 0.02 & Purity > 0.95)
table(x$SampleType)/table(qcdata$SampleType)

write.table(subset(x, SampleType == 'BE', select=c('PatientID','EndoID','Samplename','ASCAT.SCA.ratio','SCA.Ratio','Purity')), quote=F, sep='\t', row.names = F, file='lowsca_samples.txt')

#x = subset(qcdata, ASCAT.SCA.ratio < 0.02 & Purity > 0.95 & SampleType == 'BE')

# Quick check
# nrow(subset(ex.pred, Prob > 0.5 & Status == 'P'))/table(ex.pred$Status)['P']
# nrow(subset(ex.pred, Prob < 0.5 & Status == 'P'))/table(ex.pred$Status)['P'] #False negatives
# 
# nrow(subset(ex.pred, Prob > 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] # False pos
# nrow(subset(ex.pred, Prob < 0.5 & Status == 'NP'))/table(ex.pred$Status)['NP'] #

ct = cor.test(pred.qc.ex$Prob, pred.qc.ex$ASCAT.SCA.ratio_mean)
ct2 = cor.test(pred.qc$Prob, pred.qc$ASCAT.SCA.ratio_mean)
```

The risk probability is correlated with the % of the genome that's been altered as identified by Xiahong, R=`r ct$estimate` and that's true even when no patients are excluded (R=`r ct2$estimate`).

# Regression using SNP Data

```{r regression, echo=F, message=F, eval=T}
folds = 10; splits = 5
## ----- All ----- ##
cache.dir = '~/Data/Ellie/Analysis/SNP'
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
 message(paste("loading", file))
 load(file, verbose=T)
 model.performance(performance.at.1se, coefs)$plot + labs(title='SNP CV models')
} else {
  warning("Run 'snp-regression.R' first")
}
#save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp-models.Rdata',sep='/'))

#do.call(grid.arrange, c(plots))
```

```{r compare.coefs, echo=F, eval=F}
# Matching
snp.features = coefs$`0.5`

intersect(rownames(snp.features), rownames(features))

cxisft = grep('cx',rownames(snp.features))

length(which(rowSums(snp.features[,-1])/50 > 0.5))

snp.features = do.call(rbind.data.frame, c(strsplit(rownames(snp.features)[-cxisft], ':|-'), stringsAsFactors=F))
colnames(snp.features) = c('chr','start','end')

swgs.features = do.call(rbind.data.frame, c(strsplit(rownames(features), ':|-'), stringsAsFactors=F))
colnames(swgs.features) = c('chr','start','end')

snp.features = makeGRangesFromDataFrame(snp.features)
snp.features = snp.features[-which(width(snp.features) > 5e6)] # No arms

swgs.features = makeGRangesFromDataFrame(swgs.features)

dist = distanceToNearest(swgs.features,snp.features)

swgs.features[queryHits(dist)]
snp.features[subjectHits(dist)]

match = dist[elementMetadata(dist)$distance != 0]
elementMetadata(dist)$bins = round(elementMetadata(dist)$distance/5e6)

x1 = dist[elementMetadata(dist)$bins <= 2]

swgs.features[queryHits(x1)]
snp.features[subjectHits(x1)]
```

```{r loo, eval=T}

file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  stop("Run snp-regression.R first")
}

snp.roc = pROC::roc(Status~Prediction,slabels, auc=T, ci=T, of='thresholds')
roc.plot(snp.roc, title="SNP LOO Model ROC")

ggplot(cbind.data.frame('performance'=performance.at.1se, 'type'='snp'), aes(y=performance, x=type, group=type,fill=type)) + 
  geom_boxplot(outlier.colour = NA, show.legend = F) + geom_jitter(width=0.2) + geom_label(data=cbind.data.frame('performance'=range(performance.at.1se), 'type'='snp'), aes(label=round(performance,2)), nudge_x=0.1) + theme_bw() + theme(legend.position = 'none') + labs(x='', y='Classification', title='LOO Performance SNP')

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)

fit.snp = models$`1`

probs = predict(fit.snp, newx = dysplasia.df, s = performance.at.1se$`1`$lambda,type='response' )

swgs.pred.snp = cbind.data.frame('Prob'=probs[,1], 'Status'=swgs_labels[rownames(probs)])
ssroc = pROC::roc(Status~Prob, data=swgs.pred.snp, ci=T, auc=T, of='thresholds')
roc.plot(ssroc, title='sWGS prediction on SNP model')
```