### 2018-10-12 sWGS Normals.Rmd

---
title: "sWGS Normals"
author: "Sarah Killcoyne"
date: "7/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Matrix)
library(glmnet)
library(ggplot2)
library(gridExtra)
library(pander)
source('lib/load_patient_metadata.R')
source('lib/data_func.R')
source('lib/fastPCF.R')

data.files = list.files('~/Data/Ellie/QDNAseq', full.names = T)

rpi = read.patient.info('~/Data/Ellie/QDNAseq/All_patient_info.xlsx')
normals = rpi$removed
path = read_excel('~/Data/Ellie/QDNAseq/All_patient_info.xlsx', 'Normals')[,c('Patient ID', 'Hospital Research ID', 'Pathology notes','Plate Index','SLX ID 1')]
colnames(path)[c(1,4,5)] = c('Patient','Plate.Index','SLX.ID')
path$SLX.ID = sub('SLX-','', gsub(' ', '', path$SLX.ID))
normals = merge(normals, path, by=c('Patient','Plate.Index','SLX.ID'))

data.file = grep('MultisampleCopynumberBinnedAndFitted2.RData', data.files, value=T)
if (length(data.file) <= 0)
  stop("Missing necessary Rdata file from Multisample script")
load(file=data.file, verbose=T)

loc.cols = c('location','chrom','start','end')
fit.data = fit.data[,c(loc.cols,normals$Samplename)]
raw.data = raw.data[,c(loc.cols,normals$Samplename)]

exclude.file = grep('qDNAseq_blacklistedRegions.txt', data.files, value=T)
if (length(data.file) <= 0)
  stop(paste("Missing necessary exclusion file from 'qDNAseq_blacklistedRegions.txt' in",data))
blacklisted.regions = read.table(exclude.file,sep="\t",header=T,stringsAsFactors=F)

if (file.exists('/tmp/normals.Rdata')) {
  load('/tmp/normals.Rdata', verbose=T)
} else {
  plist = list()
  seg.residuals = list()
  tiled.normals = NULL; tiled.normal.arms = NULL
  for (sample in normals$Samplename) {
    binned = binSWGS(raw.data[,c(loc.cols,sample)],fit.data[,c(loc.cols,sample)],blacklisted.regions,min.probes=67,gamma2=250,metrics.file=NULL)
    plist[[sample]] = binned$plots[[1]]
    seg.residuals[[sample]] = binned$resvar[[sample]]

    tiled = tile.segmented.data(binned$segvals,chr.info=get.chr.lengths(file='/tmp/hg19_info.txt'),verbose=T)
    tiled.arms = tile.segmented.data(binned$segvals,chr.info=get.chr.lengths(file='/tmp/hg19_info.txt'),size='arms', verbose=T)
    if (is.null(tiled.normals) | is.null(tiled.normal.arms)) {
      tiled.normals = tiled
      tiled.normal.arms = tiled.arms
    } else {
      tiled.normals = cbind(tiled.normals, tiled[,sample,drop=F])
      tiled.normal.arms = cbind(tiled.normal.arms, tiled.arms[,sample,drop=F])
    }
  }
  save(tiled.normals, tiled.normal.arms, plist, seg.residuals, file='/tmp/normals.Rdata')
}

resids = sample.residual.variance(seg.residuals)
resids = merge(normals[,c('Patient','Hospital.Research.ID','Samplename')], resids, by.x='Samplename', by.y='sample')

failedQC = arrange(subset(resids, round(var1sd_median,3) > 0.012, select=c(Hospital.Research.ID,Samplename, var1sd_median,var1sd_sd, var1sd_Q1,var1sd_Q3)),  var1sd_median)

pander(normals[,c('Patient','Hospital.Research.ID','Status','Endoscopy.Year','Pathology notes')], caption='Normal samples', justify='left')
```

`r nrow(failedQC)` samples failed the variant residual QC based on a cutoff determined by the discovery cohort.

```{r, echo=F, warning=F, fig.height=30, fig.width=8}
do.call(grid.arrange, c(plist, ncol=1))
```

## Predictions

Using the model trained on the progressors and non-progressors in the initial cohort the normal samples can be predicted as well.

```{r, echo=F}
modeldir = '~/Data/Ellie/Analysis/5e6_arms_all'
file = paste(modeldir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
file = paste(modeldir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
rm(dysplasia.df,coefs,plots,labels)

select.alpha = '0.9'
fitV = models[[select.alpha]]
lambda.opt = performance.at.1se[[select.alpha]][, 'lambda']

tiled.segs = segment.matrix(tiled.normals)
tiled.arms = segment.matrix(tiled.normal.arms)

cx.score = score.cx(tiled.segs,1)

for (i in 1:ncol(tiled.segs)) 
  tiled.segs[,i] = unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
for (i in 1:ncol(tiled.arms)) 
  tiled.arms[,i] = unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])

mergedDf = subtract.arms(tiled.segs, tiled.arms)
mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))

sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
                            dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]

preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')

high = preds[preds > 0.7,,drop=F]

pander(subset(normals, Samplename %in% rownames(high), select=c('Patient','Hospital.Research.ID','Status','Endoscopy.Year','Pathology notes')), caption='Normal samples predicted to be high risk')

```