### 2019-08-16 Downsampled Predictions.Rmd
...

---
title: "Downsampled"
author: "Sarah Killcoyne"
date: "14/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)

library(tidyverse)
library(BarrettsProgressionRisk)
library(kableExtra)
source('lib/common_plots.R')

dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions'
info = '~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx'

info = readxl::read_xlsx(info)

files = list.files(dir, 'Rdata', recursive = T, full.names = T)

pfile = paste0(dir, '/ds_preds.txt')
if (file.exists(pfile)) {
  ds.preds = read_tsv(pfile)
} else {
  ds.preds = do.call(bind_rows, lapply(files, function(f) {
    load(f, verbose=T)
    bind_cols(predictions(prr), adjustRisk(prr, 'mean', 'sample')$adj.predictions %>% dplyr::select(Probability, `Relative Risk`) %>% dplyr::rename_all(funs(paste0), ' Adjusted ') )
  }))
  write_tsv(ds.preds, pfile)
}

ds.preds = ds.preds %>% dplyr::select(-Status) %>% left_join(info %>% dplyr::select(`Illumina ID`, Status), by=c('Sample' = 'Illumina ID')) %>% dplyr::mutate(Type = factor(Type, levels=c('Normal','NDBE', 'LGD', 'Tumor Adjacent BE'))) %>% arrange(Probability)

```

## Downsampled Datasets

These samples are from Annalise's cohort and the OCCAMS tumor-adjacent Barrett's cohort.  Additionally, I downsampled 40 normal blood WGS as a non-progressor test set. A total of `r nrow(info)` samples were downsampled from 30-50X WGS to an average 0.4X then run through the QDNAseq pipeline.

```{r}
info %>% dplyr::group_by(Status, Type) %>% summarise(n = length(Type)) %>% spread(Type,n) %>% kable() %>% kable_styling(full_width = F)
```

They were then predicted using the model trained on the initial training cohort.

```{r echo=F, eval=T}
myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))

ggplot(ds.preds, aes(`Relative Risk`)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1), show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
    plot.theme + labs(title='Sample predictions', y='n Samples') + theme(legend.position = 'none')

# roc = pROC::roc(Status ~ Probability, data=ds.preds, ci=T, of='thresholds', transpose=T)
# pROC::coords(roc, 'best', ret='threshold', transpose=T)
# roc.plot(roc)
```

```{r fig.height=12}
m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk','Relative Risk Adjusted'), id.vars = c('Type', 'Status', 'Risk'))

ggplot(m, aes(variable, value)) + 
  facet_wrap(~Type, ncol=2, scales = 'free_x') + 
  geom_jitter(width=0.2, aes(shape=Status, color=Risk), size=2) + 
  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
  scale_color_manual(values =  BarrettsProgressionRisk::riskColors()) +
  #scale_color_gradientn(colors = myPal,  name='') +
  scale_fill_manual(values=rev(RColorBrewer::brewer.pal(5,'RdYlBu')[c(1,2,4,5)])) +
  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + plot.theme +
  theme(axis.text.x = element_text(angle=45, hjust=1))

ds.preds %>% group_by(Status, Risk) %>% summarise(n=length(Risk)) %>% spread(Risk, n) %>% kableExtra::kable(caption = 'Sample by risk and known status') %>% kableExtra::kable_styling(full_width = F)

```