### 2019-09-10 Raw Data Comparison.Rmd
...

---
title: "Raw Data Comparison"
author: "Sarah Killcoyne"
date: "23/08/2019"
output: 
  html_document: 
    fig_height: 4
    fig_width: 6
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(BarrettsProgressionRisk)
library(reshape2)
library(gridExtra)
library(e1071)

chr.lengths = BarrettsProgressionRisk:::chrInfo()
blacklist = readr::read_tsv(system.file("extdata", "qDNAseq_blacklistedRegions.txt", package="BarrettsProgressionRisk"), col_names=T, col_types='cii')

load('~/Data/BarrettsProgressionRisk/QDNAseq/training/merged_qdnaseq_output.Rdata',verbose=T)
tfit.data = fit.data %>% as_tibble %>% mutate(chrom = factor(chrom, levels = chr.lengths$chr, ordered=T)) %>% arrange(chrom, start)
traw.data = raw.data %>% as_tibble %>% mutate(chrom = factor(chrom, levels = chr.lengths$chr, ordered=T)) %>% arrange(chrom, start)
rm(fit.data, raw.data)

val.file = '~/Data/BarrettsProgressionRisk/QDNAseq/validation/sWGS_validation_batches.xlsx'
sheets = readxl::excel_sheets(val.file)[8:13]
all.val = do.call(bind_rows, lapply(sheets, function(s) {
  readxl::read_xlsx(val.file, s) %>% dplyr::select(`Hospital Research ID`, matches('Status'), `Block ID`,`Sample Type`, `SLX-ID`, `Index Sequence`, Cohort, Batch, RA, matches('Collection')) %>% dplyr::filter(!is.na(`SLX-ID`)) %>% mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character))# %>% fncols('Collection', 'Biopsy')
})) %>% mutate(Samplename = paste(`SLX-ID`,`Index Sequence`, sep='.'))

load('~/Data/BarrettsProgressionRisk/QDNAseq/validation/merged_raw_fit.Rdata',verbose=T)
vfit.data = merged.fit %>% as_tibble %>% mutate(chrom = factor(chrom, levels = chr.lengths$chr, ordered=T)) %>% 
  arrange(chrom, start) %>% dplyr::select(location,chrom,start,end, all.val$Samplename)
vraw.data = merged.raw %>% as_tibble %>% mutate(chrom = factor(chrom, levels = chr.lengths$chr, ordered=T)) %>% 
  arrange(chrom, start) %>% dplyr::select(location,chrom,start,end, all.val$Samplename)
rm(merged.fit, merged.raw)

dim(vfit.data)
dim(tfit.data)
```

# Raw counts & fitted values

This is the raw data that gets segmented (after some processing).  It looks like there may be a slight shift in the distribution (per patient) of these values but it's not extremely high.

```{r, fig.width=12, fig.height=12}

summarise.per.pt<-function(df) {
  means = df %>% dplyr::select(-matches('location|chrom|start|end')) %>% summarise_all( list(~mean(.,na.rm=T))  ) 
  sds = df %>% dplyr::select(-matches('location|chrom|start|end')) %>% summarise_all( list(~sd(.,na.rm=T))  ) 
  vars = df %>% dplyr::select(-matches('location|chrom|start|end')) %>% summarise_all( list(~var(.,na.rm=T))  ) 
  iqrs = df %>% dplyr::select(-matches('location|chrom|start|end')) %>% summarise_all( list(~IQR(.,na.rm=T))  )

  disp = full_join(
           full_join( as_tibble(t(means), rownames='sample') %>% rename(V1='mean'), 
                      as_tibble(t(sds), rownames='sample') %>% rename(V1='sd'), by='sample'),
           full_join(as_tibble(t(vars), rownames='sample') %>% rename(V1='var'),
                     as_tibble(t(iqrs), rownames='sample') %>% rename(V1='IQR'), by='sample'), 
           by='sample') 
  return(disp)
}

t.fit.disp = summarise.per.pt(tfit.data) %>% mutate(cohort='train')
v.fit.disp = summarise.per.pt(vfit.data) %>% mutate(cohort='val')

wilcox.test(t.fit.disp$mean, v.fit.disp$mean)

fit.disp = bind_rows(t.fit.disp, v.fit.disp)

gridExtra::grid.arrange(
  ggplot(fit.disp, aes(x=cohort, y=mean, group=cohort, fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none'),
  ggplot(fit.disp, aes(x=cohort, y=sd, group=cohort, fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none'),
  ncol=2, top='Raw fitted values')

t.raw.disp = summarise.per.pt(traw.data) %>% mutate(cohort='train')
v.raw.disp = summarise.per.pt(vraw.data) %>% mutate(cohort='val')

wilcox.test(t.raw.disp$mean, v.raw.disp$mean)

raw.disp = bind_rows(t.raw.disp, v.raw.disp)

gridExtra::grid.arrange(
  ggplot(raw.disp, aes(x=cohort, y=mean, group=cohort, fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none'),
  ggplot(raw.disp, aes(x=cohort, y=sd, group=cohort, fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none'), 
  ncol=2, top='Raw count values')

rm(tfit.data,traw.data,vfit.data,vraw.data)
```

# First processing steps
```{r}
if (file.exists('prepped.Rdata')) {
  load('prepped.Rdata')
} else {
  t.prepped = BarrettsProgressionRisk:::.prepRawSWGS(traw.data, tfit.data, blacklist, verbose=T, plot=F)
  dim(t.prepped$data)

  v.prepped = BarrettsProgressionRisk:::.prepRawSWGS(vraw.data, vfit.data, blacklist, verbose=T, plot=F)
  dim(v.prepped$data)

  t.sdevs = t.prepped$sdevs
  v.sdevs = v.prepped$sdevs
  rm(t.prepped,v.prepped)

  save(t.sdevs, v.sdevs, file='prepped.Rdata')  
}

# The sdevs are used to adjust the gamma values for segmentation
wilcox.test(t.sdevs, v.sdevs)

sdevs = bind_rows(tibble('sdev'=t.sdevs, 'cohort'=rep('train',length(t.sdevs))),
          tibble('sdev'=v.sdevs, 'cohort'=rep('val',length(v.sdevs))))

# Validation has one extreme value, but this is per-sample so it should only affect that sample
ggplot(sdevs, aes(x=cohort, y=sdev, group=cohort, fill=cohort)) + ylim(0,0.5) +
  geom_jitter(width=0.1) + geom_violin(alpha=0.5) + 
  theme(legend.position = 'none') + labs(title='sdevs')

ggplot(sdevs, aes(sdev)) + geom_histogram(stat='density',bins=50, color='lightblue') + xlim(0,0.5) + facet_wrap(~cohort)
```

# Segmented Data

General distributions. These are per-patient on the segmented (but not yet binned) data. It looks like there is a shift in the per-patient distributions.

```{r, fig.height=10, fig.width=10}

t.disp = read_tsv(list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient', pattern='disp', full.names = T, recursive = T), col_types = 'cddddddd') %>% mutate(cohort='train')
v.disp = read_tsv(list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf', pattern='disp', full.names = T, recursive = T), col_types = 'cddddddd') %>% mutate(cohort='val')

v.disp %>% filter(sd > 0.15)

x = bind_rows(t.disp, v.disp)

gridExtra::grid.arrange(
  ggplot(x, aes(y=mean,x=cohort,group=cohort,fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') + labs(main='mean', x=''),
  ggplot(x, aes(y=sd,x=cohort,group=cohort,fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') + labs(main='StDev', x=''),
  ggplot(x, aes(y=var,x=cohort,group=cohort,fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') + labs(main='var', x=''),
  ggplot(x, aes(y=IQR,x=cohort,group=cohort,fill=cohort)) + geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') + labs(main='IQR', x=''),
ncol=2, top='Segmented data distribution disp')

wilcox.test(t.disp$mean, v.disp$mean)
wilcox.test(t.disp$sd, v.disp$sd)

```

Looked at the fourier periodograms for both the training and validation data.  The RHS values are based on a frequency cutoff of 0.25. This does suggest that there's more noise in the validation data overall after segmentation.

I tried decreasing this by more agressively segmenting some samples with a higher RHS, but while the RHS drops it doesn't change the overall predictions significantly.  Given the upward shift in the raw and segmented values it may simply be that the shift is affecting it.

```{r rhs}
t.rhs = read_tsv(list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', pattern='rhs', full.names = T, recursive = T), col_types = 'cdd') %>% mutate('cohort'='train')
v.rhs = read_tsv(list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/', pattern='rhs', full.names = T, recursive = T), col_types='cdd') %>% mutate('cohort'='val')

ggplot(bind_rows(t.rhs %>% dplyr::select(rhs,cohort), v.rhs %>% dplyr::select(rhs,cohort)), aes(x=cohort, y=rhs, group=cohort)) + 
  geom_jitter(width=0.2) + geom_violin(aes(fill=cohort),alpha=0.5, show.legend = F) + theme_minimal()

wilcox.test(t.rhs$rhs, v.rhs$rhs)
t.test(t.rhs$rhs, v.rhs$rhs)

```

Example
```{r loadrawval, echo=F, warning=F, message=F}
samples = v.rhs %>% arrange(-rhs) %>% dplyr::slice(1,nrow(v.rhs))

load('~/Data/BarrettsProgressionRisk/QDNAseq/validation/merged_raw_fit.Rdata', verbose=T)

fit.data = merged.fit %>% dplyr::select(location, chrom, start, end, samples$Sample)
raw.data = merged.raw %>% dplyr::select(location, chrom, start, end, samples$Sample)
rm(merged.fit, merged.raw)

```

Highest RHS
```{r highRHS}
if (file.exists('highRHS.Rdata')) {
  load('highRHS.Rdata')
} else {
  info = loadSampleInformation(tibble(Sample=samples$Sample[1], Endoscopy='2010/01/01'))

  # Default gamma is 250* the mad(std dev of the adjusted data)
  seg250 = BarrettsProgressionRisk::segmentRawData(info, raw.data[,c(1:5)], fit.data[,c(1:5)], cutoff = 0.011, gamma2=250, verbose = T)
  prr250 = predictRiskFromSegments(seg250, verbose=F)

  seg500 = BarrettsProgressionRisk::segmentRawData(info, raw.data[,c(1:5)], fit.data[,c(1:5)], cutoff = 0.011, gamma2=500, verbose = T)

  prr500 = predictRiskFromSegments(seg500, verbose=F)
  save(seg250,prr250,seg500,prr500,file='highRHS.Rdata')
}

# Gamma 250
pd250 = TSA::periodogram(seg250$seg.vals[[samples$Sample[1]]], plot=T, main='gamma 250') 
plotSegmentData(seg250) 
# number of segs
nrow(seg250$seg.vals)
# rhs
print( sum(pd250$freq[pd250$freq > 0.25]) )

# gamma 500
pd500 = TSA::periodogram(seg500$seg.vals[[samples$Sample[1]]], plot=T, main='gamma 500') 
plotSegmentData(seg500) 
# number of segs
nrow(seg500$seg.vals)
# rhs
print( sum(pd500$freq[pd500$freq > 0.25]) )

# Relative risks does drop, but not hugely
predictions(prr250)
predictions(prr500)

# These come from a NP
all.val %>% dplyr::filter(Samplename == samples$Sample[1])
```

Lowest RHS.  But again, a more aggressive segmentation isn't really altering things.  
```{r lowRHS}
if (file.exists('lowRHS.Rdata')) {
  load('lowRHS.Rdata')
} else {
  info = loadSampleInformation(tibble(Sample=samples$Sample[2], Endoscopy='2010/01/01'))

  seg250 = BarrettsProgressionRisk::segmentRawData(info, raw.data[,c(1:4,6)], fit.data[,c(1:4,6)], cutoff = 0.011, gamma2=250, verbose = F)
  prr250 = predictRiskFromSegments(seg250, verbose=F)

  seg500 = BarrettsProgressionRisk::segmentRawData(info,  raw.data[,c(1:4,6)], fit.data[,c(1:4,6)], cutoff = 0.011, gamma2=500, verbose = F)

  prr500 = predictRiskFromSegments(seg500, verbose=F)
  save(seg250,prr250,seg500,prr500,file='lowRHS.Rdata')
}

pd250 = TSA::periodogram(seg250$seg.vals[[samples$Sample[2]]], plot=T, main='gamma 250') 
plotSegmentData(seg250)
# number of segs
nrow(seg250$seg.vals)
# rhs
print( sum(pd250$freq[pd250$freq > 0.25]) )

pd500 = TSA::periodogram(seg500$seg.vals[[samples$Sample[2]]], plot=T, main='gamma 500') 
plotSegmentData(seg500)
# number of segs
nrow(seg500$seg.vals)
# rhs
print( sum(pd500$freq[pd500$freq > 0.25]) )

predictions(prr250)
predictions(prr500)

all.val %>% dplyr::filter(Samplename == samples$Sample[2])
```

# Binned Data

```{r, echo=F, message=F, warning=F}
files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06_cleaned_tiled', recursive=T, full.names = T)
training.tiles = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/multipcf_perPatient/', '5e06.*MSE', recursive=T, full.names = T)
training.resids = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/', '5e06_tiles', recursive=T, full.names = T)
val.tiles = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

files = list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/', '5e06_MSE', recursive=T, full.names = T)
val.resids = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

```

Per patient after tiling which merges the segments using a weighted mean.  Again, there's an upwards shift per-patient. 
```{r, fig.width=10}
per.pt.tile<-function(df) {
  means = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~mean(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('mean' = 'V1')
  sds = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~sd(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('sd' = 'V1')
  vars = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~var(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('var' = 'V1')
  iqrs = df [,-1] %>% t %>% as_tibble %>% summarise_all(list(~IQR(.,na.rm=T))) %>% t %>% as_tibble %>% dplyr::rename('IQR' = 'V1')

  return(add_column(bind_cols(means,sds,vars,iqrs), 'Sample'=df$Sample, .before=1))
}

t.pt = per.pt.tile(training.tiles) %>% mutate(cohort='train')
v.pt = per.pt.tile(val.tiles) %>% mutate(cohort='val')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.pt, v.pt), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean per patient'),
  ggplot(bind_rows(t.pt, v.pt), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='StdDev per patient'),
  ncol=2, top='Per patient tiled values')

```

Per 5MB bin (across patients).  There's a lot more variation in the means per bin in the validation data. This is before unit normalization, which is based on the training data mean/sd, so where the variance causes issues is starting to become more clear.

```{r}
tile.dist.measure<-function(df) {
  means = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~mean(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('mean' = 'V1')
  sds = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~sd(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('sd' = 'V1')
  iqrs = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~IQR(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('iqrs' = 'V1')
  vars = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~var(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('var' = 'V1')

  skew = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~skewness(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('skew' = 'V1')
  kurt = as_tibble(df %>% dplyr::summarise_at(vars(-Sample), list(~kurtosis(.,na.rm=T))) %>% t, rownames='bin')  %>% dplyr::rename('kurtosis' = 'V1')

  dist = means %>% full_join(sds, by='bin') %>% full_join(iqrs, by='bin') %>% full_join(vars, by='bin') %>% full_join(skew, by='bin') %>% full_join(kurt, by='bin')
  return(dist)
}

t.dist = tile.dist.measure(training.tiles) %>% mutate(cohort='train')
v.dist = tile.dist.measure(val.tiles) %>% mutate(cohort='val')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'), 
  ncol=2, top='Per bin tiled values')

gridExtra::grid.arrange(
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, skew, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Skew'),
  ggplot(bind_rows(t.dist, v.dist), aes(cohort, kurtosis, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Kurtosis'), 
  ncol=2, top='Per bin tiled values')

grid.arrange(
  ggplot(full_join(t.dist, v.dist, by='bin'), aes(mean.x,mean.y)) + geom_point() + labs(x='train', y='val', title='mean'),
  ggplot(full_join(t.dist, v.dist, by='bin'), aes(sd.x,sd.y)) + geom_point() + labs(x='train', y='val', title='SD'),
  ggplot(full_join(t.dist, v.dist, by='bin'), aes(var.x,var.y)) + geom_point() + labs(x='train', y='val', title='var'),
  ggplot(full_join(t.dist, v.dist, by='bin'), aes(skew.x,skew.y)) + geom_point() + labs(x='train', y='val', title='Skew'),
ncol=2)

```

After scaling and centering per-bin using the z scores from the training data. And sure enough, it spreads the values out considerably for the validation data.  So the question is if the validation data were adjusted would it do this?

```{r}
#load('~/Data/BarrettsProgressionRisk/Analysis/5e6_arms/model_data.Rdata')
#rm(dysplasia.df,labels,z.arms.sd,z.arms.mean,mn.cx,sd.cx)

means = apply(as.matrix(training.tiles[,-1]),2, mean, na.rm=T)
sd = apply(as.matrix(training.tiles[,-1]),2, sd, na.rm=T)

tt.norm = as.matrix(training.tiles[-1])
for (i in 1:ncol(tt.norm))
  tt.norm[,i] = BarrettsProgressionRisk:::unit.var(tt.norm[,i], means[i], sd[i])

val.norm = as.matrix(val.tiles[-1])
for (i in 1:ncol(val.norm))
  val.norm[,i] = BarrettsProgressionRisk:::unit.var(val.norm[,i], means[i], sd[i])

tt.norm = as_tibble(tt.norm) %>% add_column('Sample'=training.tiles$Sample,.before=T)
val.norm = as_tibble(val.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)

tnorm.dist = tile.dist.measure(tt.norm) %>% mutate(cohort='train')
vnorm.dist = tile.dist.measure(val.norm) %>% mutate(cohort='val')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vnorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vnorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values')
```

This strongly suggests that I need to try re-evaluate how to normalize bins.

But what if I, as a quick test, normalize the validation cohort with itself? This looks better here, but when I use this for predictions it doesn't significantly change the results.

```{r}
v.means = apply(as.matrix(val.tiles[,-1]),2, mean, na.rm=T)
v.sd = apply(as.matrix(val.tiles[,-1]),2, sd, na.rm=T)

val.int.norm = as.matrix(val.tiles[,-1])
for (i in 1:ncol(val.int.norm)) 
  val.int.norm[,i] = BarrettsProgressionRisk:::unit.var(val.int.norm[,i], v.means[i], v.sd[i])

val.int.norm = as_tibble(val.int.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)
vinorm.dist = tile.dist.measure(val.int.norm) %>% mutate(cohort='val int')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vinorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values', bottom='Val int = validation bins normalized using internal z-scores')
```

## Rank adjust validation values then scale?

This moved the non-progressors the right direction with the predictions, but there was still a 30% FPR for non-progressors.  The age-bias reappeared for the progressors though.

```{r}

find.rank<-function(r,x) {
  if ( nrow(x %>% filter(rank == r)) > 0 ) return(x %>% filter(rank == r))
  find.rank(r-1,x)
}

adj.v = val.tiles
for (n in 2:ncol(training.tiles)) {
  col = colnames(training.tiles)[n]

  rk = sapply(seq(1, length(training.tiles[[col]]), 4), function(s) mean(training.tiles[[col]][s:(s+4)], na.rm=T))

  tt.rank = tibble(val =rk, rank=rank(rk)) %>%  arrange(rank)
  vt.rank = floor(rank(val.tiles[[col]]))

  for (i in 1:length(vt.rank)) {
    r = vt.rank[i]
    if (is.na(mean(find.rank(r,tt.rank)$val,na.rm=T))) stop(paste(col, r))
    adj.v[i, col] = mean(find.rank(r,tt.rank)$val,na.rm=T)
  }
}

plot( training.tiles %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(sd)) %>% unlist,
      val.tiles %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(sd)) %>% unlist, main='SD', xlab='train', ylab='val tiles')

plot( training.tiles %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(sd)) %>% unlist,
      adj.v %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(sd)) %>% unlist, main='SD', xlab='train', ylab='rank adj val')

plot( training.tiles %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(mean)) %>% unlist,
      adj.v %>% dplyr::select(-Sample) %>% dplyr::summarise_each(funs(mean)) %>% unlist, main='mean', xlab='train', ylab='rank adj val')

# After normalization
adj.v.norm = as.matrix(adj.v[-1])
for (i in 1:ncol(adj.v.norm))
  adj.v.norm[,i] = BarrettsProgressionRisk:::unit.var(adj.v.norm[,i], means[i], sd[i])
adj.v.norm = as_tibble(adj.v.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)
adj.vnorm.dist = tile.dist.measure(adj.v.norm) %>% mutate(cohort='adj val')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vnorm.dist, adj.vnorm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vnorm.dist,adj.vnorm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values')

```

### Predictions

```{r}
preds = do.call(bind_rows, purrr::map(
  list.files('~/Data/BarrettsProgressionRisk/Analysis/validation/multipcf/predictions-rankadj/0.9/', pattern='predictions.tsv', recursive = T, full.names = T), 
  function(f) read_tsv(f, col_types = 'cddcccccccc'))) %>% mutate(Endoscopy = as.Date(Endoscopy)) %>% dplyr::select(-`Path Notes`)

preds = left_join(all.val, preds, by=c('Samplename'='Sample', 'Hospital Research ID', 'SLX-ID', 'Index Sequence', 'Block ID'))

# Non progressors
preds %>% dplyr::filter(Status == 'NP') %>% group_by(Risk) %>% tally() %>% spread(Risk, n)

# FPR is nearly 40% still
nrow(preds %>% filter(Status == 'NP' & Risk == 'High'))/nrow(preds %>% filter(Status == 'NP'))

# Progressors
preds %>% dplyr::filter(Status == 'P' & !grepl('OCCAMS',`Hospital Research ID`)) %>% group_by(Risk) %>% tally() %>% spread(Risk, n)

# FNR is still over 50% in the old progressors as well
nrow(preds %>% filter(Status == 'P' & Risk == 'Low' & !grepl('OCCAMS',`Hospital Research ID`)))/nrow(preds %>% filter(Status == 'P' &  !grepl('OCCAMS',`Hospital Research ID`)))
```

## Distributions

How does it change if I use a global variance, or a rolling-bin variance (to preserve some local info) to adjust?

```{r}
# global variance
all.tiles = bind_rows(training.tiles, val.tiles)

means = apply(all.tiles[,-1], 2, mean)
sds = apply(all.tiles[,-1], 2, sd)
global.sd = sd(as.matrix(all.tiles[,-1]))

wt.sds = sapply(sds, function(x) weighted.mean(x,global.sd,w=c(0.2)))

plot(wt.sds, sds)

tt.norm.gv = as.matrix(training.tiles[-1])
for (i in 1:ncol(tt.norm.gv))
  tt.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(tt.norm.gv[,i], means[i], wt.sds[i])

val.norm.gv = as.matrix(val.tiles[-1])
for (i in 1:ncol(tt.norm.gv))
  val.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(val.norm.gv[,i], means[i], wt.sds[i])

all.norm.gv = as.matrix(all.tiles[-1])
for (i in 1:ncol(all.norm.gv))
  all.norm.gv[,i] = BarrettsProgressionRisk:::unit.var(all.norm.gv[,i], means[i], wt.sds[i])

tt.norm.gv = as_tibble(tt.norm.gv) %>% add_column('Sample'=training.tiles$Sample,.before=T)
val.norm.gv = as_tibble(val.norm.gv) %>% add_column('Sample'=val.tiles$Sample,.before=T)
all.norm.gv = as_tibble(all.norm.gv) %>% add_column('Sample'=all.tiles$Sample,.before=T)

tt.norm.dist = tile.dist.measure(tt.norm.gv) %>% mutate(cohort='train, wt var')
vp.norm.dist = tile.dist.measure(val.norm.gv) %>% mutate(cohort='val, wt var')
all.norm.dist = tile.dist.measure(val.norm.gv) %>% mutate(cohort='all, wt var')

gridExtra::grid.arrange(
  ggplot(bind_rows(tt.norm.dist, vp.norm.dist, all.norm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tt.norm.dist, vp.norm.dist, all.norm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin wt var scale values')

```

```{r}

tx = melt(training.tiles)
vx = melt(val.tiles)

grid.arrange(ggplot(tx, aes(value)) + xlim(0.4, 1.5) + geom_histogram(bins=50, color='lightblue') + labs(title='Training'),
             ggplot(vx, aes(value)) + xlim(0.4,1.5) + geom_histogram(bins=50, color='lightblue') + labs(title='Validation'))

# The extreme values in the validation set come from the OACs
all.val %>% filter(Samplename %in% unique(vx[which(vx$value > 1.5),'Sample']))

vx = vx %>% filter(Sample %in% (all.val %>% filter( !grepl('OCCAMS',`Hospital Research ID`)) %>% dplyr::select(Samplename) %>% pull) )

tx.normal = fitdistrplus::fitdist(tx$value, 'norm')
vx.normal = fitdistrplus::fitdist(vx$value, 'norm')

```

From training data
```{r}
plot(tx.normal)
```

From validation data
```{r}
plot(vx.normal)
```

```{r}

train.P = round(pnorm(tx$value,mean=tx.normal$estimate['mean'], sd=tx.normal$estimate['sd']),4)
val.P = round(pnorm(as.matrix(val.tiles[-1]),mean=tx.normal$estimate['mean'], sd=tx.normal$estimate['sd']),4)
#val.P = round(pnorm(as.matrix(val.tiles[-1]),mean=vx.normal$estimate['mean'], sd=vx.normal$estimate['sd']),4)

p.transform = as.matrix(val.tiles[,-1])
for (i in 1:nrow(p.transform)) {
  for (j in 1:ncol(p.transform)) {
    p.transform[i,j] = mean(tx[which(train.P == val.P[i,j]),'value'])
  }
}

n = 20
hist(as.matrix(training.tiles[,-1])[,n] ,breaks=20)
hist(as.matrix(val.tiles[,-1])[,n] ,breaks=20)
hist(p.transform[,n], breaks=20)

# so now if I scale...

vp.norm = p.transform
for (i in 1:ncol(vp.norm))
  vp.norm[,i] = BarrettsProgressionRisk:::unit.var(vp.norm[,i], means[i], sd[i])
vp.norm = as_tibble(vp.norm) %>% add_column('Sample'=val.tiles$Sample,.before=T)

vp.norm.dist = tile.dist.measure(vp.norm) %>% mutate(cohort='val p')

gridExtra::grid.arrange(
  ggplot(bind_rows(tnorm.dist, vnorm.dist, vp.norm.dist), aes(cohort, mean, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='Mean'),
  ggplot(bind_rows(tnorm.dist, vnorm.dist,vp.norm.dist), aes(cohort, sd, group=cohort, fill=cohort)) + 
    geom_jitter(width=0.1) + geom_violin(alpha=0.5) + theme(legend.position = 'none') +
    labs(title='SD'),
  ncol=2, top='Per bin normalized tile values')

```