### 2019-11-19 Paper_resubmit.Rmd
generating all new plots

---
title: "Paper"
author: "Sarah Killcoyne"
date: "2019-11-13"
output: 
  html_document:
    fig_height: 5
    fig_width: 5
    toc: yes
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(BarrettsProgressionRisk)

library(ggrepel)
#library(GGally)
library(kableExtra)
library(reshape2)
library(gridExtra)
#library(CoxHD)
library(RColorBrewer)

library(OACUtils)

source('lib/load_patient_metadata.R')
source('lib/common_plots.R')

dir = '~/Data/BarrettsProgressionRisk/'
info.dir = paste0(dir,'/QDNAseq')

patient.file = list.files(info.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
demo.file = list.files(info.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)

if (length(patient.file) != 1 | length(demo.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", info.dir))

all.patient.info = read.patient.info(patient.file, demo.file, set='All')$info
patient.info = all.patient.info

patient.info = arrange(patient.info, Status, Hospital.Research.ID, Endoscopy.Year, Pathology)
sum.patient.data = as_tibble(summarise.patient.info(patient.info))

patient.info = patient.info %>% mutate( Pathology = recode(Pathology, 'BE'='NDBE'))

```

# Demographics

```{r}

cohortSummary = sum.patient.data %>% dplyr::group_by(Status) %>% dplyr::summarise(
  'Total Patients'=length(unique(Patient)),
  'Total Samples'=sum(total.samples),
  'Range follow-up (years)'=paste(range(end.year-start.year), collapse='-'),
  'Mean follow-up (years)'=paste(round(mean(end.year-start.year), 1),'+/-',round(sd(end.year-start.year),1)),
  'Range endoscopies'=paste(range(total.endos), collapse='-'),
  'Mean Endoscopies'=paste(round(mean(total.endos), 1),'+/-',round(sd(total.endos),1)),
  'Mean age'=paste(round(mean(age.diagnosis), 1),'+/-',round(sd(age.diagnosis),1)),
  'gender F:M'=paste(table(gender),collapse=':'),
  'Mean Length'=paste(round(mean(C,na.rm=T), 1),'+/-',round(sd(C,na.rm=T),1)),
  'naC'=length(which(is.na(C)))
)

cohortSummary = full_join(cohortSummary, cbind.data.frame('Status' = c('NP','P'), rbind(
      table(subset(patient.info, Status == 'NP')$Pathology),
      table(subset(patient.info, Status == 'P')$Pathology))), by='Status')

cohortSummary %>% t %>% kable(caption='Training set') %>% kable_styling()
```

```{r, echo=F}
grid.arrange(
  ggplot(patient.info, aes(p53.Status, fill=Status)) + geom_bar(position = position_dodge()) + labs(title='p53 Staining') + plot.theme,
  tableGrob( 
    rbind('Non-progressor'=summary(subset(patient.info, Status == 'NP')$p53.Status), 
           'Progressor'=summary(subset(patient.info, Status == 'P')$p53.Status))), 
  nrow=2, as.table=T, heights=c(4,1) )
```

# Copy number data set up

To use GLMs across the patients I merge all samples from all patients and overlap the copy number segments. Where segments from a sample get split, the value of the original sample is retained for each of the split samples.

## y( Progressors (1) vs Non (0) )

The labels are split such that all samples from progressor patients are (1) and all samples from non-progressors are (0).

```{r labelsPNP, echo=F, message=F, include=T}
cache.dir = paste(dir, 'Analysis/models_5e6_all/50kb', sep='/')

file = paste(cache.dir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))

load(file, verbose=F)
#dim(dysplasia.df)
#length(labels)

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=F)
rm(plots, coefs, performance.at.1se, models)

riskCols = brewer.pal(11, "RdYlBu")[c(10, 5, 1)]
```

We have `r table(labels)[1]` samples from non-progressors and `r table(labels)[2]` samples from progressors.

Example subset of the data matrix:
`r dysplasia.df[1:5, 1:5] %>% kable(caption=paste("dimensions:", paste(dim(dysplasia.df), collapse=', '))) %>% kable_styling('striped')`

## Global copy number plots

```{r echo=F, warning=F, message=F, eval=T, fig.height=12, fig.width=20}
chr.lengths = BarrettsProgressionRisk:::chrInfo() %>% dplyr::filter(chr %in% c(1:22)) %>%
  dplyr::mutate(chrom = factor(sub('chr','', chrom), levels=c(1:22), ordered=T))

arm.loc = arrange(bind_rows(
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=1, 'end'=chr.cent-cent.gap))[c('chrom','start','end')],
  (chr.lengths %>% dplyr::group_by(chrom) %>% dplyr::mutate('start'=chr.cent+cent.gap, 'end'=chr.length))[c('chrom','start','end')]), chrom)

#ddf = rbind(dysplasia.df, 'mean.value'=apply(dysplasia.df, 2, mean))
globalCNdf = as_tibble( t(dysplasia.df[,grep('^\\d+:',colnames(dysplasia.df))]), rownames = 'locs') %>% 
  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), as.numeric) %>%
  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  left_join(patient.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
  dplyr::mutate( 
    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))

n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor')$variable)),')'))

globalCN.plot = cn.mtn.plot(globalCNdf %>% filter(seg.type != 'arms'), labeller(Status=n.status, chr=label_value))

cnLegend = get.legend(globalCN.plot)
globalCN.plot + theme(legend.position = 'none')

## NDBE
nb.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(globalCNdf, Status == 'Progressor' & Pathology == 'BE')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(globalCNdf, Status == 'Non-Progressor' & Pathology == 'BE')$variable)),')'))

globalBECN.plot = cn.mtn.plot(filter(globalCNdf, seg.type != 'arms' & Pathology == 'NDBE'), labeller(Status=nb.status, chr=label_value)) + labs(title='NDBE samples only')

gp = globalCN.plot + theme(legend.position = 'none') + labs(y='mean adj. seg.', title='All samples')

gp = globalCN.plot + theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
  labs(y='',x='', title='All samples')
ggsave(filename='plots/globalCN.png', plot=gp, width=8, height=2.3, units='in', dpi=300)

gbe = globalBECN.plot +theme(legend.position = 'none', axis.text = element_blank(), strip.text.y = element_blank(), strip.text.x = element_text(size=8)) + 
  labs(y='',x='', title="Non-Dysplastic Barrett's Samples") 
ggsave(filename='plots/globalBECN.png', plot=gbe, width=8, height=2.3, units='in', dpi=300)

```

## *Complexity Scores*

```{r loaddata, message=F, warning=F, echo=F}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

cx = dysplasia.df[,'cx']
complexity.score = tibble('Status'=labels, 'cx'=cx[names(labels)], 'Samplename' = names(labels)) %>%
  mutate(Status = factor(Status, labels = c('Non-Progressor','Progressor'))) %>%
  left_join(patient.info[,c('Samplename','Pathology','Patient')],by=c('Samplename'))

tt = t.test(subset(complexity.score, Status == 'Progressor')$cx, subset(complexity.score, Status == 'Non-Progressor')$cx)
pv = ggplot(complexity.score, aes(Status, cx, fill=Status)) + 
  geom_jitter(width=0.2, color='grey39') + 
  geom_boxplot(outlier.colour = NA, alpha=0.8) +
  labs(x='', y='cx', title='Per sample cx scores', subtitle=paste('p-value=', format.pval(tt$p.value, digits=2))) + theme(legend.position = 'none') +
  plot.theme
pv
ggsave('plots/per_sample_cx.png', plot=pv + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocX, title='Per sample cx') + plot.theme

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocX = pROC::roc(Status ~ cx, data=complexity.score, auc=T, ci=T, of='thresholds', transpose = F)
roc.plot(rocX, title='Per sample cx') + plot.theme

ggsave('plots/per_sample_cx_roc.png', plot=roc.plot(rocX, title='Per sample cx') + plot.theme + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
```

# Building the model

We built the model using a 50-fold cross-validation process on the patients (rather than samples) for various lambda and alpha values to find the best fitting model.

```{r model, message=F, warning=F, echo=F, fig.height=16, fig.width=16}
folds = 10; splits = 5 
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits'))

ggsave('plots/eyebrows.png', plot=do.call(grid.arrange, c(plots, top='All samples, 10fold, 5 splits')), width=10, height=10, units='in', dpi=300)

```

```{r echo=F, fig.height=5, fig.width=5}
library(ggfortify)
autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme

ggsave('plots/glm_0.9.png', plot=autoplot(models$`0.9`) + theme(legend.position = 'none') + plot.theme, width=5, height=5, units='in', dpi=300)
```

## Model performance

```{r model_perf, message=F, warning=F, echo=F, fig.width=6, fig.height=6}
classification.rate = as_tibble(matrix(ncol=5, nrow=0, dimnames=list(c(), c('model','mean','sme','stable.coef', 'all.coef'))))

mp = model.performance(performance.at.1se, coefs, folds, splits )
mp$plot = mp$plot + labs(title=paste('All samples,', folds, 'folds,', splits, 'patient splits'))
mp$plot

ggsave('plots/performance/ridge_v_lasso.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

select.alpha = as.character(0.9)
# Genomic regions only
most.stable = grep('^\\d+', names(which(mp$stable.coeffients[[select.alpha]] >= 0.75 )), value=T)
most.stable = cbind.data.frame(do.call(rbind, strsplit(most.stable, ':|-')), most.stable)
colnames(most.stable) = c('chr','start','stop', 'name')

most.stable$chr = as.integer(as.character(most.stable$chr))
most.stable = arrange(most.stable, chr, start)

classification.rate = add_row(classification.rate, 'model'='All samples', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

mp$performance[,c('n.Coef','25%','50%','75%', '100%')] %>% kable(caption='Number of features stable in >=n% of folds') %>% kable_styling(full_width = F)
```

## Feature Stability

## Remove HGD/IMC & LGD Samples and check the model

The initial model is trained against `r nrow(dysplasia.df)` samples with `r ncol(dysplasia.df)` features. The samples include the final (diagnostic) HGD/IMC samples from progressors.

### Remove HGD/IMC

```{r noHGD, echo=F, message=F, warning=F, fig.height=16, fig.width=16, eval=T}
rm(performance.at.1se, coefs, mp)

file = paste(cache.dir, 'nohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

nohgd.coefs = coefs[[select.alpha]]
#do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2))
ggsave('plots/nohgd_eyebrow.png', plot=do.call(grid.arrange, c(no.hgd.plots, top="No HGD/IMC samples", ncol=2)), height=10, width=10, units='in', dpi=300)
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs, folds, splits)
mp$plot = mp$plot + labs(title='No HGD/IMC samples')
mp$plot
ggsave('plots/performance/ridge_v_lasso_nohgd.png', plot=mp$plot, width=5, height=5, units='in', dpi=300)

classification.rate = add_row(classification.rate, 'model'='No HGD/IMC', 
                              'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp)
```

### Remove LGD

```{r noLGD, echo=F, warning=F, message=F, fig.height=16, fig.width=16}
file = paste(cache.dir, 'nolgd.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("No model file", file, "found"))
}

#do.call(grid.arrange, c(nolgd.plots, top="No LGD/HGD/IMC samples", ncol=2))
```

```{r, echo=F, fig.width=6, fig.height=6}
mp = model.performance(performance.at.1se, coefs, folds, splits)
mp$plot = mp$plot + labs(title='No LGD/HGD/IMC samples')
mp$plot
ggsave('plots/performance/no_lgd_perf.png', plot=mp$plot, height=5, width=5, units='in', dpi=300)

classification.rate = add_row(classification.rate, 'model'='No LGD/HGD/IMC', 'mean'=mp$performance[select.alpha, 'mean'], 'sme'=mp$performance[select.alpha, 'sme'],
                              'stable.coef'=length(which(mp$stable.coeffients[[select.alpha]] >= 0.75)), 'all.coef'=length(mp$stable.coeffients[[select.alpha]]))

rm(performance.at.1se, coefs, mp)
```

## Overall

Even without dysplastic samples our model is well able to classify samples as either progressive or non.

```{r, echo=F, warning=F, message=F, fig.width=6, fig.height=6}
#classification.rate$x = 1:nrow(classification.rate)

p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p

ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

# Features

```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=5}
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=F)
rm(plots, performance.at.1se)

featureStability = rowSums(coefs[[select.alpha]][,-1])/(splits*folds)
hazards = apply( exp(t(dysplasia.df[, rownames(coefs[[select.alpha]])]) *  coefs[[select.alpha]][,1]), 1, function(x) {
  sd(x)/mean(x)
})

features = as_tibble(hazards, rownames = 'label') %>% dplyr::rename(hazards = value) %>% 
  dplyr::mutate(coef = coefs[[select.alpha]][,1], stability = featureStability) %>% arrange(-hazards)

cvRR = features

qq = quantile(features$hazards, seq(0,1,.15))
p = ggplot(features, aes(coef, hazards, col=hazards >= qq[length(qq)])) + 
  geom_point(alpha=0.8, show.legend = F) + 
  geom_text_repel(aes(x=coef, label=ifelse(hazards >= qq[length(qq)], label, '')),show.legend=F) +
  scale_color_manual(values=c('firebrick3','darkblue')) +
  labs(title='Top features by cv(RR)', x='Coefficient value', y='cvRR') + plot.theme 
p
ggsave('plots/haz_coef.png', plot=p, height=5, width=5, units='in', dpi=300)
```

`r features %>% kable(caption='Selected coefficients') %>% kable_styling(full_width=F)`

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
features = features %>% filter(label != 'cx') %>% separate(label, c('chr','start','end'), sep=':|-', remove=F) %>% 
  dplyr::mutate_at(vars(start, end), list(as.numeric)) %>% 
  mutate(chr = factor(chr,levels=c(1:22))) %>% 
  dplyr::mutate(ymax = ifelse(coef > 0, max(globalCN.plot$data$mean.value), min(globalCN.plot$data$mean.value)), gl = factor(ifelse(coef > 0, 'gain','loss'))) %>%
  left_join(chr.lengths %>% dplyr::select(chr, chr.length), by=c('chr')) %>% mutate(chr = factor(chr, levels=c(1:22), ordered=T))

custom.max<-function(cf, hazards, ymax) {
  if (cf > 0) {
    max = hazards*ymax
    max = ifelse(max > ymax, ymax, max)
  } else {
    max = hazards*ymax
    max = ifelse(max < ymax, ymax, max)
  }
  return(max)
}
features = features %>% rowwise() %>% dplyr::mutate('max'=custom.max(coef, hazards, ymax)) %>% arrange(chr)

p = globalCN.plot + geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + 
  geom_rect(data= features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
ggsave('plots/global_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)

p = globalBECN.plot +  geom_rect(data=features %>% filter(gl == 'gain'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='goldenrod2', alpha=0, show.legend=F) + geom_rect(data=features %>% filter(gl == 'loss'), aes(xmin=start, xmax=end, ymin=0, ymax=max), color='blue3', alpha=0, show.legend=F) + theme(legend.position = 'none')
ggsave('plots/ndbe_cn_coefs.png', plot=p, width=16, height=4.6, units='in', dpi=300)
```

# Nested leave-one-out predictions

Leave out each patient (both P and NP) and run a new CV fit each time then predict the samples for the patient that was left out.  Fitted model includes HGD samples.

```{r leaveoneout, echo=F, message=F, warning=F, fig.width=5, fig.height=5}
file = paste(cache.dir, 'loo_0.9.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=F)
} else {
  stop(paste("Missing model file", file))
}
loo.coefs = nzcoefs

pg.samp = pg.samp %>% mutate(Pathology = recode(Pathology, 'BE' = 'NDBE'))

names(performance.at.1se) = names(pg.samp)
df = as.data.frame(performance.at.1se)

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p
ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

#oddPt = names(pg.samp)[which.max(performance.at.1se)]
#subset(sum.patient.data, Hospital.Research.ID == oddPt)

#pg.samp[[oddPt]]$Prediction

# Not the pt with the most samples or endos
#sum.patient.data[which.max(sum.patient.data$total.samples),]
```

### Prediction ROC curves

#### LOO minus HGD

```{r loonohgd, echo=F, message=F, warning=T}
file = paste(cache.dir, 'loonohgd.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading file", file))
  load(file, verbose=T)
} else {
  stop(paste("No file", file, "found"))
}

perf = cbind.data.frame(perf=performance.at.1se, pt=unique(pg.sampNOHGD$Patient))

p = ggplot(melt(perf, id.vars='pt'), aes(x=variable,y=value, group=variable, fill=variable)) +
    geom_boxplot(fill='lightcoral', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) +
    geom_jitter(color='grey39', width=0.3) + plot.theme + 
    geom_label(data=melt(subset(perf, perf %in% range(perf) ), id.vars='pt'), aes(x=variable, y=value, label=round(value, 3)), fontface='bold', fill='lightcoral') +
    theme(legend.position = 'none') + labs(x='LOO model', y='Classification rate', title='Performance for LOO', subtitle='HGD samples excluded model')
ggsave('plots/performance/loo_perf_nohgd.png', plot=p, width=5, height=5, units='in', dpi=300)

pg.sampNOHGD = pg.sampNOHGD %>% filter(!is.na(Prediction))

rocNOHGD = pROC::roc(Status ~ Prediction, data=pg.sampNOHGD[c('Status', 'Prediction')], ci=T, of='thresholds', transpose=T)

```

```{r insilico_rocs, echo=F, message=F, warning=F, fig.height=8, fig.width=8}
load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=F)
```

```{r prediction_cutoff, echo=F, message=F, warning=F, fig.height=7, fig.width=7}
preds = pg.samp %>% dplyr::select(Status, Prediction)
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
ggsave(filename='plots/auc/training_auc.png', plot=roc.plot(roc), width=6.5, height=4.5, units='in', dpi=300)

roc$model = 'CN Model'
roc$is = 'discovery'

droc = multi.roc.plot(list(roc), colors=brewer.pal(3,'Greys')[3]) + labs(title='Discovery cohort ROC')
droc
ggsave(filename='plots/auc/discovery_roc.png', plot=droc, width=6, height=4, units='in', dpi=300)

insilico.rocs[['discovery']] = roc
df = do.call(rbind, lapply(insilico.rocs, function(r) 
    cbind.data.frame('AUC' = as.numeric(r$auc), 'Specificity'=rev(r$specificities), 'Sensitivity'=rev(r$sensitivities),'model'=r$model, 'is'=as.character(r$is) 
)))

i.thresholds = do.call(bind_rows, purrr::map(insilico.rocs, function(r) pROC::coords(r, 'best', transpose=F) %>% as_tibble %>% mutate(AUC = r$auc, model=r$model, is=as.character(r$is)))) %>%
  rename_at(vars(matches('thresh|spec|sens')), funs(str_to_title))

#summary(filter(df, model == 'in-silico')$AUC)

getPalette = colorRampPalette(brewer.pal(5, "Blues"))
p = ggplot(df, aes(Specificity, Sensitivity,color=is)) + 
  geom_line(linetype='solid', size=0.5) + 
  geom_line(data=subset(df, is == 'discovery'), color='red', size=1.5) +
  scale_x_reverse() +
  scale_color_manual(values=getPalette(length(unique(df$is)))) +
  labs(title='ROC, all splits', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme +
  theme(legend.position = 'none')
p

blues = RColorBrewer::brewer.pal(5,'Blues')
p = ggplot(df, aes(Specificity, Sensitivity)) + 
    geom_line(color=blues[2]) + 
    geom_line(data=filter(df, is == 'discovery'), color=blues[5], size=1) +
    geom_line(data=df %>% filter( (AUC == max(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    geom_line(data=df %>% filter( (AUC == min(AUC) ) & is != 'discovery'), color=blues[4], linetype = 'dotted', size=0.75) +
    scale_x_reverse() +
    geom_label(data=filter(i.thresholds, AUC %in% range(AUC) | is == 'discovery'), aes(label=paste0('AUC\n',round(AUC,2))), color='black') +
    #scale_color_brewer(palette='Blues') +
    labs(title='ROC AUC', subtitle='in-silico splits', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none', text = element_text(size=14))
p
ggsave(filename='plots/auc/splits_roc.png', plot=p, width=6, height=6, units='in', dpi=300)

# Not sure why I did this
df = cbind.data.frame('specificity'=rev(roc$specificities), 'sensitivity'=rev(roc$sensitivities))
thresholds = 
  rbind.data.frame(
    est=t(round(pROC::coords(roc, "best", transpose=T),2)),
    max=t(round(pROC::coords(roc, 0.8, transpose=T),2)),
    'No HGD'=t(round(pROC::coords(rocNOHGD, 0.6, transpose=T),2)),
    'All samples'=t(round(pROC::coords(roc, 0.6, transpose=T),2))
)
thresholds$label=rownames(thresholds)

ggplot() + 
  geom_line(data=df, aes(specificity, sensitivity), color='darkblue', size=1.5) + scale_x_reverse() +
  geom_point(data=thresholds, aes(x=specificity,y=sensitivity,color=label), size=3,shape=8, show.legend = T) +
  geom_label_repel(data=thresholds, aes(x=specificity,y=sensitivity,fill=label,label=
    paste('Threshold ', round(threshold,2), '\n(FPR ', round((1-specificity),2)*100, '%, TPR ', round(sensitivity,2)*100,'%)\nAUC:',round(roc$auc, 2)*100, '%', sep='')),show.legend = F, color='white',nudge_x=0.05) + scale_fill_manual(values=brewer.pal(4,'Dark2')) + scale_color_manual(values=brewer.pal(4,'Dark2')) +
  labs(title='CN Model ROC', x='Specificity (FPR)', y='Sensitivity (TPR)')  + plot.theme

cutoff = round(pROC::coords(roc, 'best', ret = 'threshold', transpose=T), 2)
```

### p53 ROC curve

```{r p53roc, echo=F, fig.height=2, fig.width=2}
p53samples = subset(patient.info, !is.na(p53.Status), select=c('Samplename','p53.Status','Status'))
p53roc = pROC::roc(as.numeric(p53samples$Status)-1, as.integer(p53samples$p53.Status)-1)

p53roc$model = 'p53 IHC'
roc.plot(p53roc, 'p53 IHC')
ggsave('plots/auc/p53_roc.png', plot=roc.plot(p53roc, 'p53 IHC'), width=6, height=6, units="in", limitsize=F, dpi = 300)
```

### Pathology ROC curve

Based on HGD/IMC/LGD
```{r pathroc, echo=F, fig.height=4, fig.width=12 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

pathsamples = patient.info[,c('Pathology','Status')]
pathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('LGD','HGD','IMC'), 1, 0) )
pathroc$model = 'HGD & LGD'
ggsave('plots/auc/dysplasia_auc.png', plot=roc.plot(pathroc, title='Dysplasia (HGD & LGD)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

hgdpathroc = pROC::roc((as.integer(pathsamples$Status)-1), ifelse(pathsamples$Pathology %in% c('HGD','IMC'), 1, 0) )
hgdpathroc$model = 'HGD Only'
ggsave('plots/auc/hgd_auc.png', plot=roc.plot(hgdpathroc, title='HGD Only'), width=6, height=6, units="in", limitsize=F, dpi = 300)

grid.arrange(roc.plot(pathroc, 'Pathology (LGD,HGD,IMC)'), roc.plot(hgdpathroc, 'Pathology (HGD,IMC)'), ncol=2)
```

### Per Endoscopy ROCs
```{r, echo=F, warning=F, message=F,fig.height=4, fig.width=4}
pendo = pg.samp %>% dplyr::group_by(Patient, Status, Hospital.Research.ID, PID ) %>% 
  dplyr::summarise( 'max'=max(Prediction), 'avg'=mean(Prediction), 'n'=length(PID) )

rocEndoM = pROC::roc(Status ~ max, data=pendo, ci=T, of='thresholds')
rocEndoM$model = 'Per Endoscopy Max'
ggsave('plots/auc/maxendo_auc.png', plot=roc.plot(rocEndoM, title='Max per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocEndoA = pROC::roc(Status ~ avg, data=pendo, ci=T, of='thresholds')
rocEndoA$model = 'Per Endoscopy Avg'
ggsave('plots/auc/avgendo_auc.png', plot=roc.plot(rocEndoA, title='Avg per Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

lastEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, PID) %>% arrange(desc(Endoscopy.Year), PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == max(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)

rocLastEndo = pROC::roc(Status ~ avg, data=lastEndo, ci=T, of='thresholds')
rocLastEndo$model = 'Last endoscopy'
ggsave('plots/auc/lastendo_auc.png', plot=roc.plot(rocLastEndo, title='Last endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

firstEndo = pg.samp %>% group_by(Patient, Hospital.Research.ID, Status,Endoscopy.Year, PID) %>% arrange(Endoscopy.Year, PID)  %>% dplyr::summarise(avg = mean(Prediction) ) %>% ungroup %>% group_by(Patient) %>% dplyr::filter( Endoscopy.Year == min(Endoscopy.Year) ) %>% dplyr::select(-Endoscopy.Year, -PID)

rocFirstEndo = pROC::roc(Status ~ avg, data=firstEndo, ci=T, of='thresholds')

rocFirstEndo$model = 'First endoscopy'
ggsave('plots/auc/firstendo_auc.png', plot=roc.plot(rocFirstEndo, title='First Endoscopy'), width=6, height=6, units="in", limitsize=F, dpi = 300)

endo.thresholds = bind_cols('roc'=c('Max','Avg','Last','First'),
  bind_rows(
    pROC::coords(rocEndoM, 'best', transpose=T),
    pROC::coords(rocEndoA, 'best', transpose=T),
    pROC::coords(rocLastEndo, 'best', transpose=T),
    pROC::coords(rocFirstEndo, 'best', transpose=T)
  ))

```

### Per Patient ROC
```{r echo=F, fig.height=4, fig.width=12}
# Excluding the final endoscopy
ptendo = do.call(bind_rows, lapply(unique(pg.samp$Hospital.Research.ID), function(pt) {
  x = pg.samp %>% filter(Hospital.Research.ID == pt)  
  x %>% filter(Endoscopy.Year != max(x$Endoscopy.Year))
})) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise_at(vars(Prediction), funs(max, mean), na.rm=T)

ptendo = ptendo %>% filter(!is.na(mean))

rocPtM = pROC::roc(Status ~ max, data=ptendo, ci=T, of='thresholds')
#pROC::coords(rocPtM, 'best', transpose=T)
rocPtM$model = 'Per Patient Max'
ggsave('plots/auc/maxpt_auc.png', plot=roc.plot(rocPtM, title='Max per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

rocPtA = pROC::roc(Status ~ mean, data=ptendo, ci=T, of='thresholds')
#pROC::coords(rocPtA, 'best', transpose=T))
rocPtA$model = 'Per Patient Avg'
ggsave('plots/auc/avgpt_auc.png', plot=roc.plot(rocPtA, title='Avg per Patient (excl. final)'), width=6, height=6, units="in", limitsize=F, dpi = 300)

grid.arrange(roc.plot(rocPtM, title='Max') , roc.plot(rocPtA, title='Avg'), nrow=1, top='Per Patient')
```

### Compare ROCs

```{r echo=F, fig.height=8, fig.width=6}
get.roc.stat<-function(roc, x='best') {
  auc = c( pROC::auc(roc), range(pROC::ci.auc(roc))  )
  sens = c(pROC::coords(roc, x, tranpose=T)[['sensitivity']], pROC::ci.se(roc,pROC::coords(roc, x, tranpose=T)[['specificity']], conf.level=0.95))
  spec = c(pROC::coords(roc, x, tranpose=T)[['specificity']], pROC::ci.sp(roc,pROC::coords(roc, x, tranpose=T)[['sensitivity']], conf.level=0.95))

  stat = rbind('AUC'=auc, 'Sensitivity'=sens[c(1,2,4)], 'Specificity'=spec[c(1,2,4)])
  colnames(stat) = c ('value','CI.min','CI.max')
  stat = as.data.frame(stat)
  stat$model = roc$model
  stat$Measure = rownames(stat)
  stat
}

get.auc.at<-function(roc, foc='spec', p=c(1,0.99)) {
# To find the auc at a given spec or sens
  pROC::auc(roc, partial.auc=p, partial.auc.focus=foc, partial.auc.correct=T)
}

# get.auc.at(p53roc)
# get.auc.at(hgdpathroc)
# get.auc.at(pathroc)
# get.auc.at(roc)

roc$model = "CN Model"
m = rbind.data.frame(
      get.roc.stat(p53roc),
      get.roc.stat(hgdpathroc),
      get.roc.stat(pathroc),
      get.roc.stat(roc))
m$model = factor(m$model, levels=c('p53 IHC','HGD Only','HGD & LGD','CN Model'), ordered = T)

rocList = list(roc,pathroc,hgdpathroc,p53roc)

cols = brewer.pal(4, "PRGn")
dodge = position_dodge(width=0.9)

comp = ggplot(melt(subset(m, Measure == 'AUC'), measure.vars='value'), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=Measure)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='black' ) + 
  geom_text(aes(y=(value*100)-4, label=paste(round(value,2)*100,'%',sep='')), position=dodge, color='white')  +
  scale_y_continuous(limits=c(0,100)) +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position = 'none', text = element_text(size=14), axis.text.x = element_text(angle=45, hjust=1)) + 
  labs(x='', y='AUC', title='ROC Comparison') 
comp

ggsave('plots/auc/path_p53_cn_auc.png', plot=comp, width=3, height=4, units="in", limitsize=F, dpi = 300)

```

```{r}
roc$model = "Per Sample"

m = rbind.data.frame(
      get.roc.stat(roc),
      get.roc.stat(rocEndoM),get.roc.stat(rocEndoA),
      get.roc.stat(rocPtM), get.roc.stat(rocPtA),
      get.roc.stat(rocFirstEndo), get.roc.stat(rocLastEndo))
m$model = factor(m$model)

cols = brewer.pal(11, 'Paired')
dodge = position_dodge(width=0.9)
mm = melt(subset(m, Measure == 'AUC'), measure.vars='value')
p = ggplot( arrange(mm, CI.min), aes(x=model, y=value*100, ymin=CI.min*100, ymax=CI.max*100, group=Measure, fill=model)) + 
  geom_bar(stat='identity',position=dodge) + geom_errorbar(position=dodge, width=0.3, color='grey39' ) + 
  geom_text(aes(y=CI.min*100, label=paste(round(value,2)*100,'%',sep='')), position=dodge, vjust=3)  +
  scale_fill_manual(values=cols, name='') + plot.theme + 
  theme(legend.position='none', axis.text.x=element_text(angle=45, hjust=1)) + labs(x='', y='AUC', title='ROC Comparison') 
p

ggsave('plots/auc/multi_auc.png', plot=p, height=6, width=5, units='in',dpi=300)
```

## Prediction probability distribution

```{r}
pg.samp = do.call(bind_rows,lapply(unique(pg.samp$Patient) %>% map( function(p) pg.samp %>% filter(Patient == p)), transform.by.time, info=patient.info)) 
sq = c(0, seq(1,132,12),max(pg.samp$months.before.final))

pg.samp = pg.samp %>% mutate(
  brks = cut(months.before.final, breaks=sq, 
                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'),include.lowest = T, ordered_result = T),
) %>% dplyr::mutate(brks = factor(brks, rev(levels(brks))))

ggplot(pg.samp, aes(Status, Prediction, group=Status)) + 
  facet_grid(~Pathology) + 
  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
  scale_color_gradientn(colours = myPal) +
  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
  geom_jitter(width=0.2, aes(color=Prediction)) + 
  plot.theme  + theme(legend.position = 'none') + labs(y='Probability')

ggplot(pg.samp, aes(Status, RR, group=Status)) + 
  facet_grid(~Pathology) + 
  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
  scale_color_gradientn(colours = myPal) +
  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
  geom_jitter(width=0.2, aes(color=Prediction)) + 
  plot.theme  + theme(legend.position = 'none')

```

The labels in each decile are the ratio of progessor samples that have probabilities within that region. The deciles that are greyed out show insigificant enrichment for the NP/P ratio.

```{r spacial, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
maxP = 0.5; minP = 0.3

cuts = seq(0,1,0.1)

pg.samp = pg.samp %>% mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T))
qt = pg.samp %>% group_by(quants, Status) %>% dplyr::summarise(n=length(Status) ) %>% spread(Status, n)

ft.fun<-function(NP,P) {
 f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
 cbind.data.frame('p.value'=round(f$p.value, 4))
}

qt = left_join(qt, pg.samp %>% dplyr::group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ), by='quants')

pred.confidence = qt %>% dplyr::group_by(quants) %>% 
  dplyr::mutate( 'P.ratio'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') ) %>% 
  separate(quants, c('r1','r2'), ',', remove = F) %>% 
  mutate_at(vars('r1','r2'), list(sub), pattern='\\[|\\]|\\(', replacement='') %>% mutate_at(vars(r1,r2), as.double) %>%  
  dplyr::mutate(Risk = ifelse(round(P.ratio,1)<.5, 'Low','High'))

pred.confidence = bind_cols(pred.confidence, 
                            data.frame(ci.low=qbeta(0.025, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5),
                                       ci.high=qbeta(0.975, shape1=pred.confidence$P+.5, shape2 = pred.confidence$NP+.5)))

pcfit = lm(P.ratio~mn, data=pred.confidence)
rsid = rstandard(pcfit)
#pred.confidence[which( rsid >= 1 |  rsid <= -1 ), 'Risk'] = 'Moderate'
pred.confidence = pred.confidence %>% mutate(Risk = as.character(ifelse(r1 < maxP & r2> minP, 'Moderate', Risk)))

annotation.table = pred.confidence %>% dplyr::group_by(Risk) %>% 
  dplyr::summarise( 'n(NP)'=sum(NP), 'n(P)'=sum(P), 'p.value'=format.pval(mean(p.value), 3), 'min'=min(r1), 'max'=max(r2) ) %>%
  arrange(min)

tt3 <- ttheme_minimal(
  core=list(bg_params = list(fill = riskCols, col=NA, alpha=0.2),
            fg_params=list(fontface=3)),
  colhead=list(fg_params=list(col="black", fontface=4L)),
  rowhead=list(fg_params=list(col=c('red3','green3', 'orange2'), fontface=3L)),
  padding=unit(c(5,5),'mm'))
#grid.arrange( tableGrob(annotation.table, rows=NULL, theme=tt3) )
pred.confidence$Risk = factor(pred.confidence$Risk, levels=c('Low','Moderate','High'), ordered = T)

pg.samp = left_join(pg.samp, pred.confidence %>% dplyr::select(quants, Risk), by='quants')
```

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
pcal = BarrettsProgressionRisk::showPredictionCalibration(pred.confidence %>% dplyr::rename(perc = P.ratio))
pcal
ggsave(filename='plots/pred_calib.png', plot=pcal, width=4, height=4, units='in', dpi=300)

mini = pcal + labs(x='P(P)', y='P:NP', title='') + theme(legend.position = 'none') + 
  scale_x_continuous(expand = c(0, 0), limits = c(-0.5, 1.5), breaks = cuts, labels = c(0,'','',0.3,'',0.5,'','','','',1)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(-0.5, 1.5), breaks = seq(0,1,0.25), labels = seq(0,1,0.25)) + 
  theme(axis.text = element_text(size=14))
ggsave(filename='plots/pred_calib_mini.png', plot=mini, width=4, height=4, units='in', dpi=300)

pred.confidence %>% dplyr::mutate_if(is.double, list(signif), 2) %>% kable() %>% kable_styling('striped')

per.endo = pg.samp %>% group_by(Hospital.Research.ID, Status, Patient, PID) %>% 
  dplyr::mutate('n.samples'=length(PID),'Samples'=paste(Samplename,collapse=',')) %>% 
  dplyr::summarise_at(vars(matches('Pred|RR|Year')), list(max)) %>% dplyr::select(-matches('^Sample$')) %>% 
  arrange(Patient, Endoscopy.Year, PID)  
per.endo = per.endo %>% dplyr::mutate(quants = cut(Prediction, breaks=cuts, include.lowest = T)) %>% 
  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants') %>% 
  dplyr::select(-matches('quant'))# %>% dplyr::mutate(Risk = as.character(Risk))

pg.samp %>% group_by(Status, Risk) %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
  dplyr::mutate(n=sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>%
  mutate_if(is.double, list(round),2) %>%
  kableExtra::kable(caption='Training set per-sample predictions (ratios)') %>% kableExtra::kable_styling(full_width = F)

per.endo %>%  group_by(Status, Risk)  %>% dplyr::summarise(n = length(Risk)) %>% spread(Risk, n) %>% rowwise %>% 
  dplyr::mutate(n = sum(High,Low,Moderate), Low = Low/n, Moderate = Moderate/n, High = High/n) %>% mutate_if(is.double, list(round),2) %>%
  kableExtra::kable(caption='Training set per-endoscopy predictions') %>% kableExtra::kable_styling(full_width = F)

```

```{r probhist, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) + scale_fill_distiller(palette = 'RdYlBu', name='P(P)') +
    geom_vline(xintercept=annotation.table$max, linetype='longdash', color='grey39') +
    geom_text(data=annotation.table %>% dplyr::group_by(Risk, min,max) %>% dplyr::summarise( 'samp.perc'= signif((sum(`n(NP)`, `n(P)`)/nrow(patient.info))*100, 3) ), aes(x=min+(max-min)/2, y=5, label=paste(samp.perc, '%',sep=''))) +
    plot.theme + labs(title='Sample predictions', y='n Samples', x='Probability') + theme(legend.position = 'none')
```

```{r relrisk, echo=F, fig.width=6, fig.height=6}
unadjRR = ggplot(pg.samp, aes(RR)) + geom_histogram(aes(fill=..x..), bins=10, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk') + plot.theme
unadjRR
ggsave(filename='plots/unadj_relrisk.png', plot=unadjRR, width=6, height=5, units='in', dpi=300)

predhist = ggplot(pg.samp, aes(Prediction)) + geom_histogram(aes(fill=..x..), breaks=cuts, show.legend = F) +
    scale_fill_gradientn(colors = myPal,  name='') + 
    plot.theme + labs(title='Predictions, all samples model', y='n Samples', x='Probability') 
ggsave(filename='plots/pred_hist.png', plot=predhist, width=5, height=5, units='in', dpi=300)
```

### Mountain plots using spatial info

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=20}
min.theme = theme(text=element_text(size=8), panel.background=element_blank(), strip.background=element_rect(fill="grey97"), strip.text.y = element_text(size=6),
                   strip.text.x = element_text(size=12), axis.text = element_blank(), axis.ticks = element_blank(),
                   axis.line=element_line(color='black'), panel.grid.major.y =element_line(color='grey95'), panel.grid.major.x = element_blank(),
                   panel.border = element_rect(color="grey", fill=NA, size=0.5), panel.spacing.x = unit(0, 'lines'), panel.spacing.y = unit(0.1, 'lines')   )

setOgj <- function(nl) {
  if (nl >= 0 & nl < 2) return(1)
  if (nl >= 2 & nl < 4) return(2)
  if (nl >= 4 & nl < 7) return(3)
  if (nl >= 7) return(4)
}
pg.samp = pg.samp %>% mutate(ogj = factor(unlist(purrr::map(nl, setOgj)), ordered=T))

#df = globalCNdf %>% filter(Patient == 20 & seg.type != 'arms')
#df = globalCNdf %>% filter(Patient == 51 & seg.type != 'arms')

# early/late samples
#df = subset(globalCNdf, Patient == pt)
df = left_join(globalCNdf, pg.samp[,c('Samplename','Endoscopy.Year', 'ogj', 'months.before.final', 'RR','Prediction')], by=c('variable'='Samplename')) %>% arrange(chr, start, Endoscopy.Year)

## count arms differently than segs??
df2 = df %>% dplyr::group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(RR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  ) %>% mutate(Patient = factor(Patient))
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

df2 = df2 %>% dplyr::group_by(Status, Patient) %>% dplyr::mutate(
  'early'= Endoscopy.Year == min(Endoscopy.Year),
  'late' = Endoscopy.Year == max(Endoscopy.Year)
)
df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))

el = df2 %>% filter(early | late)
el2 = el %>% dplyr::group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'

p = ggplot(df2 %>% filter(early|late), aes(Endoscopy.Year, arms, color=Patient)) + facet_grid(~Status) +
  geom_point() +  geom_line(data=subset(df2, early | late)) + theme(legend.position = 'none', text=element_text(size=11)) + 
  plot.theme + labs(x='Endoscopy Year', y='CN arm values', title='Initial vs final endoscopy', subtitle=paste('p', wtA))
p
ggsave(filename='plots/status_hetero.png', plot=p, width=8, height=6, units='in', dpi=300)

wtA = round(wilcox.test(subset(df2, early &  Status == 'Progressor')$arms, subset(df2, early & Status == 'Non-Progressor')$arms)$p.value, 3)
wtCN = round(wilcox.test(subset(df2, early & Status == 'Progressor')$CNA, subset(df2, early & Status == 'Non-Progressor')$CNA)$p.value, 2)

m = melt(subset(df2, early), measure.vars=c('CNA','arms'), id.vars=c('Status'))
p = ggplot(m, aes(Status, value, fill=Status)) + facet_wrap(~variable, scales='free_y') + 
  geom_boxplot(outlier.colour = NA) + geom_jitter(width=0.2) + 
  scale_fill_brewer(palette = 'Dark2') + 
  plot.theme + theme(legend.position = 'none', text=element_text(size=11)) +
  labs(y='', x='', title='CN values, initial endoscopy', subtitle=paste('Arms p=',wtA, '   CNA p=',wtCN, sep=''))
p
ggsave(filename='plots/hetero_initial.png', plot=p, width=6, height=4, units='in', dpi=300)

ogjLimits = levels(factor(pg.samp$ogj, ordered = T))
p2 = mtn.spatial(globalCNdf %>% filter(Patient == 20 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 20 (P)' )
p6 = mtn.spatial(globalCNdf %>% filter(Patient == 51 & seg.type != 'arms'), pg.samp, ogjLimits,'Patient 51 (NP)' )
ggsave(filename='plots/spatial/patient_20_P.png', plot=grid.arrange(p2$grob), width=10.5, height=4.5, units='in', dpi=300)
ggsave(filename='plots/spatial/patient_51_NP.png', plot=grid.arrange(p6$grob), width=10.5, height=4.5, units='in', dpi=300)

grid.arrange(p2$grob, p6$grob, ncol=1)

# for (i in 1:nrow(sum.patient.data)) {
#   print(i)
#   pt = sum.patient.data[[i, 'Patient']]
#   status = as.character(sum.patient.data[[i, 'Status']])
# 
#   pms = mtn.spatial(subset(globalCNdf, Patient == pt & seg.type != 'arms'), pg.samp, ogjLimits, paste('Patient', pt, '(', status, ')' ))
# 
#   fl = paste0('patient_', pt, '_',status,'.png')
#   ggsave(paste('plots/spatial/', fl, sep=''), plot=grid.arrange(pms$grob), width=10.5, height=4.5, units='in', dpi=300)
# }

pg.samp = pg.samp %>% dplyr::select(-ogj)

ggsave('plots/spatial/legend.png', plot=grid.arrange(p2$legend), width=2,height=2, units='in', dpi=300)
```

## How do the predictions look at each time point?

```{r, echo=F, warning=F, message=F, fig.height=10, fig.width=12}
p53Path = pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
  dplyr::group_by(Pathology,p53.Status) %>% 
  dplyr::summarise(Freq = length(Pathology)) %>% dplyr::mutate(Status = 'p53 IHC')

p53Path = full_join( pg.samp %>% dplyr::filter(Status == 'P' & !is.na(p53.Status)) %>% 
                       dplyr::group_by(Pathology) %>% 
                       dplyr::summarise(nPath = length(Pathology)), p53Path, by='Pathology') %>% 
  dplyr::mutate(ratio = round(Freq/nPath,2))

rp = plot.risk.by.path(pg.samp)
rp
ggsave(filename='plots/pred_by_path.png', plot=(rp + theme(legend.position = 'none')), width=7, height=6, units='in', dpi=300)
```

```{r p53IHC, echo=F, warning=F, message=F, fig.height=8, fig.width=8}
#p53Path$Risk[p53Path$Risk == 'High'] = 'Aberrant'
#p53Path$Risk[p53Path$Risk == 'Low'] = 'Normal'

#p53Path$p53.Status = factor(p53Path$p53.Status, levels=c('Aberrant', 'Normal'), ordered = F)

p53Path$p53.Status = factor(p53Path$p53.Status, ordered = T)
p53Path =arrange(p53Path, Pathology, p53.Status)
col = brewer.pal(11, "PRGn")[c(10,2)]
rp = ggplot(p53Path, aes(Pathology, ratio*100, group=Pathology, fill=p53.Status)) + geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=col, name='p53 IHC', labels=c('Normal','Aberrant')) + 
  geom_text(aes(label=paste('n=',nPath,sep=''), x=Pathology, y=101), nudge_y=1) +
  geom_text(aes(label=paste(ratio*100,'%',sep='')), position=position_stack(vjust = 0.5), size=5,color='white') +
  plot.theme + labs(title='Samples predicted by aberrant p53 IHC', y='% Predicted', x='Pathology') + 
  theme(legend.position = 'bottom', text = element_text(size=14))
rp
ggsave(filename='plots/p53_pred.png', plot=rp, width=6, height=5, units='in', dpi=300)
```

```{r, echo=F, warning=F, message=F, fig.width=10, fig.height=6}
all.preds = pg.samp %>% dplyr::group_by(Status, Patient, Risk) %>% tally %>% spread(Risk, n) %>% ungroup %>%
  mutate_if(is.numeric, funs( ifelse(is.na(.), 0, .) )) %>% 
  left_join(pg.samp %>% dplyr::group_by(Status, Patient) %>% tally, by=c('Status','Patient')) %>%
  dplyr::rename('n.samples'='n') %>% 
  dplyr::mutate(Patient = factor(Patient))

all.preds = melt(all.preds, id.vars=c('Status','Patient', 'n.samples'))

ggplot(all.preds, aes(Patient, value/n.samples, group=variable, fill=variable)) + 
  geom_bar(stat='identity', position = position_stack()) +
  scale_fill_manual(values=riskCols) + geom_text(aes(y=1.01, label=n.samples)) +
  facet_wrap(~Status, scales='free_x') + labs(y='Sample ratios', title='Per patient ratio of predictions') + 
  plot.theme
```

## Backwards from HGD

Clinicians always ask to see the timepoints peeled backwards to see how early prediction is possible.

```{r endos, warning=F, message=F, echo=F, fig.height=8, fig.width=6}
sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)

endos = pg.samp %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% 
  dplyr::summarise('pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
              'n.endos'=length(unique(PID)), 
              'n.samples'=length(PID),
              'pred.samples'=length(which(Risk == 'High')),
              'max.path'=max(Pathology),
              'max.prediction' = max(Prediction)
              ) %>% ungroup %>% 
  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T,
     breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>% 
  dplyr::mutate(brks = factor(brks, rev(levels(brks))))

endos = endos %>% mutate(brks = recode_factor(brks, '96-120' = '>96', '>120'='>96'))

endos.m = endos %>% dplyr::filter(Status == 'P') %>% dplyr::group_by(brks) %>% 
  dplyr::summarise( 'pred.r'=sum(pred.endo)/sum(n.endos), 'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos),'mean(P)'=mean(max.prediction),'sem(P)'=sd(max.prediction)/sqrt(length(brks)))

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
  geom_errorbar(color='grey55', width=0.2) +
  geom_text(aes(label=paste(round(pred.r*100), '%', sep='')), color='black', nudge_y=2) +
  geom_text(aes(y=5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white', nudge_y=2) + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12) ) 
rp
ggsave(filename='plots/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)

```

Using the discretized risk what is the AUC?

```{r}
risk_df = pg.samp %>% dplyr::select(Status, Risk) %>% mutate( Risk  = factor(droplevels(recode_factor(Risk, 'Moderate' = 'Low')), levels=c('Low','High'), ordered=T) )

rocRisk = pROC::roc(Status ~ Risk, data=risk_df, ci=T, of='thresholds', transpose = T)
p1 = roc.plot(rocRisk) + labs(title='ROC based on discretized risk', subtitle=paste(levels(risk_df$Risk), collapse=', '))
p1
ggsave(filename='plots/risk_auc.png', plot=p1, width=4, height=4, units='in', dpi=300)

#rocRisk2 = pROC::roc(Status ~ Risk, data=pg.samp, ci=T, of='thresholds', transpose = T)
#p2 = roc.plot(rocRisk2) + labs(title='ROC based on discretized risk', subtitle=paste(levels(pg.samp$Risk), collapse=', '))
#p2

#grid.arrange(p1,p2, ncol=1)
```

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5 }
pg.samp = pg.samp %>% dplyr::mutate(nl = factor(nl, ordered=T), 
                                    Risk = factor(Risk, levels=c('Low','Moderate','High'), ordered = T), 
                                    Patient = factor(Patient, ordered=T), 
                                    Block = factor(Block, ordered=T))

#blank.axis = theme(axis.text = element_blank(), legend.key=element_rect(fill='grey39'), panel.background=element_rect(colour = 'black'), panel.grid.major=element_blank(), panel.spacing = unit(0.2, 'lines'), panel.border = element_rect(color="black", fill=NA, size=0.5)  )

p = BarrettsProgressionRisk::patientRiskTilesPlot(pg.samp %>% filter(Patient == 6) %>% dplyr::select(Endoscopy.Year,months.before.final, Pathology,Block,Risk) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block), 'months.before.final' ) 
tilesLegend = get.legend(p)
p + theme(legend.position = 'none')

ggsave('plots/heatmaps/legend_long.png',plot=grid.arrange(tilesLegend), width=10, height=2, units='in', dpi=300)

dir.create('plots/heatmaps/tiny/P',recursive = T, showWarnings = F)
dir.create('plots/heatmaps/tiny/NP',recursive = T, showWarnings = F)
# Tiny heatmaps
plist = purrr::map(sum.patient.data$Patient, function(pt) {
  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup

  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='',title='') + theme(axis.ticks = element_blank(), axis.text = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm"))

  ht = ifelse(length(unique(patient$GEJ.Distance)) > 1, length(unique(patient$GEJ.Distance)), length(unique(patient$GEJ.Distance))+0.5)
  wd = ifelse(length(unique(patient$months.before.final)) > 1, length(unique(patient$months.before.final)), length(unique(patient$months.before.final))+0.5 )

  ggsave(paste0('plots/heatmaps/tiny/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=ht, width=length(unique(patient$months.before.final)), units='in', dpi=150, limitsize=F)

  patient = pg.samp %>% filter(Patient == pt) %>% dplyr::select(Patient,Status,Pathology,Endoscopy.Year,Block,Risk,months.before.final) %>% dplyr::rename(Endoscopy = Endoscopy.Year, GEJ.Distance = Block) %>% ungroup

    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='months',y='location') + theme(plot.title = element_text(face = 'bold', size=16))

    ggsave(paste0('plots/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F)

    return(p)
})
names(plist) = sum.patient.data$Patient

```

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=15}
a = do.call('arrangeGrob', c(plist[c('13','20','34','56','60','48')], top='Progressors'))
b = do.call('arrangeGrob', c(plist[c('15','16','21','3','45','19')], top='Non-progressors'))

ggsave('plots/ex_pt_heatmap.png', plot=grid.arrange(a,b,ncol=2), width=12, height=15, units='in', dpi=300)

table.path.risk<-function(df) {
  summary.t = as.data.frame.matrix(with(df, table(Pathology, Risk)))
  summary.t$Pathology = rownames(summary.t)

  summary.t = summary.t %>% dplyr::group_by(Pathology) %>% dplyr::mutate( 'High.r'=round(High/sum(High,Low,Moderate),3), 
                                                'Low.r'=round(Low/sum(High,Low,Moderate),3),
                                                'Moderate.r'=round(Moderate/sum(High,Low,Moderate),3))
  for (g in unique(df$Risk)) {
    for (i in 1:nrow(summary.t)) {
      summary.t[i,g] = paste(summary.t[i,g], ' (', summary.t[i,paste(g,'.r',sep='')]*100, '%)',sep='')
    }
  }
  summary.t[,c('Pathology',levels(df$Risk))]
}

summary.t = table.path.risk(pg.samp)
#summary.t %>% kableExtra::kable() %>% kableExtra::kable_styling('striped', full_width = F)
write_tsv(summary.t, path = 'plots/risk_summary.tsv')
```

```{r, echo=F, message=F, warning=F}
pg.samp = pg.samp %>% dplyr::group_by(Patient) %>% dplyr::mutate( 'pred.ratio'=round(sum(Prediction)/length(Prediction),2))

p1 = do.call('arrangeGrob', c(plist[as.character(filter(sum.patient.data,Status == 'P')$Patient)], top='Progressors'))
p2 = do.call('arrangeGrob', c(plist[as.character(filter(sum.patient.data,Status == 'NP')$Patient)], top='Non-Progressors'))

ggsave('plots/heatmaps/pred_heatmap.png', plot=grid.arrange(p2,p1,nrow=2), width=52, height=16, units="in", limitsize=F, dpi = 300)
ggsave('plots/heatmaps/pg_heatmap.png', plot=p1, width=52, height=24, units="in", limitsize=F, dpi = 300)
ggsave('plots/heatmaps/npg_heatmap.png', plot=p2, width=52, height=24, units="in", limitsize=F, dpi = 300)

#ggsave('plots/risk_legend.png', plot=grid.arrange(riskLegend), width=2,height=4, dpi=300)
#ggsave('plots/risk_legend_long.png', plot=grid.arrange(riskLegendLong), width=8,height=2, dpi=300)

```

# Chr17

## p arm
```{r, echo=T}

p11 = filter(globalCNdf, chr == 17 &  start == 1 & seg.type == 'arms')
p11Loss = p11 %>% filter(value < -1)

p11Loss %>% filter(value < -1) %>% group_by(Status) %>% tally(name='p11 Loss') %>% full_join(
  p11 %>% group_by(Status) %>% tally(name='Total Samples') ) %>% 
  kable(caption='Chr17 p11 loss') %>% kable_styling(full_width = F)

p11Loss %>% group_by(Status, Hospital.Research.ID) %>% tally

p11Path = patient.info %>% filter(Samplename %in% p11Loss$variable) %>% dplyr::group_by(Pathology) %>% tally(name='samples')

p11Path %>% dplyr::mutate('ratio'=round(samples/sum(p11Path$samples),2)) %>% 
kable(caption='Ratio of all Chr17 p11 loss samples by pathology' ) %>% kable_styling(full_width = F)
```

Samples with a loss vs all Progressor samples
Ratio of samples with p11 loss vs all progressor samples: `r length(unique(p11Loss$variable))/length(unique(filter(globalCNdf, Status == 'Progressor')$variable))`

Ratio of progressor patients with a p11 loss in at least 1 sample: `r length(unique(p11Loss$Patient))/length(subset(sum.patient.data, Status == 'P')$Patient)`

```{r}
patient.info %>% filter(Samplename %in% p11Loss$variable) %>% group_by(forcats::fct_explicit_na(p53.Status)) %>% tally %>% 
  set_names(c('p53 Status', 'n samples')) %>% kable(caption='p53 status for samples with p11 loss') %>% kable_styling(full_width = F)

pi = patient.info
pi$p11Loss = as.numeric(pi$Samplename %in% p11Loss$variable)
#plot(table(p11=pi$p11Loss, ihc=pi$p53.Status))

cor.test(pi$p11Loss, (as.numeric(pi$p53.Status)-1))
```

## Top Features

## Chr 17 Predictive TODO

One way of looking at the p53 prediction (e.g. Adam Bass) is to see how predictive the p-arm loss in chr17 is.
```{r, echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
ccgenes = read_tsv('~/Data/CosmicCensusGenes.tsv', col_types = cols(.default = col_character())) %>% 
  filter(!grepl('^X|Y',`Genome Location`)) %>% 
  separate(`Genome Location`, c('chr','start','end'),':|-', remove=F) %>%
  dplyr::filter(start != '' | end != '') %>%
  dplyr::select(`Gene Symbol`, `Genome Location`, `Role in Cancer`, `Mutation Types`, chr, start, end)

genes = ccgenes %>% dplyr::filter(grepl('TSG|oncogene', `Role in Cancer`) | `Mutation Types` != 'T') %>% 
  dplyr::select(`Gene Symbol`, `Genome Location`, chr, start, end) %>%
  dplyr::mutate(value = min(globalCNdf$value), Patient = '', chr = factor(chr,levels=levels(chr.lengths$chr),ordered=T), start = as.numeric(start), end=as.numeric(end)) %>% arrange(chr, start)

genesGR = makeGRangesFromDataFrame(genes, keep.extra.columns = T)
chGR = makeGRangesFromDataFrame(features, keep.extra.columns = T)

#genes = as.data.frame(genesGR[unique(subjectHits(ov))])
#colnames(genes)[1] = c('chr')

g17 = plot.cn.chr(globalCNdf, chrom='17', chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) + 
  labs(title="Chromosome 17") + scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
  theme(legend.position = 'none', strip.text = element_blank(), text=element_text(size=14))

ggsave(filename='plots/hazards/chr17.png', plot=g17, width=8, height=4, units='in', dpi=300)

```

```{r echo=F, warning=F, message=F, fig.height=5, fig.width=10, eval=F}
ov = findOverlaps(chGR, genesGR)
chGR$Genes = ''
for (hit in unique(queryHits(ov))) {
  chGR[hit]$Genes = paste(genesGR[subjectHits(ov[queryHits(ov) == hit])]$Gene.Symbol, collapse=',')
}

for (chr in unique(features$chr)) {  
  print(chr)
  p = plot.cn.chr(globalCNdf,  chrom=chr, info=chr.lengths, haz=features, genes=genes, label=labeller(chr=label_both, Status=n.status)) +
    scale_y_continuous(breaks = seq(-30,25,10), labels = c('-30', '','',0,'',20)) +
    theme(strip.text.x = element_blank(), text=element_text(size=14)) + labs(title=paste('Chromosome',chr))
  f = paste('chr',chr,'.png',sep='')
  ggsave(paste('plots/hazards/',f,sep=''), plot=p, width=8, height=4, units='in', dpi=300)
}

annoG = ccgenes %>% filter(grepl('CDKN2A$|TP53|MYC$|SMAD4|VEG|ERBB2|EGFR|SMYD|ARID1A|FHIT|RUNX1', `Gene Symbol`)) %>% arrange(chr) %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  mutate(ymin = round(range(globalCN.plot$data$mean.value))[1]/2, ymax = round(range(globalCN.plot$data$mean.value))[2]/2) %>%
  mutate(chr = factor(chr, levels=levels(chr.lengths$chr)), start = as.numeric(start), end = as.numeric(end))

p = globalCN.plot + #geom_rect(data=annoG, aes(xmin=start, xmax=end, ymin=ymin,ymax=ymax), color='red') +
  geom_point(data=annoG, aes(x=start, y=0), size=1.5, color='red') +
  geom_text_repel(data=annoG, aes(x=start, y=ymax/2, label=`Gene Symbol`), color='red', size=1.5 ) + theme(legend.position = 'none')

ggsave('plots/global_gene_ann.png', plot=p, width=10, height=5, units='in', dpi=300)

#(genes[,c(1:3,6,7)]) %>% kable() %>% kable_styling(full_width=F)
```

```{r, echo=F, message=F, warning=F, fig.height=5, fig.width=8, eval=F}
file = paste(cache.dir, 'chr17.predictive.Rdata', sep='/')
if (file.exists(file)) {
  message(paste("loading", file))
  load(file, verbose=T)
} else {
  stop(paste("File missing", file))
}

names(performance.at.1se)

mp = model.performance(performance.at.1se, coefs) 
mp$plot = mp$plot + labs(x='Models', title='Model performance for subsets of genomic regions') + 
  scale_colour_manual(values=c('grey39', 'grey39')) + theme(axis.text.x = element_text(angle=45,hjust=1)) +
  geom_hline(yintercept=mean(mp$performance$mean)+sd(mp$performance$mean), color='red', alpha=0.5) +
  geom_hline(yintercept=mean(mp$performance$mean)-sd(mp$performance$mean), color='red', alpha=0.5)
mp$plot
ggsave('plots/performance/performance_subsets.png', plot=mp$plot, height=5, width=9, units='in',dpi=300)

chr.info = read.table('hg19_genes.txt', sep='\t', header=T)
chr.info = chr.info[1:22,]

chr.info$info.content = chr.info$protein_coding/chr.info$length

# No correlation between information content (genes/length) and classification rate 
cor.test(mp$performance[1:22,'mean'], chr.info$info.content)

median(mp$performance$mean)
sd(mp$performance$mean)
mp$performance[17,'mean']

```

# Validation Cohort

```{r, echo=F}
#p.diag = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/QDNAseq/validation/progressor_diagnosis.xlsx') %>% 
#  dplyr::mutate(`Hospital Research ID` = gsub('/', '_', `Hospital Research ID`), `Weeks Follow Up` = as.double(difftime(`Diagnostic Endoscopy`, `Sequenced Endoscopy`, units='weeks')))
pastefun<-function(x) {
  if ( !grepl('SLX-', x) ) x = paste0('SLX-',x)
  return(x)
}

patient.file = '~/Data/Ellie/QDNAseq/validation/sWGS_validation_batches.xlsx'

oac.info = readxl::read_xlsx(patient.file, 'OAC Test Samples') %>% 
  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
  dplyr::select(-matches('SLX|Index')) %>%
  mutate(Status = factor(Status, levels=levels(pg.samp$Status)), Block = factor(Block, levels=c('E','1','2'), ordered=T))

val.info = readxl::read_xlsx(patient.file, 'Final Validation Samples') %>% 
  dplyr::mutate_at(vars(`SLX-ID`, `Block ID`), list(as.character)) %>% dplyr::filter(!is.na(`SLX-ID`)) %>%
  mutate(Samplename = paste(`SLX-ID`, `Index Sequence`, sep='.')) %>%
  separate(`Block ID`, c('PID','Block'), sep='-|[:blank:]', remove=F) %>% dplyr::rename(PathID = `Block ID`) %>%
  dplyr::select(-matches('SLX|Index'), -`Path Notes`) %>%
  mutate(Status = factor(Status, levels=levels(pg.samp$Status))) %>% ungroup %>% group_by(`Hospital Research ID`) %>% 
  dplyr::mutate(`Weeks Follow Up` = as.double(difftime(`Last Endoscopy`, `Endoscopy`, units='weeks')))

val.info$Block = sapply(strsplit(val.info$Block, ''), function(x) {
  if (is.na(x)) return(1)
  if (length(x) == 1) {
    if (!grepl('\\d', x)) return(recode(x, 'A'=1, 'B'=2, 'C'=3, 'D'=4, 'E'=5, 'F'=6,'G'=7 ))
    return(as.numeric(x))
  }
  if (length(x) > 1) {
    i = which(grepl('\\d', x))
    return(as.numeric(x[i]))
  }
})

bind_rows(val.info %>% dplyr::select(`Hospital Research ID`, Status) %>% distinct %>% group_by(Status) %>% tally %>% 
  spread(Status,n) %>% add_column(' ' = 'n Patients', .before=1),
val.info %>% group_by(Status) %>% tally %>% spread(Status,n) %>% add_column(' ' = 'n Samples', .before=1)) %>% 
  kable(caption='Validation cohort') %>% kable_styling(full_width = F)
```

```{r, eval=F}
file = paste(cache.dir, 'model_data.Rdata', sep='/')
load(file, verbose=T)

vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb'
v.tiles = do.call(bind_rows, purrr::map(list.files(vdir, '5e06_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
})) %>% dplyr::rename(Samplename = 'X1')
v.tiles = filter(v.tiles, Samplename %in% val.info$Samplename)
v.tiles = v.tiles[-which(duplicated(v.tiles$Samplename))[2],]

v.arm.tiles = do.call(bind_rows, purrr::map(list.files(vdir, 'arms_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
})) %>% dplyr::rename(Samplename = 'X1')
v.arm.tiles = filter(v.arm.tiles, Samplename %in% val.info$Samplename)
v.arm.tiles = v.arm.tiles[-which(duplicated(v.arm.tiles$Samplename))[2],]

for (i in 1:length(z.mean))
  v.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.tiles[,i+1], z.mean[i], z.sd[i])

for (i in 1:length(z.arms.mean))
  v.arm.tiles[, i+1] = BarrettsProgressionRisk:::unit.var(v.arm.tiles[,i+1], z.arms.mean[i], z.arms.sd[i])

vt = dplyr::select(v.tiles, -Samplename) %>% as.matrix; rownames(vt) = v.tiles$Samplename
at = dplyr::select(v.arm.tiles, -Samplename) %>% as.matrix; rownames(at) = v.arm.tiles$Samplename

v.df = BarrettsProgressionRisk:::subtractArms(vt, at)

valCNdf = as_tibble(t(v.df), rownames='locs') %>% set_names(c('locs',val.info$Samplename)) %>%
  separate(locs, c('chr','start','end'), sep=':|-', remove=F) %>%   
  dplyr::mutate_at(vars(start, end), as.numeric) %>%
  #dplyr::mutate( seg.type = '5MB' ) %>%
  dplyr::mutate( seg.type = ifelse( (start %in% arm.loc$start & end %in% arm.loc$end), 'arms','5MB'  ) ) %>%
  melt(id.vars=c('locs','chr','start','end', 'seg.type')) %>% as_tibble() %>%
  left_join(chr.lengths %>% dplyr::select(chrom,chr.length), by=c('chr'='chrom')) %>%
  left_join(val.info %>% dplyr::select(matches('Hospital|Patient|Samplename|Pathology|Status')), by=c('variable'='Samplename')) %>%
  dplyr::mutate(Status = factor(recode(Status, NP='Non-Progressor',P='Progressor'), ordered=T)) %>% 
  dplyr::mutate( 
    cn = factor(unlist(purrr::map(.x=value, .f=get.cn.state)),  levels=c('loss','neutral','gain'), ordered = T),
    chr = factor(chr, levels = chr.lengths$chrom, ordered=T))

n.status <- c('Progressor'=paste0('Progressor (n=',length(unique(subset(valCNdf, Status == 'Progressor')$variable)), ')'),
              'Non-Progressor'=paste0('Non-Progressor (n=',length(unique(subset(valCNdf, Status == 'Non-Progressor')$variable)),')'))

#cn.mtn.plot(filter(valCNdf, seg.type == '5MB'), labeller(Status=n.status, chr=label_value)) + ylim(-30,20)

```

## Per sample (alpha = 0.9)

```{r, fig.width=8, fig.height=12}
vdir ='~/Data/BarrettsProgressionRisk/Analysis/validation/pcf_perPatient/50kb/predictions_5e6_all'

residuals = do.call(bind_rows, purrr::map(list.files( dirname(vdir), 'residuals.tsv', full.names = T, recursive = T ), function(f) {
  read_tsv(f, col_types = 'ccddddl')
}))

failedQC = residuals %>% filter(varMAD_median > 0.008)

all.val.preds = do.call(bind_rows, purrr::map(list.files(vdir, 'predictions.tsv', full.names = T, recursive = T), function(f) {
    pd = read_tsv(f, col_types = 'cddcccDcccc') %>% dplyr::select(`Hospital Research ID`, Sample, Probability, `Relative Risk`)
    return(pd)
})) %>% mutate(quants = cut(Probability, breaks=cuts, include.lowest = T)) %>%
  left_join(pred.confidence %>% dplyr::select(quants, Risk), by='quants')

oac.preds = all.val.preds %>% filter(`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`))

all.val.preds = all.val.preds %>% filter( !Sample %in% failedQC$samplename &  
                                            !`Hospital Research ID` %in% unique(oac.info$`Hospital Research ID`) )

val.preds = left_join(all.val.preds, val.info, by=c('Hospital Research ID', 'Sample'='Samplename')) %>% dplyr::select(Patient, everything())

val.preds = val.preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','GM','ID','LGD'), ordered=T)) %>% 
  mutate(Final.Endoscopy = `Last Endoscopy`, Endoscopy.Year = Endoscopy, Block = factor(Block, ordered=T))

```

```{r}
roc = pROC::roc(Status ~ Probability, data=val.preds, auc=T, ci=T, of='thresholds',transpose=T)
roc.plot(roc) + labs(title='Validation cohort')
ggsave('plots/validation/validatation_roc.png', plot=roc.plot(roc) + labs(title='Validation cohort'), width=5, height=5, units='in',dpi=300)

totals = val.preds %>% group_by(Status) %>% tally
val.counts = val.preds %>% 
    group_by(Status, Risk) %>% tally %>% spread(Risk,n) %>% ungroup %>%
    left_join(totals, by='Status') %>% mutate_if(is.numeric, funs(./n)) %>% dplyr::select(-n)

ggplot(melt(val.counts, id.vars=c('Status')), aes(variable, value, group=Status, fill=Status)) + ylim(0,1) + 
    geom_bar(stat='identity', position='dodge') + geom_text(aes(label=round(value,2)), position = position_dodge(width=1)) +
    labs(title='Validation risk (sample ratios per NP:P)', x='', y='ratio') + theme_minimal()
```

## By path

```{r}
rp = plot.risk.by.path(val.preds) + labs(title='Validation samples predicted by pathology')  
rp
ggsave(filename='plots/validation/val_pred_by_path.png', plot=rp, width=5, height=6, units='in', dpi=300)

get.year<-function(x) {
  as.numeric(format(as.Date(x), '%Y'))
}

val.preds = dplyr::rename(val.preds, 'Samplename'='Sample')

val.preds = do.call(bind_rows, lapply(unique(val.preds$Patient) %>% 
                            map( function(p) val.preds %>% filter(Patient == p) %>% 
                                   mutate_at(vars(Endoscopy.Year, Final.Endoscopy), list(~get.year(.)) )
                                 ), transform.by.time, info=val.info) )

dir.create('plots/validation/heatmaps/P/tiny',recursive = T, showWarnings = F)
dir.create('plots/validation/heatmaps/NP/tiny',recursive = T, showWarnings = F)

for (pt in unique(val.preds$Patient)) {
#print(pt)
  patient = val.preds %>% filter(Patient == pt) %>% 
    dplyr::select(Patient, Status, Endoscopy, months.before.final, Block,Risk) %>% 
    dplyr::rename(GEJ.Distance = Block)

    p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final','rev') + theme(legend.position = 'none') + labs(x='',y='',title='') + theme(axis.ticks = element_blank(), axis.text = element_blank(), plot.margin=grid::unit(c(0,0,0,0), "mm"))

    ht = ifelse(length(unique(patient$GEJ.Distance)) > 1, length(unique(patient$GEJ.Distance)), length(unique(patient$GEJ.Distance))+0.5)
    wd = ifelse(length(unique(patient$months.before.final)) > 1, length(unique(patient$months.before.final)), length(unique(patient$months.before.final))+0.5 )

    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/tiny/pt_',pt,'.png'), plot = p, height=ht, width=length(unique(patient$months.before.final)), units='in', dpi=150, limitsize=F)

    patient = val.preds %>% filter(Patient == pt) %>% 
    dplyr::select(Patient, Status, Endoscopy, months.before.final, Pathology,Block,Risk) %>% 
    dplyr::rename(GEJ.Distance = Block)

  p = BarrettsProgressionRisk::patientRiskTilesPlot(patient, 'months.before.final', 'rev') + labs(x='months', y='location') + theme(legend.position = 'none',plot.title = element_text(face = 'bold', size=16))

    ggsave(paste0('plots/validation/heatmaps/',unique(patient$Status),'/pt_',pt,'.png'), plot = p, height=length(unique(patient$GEJ.Distance))+1, width=length(unique(patient$months.before.final))+1, units='in', dpi=300, limitsize=F) 
}

sq = c(0, 1, 1*12, 2*12, 4*12, 6*12, 8*12, 10*12, 15*12)
val.preds = val.preds %>% mutate(months.before.final = ifelse(months.before.final > sq[length(sq)], sq[length(sq)], months.before.final))

v.prog = val.preds %>% filter(Status == 'P') %>% mutate(
  brks = cut(months.before.final, breaks=sq, 
                 labels=c(c('Endpoint',apply(cbind(sq[-1], sq[-1]+11), 1, function(x) paste(x, collapse='-')))[-c(length(sq)-1,length(sq))], '>120'), include.lowest = T, ordered_result = T),
) %>% dplyr::mutate(brks = factor(brks, rev(levels(brks))))

endos = v.prog %>% dplyr::group_by(Status, months.before.final, brks, PID) %>% 
  dplyr::summarise('pred.endo'=ifelse(length(which(Risk == 'High')) > 0, 1, 0), 
              'n.endos'=length(unique(PID)), 
              'n.samples'=length(PID),
              'pred.samples'=length(which(Risk == 'High')),
              'max.path'=max(Pathology),
              'max.prediction' = max(Probability)
              ) %>% ungroup %>% 
  dplyr::mutate(brks = cut(months.before.final, include.lowest = T, ordered_result = T, breaks=sq,  labels=c('Endpoint','1-12','12-24','24-48','48-72','72-96','96-120','>120'))) %>%
  dplyr::mutate(brks = factor(brks, rev(levels(brks))))

endos = endos %>% mutate(brks = recode_factor(brks, '96-120' = '>96', '>120'='>96'))

endos.m = endos %>% dplyr::filter(Status == 'P') %>% dplyr::group_by(brks) %>% 
  dplyr::summarise( 'pred.r'=sum(pred.endo)/sum(n.endos), 'n.samp'=sum(n.samples), 'n.endo'=sum(n.endos),'mean(P)'=mean(max.prediction),'sem(P)'=sd(max.prediction)/sqrt(length(brks)))

rp = ggplot(endos.m, aes(x=brks, y=pred.r*100, ymin=(pred.r-`sem(P)`)*100, ymax=(pred.r+`sem(P)`)*100, group=brks)) +
  geom_bar(stat='identity', fill=BarrettsProgressionRisk:::riskColors()[['High']], show.legend = F) + 
  geom_errorbar(color='grey21', width=0.2) +
  geom_text(aes(y=3, label=paste(round(pred.r*100), '%', sep='')), color='white') +
  geom_text(aes(y=pred.r*100-5, label=paste('n=',n.endo,'\ns=',n.samp,sep='')), color='white') + 
  labs(x='Months prior to endpoint', y='Percent predicted', title='Predictions prior to final endoscopy', subtitle='n = #endoscopies, s = #samples') + coord_cartesian(ylim = c(0,100)) + 
  plot.theme + theme( axis.text.x=element_text(angle=45,hjust=1), text=element_text(size=12)) 
rp
ggsave(filename='plots/validation/pred_endo.png', plot=rp, width=6, height=5, units='in', dpi=300)
```

# Clinical Story

## Vignettes

```{r, echo=F, warning=F, message=F, fig.width=5, fig.height=5}
sum.patient.data = sum.patient.data %>% mutate(smoking_str = recode(smoking, 'Y'='smoker', 'N'=''))
sum.patient.data$smoking_str = replace_na(sum.patient.data$smoking_str, '')

create.desc<-function(pt,age,sex,smk,len,date) {
  str =  paste0('Patient ', pt, ': ', age, sex)
  if (smk != '' & !is.na(smk)) str = paste0(str, ' ,', smk)
  str = paste0(str,'\n')

  if (len != '' & !is.na(len)) str = paste0(str, len,'cm length ')

  str = paste0(str, 'BE diagnosed ', date)

  return(str)
}
sum.patient.data$desc = sum.patient.data %>% group_by(Patient) %>% rowwise() %>% dplyr::summarise(desc = create.desc(Patient, age.diagnosis, gender, smoking_str, M, initial.endo))

pt = 34
p34 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p34, width=6.35, height=4.5,dpi=72)

pt = 3
p3 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p3, width=6.35, height=4,dpi=72)

pt = 68
p68 = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) + 
    theme(legend.position = 'none', text = element_text(size=12))
ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p68, width=6.35, height=5,dpi=72)

for (pt in c(56,20,86,23,33,49,51,6,25)) {
  length = sum.patient.data %>% ungroup %>% filter(Patient == pt) %>% dplyr::summarise(years = final.endo-initial.endo) %>% pull
  height = as.numeric(max( pg.samp %>% ungroup %>% filter(Patient == pt) %>% dplyr::select(Block) %>% pull ) ) + 1

  #length = 6
  #height = 5

  p = vig.plot(df=subset(pg.samp, Patient == pt),info= subset(patient.info, Patient == pt),cols=riskCols) + 
    theme(legend.position = 'bottom') + labs(title = filter(sum.patient.data, Patient == pt)$desc) 
  vigLeg = get.legend(p)
  p = p + theme(legend.position = 'none', text = element_text(size=12))
  ggsave(paste0('plots/vignette/vig_',pt,'.png'), plot=p, width=length, height=height,dpi=300)
}
ggsave('plots/vignette/legend.png', vigLeg, width=8, height=1, dpi=300)

```

### Recommendations

 1.	Immediate RFA: HGD diagnosis, two or more consecutive high risk predictions
 2.	Recheck 6-12 months: One high risk prediction or aberrant p53 IHC
 3. Recheck endoscopy 12-24 months: One or more moderate risk predictions
 4. Regular surveillance 3-5 years: Two or more consecutive low risk predictions

```{r, rx, echo=F, warning=F, message=T, fig.width=8, fig.height=5}

rx<-function(Patient, brks, Prediction, Pathology, Risk, p53.Status) {
  rules = as.data.frame(matrix(ncol=0, nrow=0))
  # Consecutive
  for (i in 1:length(brks)) {
    risks = table(Risk[i:(i+1)])
    p53 = table(p53.Status[i:(i+1)])
    rule = 'None'
    if ( risks['High'] == 2 | length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0 ) {
      rule = 1
    } else if ( risks['High'] == 1 | p53['1'] > 0 ) {
      rule = 2
    } else if ( risks['Moderate'] > 0 ) {
      rule = 3
    } else if ( risks['Low'] == 2) {
      rule = 4
    }
  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
  }  
  colnames(rules) = c('t1','t2', 'rule')
  rules$Patient = Patient
  rules
}

pg.samp = pg.samp %>% mutate(p53.Status = factor(p53.Status, levels=c(0,1), ordered=T))

maxP = pg.samp %>% dplyr::group_by(Patient, Status, brks) %>% 
  dplyr::summarise( 'maxP'=max(Prediction), 'Risk'=max(Risk), 'Pathology'=max(Pathology), 'p53'=max(p53.Status,na.rm=T) )

rules = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
  r = with(arrange(subset(maxP, Patient == pt), brks), rx(pt, brks,maxP,Pathology,Risk,p53))
  r$Status = unique(subset(maxP, Patient == pt)$Status)
  r
}))

rules$t1 = factor(rules$t1, levels=levels(pg.samp$brks), ordered=T)
rules$t2 = factor(rules$t2, levels=levels(pg.samp$brks), ordered=T)

penult = filter(rules, t2 != 'Endpoint' & (t1 == rev(levels(pg.samp$brks))[2] | t2 == rev(levels(pg.samp$brks))[2] ))

penult %>% filter(Status == 'P' & rule %in% c(3,4)) %>% kable(caption='Patients that would be missed if we moved Moderate to Low risk') %>% kable_styling(full_width = F)

penult %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  left_join(penult %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
  kable(caption='Penult Endoscopy: % patients by rx rules based on the max (P) per endoscopy') %>% kable_styling(full_width=F)

endpoint = subset(rules,  t2 != 'Endpoint') %>% 
  dplyr::group_by(Patient, Status) %>% 
  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

endpoint %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  left_join(endpoint %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (P) per endoscopy') %>% kable_styling(full_width=F)

path.rx<-function(Patient, brks, Pathology) {
  rules = as.data.frame(matrix(ncol=0, nrow=0))
  # Consecutive
  for (i in 1:length(brks)) {
    #risks = table(Pathology[i:(i+1)])
    rule = 'None'
    if ( length(which(grepl('HGD|IMC', Pathology[i:(i+1)]))) > 0  |
         length(which(grepl('LGD', Pathology[i:(i+1)]))) > 1 ) {
      rule = 1
    } else if ( length(which(grepl('LGD', Pathology[i:(i+1)]))) > 0 ) {
      rule = 2
    } else if ( length(which(grepl('BE|NDBE|ID', Pathology[i:(i+1)]))) > 1  ) {
      rule = 3
    }
  rules = rbind(rules, cbind( as.character(brks[i]), as.character(brks[(i+1)]), rule))
  }  
  colnames(rules) = c('t1','t2', 'rule')
  rules$Patient = Patient
  rules
}

rules2 = do.call(rbind, lapply(unique(maxP$Patient), function(pt) {
  r = with(arrange(subset(maxP, Patient == pt), brks), path.rx(pt, brks,Pathology))
  r$Status = unique(subset(maxP, Patient == pt)$Status)
  r
}))
rules2$t1 = factor(rules2$t1, levels=levels(pg.samp$brks), ordered=T)
rules2$t2 = factor(rules2$t2, levels=levels(pg.samp$brks), ordered=T)

second = subset(rules2,  t2 != 'Endpoint') %>% 
  dplyr::group_by(Patient, Status) %>% 
  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

second %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  kable(caption='Patient recommendation by PATH alone in penult endoscopy' ) %>% kable_styling(full_width = F)

second %>% group_by(Status, rule) %>% tally %>% spread(rule,n) %>% 
  left_join(second %>% group_by(Status) %>% tally(name='total'), by='Status') %>%
  mutate_if(is.numeric, funs(  round((./total)*100,1)  )) %>% dplyr::select(-total) %>%
  kable(caption='Final endoscopy: % patients by recommendation rule based on the max (PATH) per endoscopy, not (RISK)') %>% kable_styling(full_width=F)

```