### 2019-12-24 snp/SNP_Prediction_Analysis.Rmd

---
title: "SNP Analysis"
author: "Sarah Killcoyne"
date: "6/13/2018"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F)
library(tidyverse)
library(gridExtra)
library(glmnet)
library(pROC)
#library(ggfortify)
library(BarrettsProgressionRisk)
library(pander)
library(kableExtra)
library(reshape2)

#source('~/workspace/shallowwgs_pipeline/lib/data_func.R')
source('~/workspace/shallowwgs_pipeline/lib/common_plots.R')

dir.create('plots/snp',recursive = T, showWarnings = F)

ggplot2::theme_set(theme_minimal())

chr.info = BarrettsProgressionRisk:::chrInfo(build='hg19')

model.dir = '~/Data/BarrettsProgressionRisk/Analysis/models_5e6_all/50kb'
load(paste0(model.dir,'/model_data.Rdata'), verbose = T)
load(paste0(model.dir,'/all.pt.alpha.Rdata'), verbose = F)

train.labels = labels
alpha = '0.9'
fitV = models[[alpha]]
l = performance.at.1se[[alpha]]$lambda
features = coefs[[alpha]]

swgs_labels = labels

patient.info = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/metadata_T1T2.xlsx') %>% 
  mutate( UniqueSampleID = paste(PatientID, `Timepoint Code`, sep='_'), Path.Status = Status, PatientID = as.character(PatientID))
#%>%  group_by(PatientID) %>% mutate(Path.Status = as.character(ifelse(length(which(grepl('IMC|HGD',Pathology))) > 0, 'P', Path.Status)))

path.np.ids = patient.info %>% filter(Pathology %in% c('HGD', 'IMC') & Status == 'NP') %>% dplyr::select(PatientID) %>% distinct %>% pull
patient.info = patient.info %>% mutate(Path.Status = ifelse(PatientID %in% path.np.ids, 'P', Path.Status))

patient.info %>% filter(PatientID == 524)

rm(plots,performance.at.1se,models,cvs,labels)

load('~/Data/Reid_SNP/PerPatient/allpts_ascat.Rdata', verbose=T)

qcdata = qcdata %>% as_tibble %>%  mutate(Samplename = sub('_$','',Samplename)) %>%
  separate(Samplename, c('PatientID','SampleID','EndoID','Level'), sep='_', remove=F)

qcdata[749, 'Level'] = 'BLD' # Making an assumption here but this is what it looks like

segments.list = lapply(segments.list, function(pt) {
  pt$sample = sub('\\.LogR','',pt$sample)
  pt %>% as_tibble
})

# swgs.seg.count = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb/segment_counts.tsv', col_types = 'dc')
# 
# # All segments
# seg.counts = do.call(bind_rows, purrr::map(segments.list, function(sl) {
#   sl %>% dplyr::mutate(Total = nMajor+nMinor) %>% group_by(sample, Total) %>% 
#     tally %>% #filter(Total != 2) %>% 
#     ungroup %>% group_by(sample) %>% dplyr::summarise(n = sum(n))
# }))

#plot(seg.counts$n)
#points(swgs.seg.count$`n segs`, col='red')

calc.ratio<-function(PatientID, Samplename,Ploidy) {
  smp = filter( segments.list[[ PatientID ]], sample == Samplename & chr %in% c(1:22) )
  smp = smp %>% dplyr::mutate(
      #'Total' = nAraw + nBraw,
      'Total' = nMajor + nMinor,
      'CNV' = round(Total) - round(as.numeric(Ploidy)) )
  x = filter(smp, CNV != 0 & chr %in% c(1:22))
  return(sum(x$endpos - x$startpos)/as.numeric(chr.info[22,'genome.length']))
}

qcdata = qcdata %>% mutate(
  ASCAT.SCA.ratio = purrr::pmap_dbl(list(PatientID, Samplename, Ploidy), .f=calc.ratio),
  Level = toupper(Level),
  SampleType = case_when(
    grepl('BLD|gastric', Level, ignore.case = T) ~ 'Blood/Gastric Normal',
    TRUE ~ 'BE'
))

sample.list = readxl::read_xlsx('~/Data/BarrettsProgressionRisk/Analysis/SNP/20180604_Reid_1M_SampleList.xlsx', sheet = 2) %>%
  dplyr::rename('Total.SCA' = 3) %>% mutate(SCA.Ratio = Total.SCA/max(chr.info$genome.length))

sample.info = dplyr::select(qcdata, c('PatientID','SampleID','EndoID','Level','Samplename','ASCAT.SCA.ratio'))

message(paste(length(unique(sample.info$PatientID)), 'patients listed in metadata file'))
message(paste(nrow(qcdata), 'samples available'))

sample.info = left_join(sample.info, dplyr::select(patient.info,c('PatientID','Timepoint','Timepoint Code','Status','Pathology')), by=c('PatientID','EndoID' = 'Timepoint Code')) %>%
  mutate(Status = factor(as.character( ifelse(is.na(Status) & Level %in% c('BLD','GASTRIC'), 'Normal', Status))))
head(sample.info)

get.ratio<-function(PID, SID) {
  subset(sample.list, PatientID == PID & SampleNum == SID)$SCA.Ratio
}

sample.info = sample.info %>% rowwise() %>% dplyr::mutate(
  SCA.Ratio = ifelse(Level %in% c('BLD','GASTRIC'), 0, get.ratio(PatientID, SampleID)), 
  Endoscopy = ifelse(Timepoint == 'T1', 1, 2)
)

ggplot(sample.info, aes(ASCAT.SCA.ratio, SCA.Ratio)) + geom_point() + facet_grid(~Status)

#seg.counts %>% dplyr::summarise_at(vars(-`2`), )
#seg.counts = left_join(seg.counts, dplyr::select(qcdata,c(Samplename, Ploidy, Purity, ASCAT.SCA.ratio, SampleType)), by=c('sample'='Samplename'))
# seg.counts %>% filter(ASCAT.SCA.ratio < 0.01 & Purity > 0.98)
# ggplot(filter(seg.counts, !is.na(Ploidy)), aes(n,ASCAT.SCA.ratio)) + facet_grid(~SampleType) + ylim(0,0.05) + xlim(0,200) +
#  geom_point(aes(color = ASCAT.SCA.ratio < 0.01 & Purity > 0.98)) + theme(legend.position = 'none')

ct = cor.test(sample.info$ASCAT.SCA.ratio, sample.info$SCA.Ratio)
```

# Brief processing notes

ASCAT was run on each SNP sample to obtain SCA calls. ASCAT and Xiahong's estimates show a `r round(ct$estimate,2)` correlation.

Prior to predicting the samples, in the step where I create 5Mb bins across the genome I merge the samples per endoscopy due to differences in the sampling. The sWGS had multiple biopsies aggregated per-level, while the SNP data was a single 'research' biopsy per level.

## Patient info

`r nrow(patient.info)` total endoscopies, `r patient.info %>% group_by(Status) %>% tally %>% spread(Status,n) %>% kable(caption="Split between P and NP") %>% kable_styling(full_width=F)`

Histology breakdown:

```{r, echo=F}
sample.info = sample.info %>% mutate(Pathology = recode(Pathology, IMC = 'EAC'))

sm = sample.info %>% group_by(Status, Endoscopy, Pathology) %>% dplyr::tally() %>% 
  group_by(Status, Endoscopy) %>% dplyr::mutate(freq=n/sum(n) ) %>% 
  mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','EAC'), ordered = T))

ggplot(sm %>% filter(!Status == 'Normal'), aes(Pathology, freq, group=Endoscopy, fill=factor(Endoscopy))) + 
  facet_grid(~Status) + 
  geom_col(position = position_dodge()) + ylim(0,1) + labs(title='Pathology frequency per timepoint')
```

There's a significant difference in the numbers of NDBE or HGD samples between the groups at T1.
```{r, echo=F}
#pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['HGD',]), caption='HGD')
#pander(chisq.test(table(patient.info$Pathology, patient.info$Status)['NDBE',]), caption='NDBE')

# Most progressors are HGD at timepoint 1
sample.info %>% filter(Timepoint == 'T1') %>% group_by(Status, Pathology) %>% tally %>% spread(Pathology,n) %>% kable(caption='Number of samples by pathology at T1') %>% kable_styling(full_width = F)

```

# Quick QC

Purity, ploidy and goodnessoffit from single sample (not paired) ASCAT.

`r qcdata %>% group_by(SampleType) %>% tally %>% spread(SampleType,n) %>% kable(caption='Sample numbers') %>% kable_styling(full_width=F)`

```{r, echo=F, warning=F, message=F}

qcdata = left_join(qcdata, dplyr::select(sample.info,c('Samplename','Status','Timepoint','SCA.Ratio')), by='Samplename')

ggplot(qcdata, aes(Status, ASCAT.SCA.ratio, color=Status, group=Status)) + geom_boxplot() + facet_grid(~SampleType)

low = filter(qcdata, SampleType == 'BE' & ASCAT.SCA.ratio <= median(filter(qcdata, SampleType != 'BE')$ASCAT.SCA.ratio))

low %>% group_by(Status, Timepoint) %>% tally %>% spread(Timepoint,n) %>% kable(caption='BE Samples with an ASCAT SCA ratio below the median ratio of the blood/gastric normal samples') %>%
  kable_styling(full_width = F)

qcdata = qcdata %>% mutate(SampleType = case_when(grepl('Normal',SampleType) ~ 'Blood/Gastric Normal', TRUE ~ 'BE'),)

anno = qcdata %>% group_by(SampleType) %>% dplyr::summarise(Ratio = median(ASCAT.SCA.ratio) ) %>% rowwise %>% 
  mutate(label = paste0('median(',SampleType,')','=',round(Ratio,3)))

p = ggplot(qcdata %>% filter(grepl('Normal',SampleType)), aes(Purity,ASCAT.SCA.ratio)) + 
  ylim(0,0.02) +
  geom_point(color='steelblue' ) + labs(title='Blood/Gastric Normal ASCAT values')
p
ggsave('plots/snp/normal_sca_purity.png', p, height=4, width=4, units='in')

ggplot(qcdata %>% filter(!grepl('Normal',SampleType)), aes(Purity,ASCAT.SCA.ratio)) + 
  #ylim(0,0.5) +
  geom_point(color='steelblue' ) + 
  geom_point(data = qcdata %>% filter(!grepl('Normal',SampleType) & ASCAT.SCA.ratio < 0.01 & Purity < 0.98), color='lightgreen') +
  labs(title='BE biopsy ASCAT values')

# 
# ggplot(qcdata, aes(Status, ASCAT.SCA.ratio)) + ylim(0,0.05) + 
#   facet_grid(~SampleType, scales='free_x') +
#   geom_jitter(width=0.1, aes(color=Status)) + 
#   geom_hline(data=dplyr::rename(anno, Status=SampleType), aes(yintercept = Ratio, linetype=Status), color='grey42') +
#   geom_label(data=dplyr::rename(anno, Status=SampleType), aes(x='NP', y=Ratio, label=label), alpha=0.5)

ggplot(filter(qcdata, !is.na(Ploidy)), aes(factor(round(Ploidy)), ASCAT.SCA.ratio, color=Status)) + 
  facet_grid(~SampleType, scales = 'free_x', space='free') + geom_boxplot() + labs(x='Ploidy')

# ggplot(filter(qcdata,SampleType == 'BE'), aes(Timepoint, ASCAT.SCA.ratio, fill=Timepoint, group=Timepoint)) + 
#   geom_jitter(width=0.2) + geom_boxplot(alpha=0.5) + 
#   facet_wrap(~Status, scales='free_x', ncol=2) + 
#   labs(title='SCA Ratio by Timepoint') + theme(legend.position = 'none')

ggplot(qcdata, aes(round(Purity,2))) + geom_histogram() + 
  facet_wrap(~SampleType, scales='free', ncol=1) + labs(x='Purity', title='Purity, normal vs BE')

x = 0.01
y = round(mean(filter(qcdata, SampleType != 'BE')$Purity) - sd(filter(qcdata, SampleType != 'BE')$Purity),2)

p = ggplot(qcdata, aes(Purity, ASCAT.SCA.ratio)) + facet_wrap(~SampleType, nrow=1) + 
  geom_point(aes(color=ASCAT.SCA.ratio < x & Purity > y), show.legend = F) + 
  geom_vline(xintercept=y, color='grey') + 
  geom_hline(yintercept=x, color='grey') +
  labs(title='Purity~SCA.Ratio') + theme(text = element_text(size=12))
p
ggsave('plots/snp/purity_v_sca.png', plot=p, height=4, width=4, units = 'in', dpi=300)

ggplot(qcdata, aes(Purity, Goodnessoffit, color=Status)) + geom_point() + 
         facet_wrap(~SampleType, ncol=1) + labs(title='Purity~fit, normal vs BE') + geom_vline(xintercept=y, color='grey')

ggplot(subset(qcdata, SampleType == 'BE'), aes(Ploidy, ASCAT.SCA.ratio, color=Status)) + facet_grid(~Status) + geom_point() + 
  geom_hline(yintercept = x, color='grey39', linetype = 'dashed') + labs(title='Barretts biopsies')

## PER ENDOSCOPY
qcdata.endo = qcdata %>% group_by(PatientID, EndoID, Status, Timepoint, SampleType) %>% 
  dplyr::summarise_if(is.numeric, c('mean','max','min','sd')) %>% mutate_if(is.numeric, round, digits=3)

#qcdata %>% dplyr::group_by( Status, SampleType) %>% dplyr::summarise( median(ASCAT.SCA.ratio), sd(ASCAT.SCA.ratio) )
# lowsca = filter(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_mean < 0.01 & Purity_max > 0.98 )
# nrow(lowsca)
# nrow(filter(qcdata.endo, SampleType == 'BE' & ASCAT.SCA.ratio_mean >= 0.01 & Purity_max <= 0.98))

p = ggplot(qcdata.endo, aes(Purity_mean, ASCAT.SCA.ratio_mean)) + facet_grid(~SampleType) +
  geom_point(aes(color=ASCAT.SCA.ratio_mean < 0.01 & Purity_mean > 0.95 ), show.legend = F) + 
  labs(title='Merged biopsies per-timepoint', y='mean(ASCAT.SCA.ratio)', x='mean(ASCAT purity)')
p
ggsave('plots/snp/sca_purity_per_tp.png', p, width=4, height=4, units='in')

tmp = qcdata.endo %>% filter(!is.na(Ploidy_mean)) %>% mutate(Ploidy_mean = factor(round(Ploidy_mean)))
p = ggplot(tmp, aes(Ploidy_mean, ASCAT.SCA.ratio_mean, group=Ploidy_mean)) + 
  facet_grid(~SampleType) +
  geom_jitter(width=0.2, aes(color=Ploidy_mean)) + 
  geom_boxplot(outlier.colour = NA, alpha=0.5) + 
  geom_hline(yintercept = 0.01, color='grey39') + labs(title='Merged per endoscopy', x='Mean ploidy') + 
  theme(text=element_text(size=12), legend.position = 'none')
p
ggsave('plots/snp/sca_v_ploidy.png', plot=p, height=4, width=4, units = 'in', dpi=300)

```

# Load tiled data

I use the raw ASCAT total CN values, adjusted for the sample ploidy, to create the weighted average values in each 5Mb and arm bin of the genome. Note again that at this point the tiled values are per *endoscopy* not per sample.

```{r, warning=F, echo=F, message=F, fig.height=10, fig.width=8}
# Adjust for ploidy -- this was done in the processing script
if (file.exists('~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata')) {
  load('~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata', verbose=T) 
} else {
  ptdirs = list.dirs('~/Data/Reid_SNP/PerPatient', recursive = F)
  mergedSegs = NULL;  mergedArms = NULL
  length(ptdirs)

  for (pt in ptdirs) {
    print(pt)
    if (length(list.files(pt, '*wins_tiled.txt', full.names=T)) <= 0) {
      message(paste("No tiled files for",pt))
      next
    }

    segvals = read_tsv(list.files(pt, '*wins_tiled.txt', full.names=T), col_types = cols(.default = col_double(), 'X1'= col_character())) %>% dplyr::rename(Samplename = X1)
    armvals = read_tsv(list.files(pt, '*arms_tiled.txt', full.names=T), col_types = cols(.default = col_double(), 'X1'= col_character())) %>% dplyr::rename(Samplename = X1)
    #samplenames = segvals$Samplename

    segvals = segvals %>% set_rownames(.$Samplename) %>% dplyr::select(-Samplename) %>% as.matrix 
    armvals = armvals %>% set_rownames(.$Samplename) %>% dplyr::select(-Samplename) %>% as.matrix

    #segvals[is.na(segvals)] = mean(segvals,na.rm=T)
    #armvals[is.na(armvals)] = mean(armvals,na.rm=T)

    if (is.null(segvals) | is.null(armvals))
      stop(paste("Missing values in ", pt))

    if (is.null(mergedSegs)) {
      mergedSegs = segvals
      mergedArms = armvals
    } else {
      mergedSegs = rbind(mergedSegs, segvals)    
      mergedArms = rbind(mergedArms, armvals)    
    }
  }
  nrow(mergedSegs) == nrow(mergedArms)
  #setdiff(rownames(mergedSegs), rownames(mergedArms))
  dim(mergedSegs)

  save(mergedSegs, mergedArms, file='~/Data/Reid_SNP/PerPatient/tmp_seg_pt.Rdata')
}

rownames(mergedSegs) = sub('\\.LogR','', rownames(mergedSegs))
rownames(mergedArms) = sub('\\.LogR','', rownames(mergedArms))

dim(mergedSegs)
dim(mergedArms)

segmentVariance = apply(mergedSegs, 1, var)

nm.rows = grep('BLD|GASTRIC', rownames(mergedSegs))

lowSeg = mergedSegs[grep('^1001', rownames(mergedSegs), value=T),,drop=F]
p1 = ggplot(melt(lowSeg), aes(Var2, value)) + ylim(-1,1) + facet_grid(~Var1) + geom_point() + 
  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 1001 NP', y='5Mb bins')

lowSeg = mergedSegs[grep('^403', rownames(mergedSegs), value=T),,drop=F]
p2 = ggplot(melt(lowSeg), aes(Var2, value)) +  ylim(-1,1) + facet_grid(~Var1) + geom_point() + 
  theme(axis.text.x=element_blank() ) + labs(title='Low SCA samples', subtitle='patient 403 P', y='5Mb bins')

grid.arrange(p1,p2,ncol=1)

highSeg = mergedSegs[grep('^1114', rownames(mergedSegs), value=T),,drop=F]
p3 = ggplot(melt(highSeg), aes(Var2, value)) +  ylim(-1,1) + facet_grid(~Var1) + geom_point() + 
  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 1114 NP', y='5Mb bins')

highSeg = mergedSegs[grep('^512', rownames(mergedSegs), value=T),,drop=F]
p4 = ggplot(melt(highSeg), aes(Var2, value)) +  ylim(-1,1) + facet_grid(~Var1) + geom_point() + 
  theme(axis.text.x=element_blank() ) + labs(title='Mixed SCA samples', subtitle='patient 512 P', y='5Mb bins')

grid.arrange(p1,p2,p3,p4,ncol=1)

```

## Compare the binned values to the median LRR from ASCAT

Using a random normal sample and a known BE progresser sample

```{r rnormals, echo=F, warning=F, message=F, fig.width=10, fig.height=8, eval=T}

adjust.cols<-function(df, means, sds) {
  nms = colnames(df)

  for (n in nms)
    df[,n] = BarrettsProgressionRisk:::unit.var(df[,n]+0.90, means[[n]], sds[[n]])

  return(df)  
}

randNormal = sample( grep('BLD',rownames(mergedSegs)[nm.rows], value=T), 1)
pt = unlist(strsplit(randNormal, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*', randNormal),sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

lims = range(segs$medLRR, mergedSegs[randNormal,])

grid.arrange(
  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) +
    theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt),

  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randNormal),])), aes(Var1, value)) + geom_point() +
        theme(axis.text.x=element_blank()) + labs(title='Ploidy adjusted CN, binned'),

 top='Random BLD or gastric normal', bottom=randNormal)

#randBE = '512_133R'
randBE = '1006_190V'
pt = unlist(strsplit(randBE, '_'))[1]
segs = subset(segments.list[[pt]], chr %in% c(1:22) & grepl(sub('_', '.*',randBE), sample))
segs$chr = factor(segs$chr, levels=c(1:22), ordered = T)

lims = range(segs$medLRR, mergedSegs[sub('\\.LogR','',randBE),])
grid.arrange(
  ggplot(segs) + facet_grid(~chr, space='free_x', scales='free')  +
    geom_segment(aes(x=startpos, xend=endpos, y=medLRR, yend=medLRR), size=3) +
    theme(axis.text.x=element_blank() ) + labs(title='Median LRR, ASCAT segments', subtitle=pt),

  ggplot(melt(as.matrix(mergedSegs[sub('\\.LogR','',randBE),])), aes(Var1, value)) + geom_point() +
    theme(axis.text.x=element_blank()) + labs(title='Ploidy adjusted CN, binned'),

top='BE', bottom=randBE)

```

# Set up values to predict

## Adjust per column

The ploidy adjusted values now need to be mean normalized to match the means of the training data. The training data mean is equivalent to the diploid state and the same is true of the adjusted CN data.
So the mean normalization is simply adding the difference of the means to the SNP bins.
```{r}
training = do.call(bind_rows, purrr::map(list.files('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb', '5e06_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
}))
training.segs = as.matrix(training[,-1]); rownames(training.segs) = training$X1

training = do.call(bind_rows, purrr::map(list.files('~/Data/BarrettsProgressionRisk/Analysis/pcf_perPatient/50kb', 'arms_tiled_segvals.txt', full.names = T, recursive = T), function(f) {
  read_tsv(f, col_types = cols(.default=col_double(), 'X1'=col_character()))
}))
training.arms = as.matrix(training[,-1]); rownames(training.arms) = training$X1
```

```{r raw.adjust, fig.height=16, fig.width=4, echo=T}
median(training.segs)
median(mergedSegs)

median(training.arms)
median(mergedArms)

# difference between the two is 0.98 so...
mergedSegs = mergedSegs + 0.98
mergedArms = mergedArms + 0.98

# Also, for the z-adjust values get a mean+sd from both the training and SNP cohort jointly, due to the difference in the variance
merged = rbind(mergedSegs, training.segs)
merged.mean = apply(merged, 2, mean, na.rm=T)
merged.sd = apply(merged, 2, sd, na.rm=T)

for (i in 1:ncol(mergedSegs))
 mergedSegs[,i] = BarrettsProgressionRisk:::unit.var(mergedSegs[,i], merged.mean[i], merged.sd[i])

merged.arms = rbind(mergedArms, training.arms)
merged.arms.mean = apply(merged.arms, 2, mean, na.rm=T)
merged.arms.sd = apply(merged.arms, 2, sd, na.rm=T)

for (i in 1:ncol(mergedArms))
  mergedArms[,i] = BarrettsProgressionRisk:::unit.var(mergedArms[,i], merged.arms.mean[i], merged.arms.sd[i])

#normal.ids = sample.info %>% filter(grepl('BLD|gastric', Level, ignore.case=T)) %>% dplyr::summarise(rows = paste0(PatientID, '_', EndoID)) %>% pull
#nm.rows = which(rownames(mergedSegs) %in% normal.ids)
```

```{r echo=F,message=F, fig.width=10, fig.height=8}
cx = BarrettsProgressionRisk:::scoreCX(mergedSegs, 1)

arrayDf = BarrettsProgressionRisk:::subtractArms(mergedSegs, mergedArms)
arrayDf = cbind(arrayDf, 'cx'=BarrettsProgressionRisk:::unit.var(cx, mn.cx, sd.cx))

melted = melt(arrayDf) %>% left_join(patient.info %>% dplyr::select(UniqueSampleID, Status), by=c('Var1'='UniqueSampleID'))

train.melt = melt(dysplasia.df) %>% 
  left_join(as_tibble(train.labels) %>% mutate(label = names(train.labels)), by=c('Var1'='label')) %>%
  mutate(value.y = as.factor(value.y))

lims=range(arrayDf, dysplasia.df,na.rm=T)
grid.arrange(
  ggplot(melted, aes(Var2, value,color=Status)) + geom_point() + ylim(lims) +
    geom_point(data=melt(arrayDf[,589:634]), col='red') + labs(title='SNP'),
  ggplot() + geom_point(data=train.melt, aes(Var2, value.x, color=value.y)) + ylim(lims) +
      geom_point(data=melt(dysplasia.df[,589:634]), aes(Var2, value), col='red') + labs(title='sWGS'),
nrow=1)

```

The SNP and sWGS data is comparable, but it's also clear that the variance is fairly different.  Given our experience with other data this is expected.

# Predictions

For simplicity here I'm not bothering with the 'moderate' category, so the cutoff for high ~ low is a probability of  low<0.5>=high

It's worth noting here that the relative risks are much higher, again I'd attribute this to the variance.

```{r, warning=F, fig.height=8}
prob = predict(fitV, newx=arrayDf, s=l, type='response')
rr = predict(fitV, newx=arrayDf, s=l, type='link')

preds = bind_cols('Samplename'=rownames(prob),'Prob'=prob[,1],'RR'=rr[,1]) %>%
  mutate(Risk = case_when(
    Prob < 0.5 ~ 'Low',
    Prob >= 0.5 ~ 'High' 
))

normals = filter(preds, grepl('BLD|gastric', Samplename, ignore.case=T))

preds = filter(preds, !grepl('BLD|gastric', Samplename, ignore.case=T))

patient.info = patient.info %>% mutate(Pathology = recode(Pathology, 'IMC' = 'EAC'))

preds = preds %>% left_join(patient.info %>% dplyr::select(PatientID, UniqueSampleID, Pathology, Status, Path.Status), by=c('Samplename'='UniqueSampleID'))

# NP patients who are likely P
np.wgd.pid = qcdata.endo %>% filter(Ploidy_mean > 2.7 & Status == 'NP') %>% ungroup %>% dplyr::select(PatientID) %>% pull
preds = preds %>% dplyr::mutate(Path.Status = as.character(ifelse(PatientID %in% np.wgd.pid, 'P', Path.Status)))

preds %>% group_by(Status, Risk) %>% tally %>% spread(Risk,n)

preds %>% group_by(Path.Status, Risk) %>% tally %>% spread(Risk,n)

myPal = rev(RColorBrewer::brewer.pal(11, 'RdYlBu'))
ggplot(bind_rows(preds,normals), aes(RR)) + geom_histogram(aes(fill=..x..), bins=15, show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(y='n Samples', x='Relative Risk', title='Unadjusted relative risk', subtitle='All SNP samples')

ggplot(bind_rows(preds,normals), aes(Prob)) + geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions, all samples model', y='n Samples', x='Unadjusted Probability', subtitle='All SNP samples')
```

The best false positive rate can be assessed from the blood/gastric normals.  So in this case ~17%
`r normals %>% group_by(Risk) %>% tally %>% mutate(ratio = round(n/nrow(normals),2))  %>% kable(caption='Blood & gastric normals') %>% kable_styling(full_width = F)`

We see a higher FPR rate in the full dataset. 
```{r echo=F}
preds %>% group_by(Status, Risk) %>% tally %>% spread(Risk, n) %>%
  left_join(preds %>% group_by(Status) %>% tally, by='Status') %>%
  mutate_if(is.numeric, list( ~round(./n,2))) %>% 
  dplyr::select(-n) %>% kable(caption='Ratio of samples per risk category (excluding normals)') %>% 
  kable_styling(full_width = F)
```

Despite that, we still see a fairly consistent high/low risk prediction for NP vs P.  However, HGD in NP patients predicts at a rate consistent with progression.
```{r echo=F}
preds = preds %>% mutate(Pathology = factor(Pathology, levels=c('NDBE','ID','LGD','HGD','EAC'), ordered = T))
rp = plot.risk.by.path(preds) 
rp

ggsave('plots/snp/path_risk.png', plot=rp, width=7, height=6, units='in', dpi=300)
```

### ROC 
```{r}
# roc = pROC::roc(Status ~ Prob, data=bind_rows(preds, normals %>% mutate(Status = 'NP')), auc=T, ci=T, of='thresholds',transpose=T)
# roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples')
# pROC::ci.auc(roc$auc)
# pROC::coords(roc, 'best', transpose=F)
# ggsave(filename = 'plots/snp/all_samp_roc.png', plot = roc.plot(roc) + labs(title='SNP ROC', subtitle='All samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

rocBE = pROC::roc(Status ~ Prob, data=preds, auc=T, ci=T, of='thresholds',transpose=T)
roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies')
pROC::ci.auc(rocBE$auc)
pROC::coords(rocBE, 'best', transpose=F)
ggsave(filename = 'plots/snp/be_roc.png', plot = roc.plot(rocBE) + labs(title='SNP ROC', subtitle='BE biopsies') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)
```

A quick look at the two random samples from earlier.  
`r filter(normals, Samplename == randNormal) %>% kable(caption='Random normal') %>% kable_styling(full_width=F)`

The random BE should show a very high probability, and it does...
`r filter(preds, Samplename == randBE) %>% dplyr::select(Samplename, Prob, RR, Risk) %>% kable(caption='Random BE') %>% kable_styling(full_width=F)`

## Evaluating predictions

```{r, warning=F, fig.height=8}
preds = preds %>% dplyr::group_by(Samplename) %>% dplyr::mutate(
  PatientID = unlist(strsplit(Samplename,'_'))[1],
  EndoID = unlist(strsplit(Samplename,'_'))[2]
)

pred.qc = left_join(preds, qcdata.endo %>% dplyr::select(-SampleType), by=c('PatientID', 'EndoID', 'Status'))

grid.arrange(
  ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Prob, color=Status)) + geom_point(),
  ggplot(pred.qc, aes(Ploidy_max, Prob, color=Status)) + geom_point(),
ncol=2)

ct = cor.test(pred.qc$ASCAT.SCA.ratio_mean, pred.qc$Prob)
```

The ASCAT mean SCA ratio across the genome correlates with the predicted probabilites in the BE samples, but not incredibly highly: R=`r round(ct$estimate,2)`,

```{r, warning=F, fig.height=8}
prog = filter(pred.qc, Status == 'P')
p1 = ggplot(prog, aes(Prob)) + facet_wrap(~Timepoint, scales='free_y',nrow=1) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on progressors', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))

nonprog = subset(pred.qc, Status == 'NP')
p2 = ggplot(nonprog, aes(Prob)) + facet_wrap(~Timepoint, scales='free_y', nrow=1) + 
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') + 
  labs(title='Predictions on non-progressors', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))

normals = normals %>% dplyr::mutate(Type = ifelse(grepl('BLD',Samplename), 'Blood','Gastric'))

p3 = ggplot(normals, aes(Prob)) + facet_wrap(~Type, ncol=2, scales = 'free') +
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') +
  labs(title='Predictions on Blood/Gastric normals', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))

grid.arrange(p1,p2,p3,ncol=1)

ggsave('plots/snp/probability_dists.png', plot = grid.arrange(p1,p2,p3,ncol=1), width=6, height=10, units='in', dpi=300)
```

# Compare predictions to known status
```{r roc, echo=T, eval=T}
pred.qc = pred.qc %>% mutate(Status = factor(Status))

pred.qc = pred.qc %>% group_by(Status, Samplename) %>% 
  dplyr::mutate(
    hgd.NP = (Status == 'NP' & (length(which(Pathology == 'HGD')) > 0)),
    wgd.NP = (Status == 'NP' & Ploidy_max > 2.7 ),
    lowsca = (ASCAT.SCA.ratio_mean < 0.01 & Purity_mean > 0.95)
)

pred.qc %>% filter(!lowsca | wgd.NP) %>% ungroup %>% dplyr::select(PatientID) %>% distinct

ascat.normals = qcdata %>% filter(grepl('Normal',SampleType)) %>% data.frame
ggplot(ascat.normals, aes(ASCAT.SCA.ratio, Purity, color=Status)) + geom_point()

ggplot(pred.qc, aes(ASCAT.SCA.ratio_mean, Purity_mean, color=lowsca)) + geom_point() + facet_grid(~Status) + 
  geom_vline(xintercept=0.01) + geom_hline(yintercept = 0.95)

pred.qc %>% group_by(Status, lowsca) %>% tally(name='n Samples') %>% spread(Status, `n Samples`) %>% kable(caption = 'LOWSCA high purity samples') %>% kable_styling(full_width = F)

filter(pred.qc, !lowsca & !wgd.NP) %>% group_by(Status) %>% tally(name='n Samples') %>% kable(caption='Number of samples that pass the low SCA:high purity & NP (wgd) thresholds') %>% kable_styling(full_width = F)
```

### ROC after exclusions

In this case the status has changed slightly. All patients that had a HGD/IMC path report are considered as progressors to match the sWGS cohort.  After that I include exclusions for low SCA and WGD in NP.

```{r echo=F}
# Set Path.Status = P for all patient where a sample was WGD
#wgd.ids = filter(pred.qc, wgd.NP) %>% ungroup %>% dplyr::select(PatientID) %>% distinct %>% pull 
#pred.qc = pred.qc %>% mutate(wgd.NP = PatientID %in% wgd.ids) %>% mutate(Path.Status = ifelse(wgd.NP, 'P', Path.Status) )

rocEX = pROC::roc( Path.Status ~ Prob, data=pred.qc, auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)')
pROC::ci.auc(rocEX)
ggsave(filename='plots/snp/path_status_roc.png', plot = roc.plot(rocEX) + labs(title='BE biopsies', subtitle='Path status (HGD/IMC)') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

notlow = filter(pred.qc, !lowsca)
pid = notlow %>% ungroup %>%dplyr::select(PatientID) %>% distinct %>% pull 
length(pid)
qcdata %>% filter(PatientID %in% pid & SampleType != 'Blood/Gastric Normal')

# Exclude the low SCA patients
rocpp = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples')
ggsave(filename='plots/snp/auc_ex_lowsca_roc.png', plot = roc.plot(rocpp) + labs(title='BE biopsies', subtitle='excluding low SCA samples') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Exclude all of the above
rocexall = pROC::roc(Status ~ Prob, data=filter(pred.qc, !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP')
ggsave(filename='plots/snp/auc_ex_lowsca_wgs_roc.png', plot =roc.plot(rocexall) + labs(title='BE biopsies', subtitle='excluding low SCA & WGD NP') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Timepoint 1 only
roct1 = pROC::roc(Status ~ Prob, data=filter(pred.qc, Timepoint == 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1')
ggsave(filename='plots/snp/auc_t1_roc.png', plot=roc.plot(roct1) + labs(title='BE biopsies', subtitle='Timepoint T1') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

# Timepoint 2 only
roct2 = pROC::roc(Status ~ Prob, data=subset(pred.qc, Timepoint != 'T1' & !lowsca & !wgd.NP), auc=T, ci=T, of='thresholds', transpose=T)
roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2')
ggsave(filename='plots/snp/auc_t2_roc.png', plot=roc.plot(roct2) + labs(title='BE biopsies', subtitle='Timepoint T2') + theme(text=element_text(size=12)), width=4, height=4, units='in', dpi=300)

ggplot(filter(pred.qc, !wgd.NP & !lowsca), aes(Prob)) +
  geom_histogram(aes(fill=..x..), breaks=seq(0,1,0.1) , show.legend = F) +
  scale_fill_gradientn(colors = myPal,  name='') +
  labs(title='Excluding WGD NP & low SCA', y='n Samples', x='Unadjusted Probability') + theme(text=element_text(size=12))

ct = cor.test(pred.qc$Prob, pred.qc$SCA.Ratio_mean)
ct2 = cor.test(filter(pred.qc, !wgd.NP & !hgd.NP & !lowsca)$Prob, filter(pred.qc, !wgd.NP & !hgd.NP & !lowsca)$SCA.Ratio_mean)
```

The risk probability is correlated with the % of the genome that's been altered as identified by Xiahong, at about the same rate as the correlation between ASCAT and the probabilites: R=`r ct$estimate`. And that's true even when patients are excluded, though at a slightly higher rate (R=`r ct2$estimate`).

```{r regression, echo=F, message=F, eval=F}
# Regression using SNP Data
folds = 10; splits = 5
## ----- All ----- ##
cache.dir = '~/Data/BarrettsProgressionRisk//Analysis/SNP/no-low/'
file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
if (file.exists(file)) {
 message(paste("loading", file))
 load(file, verbose=T)
 model.performance(performance.at.1se, coefs, folds, splits)$plot + labs(title='SNP CV models')
} else {
  warning("Run 'snp-regression.R' first")
}
#save(arrayDf,labels,patient.info, file=paste(cache.dir,'snp-models.Rdata',sep='/'))

#do.call(grid.arrange, c(plots))
```

```{r loo, eval=F}
file = paste(cache.dir, 'loo.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
} else {
  snp.preds = do.call(bind_rows, purrr::map(list.files(cache.dir, 'loo.Rdata', recursive = T, full.names = T), function(f) {
    load(f)
    as_tibble(preds, rownames = 'UniqueID')
  }))

  #stop("Run snp-regression.R first")
}

snp.preds = left_join(snp.preds, dplyr::select(slabels, c(UniqueID, Status)), by='UniqueID') %>% filter(!is.na(Prob))

snp.roc = pROC::roc(Status~Prob,snp.preds, auc=T, ci=T, of='thresholds')
roc.plot(snp.roc, title="SNP LOO Model ROC")

ggplot(cbind.data.frame('performance'=performance.at.1se, 'type'='snp'), aes(y=performance, x=type, group=type,fill=type)) + 
  geom_boxplot(outlier.colour = NA, show.legend = F) + geom_jitter(width=0.2) + geom_label(data=cbind.data.frame('performance'=range(performance.at.1se), 'type'='snp'), aes(label=round(performance,2)), nudge_x=0.1) + theme_bw() + theme(legend.position = 'none') + labs(x='', y='Classification', title='LOO Performance SNP')

file = paste(cache.dir, 'all.pt.alpha.Rdata', sep='/')
load(file, verbose=T)

fit.snp = models$`1`

probs = predict(fit.snp, newx = dysplasia.df, s = performance.at.1se$`1`$lambda,type='response' )

swgs.pred.snp = cbind.data.frame('Prob'=probs[,1], 'Status'=swgs_labels[rownames(probs)])
ssroc = pROC::roc(Status~Prob, data=swgs.pred.snp, ci=T, auc=T, of='thresholds')
roc.plot(ssroc, title='sWGS prediction on SNP model')
```