### 2020-04-14 sWGS Normals.Rmd

---
title: "sWGS Normals"
author: "Sarah Killcoyne"
date: "7/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(BarrettsProgressionRisk)
library(glmnet)
library(gridExtra)
library(kableExtra)
source('lib/load_patient_metadata.R')
#source('lib/data_func.R')
#source('lib/fastPCF.R')

data.dir = '~/Data/BarrettsProgressionRisk/QDNAseq'
analysis.dir = '~/Data/BarrettsProgressionRisk/Analysis'
#data.files = list.files(, full.names = T)

modeldir = paste0(analysis.dir,'/models_5e6_all/50kb')

patient.file = list.files(data.dir, pattern='All_patient_info.xlsx', recursive=T, full.names=T)
demo.file = list.files(data.dir, pattern='Demographics_full.xlsx', recursive=T, full.names=T)

rpi = read.patient.info(patient.file, demo.file)
normals = rpi$normal %>% mutate(Samplename = sub('_','-',Samplename))

data.files = list.files(paste0(analysis.dir, '/normals'), 'segment.Rdata', recursive = T, full.names = T)

resids = do.call(bind_rows, purrr::map(list.files(paste0(analysis.dir, '/normals'), 'residuals.tsv', recursive = T, full.names = T), function(f) read_tsv(f)))
resids = left_join( dplyr::select(normals, Patient, Hospital.Research.ID, Samplename, Plate.Index), resids, by=c('Samplename'='samplename'))

failedQC = filter(resids, !Pass)

normals %>% dplyr::select(-matches('Cellularity|p53|Reads|Batch|Set')) %>% kable(caption = 'Normal samples') %>% kable_styling(full_width = F)
```

`r nrow(failedQC)` samples failed the variant residual QC based on a cutoff determined by the discovery cohort.

```{r, echo=F, warning=F, fig.height=32, fig.width=12}
plist = lapply(data.files, function(f) {
  load(f)
  BarrettsProgressionRisk::plotSegmentData(segmented)
})
do.call(grid.arrange, c(plist, ncol=1))
```

## Predictions

Using the model trained on the progressors and non-progressors in the initial cohort the normal samples can be predicted as well.

```{r, echo=F}
file = paste(modeldir, 'model_data.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
file = paste(modeldir, 'all.pt.alpha.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
fit = models[['0.9']]
lambda = performance.at.1se[['0.9']]$lambda  
cvRR = BarrettsProgressionRisk:::cvRR(dysplasia.df, coefs[['0.9']])
rm(dysplasia.df,coefs,plots,labels)
file = paste(modeldir, 'loo_0.9.Rdata', sep='/')
if (!file.exists(file))
  stop(paste("Missing data file", file))
load(file, verbose=F)
rm(plots, performance.at.1se, coefs, fits, pg.samp)

be.model = BarrettsProgressionRisk:::be.model.fit(fit, lambda, 50, 5e6, z.mean, z.arms.mean, z.sd, z.arms.sd, mn.cx, sd.cx, nzcoefs, cvRR)

preds = tibble()
plist = list()
for (f in data.files) { 
  load(f)

  tiles = BarrettsProgressionRisk:::tileSamples(segmented, be.model = be.model, scale=T, MARGIN=2, verbose=F)

  prr = BarrettsProgressionRisk:::predictRisk(segmented, tiles, be.model, verbose=F)

  plist[f] = grid.arrange(BarrettsProgressionRisk::copyNumberMountainPlot(prr))

  preds = bind_rows(preds, predictions(prr))
}

paste0(analysis.dir, '/normals')
files = list.files(paste0(analysis.dir, '/normals'), '5e06_.*tiled', recursive=T, full.names = T)
tiled.segs = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

files = list.files(paste0(analysis.dir, '/normals'), 'arms_.*tiled', recursive=T, full.names = T)
tiled.arms = do.call(bind_rows, purrr::map(files, function(f) {
  read_tsv(f,col_types=c(.default=col_double()))
})) %>% dplyr::rename('Sample' = 'X1')

tiles.segs[,-1]

cx.score = BarrettsProgressionRisk::scoreCX(tiled.segs,1)

for (i in 1:ncol(tiled.segs)) 
  tiled.segs[,i] = BarrettsProgressionRisk:::unit.var(tiled.segs[,i], z.mean[i], z.sd[i])
for (i in 1:ncol(tiled.arms)) 
  tiled.arms[,i] = BarrettsProgressionRisk:::unit.var(tiled.arms[,i], z.arms.mean[i], z.arms.sd[i])

mergedDf = subtract.arms(tiled.segs, tiled.arms)
mergedDf = cbind(mergedDf, 'cx' = unit.var(cx.score, mn.cx, sd.cx))

sparsed_test_data <- Matrix(data=0, nrow=nrow(mergedDf),  ncol=ncol(mergedDf),
                            dimnames=list(rownames(mergedDf),colnames(mergedDf)), sparse=T)
for(col in colnames(mergedDf)) sparsed_test_data[,col] = mergedDf[,col]

preds = predict(fitV, newx=sparsed_test_data, s=lambda.opt, type='response')

high = preds[preds > 0.7,,drop=F]

pander(subset(normals, Samplename %in% rownames(high), select=c('Patient','Hospital.Research.ID','Status','Endoscopy.Year','Pathology notes')), caption='Normal samples predicted to be high risk')

```